### 2011-11-07 paper.tex
Converted the original wiki page as a starting point
While here:
- added a gitignore
- updated README

%\documentclass[twoside,a4paper]{report}
\documentclass[twocolumn,a4paper]{article}

\usepackage[utf8]{inputenc}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{pdflscape} %% Used for very big table
\usepackage{moreverb}
\usepackage{color}
\usepackage{listings}
\usepackage{fancyhdr}

\definecolor{codegray}{gray}{.95}
\lstset{language=C,
	numbers=left,
	tabsize=2,
	basicstyle=\scriptsize,
	stringstyle=\textrm,
	showstringspaces=false,
	frame=none,
	xleftmargin=3pt,
	backgroundcolor=\color{codegray}}

\hypersetup{
pdftitle = {Survey on Knowledge Representation Systems for Robotics},
pdfauthor = {Séverin Lemaignan, Moritz Tenorth},
pdfkeywords = {cognitive robotics, knowledge representation},
}
\title{Survey on Knowledge Representation Systems for Robotics}
\author{Séverin Lemaignan, Moritz Tenorth}

\begin{document}

\maketitle
\tableofcontents

\section{Which systems do we survey?}

Systems that:

\begin{itemize}
	\item  Run on \emph{service robot} (robots that interact with objects in a semantic-rich environment),
	\item  Ground their knowledge in the physical world (physically embedded),
	\begin{itemize}
		\item  Able to \emph{resolve entities}
		\item  Able to automatically create new object instances
	\end{itemize}

	\item  Are able to merge different knowledge modalities,
	\item  Are endowed with on-line, dynamic reasoning (not just a static dictionary)
\end{itemize}

\section{Criteria, capabilities that are compared}

\subsection{Intrinsic features}

\begin{itemize}
	\item  Expressiveness
	\begin{itemize}
		\item  Which logic formalism (DL, 2nd order\ldots{})
		\item  OWA/CWA
		\item  Representation of uncertainty
		\item  (non) monotonic reasoning
		\item  General knowledge + exception to this knowledge (blue sky/white sky)
		\item  Microtheories
		\item  Lazy evaluation
		\item  Presupposition accomodation (ability to represent facts that are only verbally hinted, but not grounded into perception)
	\end{itemize}

	\item  Explicit modeling of context of knowledge / domain of validity
	\item  Representation of change (\emph{"The pancake dough disappears into a pancake"})
	\item  Support for learning
	\item  Support for introspection?
	\item  Support for prediction/projection?
\end{itemize}

\subsection{Integration in robotic architecture}

\begin{itemize}
	\item  Knowledge acquisition possible\ldots{}
	\begin{itemize}
		\item  \ldots{}through interaction? How?
		\item  \ldots{}through the Web? How?
		\item  \ldots{}through learning? How?
		\item  \ldots{}through perception? How?
	\end{itemize}

	\item  Integration with supervision
	\begin{itemize}
		\item  Events
		\item  Relation to task planning
	\end{itemize}

	\item  Scalability and responsiveness
\end{itemize}

\subsection{Knowledge model}

\begin{itemize}
	\item  Which underlying knowledge (\emph{common-sense}, \emph{upper knowledge}\ldots{})
	\begin{itemize}
		\item  top-down approach?
	\end{itemize}

\end{itemize}

\section{Major players}

\begin{itemize}
	\item  Saffiotti/Coradeschi
	\begin{itemize}
		\item  Grounding, anchoring
		\item  \emph{PEIS Ecology} (more "SmartHome")
	\end{itemize}

	\item  Tom Kollar/Stefanie Tellex
	\begin{itemize}
		\item  Dialog grounding
	\end{itemize}

	\item  OMKRF
	\item  Deb Roy/Nikolaos Mavridis
	\item  Wrighteagle RoboCup team
	\item  Markus Vincze (Vienna)
	\item  Bielefeld
	\item  Karlsruhe
	\begin{itemize}
		\item  ARMAR
		\item  Sven Schmidt-Rohr
	\end{itemize}

	\item  Saarbrücken (SemanticMap)
	\item  Joachim Hertzberg (Osnabrueck)
	\item  DFKI Bremen
	\begin{itemize}
		\item  more "SmartHome"
	\end{itemize}

	\item  JSK
	\begin{itemize}
		\item  KnowRob + physics simulator + planning
	\end{itemize}

	\item  Frederik Heitnz, Patrick Dowerty
	\begin{itemize}
		\item  \emph{DY-KNOW} (stream-based knowledge bases)
	\end{itemize}

\end{itemize}
\end{document}

### 2011-11-07 paper.tex
Added some empty sections + 'main players' list -> table

@@
 %\documentclass[twoside,a4paper]{report}
-\documentclass[twocolumn,a4paper]{article}
+\documentclass[a4paper]{article}
 
@@
 \usepackage{moreverb}
-\usepackage{color}
+\usepackage[table]{xcolor}
 \usepackage{listings}
 \usepackage{fancyhdr}
+\usepackage{fixme}
 
 \definecolor{codegray}{gray}{.95}
+\definecolor{palegray}{gray}{.65}
 \lstset{language=C,
@@
 
-\hypersetup{
-pdftitle = {Survey on Knowledge Representation Systems for Robotics},
-pdfauthor = {Séverin Lemaignan, Moritz Tenorth},
-pdfkeywords = {cognitive robotics, knowledge representation},
-}
-\title{Survey on Knowledge Representation Systems for Robotics}
+\title{Current State of Knowledge Representation Systems for Robotics}
 \author{Séverin Lemaignan, Moritz Tenorth}
@@
 
-\section{Which systems do we survey?}
+\section{Introduction}
+\label{sect|intro}
 
-Systems that:
+\subsection{Definition of inclusion criteria}
+
+We have setup the following criteria to select the knowledge representation
+systems to include in this survey.
+
+Those systems must:
 
 \begin{itemize}
-	\item  Run on \emph{service robot} (robots that interact with objects in a semantic-rich environment),
+	\item  Run on \emph{service robot} (robots that interact with objects in a
+	semantic-rich environment),
+
 	\item  Ground their knowledge in the physical world (physically embedded),
@@
 
-	\item  Are able to merge different knowledge modalities,
-	\item  Are endowed with on-line, dynamic reasoning (not just a static dictionary)
+	\item  Be able to merge different knowledge modalities,
+	\item  Be endowed with on-line, dynamic reasoning (not just a static
+	dictionary)
+
 \end{itemize}
 
-\section{Criteria, capabilities that are compared}
+\subsection{Methodology}
+\label{sect|methodology}
+
+\fxnote{Mention here that each team was proposed to complete/amend the description of their
+system}
+
+\section{Comparison criteria}
+\label{sect|comparison-criteria}
 
 \subsection{Intrinsic features}
+\label{sect|intrinsic-features}
 
@@
 
-\subsection{Integration in robotic architecture}
+\subsection{Integration in a larger robotic architecture}
 
@@
 
-\subsection{Knowledge model}
+\subsection{Underlying knowledge model}
 
@@
 
-\section{Major players}
+\section{Surveyed systems}
 
-\begin{itemize}
-	\item  Saffiotti/Coradeschi
-	\begin{itemize}
-		\item  Grounding, anchoring
-		\item  \emph{PEIS Ecology} (more "SmartHome")
-	\end{itemize}
+Table \ref{table|surveyed-systems} presents the fifteen knowledge representation systems surveyed
+in this article.
 
-	\item  Tom Kollar/Stefanie Tellex
-	\begin{itemize}
-		\item  Dialog grounding
-	\end{itemize}
+This section briefly presents each of them.
 
-	\item  OMKRF
-	\item  Deb Roy/Nikolaos Mavridis
-	\item  Wrighteagle RoboCup team
-	\item  Markus Vincze (Vienna)
-	\item  Bielefeld
-	\item  Karlsruhe
-	\begin{itemize}
-		\item  ARMAR
-		\item  Sven Schmidt-Rohr
-	\end{itemize}
+\begin{landscape}
+\begin{table}
+\begin{center}
+\rowcolors{2}{lightgray}{codegray}
 
-	\item  Saarbrücken (SemanticMap)
-	\item  Joachim Hertzberg (Osnabrueck)
-	\item  DFKI Bremen
-	\begin{itemize}
-		\item  more "SmartHome"
-	\end{itemize}
+\begin{tabularx}{\textwidth}{llp{2cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
+\hiderowcolors
+{\bf Name} & {\bf Authors (Institution)} & {\bf Surveyed version} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} \\
+\hline
+\showrowcolors
+{\sc KnowRob} & Tenorth (TU Munich) & & & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) \\
+ORO & Lemaignan (LAAS-CNRS) & 0.8.0 & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} \\
+PEIS Ecology & Saffiotti, Coradeschi \\
+ & Kollar, Tellex \\
+OMKRF \\
+ & Roy, Mavridis \\
+ & Wrighteagle \\
+ & Vincze \\
+ & (Bielfeld) \\
+ARMAR & Schmidt-Rohr (Karlsruhe) \\
+ & (Saarbrücken) \\
+ & Hertzberg (Osnabrück) \\
+ & (DFKI Bremen) \\
+ (based on {\sc KnowRob} & (JSK) \\
+DY-KNOW & Heitnz, Dowerty \\
 
-	\item  JSK
-	\begin{itemize}
-		\item  KnowRob + physics simulator + planning
-	\end{itemize}
+\hline
 
-	\item  Frederik Heitnz, Patrick Dowerty
-	\begin{itemize}
-		\item  \emph{DY-KNOW} (stream-based knowledge bases)
-	\end{itemize}
+\end{tabularx}
+\end{center}
+\caption{List of surveyed systems}
+\label{table|surveyed-systems}
+\end{table}
+\end{landscape}
 
-\end{itemize}
+%\begin{itemize}
+%	\item  Saffiotti/Coradeschi
+%	\begin{itemize}
+%		\item  Grounding, anchoring
+%		\item  \emph{PEIS Ecology} (more "SmartHome")
+%	\end{itemize}
+%
+%	\item  Tom Kollar/Stefanie Tellex
+%	\begin{itemize}
+%		\item  Dialog grounding
+%	\end{itemize}
+%
+%	\item  OMKRF
+%	\item  Deb Roy/Nikolaos Mavridis
+%	\item  Wrighteagle RoboCup team
+%	\item  Markus Vincze (Vienna)
+%	\item  Bielefeld
+%	\item  Karlsruhe
+%	\begin{itemize}
+%		\item  ARMAR
+%		\item  Sven Schmidt-Rohr
+%	\end{itemize}
+%
+%	\item  Saarbrücken (SemanticMap)
+%	\item  Joachim Hertzberg (Osnabrueck)
+%	\item  DFKI Bremen
+%	\begin{itemize}
+%		\item  more "SmartHome"
+%	\end{itemize}
+%
+%	\item  JSK
+%	\begin{itemize}
+%		\item  KnowRob + physics simulator + planning
+%	\end{itemize}
+%
+%	\item  Frederik Heitnz, Patrick Dowerty
+%	\begin{itemize}
+%		\item  \emph{DY-KNOW} (stream-based knowledge bases)
+%	\end{itemize}
+%
+%\end{itemize}
+%
+
+\section{Conclusion}
+\label{sect|conclusion}
+
+\subsection{Main groups}
+
+\subsection{What is not successfully covered by current systems}
+
+\subsection{Future challenges}
+\label{sect|future-challenges}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\section*{Acknowledgements}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bibliographystyle{spbasic}
+\bibliography{biblio}
+
 \end{document}

### 2011-11-21 paper.tex
Added notes on PEIS Ecology. Prbl: no KRS...?

@@
 ORO & Lemaignan (LAAS-CNRS) & 0.8.0 & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} \\
-PEIS Ecology & Saffiotti, Coradeschi \\
+PEIS Ecology & Saffiotti (Örebro Univ.) & & \url{www.aass.oru.se/~peis} & {\sc C}\\
  & Kollar, Tellex \\
@@
 \end{landscape}
+
+\subsection{PEIS Ecology}
+\label{sect|peis-ecology}
+
+{\sc PEIS Ecology} is a software \emph{ecosystem} that aim to binds autonomous
+robotics with ambient intelligence (network of sensors). \emph{PEIS} stands for
+\emph{Physically Embedded Intelligent System}: every robots or intelligent
+device in the environment is abstracted as a PEIS.
+
+Each PEIS component is running a \emph{PEIS Kernel} instance. Communication between instance rely on a custom P2P communication protocol.
+
+\fixme{Very good website, lot of doc, examples, but could not find something directly related to knowledge representation...}
 

### 2011-11-21 paper.tex
Fixed biblio
While here, imported my whole biblio. Easier for now
Started surveying PEIS Ecology.
Added the PEIS architecture diagram, from [Daoutis2009].

While here, added a table of knowledge sources per project

--- 2011-11-07 paper.tex
@@
 \usepackage{fancyhdr}
-\usepackage{fixme}
+\usepackage[draft]{fixme}
 
@@
 \author{Séverin Lemaignan, Moritz Tenorth}
+
+\graphicspath{{figs/}}
 
@@
 
-\begin{tabularx}{\textwidth}{llp{2cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
+%\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
+\begin{tabular}{lllllll}
 \hiderowcolors
-{\bf Name} & {\bf Authors (Institution)} & {\bf Surveyed version} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} \\
+{\bf Project} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} & Main reference \\
 \hline
 \showrowcolors
-{\sc KnowRob} & Tenorth (TU Munich) & & & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) \\
-ORO & Lemaignan (LAAS-CNRS) & 0.8.0 & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} \\
-PEIS Ecology & Saffiotti, Coradeschi \\
+{\sc KnowRob} & Tenorth (TU Munich) & & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) & \cite{Tenorth2009a} \\
+ORO & Lemaignan (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
+PEIS Ecology & Daoutis, Coradeshi, Loutfi, Saffiotti (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & OWL-Full, 2nd order logics & & \cite{Daoutis2009} \\
  & Kollar, Tellex \\
@@
 
-\end{tabularx}
+%\end{tabularx}
+\end{tabular}
 \end{center}
@@
 
-%\begin{itemize}
-%	\item  Saffiotti/Coradeschi
-%	\begin{itemize}
-%		\item  Grounding, anchoring
-%		\item  \emph{PEIS Ecology} (more "SmartHome")
-%	\end{itemize}
-%
-%	\item  Tom Kollar/Stefanie Tellex
-%	\begin{itemize}
-%		\item  Dialog grounding
-%	\end{itemize}
-%
-%	\item  OMKRF
-%	\item  Deb Roy/Nikolaos Mavridis
-%	\item  Wrighteagle RoboCup team
-%	\item  Markus Vincze (Vienna)
-%	\item  Bielefeld
-%	\item  Karlsruhe
-%	\begin{itemize}
-%		\item  ARMAR
-%		\item  Sven Schmidt-Rohr
-%	\end{itemize}
-%
-%	\item  Saarbrücken (SemanticMap)
-%	\item  Joachim Hertzberg (Osnabrueck)
-%	\item  DFKI Bremen
-%	\begin{itemize}
-%		\item  more "SmartHome"
-%	\end{itemize}
-%
-%	\item  JSK
-%	\begin{itemize}
-%		\item  KnowRob + physics simulator + planning
-%	\end{itemize}
-%
-%	\item  Frederik Heitnz, Patrick Dowerty
-%	\begin{itemize}
-%		\item  \emph{DY-KNOW} (stream-based knowledge bases)
-%	\end{itemize}
-%
-%\end{itemize}
-%
+\subsection{PEIS Ecology}
+\label{sect|peis-ecology}
+
+Reference paper:~\cite{Daoutis2009}
+
+{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim to binds autonomous
+robotics with ambient intelligence (network of sensors). \emph{PEIS} stands for
+\emph{Physically Embedded Intelligent System}: every robots or intelligent
+device in the environment is abstracted as a PEIS.
+
+Each PEIS component is running a \emph{PEIS Kernel} instance. Communication
+between instance relies on a custom P2P communication protocol.
+
+In short:
+* Roughly the same kind of ambition and architecture.
+* Way better than us on the perception side. 
+* They try the "2nd order logic" approach. I'd be curious to know what kind of reasoning capability they actually have.
+* Their tests are conducted in a static environment. They don't take other agents into consideration (nothing like perspective taking or rich disambiguation).
+* No task planning at all (or at least, not discussed in this paper)
+
+More in details:
+- object identification based on viewpoint independent SIFT features
+- formalized anchoring system that explicitely match percieved attributes to predicates
+- Cyc predicates
+- ground 12 colors, based on a paper on color perception. Could be useful for us.
+- idem, they cite a paper on what spatial relations to compute
+- location of objects based on a previously provided semantic map (but not much on this semantic map)
+- two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
+- uses directly Cyc (ie, 250 000 common sense concepts...), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
+Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
+- disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
+- template based natural language
+- experiment conducted in a "smart" indoor environmement + simple robot
+
+\subsubsection{Intrinsic features}
+\label{sect|peis-intrinsic-features}
+
+\subsubsection{Integration in the PEIS Ecology architecture}
+\label{sect|peis-integration}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\columnwidth]{peis-architecture.pdf}
+	\caption{The PEIS knowledge representation system, taken from~\cite{Daoutis2009}}
+	\label{fig|peis-architecture}
+\end{figure}
+
+The PEIS KR\&R system is deeply integrated to the general PEIS Ecology environment.
+
+\begin{table}
+\begin{center}
+\rowcolors{2}{lightgray}{codegray}
+
+\begin{tabular}{ll}
+\hiderowcolors
+{\bf Project} & {\bf Common-sense knowledge source} \\
+\hline
+\showrowcolors
+{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
+ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
+PEIS Ecology & {\sc ReseachCyc} \\
+ & Kollar, Tellex \\
+OMKRF \\
+ & Roy, Mavridis \\
+ & Wrighteagle \\
+ & Vincze \\
+ & (Bielfeld) \\
+ARMAR & Schmidt-Rohr (Karlsruhe) \\
+ & (Saarbrücken) \\
+ & Hertzberg (Osnabrück) \\
+ & (DFKI Bremen) \\
+ (based on {\sc KnowRob} & (JSK) \\
+DY-KNOW & Heitnz, Dowerty \\
+
+\hline
+
+\end{tabular}
+\end{center}
+\caption{Underlying knowledge sources for each project}
+\label{table|knowledge-sources}
+\end{table}
 
@@
 
-\bibliographystyle{spbasic}
+\bibliographystyle{ieeetr}
 \bibliography{biblio}

### 2011-11-21 paper.tex
Added a Comparison Criteria section: Strategies to deal with physical world
Added 'Interaction with humans' to the Comparison Criteria
Tune column widths of the main table

@@
 
-\subsection{Integration in a larger robotic architecture}
+\subsection{Strategies to deal with the physical world}
 
@@
 	\end{itemize}
+	\item Grounding/anchoring strategies
+	\item Ability to automatically create new object instances
+	\item Ability to merge modalities
+\end{itemize}
 
-	\item  Integration with supervision
+\subsection{Integration in a larger robotic architecture}
+
+\begin{itemize}
+
+	\item  Integration with executive layers
 	\begin{itemize}
@@
 	\end{itemize}
+
+	\item Interaction with humans
+
+	Both as developers (KRS as a tool to make robot knowledge explicit to the
+	robot designer) and end-users (like natural language processing,...)
 
@@
 %\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
-\begin{tabular}{lllllll}
+\begin{tabular}{p{2.7cm}p{4cm}lp{2.5cm}p{4cm}lp{1.5cm}}
 \hiderowcolors
@@
 ORO & Lemaignan (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
-PEIS Ecology & Daoutis, Coradeshi, Loutfi, Saffiotti (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & OWL-Full, 2nd order logics & & \cite{Daoutis2009} \\
+PEIS Ecology & Daoutis, Coradeshi, Loutfi, Saffiotti (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
  & Kollar, Tellex \\
@@
 \end{landscape}
+
+\subsection{KnowRob}
+\label{sect|knowrob}
+
+\subsection{ORO}
+\label{sect|oro}
 

### 2011-11-21 paper.tex
Completed a bit the tables

@@
 PEIS Ecology & Daoutis, Coradeshi, Loutfi, Saffiotti (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
+DY-KNOW & Heintz, Dowerty (Linköping Univ.) & & & & & \cite{Heintz2004} \\
+OMKRF & Suh et al. (Hanyang Univ.) & & & & & \cite{Suh2007} \\
+GSM & Mavridis, Roy (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
  & Kollar, Tellex \\
-OMKRF \\
- & Roy, Mavridis \\
  & Wrighteagle \\
  & Vincze \\
- & (Bielfeld) \\
-ARMAR & Schmidt-Rohr (Karlsruhe) \\
+ & (Bielfeld Univ.) \\
+ARMAR & Schmidt-Rohr (Karlsruhe TH) \\
  & (Saarbrücken) \\
- & Hertzberg (Osnabrück) \\
+ & Hertzberg (Osnabrück Univ.) \\
  & (DFKI Bremen) \\
  (based on {\sc KnowRob} & (JSK) \\
-DY-KNOW & Heitnz, Dowerty \\
 
@@
 PEIS Ecology & {\sc ReseachCyc} \\
+DY-KNOW & Heitnz, Dowerty \\
+OMKRF & Suh et al. (Hanyang Univ.) \\
+GSM & Mavridis, Roy (MIT MediaLab) \\
  & Kollar, Tellex \\
-OMKRF \\
- & Roy, Mavridis \\
  & Wrighteagle \\
@@
  (based on {\sc KnowRob} & (JSK) \\
-DY-KNOW & Heitnz, Dowerty \\
 

### 2011-11-21 paper.tex
More on PEIS Ecology

@@
 
-In short:
-* Roughly the same kind of ambition and architecture.
-* Way better than us on the perception side. 
-* They try the "2nd order logic" approach. I'd be curious to know what kind of reasoning capability they actually have.
-* Their tests are conducted in a static environment. They don't take other agents into consideration (nothing like perspective taking or rich disambiguation).
-* No task planning at all (or at least, not discussed in this paper)
-
 More in details:
@@
 
-\subsubsection{Integration in the PEIS Ecology architecture}
-\label{sect|peis-integration}
+\paragraph{Expressiveness} The PEIS Knowledge representation system relies on
+the {\sc ReseachCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
+allows to represent first order logic sentences and has extensions for modal logics and higher order logics.
+
+\fxfatal{Is modal logics and higher order logics actually used in PEIS?}
+
+As a system relying on {\sc CycL}, contexts can be expressed as
+\emph{microtheories}: the truth or falsity of a set of statement depends of the
+\emph{microtheory} in which these statements are evaluated.
+
+\fxfatal{OWA/CWA?}
+
+\subsection{Anchoring strategies}
+\label{sect|peis-anchoring}
 
@@
 	\caption{The PEIS knowledge representation system, taken from~\cite{Daoutis2009}}
-	\label{fig|peis-architecture}
+	\label{fig|peis-archi}
 \end{figure}
 
-The PEIS KR\&R system is deeply integrated to the general PEIS Ecology environment.
+The PEIS KR\&R system is deeply integrated to the general PEIS Ecology
+\emph{smart} environment. Figure~\ref{fig|peis-archi} gives an overview of the
+interactions between PEIS knowledge processing layers.
+
+\paragraph{Knowledge Acquisition} The primary source for knowledge acquisition
+is perception.  The PEIS ecosystem provides a SIFT-based object recognizer used
+in conjunction with ceiling cameras for object localization.  Other perceptual
+modalities are available (like human tracking, ambient environment monitoring).
+
+A template-based natural language parsing system may also be used to add new
+assertions to the system.
+
+\paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
+finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
+\Phi \times D(\Phi)$, where $\mathcal{P}$ is a set of predicate symbols, $\Phi$
+a set of percept's attributes, and $D(\Phi)$ the domain of these attributes.
+
+In the current implementation, object category (returned by the SIFT
+classifier), color, location, spatial relations (both topological -- \emph{at},
+\emph{near} -- and relative to the robot -- \emph{left}, \emph{behind}, etc.)
+and visibility are the five classes of extracted attributes.
+
+\subsubsection{Integration in the robot architecture}
+\label{sect|peis-integration}
+
+The PEIS framework offers through the \emph{PEIS middleware} a practical way to
+insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
+module can be seemlessly integrated into the PEIS ecosystem.
+
+\subsubsection{Notable experiments}
+\label{sect|peis-expe}
 

### 2011-11-21 paper.tex
Added new drafts of sections for the introduction: how to evaluate KRS?

@@
 
+\end{itemize}
+
+\subsection{Evaluation of a knowledge representation system}
+\label{sect|evaluation}
+
+\subsubsection{Desired features from other fields}
+\label{sect|evaluation-literature}
+
+\paragraph{General views on what is missing to AI} \cite{McCarthy2007}
+
+\paragraph{Natural language processing in situated context}
+In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
+challenges to be tackled: cross-modal representation systems, association of
+words with perceptual and action categories, modeling of context, figuring out
+the right granularity of models, integrating temporal modeling and planning,
+ability to match past (learned) experiences with the current interaction and
+ability to take into account the human perspective.
+
+\subsubsection{Pyschological tests}
+\label{sect|evaluation-tests}
+
+\begin{itemize}
+	\item Language comprehension tests: Token Test~\cite{DiSimoni1978}
 \end{itemize}
@@
 \paragraph{Expressiveness} The PEIS Knowledge representation system relies on
-the {\sc ReseachCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
+the {\sc ResearchCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
 allows to represent first order logic sentences and has extensions for modal logics and higher order logics.
@@
 ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
-PEIS Ecology & {\sc ReseachCyc} \\
+PEIS Ecology & {\sc ResearchCyc} \\
 DY-KNOW & Heitnz, Dowerty \\

### 2011-11-22 paper.tex
Reorganized the 'Comparison criteria' section in subsections and paragraphs

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Comparison criteria}
@@
 
-\begin{itemize}
-	\item  Expressiveness
-	\begin{itemize}
-		\item  Which logic formalism (DL, 2nd order\ldots{})
-		\item  OWA/CWA
-		\item  Representation of uncertainty
-		\item  (non) monotonic reasoning
-		\item  General knowledge + exception to this knowledge (blue sky/white sky)
-		\item  Microtheories
-		\item  Lazy evaluation
-		\item  Presupposition accomodation (ability to represent facts that are only verbally hinted, but not grounded into perception)
-	\end{itemize}
+\subsubsection{Expressiveness}
+\label{sect|expressiveness}
 
-	\item  Explicit modeling of context of knowledge / domain of validity
-	\item  Representation of change (\emph{"The pancake dough disappears into a pancake"})
-	\item  Support for learning
-	\item  Support for introspection?
-	\item  Support for prediction/projection?
-\end{itemize}
+\paragraph{Logics formalism}
+
+Which logic formalism (DL, 2nd order\ldots{})
+
+\paragraph{Open World Assumption or Close World Assumption}
+
+\paragraph{(non) monotonic reasoning}
+
+\paragraph{General knowledge + exception to this knowledge (blue sky/white sky)}
+
+\paragraph{Representation of uncertainty}
+
+\paragraph{Representation of time}
+
+\paragraph{Representation of change}
+
+E.g. \emph{"The pancake dough disappears into a pancake"}
+
+\paragraph{Context and Microtheories}
+Explicit modeling of context of knowledge / domain of validity
+
+\subsubsection{Presupposition accomodation}
+\label{sect|presupposition-accomodation}
+
+\emph{Presupposition accomodation} is the ability to acquire and represent facts that are not
+grounded into perception. A typical example is a human telling the robot that
+\emph{"there is big blue sphere behind you!"}.
+
+\subsubsection{Lazy evaluation}
+\label{sect|lazy-evaluation}
+
+\subsubsection{Introspection}
+\label{sect|introspection}
+Support for introspection?
+
+\subsubsection{Prediction and projection tasks}
+\label{sect|prediction-projection}
+
+Levesque~\cite{Levesque2007} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
+
+\paragraph{Projection task}: determining whether or not some condition while hold after a sequence of actions.
+
+\paragraph{Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+
+\subsubsection{Learning}
+\label{sect|learning}
 
@@
 
-\begin{itemize}
-	\item  Knowledge acquisition possible\ldots{}
-	\begin{itemize}
-		\item  \ldots{}through interaction? How?
-		\item  \ldots{}through the Web? How?
-		\item  \ldots{}through learning? How?
-		\item  \ldots{}through perception? How?
-	\end{itemize}
-	\item Grounding/anchoring strategies
-	\item Ability to automatically create new object instances
-	\item Ability to merge modalities
-\end{itemize}
+\subsubsection{Knowledge acquisition}
+\label{sect|knowledge-acquisition}
+
+\paragraph{Perception}
+\paragraph{Interaction}
+\paragraph{External sources (Web, upper ontologies, ...)}
+\paragraph{Learning}
+
+\subsubsection{Grounding/anchoring strategies}
+\label{sect|grounding}
+
+\subsubsection{Ability to automatically create new object instances}
+\label{sect|new-instances}
+
+\subsubsection{Ability to merge modalities}
+\label{sect|modalities-merging}
 
 \subsection{Integration in a larger robotic architecture}
+\label{sect|integration-robot}
 
-\begin{itemize}
+\subsubsection{Integration with executive layers}
+\label{sect|integration-executive-layers}
 
-	\item  Integration with executive layers
-	\begin{itemize}
-		\item  Events
-		\item  Relation to task planning
-	\end{itemize}
+\paragraph{Events}
 
-	\item Interaction with humans
+\paragraph{Relation to task planning}
+
+\subsubsection{Interaction with humans}
+\label{sect|hri}
 
@@
 
-	\item  Scalability and responsiveness
-\end{itemize}
+\subsubsection{Scalability and responsiveness}
+\label{sect|scalability}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Surveyed systems}
@@
 
-\subsection{Anchoring strategies}
+\subsubsection{Anchoring strategies}
 \label{sect|peis-anchoring}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}

### 2011-11-22 paper.tex
Added some content to the introduction

@@
 
+\newcommand{\ie}{{\textit{i.e.~}}}
+\newcommand{\cf}{{\textit{cf~}}}
+\newcommand{\eg}{{\textit{e.g.~}}}
+
 \begin{document}
@@
 
+The idea of \emph{Cognitive robotics} was coined in the early 1990s by Reiter.
+In a chapter on that subject in \emph{Fondations of Artifical
+Intelligence}~\cite{Levesque2007}, Levesque reminds about the manifesto they
+wrote together in 1998:
+
+\begin{quotation}
+
+	Central to this effort is to develop an understanding of the relationship
+	between the knowledge, the perception, and the action of such a robot. The
+	sorts of questions we want to be able to answer are
+
+	\begin{itemize}
+
+		\item to execute a program, what information does a robot need to have
+		at the outset vs. the information that it can acquire \emph{en route}
+		by perceptual means?
+
+		\item what does the robot need to know about its environment vs. what
+		need only be known by the designer?
+
+		\item when should a robot use perception to find out if something is
+		true as opposed to reasoning about what it knows was true in the past?
+
+		\item when should the inner workings of an action be available to the
+		robot for reasoning and when should the action be considered primitive
+		or atomic?
+
+	\end{itemize}
+
+	and so on. With respect to robotics, our goal (like that of many in AI) is
+	\emph{high-level robotic control}: develop a system that is capable of
+	generating actions in the world that are appropriate as a function of some
+	current set of beliefs and desires.
+
+\end{quotation}
+
+This survey focuses on the knowledge issue: we aims at establishing the current
+landscape of approaches to the knowledge representation problem in the research
+community.
+
+We have identified over ten projects that are designed for explicit knowledge
+manipulation within a robot. By knowledge manipulation, we mainly mean:
+
+\begin{itemize}
+
+	\item symbolic representation of assertions, be it static statements on the
+	world or spatio-temporal events,
+
+	\item reasoning that this representation enables,
+
+\end{itemize}
+
+We will also try, for each project, to make clear
+
+\begin{itemize}
+
+	\item how new facts can be acquired, \ie how informations from perception
+	or interaction are turned into knowledge,
+
+	\item how, in return, symbolic concept are \emph{anchored} into the robot
+	sensori-motor space,
+
+	\item how the knowledge representation system integrates with the
+	decisional layers of the robot, and if this leads to better robotic
+	control.
+
+\end{itemize}
+
 \subsection{Definition of inclusion criteria}
+\label{sect|inclusion-criteria}
 
-We have setup the following criteria to select the knowledge representation
-systems to include in this survey.
+Every robotic system has, implicitely or not, some knowledge representation
+systems. It may range from a simple state vector to an explicit symbolic
+knowledge base.  This survey focuses on the right end of this spectrum:
+symbolic systems, suited for abstract reasoning.
 
-Those systems must:
+Besides, we have decided to restraint the set of systems to those actually
+implemented on robots, and used in semantic-rich environments (\ie , dynamic,
+partially unknown environments with a large range of different entities which
+may have interactions). The typical scenario that would involve such robots is
+a service robot in a human-friendly environment like a kitchen.
+
+More precisely, we have established the following criteria to select the
+knowledge representation systems to include in this survey. Those systems must:
 
@@
 
-\subsubsection{Desired features from other fields}
+\subsubsection{Desirable features of a knowledge representation system}
 \label{sect|evaluation-literature}
@@
 
-\paragraph{Open World Assumption or Close World Assumption}
+\paragraph{Open World and Close World Assumptions}
+
+\fixme{cf Levesque~\cite{Levesque2007}, section 24.3.2}
 

### 2011-11-22 paper.tex
Removed Kollar&Tellex. Formatting of tables
Kollar, Tellex -> language grounding, no real knowledge representation system

@@
 \showrowcolors
-{\sc KnowRob} & Tenorth (TU Munich) & & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) & \cite{Tenorth2009a} \\
-ORO & Lemaignan (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
-PEIS Ecology & Daoutis, Coradeshi, Loutfi, Saffiotti (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
-DY-KNOW & Heintz, Dowerty (Linköping Univ.) & & & & & \cite{Heintz2004} \\
-OMKRF & Suh et al. (Hanyang Univ.) & & & & & \cite{Suh2007} \\
-GSM & Mavridis, Roy (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
- & Kollar, Tellex \\
+{\sc KnowRob} & Tenorth (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) & \cite{Tenorth2009a} \\
+ORO & Lemaignan \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
+PEIS KR\&R & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
+DY-KNOW & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
+OMKRF & Suh et al. \par (Hanyang Univ.) & & & & & \cite{Suh2007} \\
+GSM & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
+NKRL & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
  & Wrighteagle \\
@@
 PEIS Ecology & {\sc ResearchCyc} \\
-DY-KNOW & Heitnz, Dowerty \\
-OMKRF & Suh et al. (Hanyang Univ.) \\
-GSM & Mavridis, Roy (MIT MediaLab) \\
- & Kollar, Tellex \\
- & Wrighteagle \\
- & Vincze \\
- & (Bielfeld) \\
-ARMAR & Schmidt-Rohr (Karlsruhe) \\
- & (Saarbrücken) \\
- & Hertzberg (Osnabrück) \\
- & (DFKI Bremen) \\
- (based on {\sc KnowRob} & (JSK) \\
+DY-KNOW & \\
+OMKRF & \\
+GSM &  \\
+NKLR &  None \\
+Wrighteagle & \\
+Vincze & \\
+(Bielfeld) & \\
+ARMAR &  \\
+(Saarbrücken) & \\
+Hertzberg (Osnabrück) & \\
+(DFKI Bremen) & \\
+(JSK) & \\
 

### 2011-11-22 paper.tex
Formatting in the PEIS section

@@
 
-\subsection{PEIS Ecology}
+\subsection{PEIS KR\&R}
 \label{sect|peis-ecology}
-
-Reference paper:~\cite{Daoutis2009}
 
@@
 
-Each PEIS component is running a \emph{PEIS Kernel} instance. Communication
+Each PEIS physical component is running a \emph{PEIS Kernel} instance. Communication
 between instance relies on a custom P2P communication protocol.
 
-More in details:
-- object identification based on viewpoint independent SIFT features
-- formalized anchoring system that explicitely match percieved attributes to predicates
-- Cyc predicates
-- ground 12 colors, based on a paper on color perception. Could be useful for us.
-- idem, they cite a paper on what spatial relations to compute
-- location of objects based on a previously provided semantic map (but not much on this semantic map)
-- two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
-- uses directly Cyc (ie, 250 000 common sense concepts...), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
-Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
-- disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
-- template based natural language
-- experiment conducted in a "smart" indoor environmement + simple robot
+The PEIS architecture allows for adding new abilities through software components sharing the common \emph{tuple space}.
+
+We survey here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS KR\&R}, that includes symbolic representation and reasoning.
+
+% More in details:
+% - object identification based on viewpoint independent SIFT features
+% - formalized anchoring system that explicitely match percieved attributes to predicates
+% - Cyc predicates
+% - ground 12 colors, based on a paper on color perception. Could be useful for us.
+% - idem, they cite a paper on what spatial relations to compute
+% - location of objects based on a previously provided semantic map (but not much on this semantic map)
+% - two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
+% - uses directly Cyc (ie, 250 000 common sense concepts...), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
+% Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
+% - disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
+% - template based natural language
+% - experiment conducted in a "smart" indoor environmement + simple robot
 

### 2011-11-22 paper.tex
Added first content for NKLR

@@
 
+\subsection{NKRL}
+\label{sect|nkrl}
+
+\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
+While this language is developped since a long time by Zarri~\cite{Zarri1997,
+Zarri2008}, recent research direction include application to the robotic
+field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
+system, as it is primarily a language. However, it is used as the
+representation and reasoning mechanism for robots by Sabri et al.
+
+\subsubsection{Intrinsic language features}
+\label{sect|nkrl-intrinsic-features}
+
+\paragraph{Expressiveness}
+
+\subsubsection{Integration with physical world and in the robot architecture}
+\label{sect|nkrl-integration}
+
+...seem to be mostly WIP...
+
 \begin{table}

### 2011-11-22 krs-survey.tex
Added some 'desirable features' for KRS from MCCarthy
While here, renamed paper.tex into krs-survey.tex

--- 2011-11-22 paper.tex
@@
 
-\paragraph{General views on what is missing to AI} \cite{McCarthy2007}
+Several authors from fields that are connex to the robotics have previously listed desirable features of systems aiming at cognitive abilities as rich as possible.
+
+\paragraph{McCarthy: Towards human-level AI}
+
+In~\cite{McCarthy2007}, McCarthy lists the challenges he identifies on the road
+to a \emph{human-level AI}.
+
+\begin{itemize}
+
+	\item the ability to \emph{"operate successfully in the common sense
+	informatic situation"},
+
+	\item the necessity of relying on mathematical logic, as the most fruitful
+	formalism for machine intelligence,
+
+	\item the ability to deal with \emph{approximate concepts and approximate
+	theories} (that would include representing them, and reasoning with them),
+
+	\item nonmonotonic reasoning \fxfatal{give here a clear example},
+
+	\item what McCarthy calls \emph{Elaboration Tolerance}, \ie, the ability to
+	extend \emph{on requirement} the closed domain of interpretation for a
+	given assertion,
+
+	\item the ability to formalize and reason about contexts,
+
+	\item reasoning about events, and in particular, actions,
+
+	\item the capacity of introspection,
+
+	\item and finally, he points the issue of giving computer the right
+	heuristics for decision making.
+
+\end{itemize}
+
+Knowledge representation systems are concerned by most, if not all, of this
+points.
 
 \paragraph{Natural language processing in situated context}
+
 In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
@@
 Which logic formalism (DL, 2nd order\ldots{})
+
+On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3
 

### 2011-11-23 krs-survey.tex
Added section on the CoSy CAST approach

@@
 NKRL & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
+CAST Proxies & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & & Amodal proxies & \cite{Jacobsson2008} \\
  & Wrighteagle \\
@@
 ARMAR & Schmidt-Rohr (Karlsruhe TH) \\
- & (Saarbrücken) \\
  & Hertzberg (Osnabrück Univ.) \\
@@
 
+\subsection{CAST Knowledge model}
+\label{sect|cast}
+
+CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
+comprehensive toolkit aimed at builduing cognitive architectures for robots
+through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
+expose a central knowledge base as seen in previous works. It instead
+represents knowledge as unrooted \emph{proxies}. Thoses proxies are formally
+defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
+a set of instanciated features (like $\phi^{Colour}_{red}$) and $u_p$ a
+\emph{proxies union} that form an equivalence class corresponding to one
+entity.
+
+A union of proxies forms a global amodal representation of an entity, that can
+be explicitely shared and manipulated. Being not centralized, the knowledge
+model can be qualified of \emph{ubiquitous}. Futhermore, knowledge source in
+the CAS architecture is tightly bound to the on-line grounding process (be it
+grounded in perception or in dialogue). While nothing seems to prevent it, no
+{\it a priori} knowledge (including common-sense knowledge) is used.
+
+Knowledge sharing is ensured by the event mechanism of CAST: modules can
+monitor proxies for alteration by other modules. Jacobsson et al. mention how
+this can apply to reinforcement learning: the vision module creates a proxy for
+an orange object. This proxy get monitored by a learning module. In parallel,
+the proxy is bound to an union by the natural language understanding module
+that add new a feature like \emph{"this object is a fruit"}. The learning
+module is called back, and can add this new information to its model.
+
+In the presented implementation, the CAST knowledge model does no allow for
+effectively representing actions or temporal information.\fxfatal{What about reasoning? can they retrieve for example 'all proxies for colorful objects'?}
+
+%%%%%%%%%% Underlying knowledge model table %%%%%%%%
 \begin{table}
@@
 NKLR &  None \\
+CAST Proxies &  None \\
 Wrighteagle & \\
@@
 ARMAR &  \\
-(Saarbrücken) & \\
 Hertzberg (Osnabrück) & \\

### 2011-11-23 krs-survey.tex
Added a 'Summary of Approaches' section, with hints to the future content
Also changed/added a few other minor sections

@@
 
+\subsubsection{Benchmarking}
+\label{sect|benchmarking}
+
+\cite{Hawes2008}: benchmarking of "Information-Processing Architectures on Intelligent Systems"
 \subsubsection{Pyschological tests}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Comparison criteria}
-\label{sect|comparison-criteria}
+\section{Compared features}
+\label{sect|compared-features}
 
@@
 \label{sect|lazy-evaluation}
+
+\subsubsection{Agent-dependent models}
+\label{sect|multiple-models}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
+\section{Summary of Approaches}
+\label{sect|summary}
+
+This section tries to summary the various approaches we surveyed in the
+previous section.  This summary does not exactly follow the list of compared
+features presented in section~\ref{sect|compared-features}. We have organized
+it along four axis: \emph{what can be represented?}, \emph{How knowledge is
+created and grounded?}, \emph{What can be done with the knowledge?} and
+\emph{How to use knowledge in the whole, larger robot architecture?}.
+
+\subsection{Expressiveness: what can we represent in the current state-of-the-art?}
+\label{sect|summary-expressiveness}
+
+\begin{itemize}
+	\item expressiveness
+\end{itemize}
+
+\subsection{Creating and grounding knowledge in the physical world}
+\label{sect|summary-grounding}
+
+\begin{itemize}
+	\item which knowledge modalities can be merged
+	\item which grounding strategies
+\end{itemize}
+
+\subsection{Exploiting knowledge}
+\label{sect|summary-knowledge-sources-reasoning}
+
+\begin{itemize}
+	\item which common-sense knowledge is used
+	\item what kind of reasoning is made
+	\item introspection
+\end{itemize}
+
+\subsection{Using a knowledge representation system in a larger robotic cognitive architecture}
+\label{sect|summary-integration}
+
+\begin{itemize}
+	\item a priori evaluation/lazy evaluation and trade-offs (relations to events, scalability, detection of inconsistencies,...)
+	\item relation with planning (action representation, temporal representation...)
+	\item API/bindings/integration into executive layers
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Towards the next generation of Knowledge Representation Systems for Robotics}
 \label{sect|conclusion}
 
-\subsection{Main groups}
+\subsection{Current shotcomings: what is not successfully covered by current systems}
 
-\subsection{What is not successfully covered by current systems}
-
-\subsection{Future challenges}
-\label{sect|future-challenges}
+\subsection{Towards a new design}
 

### 2011-11-23 krs-survey.tex
Added a 'Category' column tot the main table

@@
 %\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
-\begin{tabular}{p{2.7cm}p{4cm}lp{2.5cm}p{4cm}lp{1.5cm}}
+\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{2.8cm}p{1.5cm}}
 \hiderowcolors
-{\bf Project} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} & Main reference \\
+{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} & Main reference \\
 \hline
 \showrowcolors
-{\sc KnowRob} & Tenorth (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom ({\sc Prolog}) & \cite{Tenorth2009a} \\
-ORO & Lemaignan \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
-PEIS KR\&R & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
-DY-KNOW & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
-OMKRF & Suh et al. \par (Hanyang Univ.) & & & & & \cite{Suh2007} \\
-GSM & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
-NKRL & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
-CAST Proxies & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & & Amodal proxies & \cite{Jacobsson2008} \\
- & Wrighteagle \\
- & Vincze \\
- & (Bielfeld Univ.) \\
-ARMAR & Schmidt-Rohr (Karlsruhe TH) \\
- & Hertzberg (Osnabrück Univ.) \\
- & (DFKI Bremen) \\
- (based on {\sc KnowRob} & (JSK) \\
+{\sc KnowRob} & KRS & Tenorth (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
+ORO & KRS & Lemaignan \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
+PEIS KR\&R & KRS & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
+CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
+Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
+NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
+GSM & & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
+OMKRF & & Suh et al. \par (Hanyang Univ.) & & & & & \cite{Suh2007} \\
+DY-KNOW & & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
+ & & Wrighteagle \\
+ & & Vincze \\
+ & & (Bielfeld Univ.) \\
+ARMAR & & Schmidt-Rohr (Karlsruhe TH) \\
+ & & Hertzberg (Osnabrück Univ.) \\
+ & & (DFKI Bremen) \\
+ (based on {\sc KnowRob} & & (JSK) \\
 
@@
 \end{center}
-\caption{List of surveyed systems}
+\caption{List of surveyed systems.
+Projects are listed by category (\emph{KRS} for systems that are explicit knowledge representation and reasoning modules, \emph{Ubiquitous} for systems where knowledge processing is fully distributed, \emph{Language} for languages used as KRS on robots), then names.}
 \label{table|surveyed-systems}

### 2011-11-23 krs-survey.tex
Reformatted the intro to remoze itemization

@@
 \usepackage{listings}
+\usepackage{paralist}
 \usepackage{fancyhdr}
@@
 
-The idea of \emph{Cognitive robotics} was coined in the early 1990s by Reiter.
+The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
 In a chapter on that subject in \emph{Fondations of Artifical
@@
 landscape of approaches to the knowledge representation problem in the research
-community.
+community, at summarizing the major trends and identifying the possible
+shortcomings.
 
-We have identified over ten projects that are designed for explicit knowledge
-manipulation within a robot. By knowledge manipulation, we mainly mean:
+We have identified over ten projects that either are explicitely advertised as
+knowledge representation systems or involve explicit knowledge manipulation
+within a robot. By knowledge manipulation, we mean symbolic representation of
+assertions, be it static statements on the world or spatio-temporal events, and
+reasoning on this representation.
 
-\begin{itemize}
-
-	\item symbolic representation of assertions, be it static statements on the
-	world or spatio-temporal events,
-
-	\item reasoning that this representation enables,
-
-\end{itemize}
-
-We will also try, for each project, to make clear
-
-\begin{itemize}
-
-	\item how new facts can be acquired, \ie how informations from perception
-	or interaction are turned into knowledge,
-
-	\item how, in return, symbolic concept are \emph{anchored} into the robot
-	sensori-motor space,
-
-	\item how the knowledge representation system integrates with the
-	decisional layers of the robot, and if this leads to better robotic
-	control.
-
-\end{itemize}
+We also try, for each project, to make clear \begin{inparaenum} \item how new
+facts can be acquired, \ie how informations from perception or interaction are
+turned into knowledge, \item how, in return, symbolic concept are
+\emph{anchored} into the robot sensori-motor space, \item how the knowledge
+representation system integrates with the decisional layers of the robot, and
+how this leads to better robotic control.\end{inparaenum}
 

### 2011-11-24 krs-survey.tex
Filled sections on Formal logics and CWA/OWA

@@
 In a chapter on that subject in \emph{Fondations of Artifical
-Intelligence}~\cite{Levesque2007}, Levesque reminds about the manifesto they
+Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
 wrote together in 1998:
@@
 
-\paragraph{Logics formalism}
+\paragraph{Logic formalism}
 
-Which logic formalism (DL, 2nd order\ldots{})
+The main role of a knowledge represenation system is to provide an adequate
+representation system to store facts and concepts that can be informally
+described in natural language.
 
-On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3
+Formal logic aims at providing such a representation system with the added
+value of providing a tractable support for inference and reasoning.
+
+Most (but not all) of the systems we survey rely on a particular logic
+formalism. The choice of the formalism has a strong impact, on one side, on the
+range of ideas that can be expressed conveniently (\emph{practical
+expressiveness}) or at all (\emph{theoritical expressiveness}), on the other
+side, on the ability to solve the inference problem (called
+\emph{decidability}: is a given logical sentence true in my model?) in a
+tractable manner.
+
+A large number of logic formalism do exist, we shall summarize below the most
+relevant for system actually deployed in current robotic architectures.
+
+\emph{Predicate logic} is the family of logic formalisms the most commonly
+found in knowledge representation. It distinguishes itself from the simpler
+\emph{propositional logic} by the use of quantification to increase generality.
+\emph{First-order logic} (FOL) is the subpart of \emph{predicate logic} where the
+objects of \emph{predicates} (or \emph{formulae}) are simple \emph{terms},
+while in \emph{higher-order logics}, predicates can be themselves objects of
+other predicates.
+
+The family of \emph{description logics}~\cite{Baader2008} also play an
+important role. It is a subset of the first-order logic, with some extensions
+in second-order logic. Description logics are notable because most of them are
+known to be decidable. In description logic, axioms are build from
+\emph{concepts}, \emph{roles} (that are unary or binary predicates) and
+\emph{individuals}. \emph{Modal logics}, that allow for statement qualification
+(like \emph{possibility} or \emph{necessity}) have been shown to be closely
+related to description logics.
+
+\fxfatal{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
+
+One last class of logics that is of particular relevance for robotic
+applications is the \emph{probabilistic logics} or \emph{bayesian logics}.
+These logics provide a formal framework to reason on propositions whose truth
+or falsity is uncertain.
+
+\paragraph{Some examples}
+
+\emph{The robot knows that a blue bottle is laying on the table.}
+
+\emph{The robot knows that the human knows about the position of the bottle,
+but the robot does not know what the human actually know about it.}
 
@@
 
-\fixme{cf Levesque~\cite{Levesque2007}, section 24.3.2}
+The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumption names a
+modelling choice on the \emph{completeness} of a knowledge domain. In the close
+world assumption, a proposition that can not be proven true is assumed to be
+false (\emph{negation by failure}), while in the open world assumption, a
+proposition may be considered either true, false or unknown.
+
+This distinction is important in robotics were the robot may have to manipulate
+concepts with only partial knowledge on them. For instance, let imagine a robot
+that sees a bottle over a table, whose bottom is hidden by another object. The
+robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
+representation system relying on the closed world assumption would then assume
+the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
+whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
+table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
+a simple, concerete example of consequences of the CWA/OWA choice on reasoning.
+
+\begin{table}
+	\begin{center}
+	\begin{tabular}{ll}
+	{\bf Action} & {\bf Part involved} \\
+	\hline
+	{\tt PickSoftly} & hand \\
+	{\tt PickAndPlace} & arm, hand \\
+	{\tt MoveArm} & arm \\
+	\hline
+	\end{tabular}
+	\end{center}
+	\caption{Assuming the question is: \emph{select actions that do not require
+	to move the arm}, a CWA reasoner would return {\tt PickSoftly} whereas an
+	OWA reasoner would not return anything if the {\tt PickSoftly} action is
+	not explicitely said not to involve the arm.}
+	\label{table|cwa-owa-example}
+\end{table}
+
+Domains constrained with the closed world assumption lead to more tractable
+inference problems, and allow for instance the use of logic languages like
+Prolog. Thus, several approaches exists to \emph{locally close} a domain (\cf
+Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those).
 
 \paragraph{(non) monotonic reasoning}
+
+\paragraph{Knowledge alteration possibilities}
+
+In DL -> always possible to modify the ABox, not always possible to alter the TBox
 
@@
 
-\subsubsection{Agent-dependent models}
-\label{sect|multiple-models}
+\subsubsection{Possible-Worlds and Agent-dependent models}
+\label{sect|possible-worlds}
+
+\cite{Levesque2008}, p. 4
 
@@
 
-Levesque~\cite{Levesque2007} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
+Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
 

### 2011-11-24 krs-survey.tex
Could not find much from Bielefeld...?

@@
  & & Vincze \\
- & & (Bielfeld Univ.) \\
 ARMAR & & Schmidt-Rohr (Karlsruhe TH) \\
@@
 \end{landscape}
+
+\fxfatal{Bielefeld -> could not find much...}
+\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}
 
@@
 Vincze & \\
-(Bielfeld) & \\
 ARMAR &  \\

### 2011-11-25 krs-survey.tex
Filled section on monotonicity

@@
 \emph{concepts}, \emph{roles} (that are unary or binary predicates) and
-\emph{individuals}. \emph{Modal logics}, that allow for statement qualification
-(like \emph{possibility} or \emph{necessity}) have been shown to be closely
+\emph{individuals}. The W3C OWL-DL standard is a widely-used language to describe
+domain with the description logic.
+
+\emph{Modal logics}, that allow for statement qualification
+like \emph{possibility} or \emph{necessity}, have been shown to be closely
 related to description logics.
@@
 These logics provide a formal framework to reason on propositions whose truth
-or falsity is uncertain.
+or falsity is uncertain. We elaborate below on the representation of uncertainty.
 
@@
 
-\paragraph{(non) monotonic reasoning}
+\paragraph{(Non) Monotonic Reasoning}
+
+\emph{Monotonic reasoning} means that addition of new assertions to a knowledge base
+can only extend the set of assertions that can be inferred, while a
+\emph{non-monotonic} reasoning scheme may lead to retraction of facts.
+McCarthy coined a famous example to illustrate the need of non-monotonic reasoning:
+
+\begin{quotation}
+Consider putting an axiom in a common sense database asserting that birds can
+fly. Clearly the axiom must be qualified in some way since penguins, dead birds
+and birds whose feet are encased in concrete can't fly. A careful construction
+of the axiom might succeed in including the exceptions of penguins and dead
+birds, but clearly we can think up as many additional exceptions like birds
+with their feet encased in concrete as we like. Formalized nonmonotonic
+reasoning provides a way of saying that a bird can fly unless there
+is an abnormal circumstance and reasoning that only the abnormal circumstances
+whose existence follows from the facts being taken into account will be
+considered.
+\end{quotation}
+
+\emph{Default logic} is one of the formal logic that account for representing
+general truth and exceptions to it. However, due to computational complexity of
+these model (most of inferences in default logic are known to be $NP$-complete
+problem), classical logics and most of the existing reasoners do not allow
+non-monotonic reasoning. For instance, the SWRL rule language, usually
+associated to the OWL-DL ontology language, do not allow non-monotonic
+reasoning.
+
+\fxfatal{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
+languages? reasoner?}
+
+One important exception if the \emph{Negation as failure} inference rule, as
+implemented by {\sc Prolog} for instance, that allows for non-monotonicity, but
+only \emph{within the closed world assumption}.
+
+\fxfatal{Give here an example of non-monotonic reasoning with Prolog}
+
+A monotonic system does not theoritically allow for knowledge retractation,
+which is an important issue in the robotic context where the world model is
+likely to be often altered.  However it is a practical issue only if the
+reasoning process is \emph{continuous} during the whole robot's activity
+lifespan. It is often possible to stop the reasoner, alter the knowledge, and
+restart the inference process on a new domain.
 
@@
 In DL -> always possible to modify the ABox, not always possible to alter the TBox
-
-\paragraph{General knowledge + exception to this knowledge (blue sky/white sky)}
 

### 2011-11-25 krs-survey.tex
Changed layout to 2 columns
Merged the 'representation of change' paragraph with non-monotonic reasoning

@@
 %\documentclass[twoside,a4paper]{report}
-\documentclass[a4paper]{article}
+\documentclass[a4paper, twocolumn]{article}
 
@@
 \usepackage[draft]{fixme}
+\fxsetup{footnote}
+\fxsetup{nomargin}
 
@@
 
+Another important application of non-monotonic reasoning is representation of
+change: for example, to make an omelette, you need to crack eggs and wipe them.
+The eggs disappear and are replaced by an omelette:
+
+$Egg(a) \wedge Egg(b) \wedge MakeOmelette(a, b, c) \to \lnot Egg(a) \wedge
+\lnot Egg(b) \wedge Omelette(c)$
+
+The insertion of the proposition $MakeOmelette(a, b, c)$ leads to retraction of
+other facts. This rule requires non-monotonic reasoning to be applied.
+
 \emph{Default logic} is one of the formal logic that account for representing
@@
 
-\paragraph{Representation of change}
-
-E.g. \emph{"The pancake dough disappears into a pancake"}
-
 \paragraph{Context and Microtheories}
 Explicit modeling of context of knowledge / domain of validity
-
-\subsubsection{Presupposition accomodation}
-\label{sect|presupposition-accomodation}
-
-\emph{Presupposition accomodation} is the ability to acquire and represent facts that are not
-grounded into perception. A typical example is a human telling the robot that
-\emph{"there is big blue sphere behind you!"}.
 
@@
 
+\subsubsection{Learning}
+\label{sect|learning}
+
+\subsection{Strategies to deal with the physical world}
+
+\subsubsection{Knowledge acquisition and modalities merging}
+\label{sect|knowledge-acquisition}
+
+\paragraph{Perception}
+\paragraph{Interaction}
+\paragraph{External sources (Web, upper ontologies, ...)}
+\paragraph{Learning}
+
+\subsubsection{Ability to automatically create new object instances}
+\label{sect|new-instances}
+
+\subsubsection{Grounding/anchoring strategies}
+\label{sect|grounding}
+
+\subsubsection{Presupposition accomodation}
+\label{sect|presupposition-accomodation}
+
+\emph{Presupposition accomodation} is the ability for the system to
+automatically create a context allowing to make sense of a proposition.
+
+For instance, we can imagine a human telling a robot \emph{Please get me the
+bottle that is behind you}. If the robot has not yet see what is behind it, it
+needs to assume (and represents in its knowledge model) that a undefined bottle
+can be found somewhere in the half of space behind it.
+
+A knowledge representation system able to copte with presupposition
+accomodation would be able to take into account this (usually under-defined)
+information that is not grounded into perception for later inferences.
+
+This ability to imagine a physically state of the world that is not actually
+perceived can be seen as the converse of the grounding ability.
+
+Note also that presupposition accomodation implies a bidirectional link of the
+symbolic knowledge model with a geometric (or physical) model of the
+environment. This article focuses on symbolic knowledge representation systems,
+but we shall mention when a KRS explicitely provides support for presupposition
+accomodation.
+
+%%%%%
+\subsection{Integration in robotic architectures}
+\label{sect|integration-robot}
+
 \subsubsection{Prediction and projection tasks}
@@
 
-\subsubsection{Learning}
-\label{sect|learning}
-
-\subsection{Strategies to deal with the physical world}
-
-\subsubsection{Knowledge acquisition}
-\label{sect|knowledge-acquisition}
-
-\paragraph{Perception}
-\paragraph{Interaction}
-\paragraph{External sources (Web, upper ontologies, ...)}
-\paragraph{Learning}
-
-\subsubsection{Grounding/anchoring strategies}
-\label{sect|grounding}
-
-\subsubsection{Ability to automatically create new object instances}
-\label{sect|new-instances}
-
-\subsubsection{Ability to merge modalities}
-\label{sect|modalities-merging}
-
-\subsection{Integration in a larger robotic architecture}
-\label{sect|integration-robot}
+\paragraph{Relation to task planning}
 
@@
 \paragraph{Events}
-
-\paragraph{Relation to task planning}
 
@@
 
-\begin{tabular}{ll}
+\begin{tabular}{lp{4cm}}
 \hiderowcolors
-{\bf Project} & {\bf Common-sense knowledge source} \\
+{\bf Project} & {\bf Common-sense \par knowledge source} \\
 \hline

### 2011-11-25 krs-survey.tex
Added paragraph on presupposition accomodation.
While here, moved the 'projection task' paragraph to the 'integration in robots' section

--- 2011-11-25 krs-survey.tex
@@
 
-\subsubsection{Presupposition accomodation}
-\label{sect|presupposition-accomodation}
-
-\emph{Presupposition accomodation} is the ability to acquire and represent facts that are not
-grounded into perception. A typical example is a human telling the robot that
-\emph{"there is big blue sphere behind you!"}.
-
 \subsubsection{Lazy evaluation}
@@
 
+\subsubsection{Learning}
+\label{sect|learning}
+
+\subsection{Strategies to deal with the physical world}
+
+\subsubsection{Knowledge acquisition and modalities merging}
+\label{sect|knowledge-acquisition}
+
+\paragraph{Perception}
+\paragraph{Interaction}
+\paragraph{External sources (Web, upper ontologies, ...)}
+\paragraph{Learning}
+
+\subsubsection{Ability to automatically create new object instances}
+\label{sect|new-instances}
+
+\subsubsection{Grounding/anchoring strategies}
+\label{sect|grounding}
+
+\subsubsection{Presupposition accomodation}
+\label{sect|presupposition-accomodation}
+
+\emph{Presupposition accomodation} is the ability to acquire and represent
+facts that are not grounded into perception. A typical example is a human
+telling the robot that \emph{"big blue sphere is behind you!"}.
+
+A knowledge representation system able to copte with presupposition
+accomodation would be able to take into account this (usually under-defined)
+information in later inference.
+
+This ability to imagine a physically state of the world that is not actually
+perceived can be seen as the converse of the grounding ability.
+
+Note also that presupposition accomodation implies a bidirectional link of the
+symbolic knowledge model with a geometric (or physical) model of the
+environment. This article focuses on symbolic knowledge representation systems,
+but we shall mention when a KRS explicitely provides support for presupposition
+accomodation.
+
+%%%%%
+\subsection{Integration in robotic architectures}
+\label{sect|integration-robot}
+
 \subsubsection{Prediction and projection tasks}

### 2011-11-25 krs-survey.tex
Rewrite paragraph on presupposition accomodation to be more accurate

--- 2011-11-25 krs-survey.tex
@@
 
-Another important application of non-monotonic reasoning is representation of
-change: for example, to make an omelette, you need to crack eggs and wipe them.
-The eggs disappear and are replaced by an omelette:
-
-$Egg(a) \wedge Egg(b) \wedge MakeOmelette(a, b, c) \to \lnot Egg(a) \wedge
-\lnot Egg(b) \wedge Omelette(c)$
-
-The insertion of the proposition $MakeOmelette(a, b, c)$ leads to retraction of
-other facts. This rule requires non-monotonic reasoning to be applied.
-
 \emph{Default logic} is one of the formal logic that account for representing
@@
 \paragraph{Representation of time}
+
+\paragraph{Representation of change}
+
+E.g. \emph{"The pancake dough disappears into a pancake"}
 

### 2011-11-30 krs-survey.tex
Reorganized criteria lists following discussion with Michael
Also started to fill some criteria subsections.

--- 2011-11-25 krs-survey.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Compared features}
+\section{Compared Features}
 \label{sect|compared-features}
 
-\subsection{Intrinsic features}
+\subsection{What Can be Represented? The Expressiveness Question}
 \label{sect|intrinsic-features}
 
-\subsubsection{Expressiveness}
-\label{sect|expressiveness}
+%\subsubsection{Expressiveness}
+%\label{sect|expressiveness}
 
-\paragraph{Logic formalism}
+\subsubsection{Introduction: main logic formalisms}
 
@@
 A large number of logic formalism do exist, we shall summarize below the most
-relevant for system actually deployed in current robotic architectures.
+relevant for systems actually deployed in current robotic architectures.
 
@@
 
-\paragraph{Open World and Close World Assumptions}
+\subsubsection{Open World and Close World Assumptions}
 
@@
 
-\paragraph{(Non) Monotonic Reasoning}
+\subsubsection{Representation of uncertainty}
+
+Sources of uncertainty for a robot are two-fold: uncertainty \emph{intrinsic}
+to facts (like \emph{``It may rain tomorrow''}), uncertainty caused by
+imperfect perception of the world (\emph{``Is the bottle really on the
+table?''}). Most logics do not account explicitely for uncertainty. It must be
+either relied on specific logics (like Bayesian logics) or on extensions of
+classical logics.
+
+\subsubsection{Representation of time}
+
+As an agent acting at human-like time scale and dealing with temporal concepts
+(like actions), a robot may want to represent, and possibly to reason, about
+time. Time representation is split into two distinct abilities: representing
+time points (both in the past -- which is roughly equivalent to assignment of
+timestamps to events the robot perceives -- and in the future), and
+representing \emph{passing time} (durations, timespans) like in \emph{``the
+eggs will be cooked in 10 min''}.
+
+We call a system that do not account for time (\ie that permanently lives in
+present) \emph{atemporal}.
+
+\subsubsection{Context modeling}
+
+\emph{Knowledge is contextualized information}\fxfatal{Find someone respectable
+how said that :-)}: it is essential for the robot to associate the facts it
+represents to a \emph{context}. The context carries the keys for the
+interpretation of the information. It covers the \emph{domain of validity} of
+the facts, the \emph{common-sense} knowledge required to fill the gaps in the
+representation\fxfatal{give an example}, \fxfatal{What more?}.
+
+\subsubsection{Possible-Worlds and representing what others know}
+\label{sect|possible-worlds}
+
+Linked to the context representation, but seen from another angle, knowledge
+representation systems may provide explicit ways to model other point of view
+on the world. This ability is often refered as the \emph{perspective taking}
+ability.
+
+\cite{Levesque2008}, p. 4
+
+\subsubsection{Introspection: knowledge about the knowledge}
+\label{sect|introspection}
+Support for introspection?
+
+\subsubsection{Modifying its knowledge}
+
+In DL -> always possible to modify the ABox, not always possible to alter the TBox
+
+%%%%%%%%%%
+\subsection{Reasoning Techniques}
+
+\subsubsection{Prediction and projection tasks}
+\label{sect|prediction-projection}
+
+Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
+
+\paragraph{Projection task}: determining whether or not some condition while hold after a sequence of actions.
+
+\paragraph{Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+
+\subsubsection{Reasoning in uncertainty}
+
+\subsubsection{(Non) Monotonic Reasoning}
 
@@
 restart the inference process on a new domain.
+\fxfatal{Rephrase to emphazise that when new evidences appear, it is anyway often a
+good idea to restart the reasoner.}
 
-\paragraph{Knowledge alteration possibilities}
-
-In DL -> always possible to modify the ABox, not always possible to alter the TBox
-
-\paragraph{Representation of uncertainty}
-
-\paragraph{Representation of time}
-
-\paragraph{Context and Microtheories}
-Explicit modeling of context of knowledge / domain of validity
-
-\subsubsection{Lazy evaluation}
-\label{sect|lazy-evaluation}
-
-\subsubsection{Possible-Worlds and Agent-dependent models}
-\label{sect|possible-worlds}
-
-\cite{Levesque2008}, p. 4
-
-\subsubsection{Introspection}
-\label{sect|introspection}
-Support for introspection?
-
-\subsubsection{Learning}
-\label{sect|learning}
-
-\subsection{Strategies to deal with the physical world}
-
-\subsubsection{Knowledge acquisition and modalities merging}
-\label{sect|knowledge-acquisition}
-
-\paragraph{Perception}
-\paragraph{Interaction}
-\paragraph{External sources (Web, upper ontologies, ...)}
-\paragraph{Learning}
-
-\subsubsection{Ability to automatically create new object instances}
-\label{sect|new-instances}
-
-\subsubsection{Grounding/anchoring strategies}
-\label{sect|grounding}
+\fxfatal{Mention that the 'change of world' issue can also be dealt with appropriate
+time representation.}
+\fxfatal{Mention that probabilistic reasoning lead to implicit non-monotonic reasoning}
 
@@
 
-%%%%%
-\subsection{Integration in robotic architectures}
+\subsubsection{Learning}
+\label{sect|learning}
+
+%%%%%%%%%%%%%%%%%
+\subsection{Acquiring Knowledge}
+
+\subsubsection{Knowledge acquisition and modalities merging}
+\label{sect|knowledge-acquisition}
+
+\paragraph{Perception}
+\paragraph{Interaction}
+\paragraph{External sources (Web, upper ontologies, ...)}
+\paragraph{Learning}
+
+\subsubsection{Grounding/anchoring strategies}
+\label{sect|grounding}
+
+\subsubsection{Ability to automatically create new object instances}
+\label{sect|new-instances}
+
+%%%%%%%%%%%%%%%%%
+\subsection{Practical Integration in Robotic Architectures}
 \label{sect|integration-robot}
 
-\subsubsection{Prediction and projection tasks}
-\label{sect|prediction-projection}
+\subsubsection{Lazy evaluation}
+\label{sect|lazy-evaluation}
 
-Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
-
-\paragraph{Projection task}: determining whether or not some condition while hold after a sequence of actions.
-
-\paragraph{Legality task}: determining whether a sequence of action can be performed starting in some initial state.
-
-\paragraph{Relation to task planning}
+\subsubsection{The planning task}
 
@@
 
+\paragraph{Language integration}
+
 \paragraph{Events}
 
-\subsubsection{Interaction with humans}
-\label{sect|hri}
+\subsubsection{Monitoring and debugging}
+\label{sect|debugging}
 
-Both as developers (KRS as a tool to make robot knowledge explicit to the
-robot designer) and end-users (like natural language processing,...)
-
-\subsubsection{Scalability and responsiveness}
+\subsubsection{Is it fast enough? Scalability and responsiveness}
 \label{sect|scalability}

### 2011-11-30 criteria.tex
Added a list of criteria as an independant tex file

--- 2011-11-30 krs-survey.tex
@@
 %\documentclass[twoside,a4paper]{report}
-\documentclass[a4paper, twocolumn]{article}
+\documentclass[a4paper]{article}
 
@@
 
-\definecolor{codegray}{gray}{.95}
-\definecolor{palegray}{gray}{.65}
-\lstset{language=C,
-	numbers=left,
-	tabsize=2,
-	basicstyle=\scriptsize,
-	stringstyle=\textrm,
-	showstringspaces=false,
-	frame=none,
-	xleftmargin=3pt,
-	backgroundcolor=\color{codegray}}
-
-\title{Current State of Knowledge Representation Systems for Robotics}
-\author{Séverin Lemaignan, Moritz Tenorth}
-
-\graphicspath{{figs/}}
-
 \newcommand{\ie}{{\textit{i.e.~}}}
@@
 
-\maketitle
-\tableofcontents
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sect|intro}
-
-The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
-In a chapter on that subject in \emph{Fondations of Artifical
-Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
-wrote together in 1998:
-
-\begin{quotation}
-
-	Central to this effort is to develop an understanding of the relationship
-	between the knowledge, the perception, and the action of such a robot. The
-	sorts of questions we want to be able to answer are
+\section{Compared Features}
 
 \begin{itemize}
-
-		\item to execute a program, what information does a robot need to have
-		at the outset vs. the information that it can acquire \emph{en route}
-		by perceptual means?
-
-		\item what does the robot need to know about its environment vs. what
-		need only be known by the designer?
-
-		\item when should a robot use perception to find out if something is
-		true as opposed to reasoning about what it knows was true in the past?
-
-		\item when should the inner workings of an action be available to the
-		robot for reasoning and when should the action be considered primitive
-		or atomic?
-
-	\end{itemize}
-
-	and so on. With respect to robotics, our goal (like that of many in AI) is
-	\emph{high-level robotic control}: develop a system that is capable of
-	generating actions in the world that are appropriate as a function of some
-	current set of beliefs and desires.
-
-\end{quotation}
-
-This survey focuses on the knowledge issue: we aims at establishing the current
-landscape of approaches to the knowledge representation problem in the research
-community, at summarizing the major trends and identifying the possible
-shortcomings.
-
-We have identified over ten projects that either are explicitely advertised as
-knowledge representation systems or involve explicit knowledge manipulation
-within a robot. By knowledge manipulation, we mean symbolic representation of
-assertions, be it static statements on the world or spatio-temporal events, and
-reasoning on this representation.
-
-We also try, for each project, to make clear \begin{inparaenum} \item how new
-facts can be acquired, \ie how informations from perception or interaction are
-turned into knowledge, \item how, in return, symbolic concept are
-\emph{anchored} into the robot sensori-motor space, \item how the knowledge
-representation system integrates with the decisional layers of the robot, and
-how this leads to better robotic control.\end{inparaenum}
-
-\subsection{Definition of inclusion criteria}
-\label{sect|inclusion-criteria}
-
-Every robotic system has, implicitely or not, some knowledge representation
-systems. It may range from a simple state vector to an explicit symbolic
-knowledge base.  This survey focuses on the right end of this spectrum:
-symbolic systems, suited for abstract reasoning.
-
-Besides, we have decided to restraint the set of systems to those actually
-implemented on robots, and used in semantic-rich environments (\ie , dynamic,
-partially unknown environments with a large range of different entities which
-may have interactions). The typical scenario that would involve such robots is
-a service robot in a human-friendly environment like a kitchen.
-
-More precisely, we have established the following criteria to select the
-knowledge representation systems to include in this survey. Those systems must:
+    \item{\bf What Can be Represented? The Expressiveness Question}
 
     \begin{itemize}
-	\item  Run on \emph{service robot} (robots that interact with objects in a
-	semantic-rich environment),
+        \item{\bf Introduction: main logic formalisms}
 
-	\item  Ground their knowledge in the physical world (physically embedded),
-	\begin{itemize}
-		\item  Able to \emph{resolve entities}
-		\item  Able to automatically create new object instances
-	\end{itemize}
+    [...]A large number of logic formalism do exist, we shall summarize below the most
+    relevant for systems actually deployed in current robotic architectures.[...]
 
-	\item  Be able to merge different knowledge modalities,
-	\item  Be endowed with on-line, dynamic reasoning (not just a static
-	dictionary)
-
-\end{itemize}
-
-\subsection{Evaluation of a knowledge representation system}
-\label{sect|evaluation}
-
-\subsubsection{Desirable features of a knowledge representation system}
-\label{sect|evaluation-literature}
-
-Several authors from fields that are connex to the robotics have previously listed desirable features of systems aiming at cognitive abilities as rich as possible.
-
-\paragraph{McCarthy: Towards human-level AI}
-
-In~\cite{McCarthy2007}, McCarthy lists the challenges he identifies on the road
-to a \emph{human-level AI}.
-
-\begin{itemize}
-
-	\item the ability to \emph{"operate successfully in the common sense
-	informatic situation"},
-
-	\item the necessity of relying on mathematical logic, as the most fruitful
-	formalism for machine intelligence,
-
-	\item the ability to deal with \emph{approximate concepts and approximate
-	theories} (that would include representing them, and reasoning with them),
-
-	\item nonmonotonic reasoning \fxfatal{give here a clear example},
-
-	\item what McCarthy calls \emph{Elaboration Tolerance}, \ie, the ability to
-	extend \emph{on requirement} the closed domain of interpretation for a
-	given assertion,
-
-	\item the ability to formalize and reason about contexts,
-
-	\item reasoning about events, and in particular, actions,
-
-	\item the capacity of introspection,
-
-	\item and finally, he points the issue of giving computer the right
-	heuristics for decision making.
-
-\end{itemize}
-
-Knowledge representation systems are concerned by most, if not all, of this
-points.
-
-\paragraph{Natural language processing in situated context}
-
-In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
-challenges to be tackled: cross-modal representation systems, association of
-words with perceptual and action categories, modeling of context, figuring out
-the right granularity of models, integrating temporal modeling and planning,
-ability to match past (learned) experiences with the current interaction and
-ability to take into account the human perspective.
-
-\subsubsection{Benchmarking}
-\label{sect|benchmarking}
-
-\cite{Hawes2008}: benchmarking of "Information-Processing Architectures on Intelligent Systems"
-\subsubsection{Pyschological tests}
-\label{sect|evaluation-tests}
-
-\begin{itemize}
-	\item Language comprehension tests: Token Test~\cite{DiSimoni1978}
-\end{itemize}
-
-\subsection{Methodology}
-\label{sect|methodology}
-
-\fxnote{Mention here that each team was proposed to complete/amend the description of their
-system}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Compared Features}
-\label{sect|compared-features}
-
-\subsection{What Can be Represented? The Expressiveness Question}
-\label{sect|intrinsic-features}
-
-%\subsubsection{Expressiveness}
-%\label{sect|expressiveness}
-
-\subsubsection{Introduction: main logic formalisms}
-
-The main role of a knowledge represenation system is to provide an adequate
-representation system to store facts and concepts that can be informally
-described in natural language.
-
-Formal logic aims at providing such a representation system with the added
-value of providing a tractable support for inference and reasoning.
-
-Most (but not all) of the systems we survey rely on a particular logic
-formalism. The choice of the formalism has a strong impact, on one side, on the
-range of ideas that can be expressed conveniently (\emph{practical
-expressiveness}) or at all (\emph{theoritical expressiveness}), on the other
-side, on the ability to solve the inference problem (called
-\emph{decidability}: is a given logical sentence true in my model?) in a
-tractable manner.
-
-A large number of logic formalism do exist, we shall summarize below the most
-relevant for systems actually deployed in current robotic architectures.
-
-\emph{Predicate logic} is the family of logic formalisms the most commonly
-found in knowledge representation. It distinguishes itself from the simpler
-\emph{propositional logic} by the use of quantification to increase generality.
-\emph{First-order logic} (FOL) is the subpart of \emph{predicate logic} where the
-objects of \emph{predicates} (or \emph{formulae}) are simple \emph{terms},
-while in \emph{higher-order logics}, predicates can be themselves objects of
-other predicates.
-
-The family of \emph{description logics}~\cite{Baader2008} also play an
-important role. It is a subset of the first-order logic, with some extensions
-in second-order logic. Description logics are notable because most of them are
-known to be decidable. In description logic, axioms are build from
-\emph{concepts}, \emph{roles} (that are unary or binary predicates) and
-\emph{individuals}. The W3C OWL-DL standard is a widely-used language to describe
-domain with the description logic.
-
-\emph{Modal logics}, that allow for statement qualification
-like \emph{possibility} or \emph{necessity}, have been shown to be closely
-related to description logics.
-
-\fxfatal{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
-
-One last class of logics that is of particular relevance for robotic
-applications is the \emph{probabilistic logics} or \emph{bayesian logics}.
-These logics provide a formal framework to reason on propositions whose truth
-or falsity is uncertain. We elaborate below on the representation of uncertainty.
-
-\paragraph{Some examples}
-
-\emph{The robot knows that a blue bottle is laying on the table.}
-
-\emph{The robot knows that the human knows about the position of the bottle,
-but the robot does not know what the human actually know about it.}
-
-\subsubsection{Open World and Close World Assumptions}
+        \item{\bf Open World and Close World Assumptions}
 
@@
     false (\emph{negation by failure}), while in the open world assumption, a
-proposition may be considered either true, false or unknown.
+    proposition may be considered either true, false or unknown. [...]
 
-This distinction is important in robotics were the robot may have to manipulate
-concepts with only partial knowledge on them. For instance, let imagine a robot
-that sees a bottle over a table, whose bottom is hidden by another object. The
-robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
-representation system relying on the closed world assumption would then assume
-the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
-whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
-table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
-a simple, concerete example of consequences of the CWA/OWA choice on reasoning.
+        \item{\bf Representation of uncertainty}
 
-\begin{table}
-	\begin{center}
-	\begin{tabular}{ll}
-	{\bf Action} & {\bf Part involved} \\
-	\hline
-	{\tt PickSoftly} & hand \\
-	{\tt PickAndPlace} & arm, hand \\
-	{\tt MoveArm} & arm \\
-	\hline
-	\end{tabular}
-	\end{center}
-	\caption{Assuming the question is: \emph{select actions that do not require
-	to move the arm}, a CWA reasoner would return {\tt PickSoftly} whereas an
-	OWA reasoner would not return anything if the {\tt PickSoftly} action is
-	not explicitely said not to involve the arm.}
-	\label{table|cwa-owa-example}
-\end{table}
+	Sources of uncertainty for a robot are two-fold: uncertainty
+	\emph{intrinsic} to facts (like \emph{``It may rain tomorrow''}),
+	uncertainty caused by imperfect perception of the world (\emph{``Is the
+	bottle really on the table?''}). Most logics do not account explicitely for
+	uncertainty. It must be either relied on specific logics (like Bayesian
+	logics) or on extensions of classical logics.
 
-Domains constrained with the closed world assumption lead to more tractable
-inference problems, and allow for instance the use of logic languages like
-Prolog. Thus, several approaches exists to \emph{locally close} a domain (\cf
-Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those).
+        \item{\bf Representation of time}
 
-\subsubsection{Representation of uncertainty}
+	As an agent acting at human-like time scale and dealing with temporal
+	concepts (like actions), a robot may want to represent, and possibly to
+	reason, about time. Time representation is split into two distinct
+	abilities: representing time points (both in the past -- which is roughly
+	equivalent to assignment of timestamps to events the robot perceives -- and
+	in the future), and representing \emph{passing time} (durations, timespans)
+	like in \emph{``the eggs will be cooked in 10 min''}.
 
-Sources of uncertainty for a robot are two-fold: uncertainty \emph{intrinsic}
-to facts (like \emph{``It may rain tomorrow''}), uncertainty caused by
-imperfect perception of the world (\emph{``Is the bottle really on the
-table?''}). Most logics do not account explicitely for uncertainty. It must be
-either relied on specific logics (like Bayesian logics) or on extensions of
-classical logics.
+	We call a system that do not account for time (\ie that permanently lives
+	in present) \emph{atemporal}.
 
-\subsubsection{Representation of time}
+        \item{\bf Context modeling}
 
-As an agent acting at human-like time scale and dealing with temporal concepts
-(like actions), a robot may want to represent, and possibly to reason, about
-time. Time representation is split into two distinct abilities: representing
-time points (both in the past -- which is roughly equivalent to assignment of
-timestamps to events the robot perceives -- and in the future), and
-representing \emph{passing time} (durations, timespans) like in \emph{``the
-eggs will be cooked in 10 min''}.
+	\emph{Knowledge is contextualized information}\fxfatal{Find someone
+	respectable how said that :-)}: it is essential for the robot to associate
+	the facts it represents to a \emph{context}. The context carries the keys
+	for the interpretation of the information. It covers the \emph{domain of
+	validity} of the facts, the \emph{common-sense} knowledge required to
+	fill the gaps in the representation\fxfatal{give an example}, \fxfatal{What more?}.
 
-We call a system that do not account for time (\ie that permanently lives in
-present) \emph{atemporal}.
+        \item{\bf Possible-Worlds and representing what others know}
 
-\subsubsection{Context modeling}
-
-\emph{Knowledge is contextualized information}\fxfatal{Find someone respectable
-how said that :-)}: it is essential for the robot to associate the facts it
-represents to a \emph{context}. The context carries the keys for the
-interpretation of the information. It covers the \emph{domain of validity} of
-the facts, the \emph{common-sense} knowledge required to fill the gaps in the
-representation\fxfatal{give an example}, \fxfatal{What more?}.
-
-\subsubsection{Possible-Worlds and representing what others know}
-\label{sect|possible-worlds}
-
-Linked to the context representation, but seen from another angle, knowledge
-representation systems may provide explicit ways to model other point of view
-on the world. This ability is often refered as the \emph{perspective taking}
-ability.
+	Linked to the context representation, but seen from another angle,
+	knowledge representation systems may provide explicit ways to model other
+	point of view on the world. This ability is often refered as the
+	\emph{perspective taking} ability.
 
@@
 
-\subsubsection{Introspection: knowledge about the knowledge}
-\label{sect|introspection}
-Support for introspection?
+        \item{\bf Introspection: knowledge about the knowledge}
 
-\subsubsection{Modifying its knowledge}
+        \item{\bf Modifying its knowledge}
 
@@
 
-%%%%%%%%%%
-\subsection{Reasoning Techniques}
+    \end{itemize}
 
-\subsubsection{Prediction and projection tasks}
-\label{sect|prediction-projection}
+    \item{\bf Reasoning Techniques}
+
+    \begin{itemize}
+            \item{\bf Prediction and projection tasks}
 
@@
 
-\paragraph{Projection task}: determining whether or not some condition while hold after a sequence of actions.
+        \begin{itemize}
+            \item{\bf Projection task}: determining whether or not some condition while hold after a sequence of actions.
+            \item{\bf Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+        \end{itemize}
 
-\paragraph{Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+        \item{\bf Reasoning in uncertainty}
 
-\subsubsection{Reasoning in uncertainty}
-
-\subsubsection{(Non) Monotonic Reasoning}
+        \item{\bf (Non) Monotonic Reasoning}
 
@@
         \emph{non-monotonic} reasoning scheme may lead to retraction of facts.
-McCarthy coined a famous example to illustrate the need of non-monotonic reasoning:
+        [...]
 
-\begin{quotation}
-Consider putting an axiom in a common sense database asserting that birds can
-fly. Clearly the axiom must be qualified in some way since penguins, dead birds
-and birds whose feet are encased in concrete can't fly. A careful construction
-of the axiom might succeed in including the exceptions of penguins and dead
-birds, but clearly we can think up as many additional exceptions like birds
-with their feet encased in concrete as we like. Formalized nonmonotonic
-reasoning provides a way of saying that a bird can fly unless there
-is an abnormal circumstance and reasoning that only the abnormal circumstances
-whose existence follows from the facts being taken into account will be
-considered.
-\end{quotation}
-
-Another important application of non-monotonic reasoning is representation of
-change: for example, to make an omelette, you need to crack eggs and wipe them.
-The eggs disappear and are replaced by an omelette:
-
-$Egg(a) \wedge Egg(b) \wedge MakeOmelette(a, b, c) \to \lnot Egg(a) \wedge
-\lnot Egg(b) \wedge Omelette(c)$
-
-The insertion of the proposition $MakeOmelette(a, b, c)$ leads to retraction of
-other facts. This rule requires non-monotonic reasoning to be applied.
-
-\emph{Default logic} is one of the formal logic that account for representing
-general truth and exceptions to it. However, due to computational complexity of
-these model (most of inferences in default logic are known to be $NP$-complete
-problem), classical logics and most of the existing reasoners do not allow
-non-monotonic reasoning. For instance, the SWRL rule language, usually
-associated to the OWL-DL ontology language, do not allow non-monotonic
-reasoning.
-
-\fxfatal{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
-languages? reasoner?}
-
-One important exception if the \emph{Negation as failure} inference rule, as
-implemented by {\sc Prolog} for instance, that allows for non-monotonicity, but
-only \emph{within the closed world assumption}.
-
-\fxfatal{Give here an example of non-monotonic reasoning with Prolog}
-
-A monotonic system does not theoritically allow for knowledge retractation,
-which is an important issue in the robotic context where the world model is
-likely to be often altered.  However it is a practical issue only if the
-reasoning process is \emph{continuous} during the whole robot's activity
-lifespan. It is often possible to stop the reasoner, alter the knowledge, and
-restart the inference process on a new domain.
-\fxfatal{Rephrase to emphazise that when new evidences appear, it is anyway often a
-good idea to restart the reasoner.}
-
-\fxfatal{Mention that the 'change of world' issue can also be dealt with appropriate
-time representation.}
-\fxfatal{Mention that probabilistic reasoning lead to implicit non-monotonic reasoning}
-
-\subsubsection{Presupposition accomodation}
-\label{sect|presupposition-accomodation}
+        \item{\bf Presupposition accomodation}
 
         \emph{Presupposition accomodation} is the ability for the system to
-automatically create a context allowing to make sense of a proposition.
+        automatically create a context allowing to make sense of a proposition.[...]
 
-For instance, we can imagine a human telling a robot \emph{Please get me the
-bottle that is behind you}. If the robot has not yet see what is behind it, it
-needs to assume (and represents in its knowledge model) that a undefined bottle
-can be found somewhere in the half of space behind it.
-
-A knowledge representation system able to copte with presupposition
-accomodation would be able to take into account this (usually under-defined)
-information that is not grounded into perception for later inferences.
-
-This ability to imagine a physically state of the world that is not actually
-perceived can be seen as the converse of the grounding ability.
-
-Note also that presupposition accomodation implies a bidirectional link of the
-symbolic knowledge model with a geometric (or physical) model of the
-environment. This article focuses on symbolic knowledge representation systems,
-but we shall mention when a KRS explicitely provides support for presupposition
-accomodation.
-
-\subsubsection{Learning}
-\label{sect|learning}
+        \item{\bf Learning}
+    \end{itemize}
 
 %%%%%%%%%%%%%%%%%
-\subsection{Acquiring Knowledge}
+    \item{\bf Acquiring Knowledge}
 
-\subsubsection{Knowledge acquisition and modalities merging}
-\label{sect|knowledge-acquisition}
+    \begin{itemize}
+        \item{\bf Knowledge acquisition and modalities merging}
+        \begin{itemize}
+            \item{\bf Perception}
+            \item{\bf Interaction}
+            \item{\bf External sources (Web, upper ontologies, ...)}
+            \item{\bf Learning}
+        \end{itemize}
 
-\paragraph{Perception}
-\paragraph{Interaction}
-\paragraph{External sources (Web, upper ontologies, ...)}
-\paragraph{Learning}
+        \item{\bf Grounding/anchoring strategies}
 
-\subsubsection{Grounding/anchoring strategies}
-\label{sect|grounding}
-
-\subsubsection{Ability to automatically create new object instances}
-\label{sect|new-instances}
+        \item{\bf Ability to automatically create new object instances}
+    \end{itemize}
 
 %%%%%%%%%%%%%%%%%
-\subsection{Practical Integration in Robotic Architectures}
-\label{sect|integration-robot}
+\item{\bf Practical Integration in Robotic Architectures}
 
-\subsubsection{Lazy evaluation}
-\label{sect|lazy-evaluation}
+    \begin{itemize}
+        \item{\bf Lazy evaluation}
 
-\subsubsection{The planning task}
+        \item{\bf The planning task}
 
-\subsubsection{Integration with executive layers}
-\label{sect|integration-executive-layers}
+        \item{\bf Integration with executive layers}
 
-\paragraph{Language integration}
+        \begin{itemize}
+            \item{\bf Language integration}
+            \item{\bf Events}
+        \end{itemize}
 
-\paragraph{Events}
+        \item{\bf Monitoring and debugging}
 
-\subsubsection{Monitoring and debugging}
-\label{sect|debugging}
+        \item{\bf Is it fast enough? Scalability and responsiveness}
+    \end{itemize}
 
-\subsubsection{Is it fast enough? Scalability and responsiveness}
-\label{sect|scalability}
-
-\subsection{Underlying knowledge model}
+\begin{itemize}
+    \item{\bf Underlying knowledge model}
 
@@
     \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Surveyed systems}
-
-Table \ref{table|surveyed-systems} presents the fifteen knowledge representation systems surveyed
-in this article.
-
-This section briefly presents each of them.
-
-\begin{landscape}
-\begin{table}
-\begin{center}
-\rowcolors{2}{lightgray}{codegray}
-
-%\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
-\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{2.8cm}p{1.5cm}}
-\hiderowcolors
-{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} & Main reference \\
-\hline
-\showrowcolors
-{\sc KnowRob} & KRS & Tenorth (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
-ORO & KRS & Lemaignan \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
-PEIS KR\&R & KRS & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
-CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
-Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
-NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
-GSM & & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
-OMKRF & & Suh et al. \par (Hanyang Univ.) & & & & & \cite{Suh2007} \\
-DY-KNOW & & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
- & & Wrighteagle \\
- & & Vincze \\
-ARMAR & & Schmidt-Rohr (Karlsruhe TH) \\
- & & Hertzberg (Osnabrück Univ.) \\
- & & (DFKI Bremen) \\
- (based on {\sc KnowRob} & & (JSK) \\
-
-\hline
-
-%\end{tabularx}
-\end{tabular}
-\end{center}
-\caption{List of surveyed systems.
-Projects are listed by category (\emph{KRS} for systems that are explicit knowledge representation and reasoning modules, \emph{Ubiquitous} for systems where knowledge processing is fully distributed, \emph{Language} for languages used as KRS on robots), then names.}
-\label{table|surveyed-systems}
-\end{table}
-\end{landscape}
-
-\fxfatal{Bielefeld -> could not find much...}
-\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}
-
-\subsection{KnowRob}
-\label{sect|knowrob}
-
-\subsection{ORO}
-\label{sect|oro}
-
-\subsection{PEIS KR\&R}
-\label{sect|peis-ecology}
-
-{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim to binds autonomous
-robotics with ambient intelligence (network of sensors). \emph{PEIS} stands for
-\emph{Physically Embedded Intelligent System}: every robots or intelligent
-device in the environment is abstracted as a PEIS.
-
-Each PEIS physical component is running a \emph{PEIS Kernel} instance. Communication
-between instance relies on a custom P2P communication protocol.
-
-The PEIS architecture allows for adding new abilities through software components sharing the common \emph{tuple space}.
-
-We survey here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS KR\&R}, that includes symbolic representation and reasoning.
-
-% More in details:
-% - object identification based on viewpoint independent SIFT features
-% - formalized anchoring system that explicitely match percieved attributes to predicates
-% - Cyc predicates
-% - ground 12 colors, based on a paper on color perception. Could be useful for us.
-% - idem, they cite a paper on what spatial relations to compute
-% - location of objects based on a previously provided semantic map (but not much on this semantic map)
-% - two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
-% - uses directly Cyc (ie, 250 000 common sense concepts...), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
-% Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
-% - disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
-% - template based natural language
-% - experiment conducted in a "smart" indoor environmement + simple robot
-
-\subsubsection{Intrinsic features}
-\label{sect|peis-intrinsic-features}
-
-\paragraph{Expressiveness} The PEIS Knowledge representation system relies on
-the {\sc ResearchCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
-allows to represent first order logic sentences and has extensions for modal logics and higher order logics.
-
-\fxfatal{Is modal logics and higher order logics actually used in PEIS?}
-
-As a system relying on {\sc CycL}, contexts can be expressed as
-\emph{microtheories}: the truth or falsity of a set of statement depends of the
-\emph{microtheory} in which these statements are evaluated.
-
-\fxfatal{OWA/CWA?}
-
-\subsubsection{Anchoring strategies}
-\label{sect|peis-anchoring}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\columnwidth]{peis-architecture.pdf}
-	\caption{The PEIS knowledge representation system, taken from~\cite{Daoutis2009}}
-	\label{fig|peis-archi}
-\end{figure}
-
-The PEIS KR\&R system is deeply integrated to the general PEIS Ecology
-\emph{smart} environment. Figure~\ref{fig|peis-archi} gives an overview of the
-interactions between PEIS knowledge processing layers.
-
-\paragraph{Knowledge Acquisition} The primary source for knowledge acquisition
-is perception.  The PEIS ecosystem provides a SIFT-based object recognizer used
-in conjunction with ceiling cameras for object localization.  Other perceptual
-modalities are available (like human tracking, ambient environment monitoring).
-
-A template-based natural language parsing system may also be used to add new
-assertions to the system.
-
-\paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
-finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
-\Phi \times D(\Phi)$, where $\mathcal{P}$ is a set of predicate symbols, $\Phi$
-a set of percept's attributes, and $D(\Phi)$ the domain of these attributes.
-
-In the current implementation, object category (returned by the SIFT
-classifier), color, location, spatial relations (both topological -- \emph{at},
-\emph{near} -- and relative to the robot -- \emph{left}, \emph{behind}, etc.)
-and visibility are the five classes of extracted attributes.
-
-\subsubsection{Integration in the robot architecture}
-\label{sect|peis-integration}
-
-The PEIS framework offers through the \emph{PEIS middleware} a practical way to
-insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
-module can be seemlessly integrated into the PEIS ecosystem.
-
-\subsubsection{Notable experiments}
-\label{sect|peis-expe}
-
-\subsection{NKRL}
-\label{sect|nkrl}
-
-\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
-While this language is developped since a long time by Zarri~\cite{Zarri1997,
-Zarri2008}, recent research direction include application to the robotic
-field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
-system, as it is primarily a language. However, it is used as the
-representation and reasoning mechanism for robots by Sabri et al.
-
-\subsubsection{Intrinsic language features}
-\label{sect|nkrl-intrinsic-features}
-
-\paragraph{Expressiveness}
-
-\subsubsection{Integration with physical world and in the robot architecture}
-\label{sect|nkrl-integration}
-
-...seem to be mostly WIP...
-
-\subsection{CAST Knowledge model}
-\label{sect|cast}
-
-CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
-comprehensive toolkit aimed at builduing cognitive architectures for robots
-through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
-expose a central knowledge base as seen in previous works. It instead
-represents knowledge as unrooted \emph{proxies}. Thoses proxies are formally
-defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
-a set of instanciated features (like $\phi^{Colour}_{red}$) and $u_p$ a
-\emph{proxies union} that form an equivalence class corresponding to one
-entity.
-
-A union of proxies forms a global amodal representation of an entity, that can
-be explicitely shared and manipulated. Being not centralized, the knowledge
-model can be qualified of \emph{ubiquitous}. Futhermore, knowledge source in
-the CAS architecture is tightly bound to the on-line grounding process (be it
-grounded in perception or in dialogue). While nothing seems to prevent it, no
-{\it a priori} knowledge (including common-sense knowledge) is used.
-
-Knowledge sharing is ensured by the event mechanism of CAST: modules can
-monitor proxies for alteration by other modules. Jacobsson et al. mention how
-this can apply to reinforcement learning: the vision module creates a proxy for
-an orange object. This proxy get monitored by a learning module. In parallel,
-the proxy is bound to an union by the natural language understanding module
-that add new a feature like \emph{"this object is a fruit"}. The learning
-module is called back, and can add this new information to its model.
-
-In the presented implementation, the CAST knowledge model does no allow for
-effectively representing actions or temporal information.\fxfatal{What about reasoning? can they retrieve for example 'all proxies for colorful objects'?}
-
-%%%%%%%%%% Underlying knowledge model table %%%%%%%%
-\begin{table}
-\begin{center}
-\rowcolors{2}{lightgray}{codegray}
-
-\begin{tabular}{lp{4cm}}
-\hiderowcolors
-{\bf Project} & {\bf Common-sense \par knowledge source} \\
-\hline
-\showrowcolors
-{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
-ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
-PEIS Ecology & {\sc ResearchCyc} \\
-DY-KNOW & \\
-OMKRF & \\
-GSM &  \\
-NKLR &  None \\
-CAST Proxies &  None \\
-Wrighteagle & \\
-Vincze & \\
-ARMAR &  \\
-Hertzberg (Osnabrück) & \\
-(DFKI Bremen) & \\
-(JSK) & \\
-
-\hline
-
-\end{tabular}
-\end{center}
-\caption{Underlying knowledge sources for each project}
-\label{table|knowledge-sources}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary of Approaches}
-\label{sect|summary}
-
-This section tries to summary the various approaches we surveyed in the
-previous section.  This summary does not exactly follow the list of compared
-features presented in section~\ref{sect|compared-features}. We have organized
-it along four axis: \emph{what can be represented?}, \emph{How knowledge is
-created and grounded?}, \emph{What can be done with the knowledge?} and
-\emph{How to use knowledge in the whole, larger robot architecture?}.
-
-\subsection{Expressiveness: what can we represent in the current state-of-the-art?}
-\label{sect|summary-expressiveness}
-
-\begin{itemize}
-	\item expressiveness
 \end{itemize}
-
-\subsection{Creating and grounding knowledge in the physical world}
-\label{sect|summary-grounding}
-
-\begin{itemize}
-	\item which knowledge modalities can be merged
-	\item which grounding strategies
 \end{itemize}
-
-\subsection{Exploiting knowledge}
-\label{sect|summary-knowledge-sources-reasoning}
-
-\begin{itemize}
-	\item which common-sense knowledge is used
-	\item what kind of reasoning is made
-	\item introspection
-\end{itemize}
-
-\subsection{Using a knowledge representation system in a larger robotic cognitive architecture}
-\label{sect|summary-integration}
-
-\begin{itemize}
-	\item a priori evaluation/lazy evaluation and trade-offs (relations to events, scalability, detection of inconsistencies,...)
-	\item relation with planning (action representation, temporal representation...)
-	\item API/bindings/integration into executive layers
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Towards the next generation of Knowledge Representation Systems for Robotics}
-\label{sect|conclusion}
-
-\subsection{Current shotcomings: what is not successfully covered by current systems}
-
-\subsection{Towards a new design}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section*{Acknowledgements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2011-12-01 krs-survey.tex
Typos

@@
 
-Several authors from fields that are connex to the robotics have previously listed desirable features of systems aiming at cognitive abilities as rich as possible.
+Several authors from fields that are connected to the robotics have previously listed desirable features of systems aiming at cognitive abilities as rich as possible.
 
@@
 
-	\item nonmonotonic reasoning \fxfatal{give here a clear example},
+	\item non-monotonic reasoning \fxfatal{give here a clear example},
 
@@
 \cite{Hawes2008}: benchmarking of "Information-Processing Architectures on Intelligent Systems"
-\subsubsection{Pyschological tests}
+\subsubsection{Psychological tests}
 \label{sect|evaluation-tests}
@@
 
-The main role of a knowledge represenation system is to provide an adequate
+The main role of a knowledge representation system is to provide an adequate
 representation system to store facts and concepts that can be informally
@@
 range of ideas that can be expressed conveniently (\emph{practical
-expressiveness}) or at all (\emph{theoritical expressiveness}), on the other
+expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
 side, on the ability to solve the inference problem (called
@@
 One last class of logics that is of particular relevance for robotic
-applications is the \emph{probabilistic logics} or \emph{bayesian logics}.
+applications is the \emph{probabilistic logics} or \emph{Bayesian logics}.
 These logics provide a formal framework to reason on propositions whose truth
@@
 table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
-a simple, concerete example of consequences of the CWA/OWA choice on reasoning.
+a simple, concrete example of consequences of the CWA/OWA choice on reasoning.
 
@@
 	OWA reasoner would not return anything if the {\tt PickSoftly} action is
-	not explicitely said not to involve the arm.}
+	not explicitly said not to involve the arm.}
 	\label{table|cwa-owa-example}
@@
 imperfect perception of the world (\emph{``Is the bottle really on the
-table?''}). Most logics do not account explicitely for uncertainty. It must be
+table?''}). Most logics do not account explicitly for uncertainty. It must be
 either relied on specific logics (like Bayesian logics) or on extensions of
@@
 birds, but clearly we can think up as many additional exceptions like birds
-with their feet encased in concrete as we like. Formalized nonmonotonic
+with their feet encased in concrete as we like. Formalized non-monotonic
 reasoning provides a way of saying that a bird can fly unless there
@@
 
-A monotonic system does not theoritically allow for knowledge retractation,
+A monotonic system does not theoretically allow for knowledge retractation,
 which is an important issue in the robotic context where the world model is
@@
 restart the inference process on a new domain.
-\fxfatal{Rephrase to emphazise that when new evidences appear, it is anyway often a
+\fxfatal{Rephrase to emphasize that when new evidences appear, it is anyway often a
 good idea to restart the reasoner.}
@@
 
-\subsubsection{Presupposition accomodation}
+\subsubsection{Presupposition accommodation}
 \label{sect|presupposition-accomodation}
 
-\emph{Presupposition accomodation} is the ability for the system to
+\emph{Presupposition accommodation} is the ability for the system to
 automatically create a context allowing to make sense of a proposition.

### 2011-12-01 criteria.tex
Updated criteria following latest discussion with Michael

@@
 \begin{document}
+
+\section{Survey Inclusion criteria}
+
+\begin{itemize}
+\item  Run on \emph{service robot} (robots that interact with objects in a
+semantic-rich environment),
+
+\item  Ground their knowledge in the physical world (physically embedded),
+\begin{itemize}
+\item  Able to \emph{resolve entities}
+\item  Able to automatically create new object instances
+\end{itemize}
+
+\item  Be able to merge different knowledge modalities,
+\item  Be endowed with on-line, dynamic reasoning (not just a static
+		dictionary)
+
+\end{itemize}
 
@@
 
-	As an agent acting at human-like time scale and dealing with temporal
-	concepts (like actions), a robot may want to represent, and possibly to
-	reason, about time. Time representation is split into two distinct
+	[...] Time representation is split into two distinct
 	abilities: representing time points (both in the past -- which is roughly
@@
 	like in \emph{``the eggs will be cooked in 10 min''}.
+
+    Mention here \emph{Chronicles}.
 
@@
 	knowledge representation systems may provide explicit ways to model other
-	point of view on the world. This ability is often refered as the
+	point of view on the world. This ability is often referred as the
 	\emph{perspective taking} ability.
@@
 
-        \item{\bf Introspection: knowledge about the knowledge}
-
-        \item{\bf Modifying its knowledge}
-
-        In DL -> always possible to modify the ABox, not always possible to alter the TBox
+        \item{\bf Introspection: What can I do?}
+		\begin{itemize}
+			\item{\bf General introspection}
+			The ability for the robot to represent its own capabilities.
+			\item{\bf Meta-cognition: knowledge about the knowledge}
+		\end{itemize}
 
@@
     \begin{itemize}
-            \item{\bf Prediction and projection tasks}
 
-            Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
+        \item{\bf Standard Reasoning Techniques}
+		\begin{itemize}
 
-        \begin{itemize}
-            \item{\bf Projection task}: determining whether or not some condition while hold after a sequence of actions.
-            \item{\bf Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+			\item{\bf Inferences and backtracking}
+			\item{\bf Classification}			
+			\item{\bf Inconsistencies detection}
+
 		\end{itemize}
 
-        \item{\bf Reasoning in uncertainty}
+        \item{\bf Reasoning with uncertainty}
 
@@
 
-        \item{\bf Presupposition accomodation}
+        \item{\bf Presupposition accommodation}
 
-        \emph{Presupposition accomodation} is the ability for the system to
+        \emph{Presupposition accommodation} is the ability for the system to
         automatically create a context allowing to make sense of a proposition.[...]
+
+			\item{\bf Prediction and projection tasks}
+
+			Levesque~\cite{Levesque2008} distinguish two main tasks, the
+			\emph{projection task} and the \emph{legality task}.
+
+		\begin{itemize}
+
+		\item{\bf Projection task}: determining whether or not
+		some condition while hold after a sequence of actions.
+
+		\item{\bf Legality task}: determining whether a sequence of action can be
+		performed starting in some initial state.
+
+		\item{\bf Diagnosis}
+
+		\end{itemize}
+
+		\item{\bf Planning}
+
+		\item{\bf Modifying its knowledge}
+
+        In DL -> always possible to modify the ABox, not always possible to alter the TBox
 

### 2011-12-01 krs-survey.tex
Updated criteria following latest discussion with Michael

@@
 
+\fxfatal{Discuss time chronicles~\cite{Ghallab1996}}
+
 We call a system that do not account for time (\ie that permanently lives in
@@
 representation systems may provide explicit ways to model other point of view
-on the world. This ability is often refered as the \emph{perspective taking}
+on the world. This ability is often referred as the \emph{perspective taking}
 ability.
@@
 
-\subsubsection{Introspection: knowledge about the knowledge}
+\subsubsection{Introspection: what can I do?}
 \label{sect|introspection}
-Support for introspection?
 
-\subsubsection{Modifying its knowledge}
+\paragraph{General introspection}
 
-In DL -> always possible to modify the ABox, not always possible to alter the TBox
+Modelling of the robot capabilities.
+
+\paragraph{Meta-cognition: knowledge on the knowledge}
+
+Narrower, more technical dimension of the introspection.
 
@@
 
-\subsubsection{Prediction and projection tasks}
-\label{sect|prediction-projection}
+\subsubsection{Standard reasoning techniques}
 
-Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.
+Backtracking,...
 
-\paragraph{Projection task}: determining whether or not some condition while hold after a sequence of actions.
+\paragraph{Decidability}
 
-\paragraph{Legality task}: determining whether a sequence of action can be performed starting in some initial state.
+...
 
-\subsubsection{Reasoning in uncertainty}
+\subsubsection{Reasoning with uncertainty}
 
@@
 A knowledge representation system able to copte with presupposition
-accomodation would be able to take into account this (usually under-defined)
+accommodation would be able to take into account this (usually under-defined)
 information that is not grounded into perception for later inferences.
@@
 
-Note also that presupposition accomodation implies a bidirectional link of the
+Note also that presupposition accommodation implies a bidirectional link of the
 symbolic knowledge model with a geometric (or physical) model of the
 environment. This article focuses on symbolic knowledge representation systems,
-but we shall mention when a KRS explicitely provides support for presupposition
-accomodation.
+but we shall mention when a KRS explicitly provides support for presupposition
+accommodation.
+
+\subsubsection{Prediction and projection and diagnosis tasks}
+\label{sect|prediction-projection}
+
+Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection
+task} and the \emph{legality task}.
+
+\paragraph{Projection task}: determining whether or not some condition while
+hold after a sequence of actions.
+
+\paragraph{Legality task}: determining whether a sequence of action can be
+performed starting in some initial state.
+
+\paragraph{Diagnosis}: this corresponds to the ability to rewind on past events
+in case of failure to provide possible explanation. This can be seen as the
+temporal reverse of the projection task.
+
+\subsubsection{Planning}
+\label{sect|planning}
+
+Making decision based on prediction
+
+\subsubsection{Modifying the knowledge structure}
+
+All systems allow to modify the ABox, not always possible to alter the TBox
 
@@
 
-\subsubsection{The planning task}
-
 \subsubsection{Integration with executive layers}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Underlying knowledge model}

### 2011-12-08 krs-survey.tex
Added a short 'article overview'

@@
 
+\subsection{Article overview}
+\label{sect|overview}
+
+The paper is organized in four main sections.
+
+The next section proposes a prototypical scenario in a rich semantic
+environment. We will not survey knowledge representation systems based on this
+scenario: we merely use it to ground the selection and illustrate the set
+features (Section~\ref{sect|features}) that define the dimensions of a
+knowledge representation system for robotics.
+
+Section~\ref{sect|surveyed-systems} then presents each of the systems we
+examinated, and underlines their specificities.
+
+We try in section~\ref{sect|summary} to summarize the current approaches and to
+relate them to the previously identified set of features. This leads us to
+identify several new research directions that are not sufficiently addressed by
+the current state of the art.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The Cognitive Scenario}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Compared Features}
-\label{sect|compared-features}
+\label{sect|features}
 

### 2011-12-08 krs-survey.tex
After discussion with Moritz and Michael, removed subsections on benchmarking
Reworked organization of features list after talk with Moritz & Michael

@@
 \title{Current State of Knowledge Representation Systems for Robotics}
-\author{Séverin Lemaignan, Moritz Tenorth}
+\author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz}
 
@@
 
-\subsection{Evaluation of a knowledge representation system}
-\label{sect|evaluation}
-
-\subsubsection{Desirable features of a knowledge representation system}
+\subsubsection{What to expect from a knowledge representation system}
 \label{sect|evaluation-literature}
 
-Several authors from fields that are connected to the robotics have previously listed desirable features of systems aiming at cognitive abilities as rich as possible.
+Several authors from fields that are connected to the robotics have previously
+listed desirable features of systems aiming at cognitive abilities as rich as
+possible.
 
@@
 
-\subsubsection{Benchmarking}
-\label{sect|benchmarking}
-
 \cite{Hawes2008}: benchmarking of "Information-Processing Architectures on Intelligent Systems"
-\subsubsection{Psychological tests}
-\label{sect|evaluation-tests}
-
-\begin{itemize}
-	\item Language comprehension tests: Token Test~\cite{DiSimoni1978}
-\end{itemize}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Cognitive Scenario}
+\section{A Prototypical Scenario}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Compared Features}
+\section{A taxonomy of features of knowledge representation systems}
 \label{sect|features}
 
-\subsection{What Can be Represented? The Expressiveness Question}
-\label{sect|intrinsic-features}
-
-%\subsubsection{Expressiveness}
-%\label{sect|expressiveness}
+\subsection{Expressiveness: What Can be Represented?}
+\label{sect|expressiveness}
 
@@
 
-\subsubsection{Representation of time}
+\subsubsection{Meta-cognition: knowledge on the knowledge}
+
+Narrower, more technical dimension of the introspection.
+
+%%%%%%%%%%
+\subsection{How things are represented?}
+\label{sect|higher-level-domain-representation}
+
+\subsubsection{Spatio-Temporal Representations}
+
+\paragraph{Representation of time}
 
@@
 present) \emph{atemporal}.
+
+\paragraph{Representation of space}
+
+\paragraph{Representation of events and actions}
 
@@
 
-\subsubsection{Introspection: what can I do?}
+\subsubsection{Introspection: Who am I? What can I do?}
 \label{sect|introspection}
 
-\paragraph{General introspection}
+\paragraph{The introspective robot}
 
-Modelling of the robot capabilities.
-
-\paragraph{Meta-cognition: knowledge on the knowledge}
-
-Narrower, more technical dimension of the introspection.
+\paragraph{Modelling of the robot capabilities}
 
@@
 \subsection{Reasoning Techniques}
+\label{sect|reasoning}
 
@@
 ...
+
+\subsubsection{Lazy evaluation}
+\label{sect|lazy-evaluation}
 
@@
 
-\subsubsection{Prediction and projection and diagnosis tasks}
+\subsubsection{Prediction, projection and diagnosis tasks}
 \label{sect|prediction-projection}
@@
 
-\subsubsection{Lazy evaluation}
-\label{sect|lazy-evaluation}
+\subsubsection{Integration with sensori-motor layers}
+\label{sect|integration-sensorimotor}
+
+Ability to ``listen'' to the robot internal structures.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Underlying knowledge model}
+\subsection{Knowledge instanciation}
+
+How much knowledge is available? Which content? How big is the knowledge base?
 
@@
 \section{Surveyed systems}
+\label{sect|surveyed-systems}
 

### 2011-12-08 krs-survey.tex
Added references to Chen (WrightEagle team), Vincze, Lozano-Perez

@@
 \label{sect|features}
+
+\fxnote{Should we also mention the general dynamics of the knowledge representation system? ie, the
+system from the point of view of an on-going cognitive process. See \cite{Henderson2011}}
 
@@
 \showrowcolors
-{\sc KnowRob} & KRS & Tenorth (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
-ORO & KRS & Lemaignan \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
+{\sc KnowRob} & KRS & Tenorth, Beetz \par (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
+ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
 PEIS KR\&R & KRS & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
@@
 DY-KNOW & & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
- & & Wrighteagle \\
- & & Vincze \\
+Ke Jia & & Chen et al. \par (Univ. of Science and Technology of China) & & & & & \cite{Chen2010} \\
+ & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\
+ & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\
 ARMAR & & Schmidt-Rohr (Karlsruhe TH) \\
@@
 previous section.  This summary does not exactly follow the list of compared
-features presented in section~\ref{sect|compared-features}. We have organized
+features presented in section~\ref{sect|features}. We have organized
 it along four axis: \emph{what can be represented?}, \emph{How knowledge is
@@
 \label{sect|summary-grounding}
+
+\fxnote{Mention symmetry and research of invariants \cite{Henderson2011} as a process to create knowledge?}
 

### 2011-12-12 krs-survey.tex
Added most of the 'scenario' section

@@
 
+\fxnote{Add nice illustrations from MORSE}
+
+In order to illustrate what we consider to be the required features of a
+knowledge representation system for robotics, we propose the following
+scenario.
+
+While devising a scenario inevitably constraints the breadth of abilities we
+examine to the ones actually needed for this specific scenario, we insist on
+the fact that our main purpose is to demonstrate on a real case what
+representational features are desirable.
+
+We entitle our scenario ``the Brownie scenario''.
+
+Robi and Roba are two service robots. They are considered to be able to freely
+move and pick objects, but they are not expected to have the exact same
+hardware and softwares architectures, and especially, they are not expected to
+have the same knowledge representation system. These robots cooperate with a
+human in a kitchen environment.
+
+The main task of the scenario is the joint realization of a brownie, that
+starts with the human injunction ``Let's make a brownie for tonight!''.
+
+The scenario is successful if the task is achieved (the brownie is made) and:
+
+\begin{itemize}
+
+	\item it took less time that it would have require for the human alone,
+
+	\item it didn't require a heavier cognitive involvement from the human that
+	what would have been required without the robots' help.
+
+\end{itemize}
+
+We voluntarily do not try to detail the subtasks of the scenario and to define
+how they are shared amongst agents. In our analysis we focus on the general,
+\textit{a priori} representational needs of the scenario.
+
+A ``first-order'' analysis of this task leads to a first list of categories of
+required features:
+
+\begin{enumerate}
+
+	\item Representation abilities related to the execution of a complex
+	spatio-temporal task,
+
+	\item Representation abilities related to cooperation with other agents.
+
+\end{enumerate}
+
+% Representation of a complex task
+The first category can be further refined: to prepare and bake a brownie, the
+robot first needs to make sense of the very term \emph{brownie}: what is it?
+what is it used for? what is it made of? etc. We call this knowledge
+\emph{common-sense knowledge} and the robot must be able not only to represent
+it, but also to have some kind of access to it (for instance through a initial
+set of facts that are made available at startup, or via access to a Web-based
+knowledge base like Wikipedia, etc.)
+
+In conjunction with the representation of the action \emph{make} (and it's
+possibly attached common-knowledge), this should lead the robot to build and
+represent a \emph{context}: we are in a scenario involving cooking. Actions
+related to cooking often take place in the kitchen, cooking requires ingredients, 
+utensils and a procedure that may be provided by a recipe, etc.
+
+This last sentence implies several other features for our knowledge
+representation system: ``cooking often takes place in the kitchen'' implies
+that representation of both uncertainty and likelihood is desirable. The fact
+that cooking is associated to a place further implies that the system models
+locations and is able to attach \emph{thematic relations} to concept
+(here, the likely location of the cooking action).
+
+``cooking requires ingredients that may be provided by a recipe'' hints about a
+very common feature available in most knowledge representation systems:
+\emph{reasoning}. The robot \emph{infers} that cooking may require a recipe
+since a list of ingredients is a pre-requisite of the cooking action, and a
+recipe may provided such a list. If we omit the ``may'', this is a typical
+example of first-order logic reasoning. Many other reasoning techniques exist
+(including probabilistic ones -- ones able to deal with the ``may''), we shall
+illustrate some of them further in this scenario.
+
+As mentioned, a recipe often provides a procedure (or a \emph{plan}). The robot
+should be able to store this plan in a way that allow later execution.  The
+plan is likely to contain \emph{spatio-temporal constraints} (like ``put
+the brownie in the oven for 20 min'') that must as well be appropriately
+represented.
+
+Moreover, procedures are often \emph{underspecified}: we can expect the recipe
+to provide a cooking duration, but we usually do not expect the recipe to tell
+us to first open the oven door, and then put the cake into it, since it is
+self-evident that the door must first be opened. Such underspecification should
+be detectable, representable, and ideally completable by the knowledge
+representation system.
+
+Besides representation of space and time, we may have
+
+% Representation feature that enable cooperation
+Then, we want our three agents to cooperate. This, in turn, leads to a very
+different set of cognitive abilities.
+
+Cooperation in our scenario can intervene at many levels: for instance, an
+agent may want to inform another one about the number of eggs that are
+necessary for the brownie. It makes sense to help in this way only if the first
+agent knows that the recipient agent both needs the information and does not
+know it. This in turn requires the robot to be able to model the knowledge of
+the other agents: to think \emph{from the perspective} of another agent (this
+idea is related to the so-called Theory of Mind~\cite{...}).
+
+Ability to communicate is one important pre-requisite to collaboration.
+Communication in general~\cite{Jakobson} requires the addresser and the
+addressee to share a common interpretative framework (shared common-sense
+knowledge -- or cultural background -- and a shared context). In our scenario,
+the agents are working in a kitchen. This element of context does not however
+suffice if, for example, an agent asks another agent to ``give {[him]} the
+bowl''. Besides the symbol ``bowl'', which physical entity are we actually
+talking about? If we want to act on the world, this \emph{grounding} operation
+is essential, and may be tightly bound to the underlying knowledge
+representation system. Note that grounding usually denotes the {\it top-down}
+operation: from the symbol to the percept. It is however commonly bundled with
+the converse operation, that retrieve (or create) symbols from perception.
+
+A related ability is called \emph{pre-supposition accomodation}: if one of the
+agent moves behind another one, with the brownie dough in its arm, and says
+``be careful, I'm behind you!'', we want the first agent to be able to
+represent both symbolically and geometrically (because, for instance, if the
+agent want to move, it must take into account the new obstacle) something that
+was not directly perceived. A knowledge representation system may be able to
+provide structures to represent and reason about such cases.
+
+Also central to cooperation are the notions of \emph{joint
+intentions} and \emph{joint goals}~\cite{Tomasello2005, Bratman2009}: to help
+the human during the cooking session, the robots need to store at least how far
+they are into the recipe, what is the next step the human is likely to go for,
+how task are currently split between agents, what action is currently
+blocking the procedure, etc. This knowledge should let the robot identify the
+intentions of other agents and create accordingly joint goals. Hence, a
+knowledge representation system aiming at dealing with cooperative behaviours
+is likely to have goal management structures taking explicitly into account
+other agents' actions and goals.
+
+In order to effectively share tasks, the robot must also know what it is
+capable of: \emph{capability introspection} (both in term of general capability
+and of immediate ability) is thus often desirable. More general introspection
+(like the ability to tell ``who I am'' or ``what do I think of'' does not
+appear to be necessary in our scenario. This may however enrich the human-robot
+interaction.
+
+Last but not least, our scenario assumes implicitly \emph{natural interaction}
+between humans and robots (as showed by the casual style of the
+order ``Let's make a brownie!''). While natural language processing {\it
+per-se} is usually out of the scope of a knowledge representation system, we
+may want to be sure these systems successfully interoperate, \ie that the
+knowledge representation system provides efficient support to the natural
+language processing module (for instance by adopting models and vocabulary that
+are both well suited for machine processing and remain as close as possible to
+the humans own structures and vocabulary).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A taxonomy of features of knowledge representation systems}
+\section{A typology of features of knowledge representation systems}
 \label{sect|features}
 
-\fxnote{Should we also mention the general dynamics of the knowledge representation system? ie, the
-system from the point of view of an on-going cognitive process. See \cite{Henderson2011}}
+The previous section detailed an imaginary cooking scenario. We identified, in
+no particular order, numerous desirable features for a knowledge representation
+system for robotics.
+
+This section proposes a more formal typology of desirable features for a such a
+system. For each feature, we provide a short definition along with links to
+relevant literature.
+
+\fxnote{Should we also mention the general dynamics of the knowledge
+representation system? ie, the system from the point of view of an on-going
+cognitive process. See \cite{Henderson2011}}
 
@@
 
-\subsubsection{Representation of uncertainty}
+\subsubsection{Representation of uncertainty and likelihood}
 
@@
 
-\subsubsection{Spatio-Temporal Representations}
+\subsubsection{Role Representations}
+
+Spatio-Temporal Representations:
 

### 2011-12-13 krs-survey.tex
Added a picture of the Brownie scenario, made in MORSE

@@
 
-\fxnote{Add nice illustrations from MORSE}
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
+	\caption{A representation of the scenario, in the MORSE simulator}
+	\label{fig|scenario}
+\end{figure*}
 
@@
 
-We entitle our scenario ``the Brownie scenario''.
+We entitle our scenario ``the Brownie scenario'' (Fig.~\ref{fig|scenario}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A typology of features of knowledge representation systems}
+\section{A typology of features of KRS for robotics}
 \label{sect|features}

### 2011-12-13 krs-survey.tex
Proof-read and extended the scenario part

@@
 In order to illustrate what we consider to be the required features of a
-knowledge representation system for robotics, we propose the following
-scenario.
+knowledge representation system for robotics, we introduce in this section a
+demonstration scenario.
 
 While devising a scenario inevitably constraints the breadth of abilities we
-examine to the ones actually needed for this specific scenario, we insist on
-the fact that our main purpose is to demonstrate on a real case what
-representational features are desirable.
+examine, we insist on the fact that our main purpose here is to ground into a
+real case the set of representational features we consider as desirable.  We
+will \emph{not} evaluate any of the systems we survey in this article by
+judging their applicability to this particular scenario.
 
@@
 move and pick objects, but they are not expected to have the exact same
-hardware and softwares architectures, and especially, they are not expected to
+hardware and softwares architectures, and specially, they are not expected to
 have the same knowledge representation system. These robots cooperate with a
@@
 
-The main task of the scenario is the joint realization of a brownie, that
-starts with the human injunction ``Let's make a brownie for tonight!''.
-
-The scenario is successful if the task is achieved (the brownie is made) and:
+The main task of the scenario is the joint realization of a brownie, initiated
+by the human injunction ``Let's make a brownie for tonight!''.
+
+The scenario is successful if the task is achieved (the brownie is baked) and:
 
@@
 
-We voluntarily do not try to detail the subtasks of the scenario and to define
+We voluntarily do not detail the subtasks of the scenario, neither we define
 how they are shared amongst agents. In our analysis we focus on the general,
@@
 
-A ``first-order'' analysis of this task leads to a first list of categories of
-required features:
+A ``first-order'' analysis of this task leads to a first partition of the
+required representation abilities:
 
@@
 % Representation of a complex task
-The first category can be further refined: to prepare and bake a brownie, the
+We can further refine these categories: to prepare and bake a brownie, the
 robot first needs to make sense of the very term \emph{brownie}: what is it?
@@
 
-In conjunction with the representation of the action \emph{make} (and it's
-possibly attached common-knowledge), this should lead the robot to build and
+Bound to the action \emph{make}, this should lead the robot to build and
 represent a \emph{context}: we are in a scenario involving cooking. Actions
@@
 since a list of ingredients is a pre-requisite of the cooking action, and a
-recipe may provided such a list. If we omit the ``may'', this is a typical
+recipe may provide such a list. If we omit the ``may'', this is a typical
 example of first-order logic reasoning. Many other reasoning techniques exist
 (including probabilistic ones -- ones able to deal with the ``may''), we shall
-illustrate some of them further in this scenario.
-
-As mentioned, a recipe often provides a procedure (or a \emph{plan}). The robot
+illustrate some of them later in this scenario.
+
+We mentioned that a recipe often provides a procedure (or a \emph{plan}). The robot
 should be able to store this plan in a way that allow later execution.  The
@@
 
-Moreover, procedures are often \emph{underspecified}: we can expect the recipe
+To make decision, a robot may also want to \emph{predict} the state of the world
+after some action (``if I leave the cake 2h in the oven, it will be burned'').
+Such ability to project itself in future or, generally speaking, in other
+possible state of the world is related to several cognitive ability and
+reasoning techniques: \emph{planning}, \emph{representation of possible worlds}
+and \emph{non-monotonic reasoning}, in addition to common-sense knowledge and
+\emph{physics-based reasoning}.
+
+Procedures are in addition often \emph{underspecified}: we can expect the recipe
 to provide a cooking duration, but we usually do not expect the recipe to tell
 us to first open the oven door, and then put the cake into it, since it is
-self-evident that the door must first be opened. Such underspecification should
-be detectable, representable, and ideally completable by the knowledge
-representation system.
-
-Besides representation of space and time, we may have
+self-evident that the door must first be opened to put the cake in the oven.
+Such underspecification should be detectable, representable, and ideally
+completable by the knowledge representation system\footnote{Note that we do not
+assume here the {\it knowledge representation system} to be a single software
+component: it may well be the result of the aggregation of several modules
+working together}.
 
 % Representation feature that enable cooperation
-Then, we want our three agents to cooperate. This, in turn, leads to a very
-different set of cognitive abilities.
-
-Cooperation in our scenario can intervene at many levels: for instance, an
+Then, we want our three agents to cooperate. This, in turn, leads to another
+set of cognitive abilities.
+
+Cooperation in our scenario can intervene at many places. For instance, an
 agent may want to inform another one about the number of eggs that are
-necessary for the brownie. It makes sense to help in this way only if the first
-agent knows that the recipient agent both needs the information and does not
-know it. This in turn requires the robot to be able to model the knowledge of
-the other agents: to think \emph{from the perspective} of another agent (this
-idea is related to the so-called Theory of Mind~\cite{...}).
+necessary for the brownie. This \emph{helping} behaviour makes sense only if
+the first agent knows that the recipient agent both needs the information but
+does not know it. This in turn requires the robot to be able to model the
+knowledge of the other agents: to think \emph{from the perspective} of another
+agent (this idea is related to the so-called Theory of Mind~\cite{...}).
 
@@
 bowl''. Besides the symbol ``bowl'', which physical entity are we actually
-talking about? If we want to act on the world, this \emph{grounding} operation
-is essential, and may be tightly bound to the underlying knowledge
+talking about? If we want to act on the world, this so-called \emph{grounding}
+operation is essential, and may be tightly bound to the underlying knowledge
 representation system. Note that grounding usually denotes the {\it top-down}
@@
 intentions} and \emph{joint goals}~\cite{Tomasello2005, Bratman2009}: to help
-the human during the cooking session, the robots need to store at least how far
+the human during the cooking session, the robots need to track how far
 they are into the recipe, what is the next step the human is likely to go for,
@@
 per-se} is usually out of the scope of a knowledge representation system, we
-may want to be sure these systems successfully interoperate, \ie that the
+may want to be sure these systems may successfully interoperate, \ie that the
 knowledge representation system provides efficient support to the natural
@@
 concepts with only partial knowledge on them. For instance, let imagine a robot
-that sees a bottle over a table, whose bottom is hidden by another object. The
+that sees a bottle on a table, whose bottom is hidden by another object. The
 robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
@@
 temporal reverse of the projection task.
+
+\subsubsection{Physics-based reasoning}
+\label{sect|physics}
 

### 2012-04-04 krs-survey.tex
Added an abstract
While here:
- changed the title to emphasis the 'requirements of KRS' aspect

@@
 
-\title{Current State of Knowledge Representation Systems for Robotics}
+\title{Requirements and Current State of Knowledge Representation Systems for Service Robotics}
 \author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz}
@@
 \tableofcontents
+
+\begin{abstract}
+
+The aim of this contribution is two-fold: we first propose a typology of
+knowledge representation systems for service robotics that covers intrinsic
+features like expressiveness or reasoning techniques as well as extrinsic
+features like knowledge acquisition and integration into pre-existent robotic
+architectures.
+
+We then survey the current knowledge representation system in the community to
+identify which knowledge-bound cognitive abilities are today well understood
+and implemented, and which would require more efforts.
+
+We hope this article may bring a clearer picture of the challenges of knowledge
+representation and manipulation for the robotics research community.
+
+\end{abstract}
 

### 2012-04-04 krs-survey.tex
Completed a bit the 'Previous Work' section
While here, polished the introduction

@@
 
+\subsection{Knowledge as the Root of Cognitive Abilities}
+\label{sect|cognitive-abilities}
+
 The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
@@
     Central to this effort is to develop an understanding of the relationship
-	between the knowledge, the perception, and the action of such a robot. The
+    between the knowledge, the perception, and the action of \[...\] a robot. The
     sorts of questions we want to be able to answer are
@@
 
-This survey focuses on the knowledge issue: we aims at establishing the current
-landscape of approaches to the knowledge representation problem in the research
-community, at summarizing the major trends and identifying the possible
-shortcomings.
-
-We have identified over ten projects that either are explicitely advertised as
+Those questions have in common one prerequisite to be answered: the robot must
+be able to manipulate \emph{explicitely} knowledge, and hence, need a wayto
+represent the knowledge.
+
+This survey focuses on this knowledge representation issue: we aims at first
+establishing a as-exhaustive-as-possible typology of representation needs for
+robotics, and then at painting the current landscape of approaches to the
+knowledge representation problem in the research community, in order to
+summarize the major trends and to identify the possible shortcomings.
+
+We have classified the features of knowledge representation systems (KRS) into
+six broad categories: \emph{what can be represented?} or the
+\emph{expresiveness} question ; \emph{how things are represented?} ; \emph{what
+reasoning techniques are offered?} ; \emph{how knowledge is acquired and
+anchored?} ; \emph{how the knowledge is made available and used by the other
+components of the robot?} and \emph{which actual knowledge is stored?} or the
+\emph{knowledge instanciation} question.
+
+We have identified over ten projects that are either explicitely advertised as
 knowledge representation systems or involve explicit knowledge manipulation
@@
 assertions, be it static statements on the world or spatio-temporal events, and
-reasoning on this representation.
-
-We also try, for each project, to make clear \begin{inparaenum} \item how new
-facts can be acquired, \ie how informations from perception or interaction are
-turned into knowledge, \item how, in return, symbolic concept are
-\emph{anchored} into the robot sensori-motor space, \item how the knowledge
-representation system integrates with the decisional layers of the robot, and
-how this leads to better robotic control.\end{inparaenum}
-
-\subsection{Definition of inclusion criteria}
+reasoning on this representation. Proeminent features of each of these systems
+illustrate how the various features of a knowledge representation system are
+currently implemented.
+
+We have also tried, for each project, to make clear \begin{inparaenum} \item
+how informations from perception or interaction are turned into knowledge,
+\item how, in return, symbolic concept are \emph{anchored} into the robot
+sensori-motor space, \item how the knowledge representation system integrates
+with the decisional layers of the robot, and how this can improve or ease
+robotic control.\end{inparaenum}
+
+\subsection*{Survey Inclusion Criteria}
 \label{sect|inclusion-criteria}
@@
 
-More precisely, we have established the following criteria to select the
-knowledge representation systems to include in this survey. Those systems must:
-
-\begin{itemize}
-	\item  Run on \emph{service robot} (robots that interact with objects in a
-	semantic-rich environment),
-
-	\item  Ground their knowledge in the physical world (physically embedded),
-	\begin{itemize}
-		\item  Able to \emph{resolve entities}
-		\item  Able to automatically create new object instances
-	\end{itemize}
-
-	\item  Be able to merge different knowledge modalities,
-	\item  Be endowed with on-line, dynamic reasoning (not just a static
-	dictionary)
-
-\end{itemize}
-
-\subsubsection{What to expect from a knowledge representation system}
+We have limited ourselves to systems that
+\begin{inparaenum} 
+    \item  run on \emph{service robot} (that is, robots that interact with 
+    objects in a semantic-rich environment primarly designed for humans),
+    \item  ground the knowledge in the physical world (physically embedded
+    systems able to assess their environment),
+    \item  are able to merge different knowledge modalities,
+    \item  are endowed with on-line, dynamic reasoning (\ie not a simple static
+    database).
+\end{inparaenum}
+
+\subsection{Previous Work}
 \label{sect|evaluation-literature}
 
-Several authors from fields that are connected to the robotics have previously
-listed desirable features of systems aiming at cognitive abilities as rich as
-possible.
-
-\paragraph{McCarthy: Towards human-level AI}
-
-In~\cite{McCarthy2007}, McCarthy lists the challenges he identifies on the road
-to a \emph{human-level AI}.
+To the extend of our knowledge, no previous published work attempted to identify
+knowledge-bound cognitive abilities required by service robotics.
+
+However, several authors from fields that are connected to the robotics have
+previously listed desirable features of artificial systems aiming at cognitive
+abilities as rich as possible.
+
+In~\cite{McCarthy2007}, McCarthy recently listed the challenges he identifies
+on the road to a \emph{human-level AI}.
 
@@
 
-	\item non-monotonic reasoning \fxfatal{give here a clear example},
-
-	\item what McCarthy calls \emph{Elaboration Tolerance}, \ie, the ability to
-	extend \emph{on requirement} the closed domain of interpretation for a
+	\item non-monotonic reasoning,
+
+	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
+	extend \emph{on demand} the closed domain of interpretation for a
 	given assertion,
@@
 
-Knowledge representation systems are concerned by most, if not all, of this
-points.
-
-\paragraph{Natural language processing in situated context}
-
-In~\cite{Roy2005}, Roy and Reiter summarize what they see as the main
-challenges to be tackled: cross-modal representation systems, association of
+Several surveys compare global cognitive architectures (\cite{Vernon2007, Chong2009}).
+Vernon et al.~\cite{Vernon2007} split these architectures into two broad
+categories: the \emph{cognitivist} ones (where cognition is considered as an
+explicit computation problem, often based on symbol manipulation), and the
+\emph{emergent} ones (where cognition only exists as a result of the
+interaction of the system with its environment). The approaches presented in
+this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
+that aim at making knowledge explicit within the robot architecture. Vernon et
+al. propose twelve \emph{characteristics of cognitive system} to compare
+architectures. Amongst them, they mention the \emph{inter-agent epistemology} (how
+the structure of the world is captured in a representation and shared), the
+relation to \emph{embodiement}, the ability to \emph{anticipate} and to
+\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
+level of the whole robotic architecture, these features also translate into
+knowledge representation strategies and are relevant to our study.
+
+Coming from the perspective of natural language processing in situated context,
+Roy and Reiter, in~\cite{Roy2005}, summarize what they see as the main
+challenges to be tackled by KRS: cross-modal representation systems, association of
 words with perceptual and action categories, modeling of context, figuring out
@@
 
-\cite{Hawes2008}: benchmarking of "Information-Processing Architectures on Intelligent Systems"
-
-\subsection{Methodology}
-\label{sect|methodology}
-
-\fxnote{Mention here that each team was proposed to complete/amend the description of their
-system}
+Knowledge representation systems in robotics are concerned by most, if not all,
+of these points, and we shall reuse them later, in slightly reformulated ways.
 
@@
 scenario: we merely use it to ground the selection and illustrate the set
-features (Section~\ref{sect|features}) that define the dimensions of a
+features that define the dimensions of a
 knowledge representation system for robotics.
+
+Section~\ref{sect|features} presents these features in detail, and organize
+them into a typology.
 
@@
 
-We try in section~\ref{sect|summary} to summarize the current approaches and to
-relate them to the previously identified set of features. This leads us to
+We finally summarize in section~\ref{sect|summary} the current approaches in
+relation to the previously identified set of features. This leads us to
 identify several new research directions that are not sufficiently addressed by

### 2012-04-04 krs-survey.tex
Light edition of the scenario section

@@
 	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
-	\caption{A representation of the scenario, in the MORSE simulator}
+	\caption{A possible representation of the scenario, in the MORSE simulator}
 	\label{fig|scenario}
@@
 
-A ``first-order'' analysis of this task leads to a first partition of the
+A ``first-order'' analysis of this task leads to a rough partition of the
 required representation abilities:
@@
 
-We mentioned that a recipe often provides a procedure (or a \emph{plan}). The robot
-should be able to store this plan in a way that allow later execution.  The
-plan is likely to contain \emph{spatio-temporal constraints} (like ``put
-the brownie in the oven for 20 min'') that must as well be appropriately
-represented.
+We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
+robot should be able to store this plan in a way that allow later execution.
+The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
+the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
+must as well be appropriately represented.
 
 To make decision, a robot may also want to \emph{predict} the state of the world
-after some action (``if I leave the cake 2h in the oven, it will be burned'').
+after some action (``if I leave the cake 2h in the oven, it will burn'').
 Such ability to project itself in future or, generally speaking, in other
@@
 and \emph{non-monotonic reasoning}, in addition to common-sense knowledge and
-\emph{physics-based reasoning}.
+\emph{physics-based} reasoning (that allows to predict that an egg is likely to
+break if dropped).
 
@@
 assume here the {\it knowledge representation system} to be a single software
-component: it may well be the result of the aggregation of several modules
-working together}.
+component: it may well be the result of the aggregation of several logical
+components working together}.
 
@@
 knowledge of the other agents: to think \emph{from the perspective} of another
-agent (this idea is related to the so-called Theory of Mind~\cite{...}).
+agent (this idea is related to the so-called Theory of Mind~\cite{Leslie2000}).
 
 Ability to communicate is one important pre-requisite to collaboration.
-Communication in general~\cite{Jakobson} requires the addresser and the
-addressee to share a common interpretative framework (shared common-sense
+Communication in general~\cite{Jakobson1960} requires the addresser and the
+addressee to share a common interpretative framework (a shared common-sense
 knowledge -- or cultural background -- and a shared context). In our scenario,
@@
 suffice if, for example, an agent asks another agent to ``give {[him]} the
-bowl''. Besides the symbol ``bowl'', which physical entity are we actually
+bowl''. Behind the symbol ``bowl'', which physical entity are we actually
 talking about? If we want to act on the world, this so-called \emph{grounding}
@@
 (like the ability to tell ``who I am'' or ``what do I think of'' does not
-appear to be necessary in our scenario. This may however enrich the human-robot
-interaction.
+appear to be necessary in our scenario. This may however be required in more
+general human-robot interaction.
 

### 2012-04-04 krs-survey.tex
Various minor changes

@@
 may have interactions). The typical scenario that would involve such robots is
-a service robot in a human-friendly environment like a kitchen.
+a service robot in a human-friendly environment like a kitchen (as described in
+section~\ref{sect|scenario}).
 
@@
     \item  are able to merge different knowledge modalities,
-    \item  are endowed with on-line, dynamic reasoning (\ie not a simple static
-    database).
+    \item  are able of on-line, dynamic knowledge acquisition and reasoning 
+    (\ie not a simple static database).
 \end{inparaenum}
+
+This exclude \textit{de facto} classical cognitive systems like {\sc Soar} or
+{\sc ACT-R} for which we could not find recent implementations on full-scale
+service robots.
 
@@
 \section{A Prototypical Scenario}
+\label{sect|scenario}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A typology of features of KRS for robotics}
+\section{A Typology of Knowledge Representation Systems Requirements for Robotics}
 \label{sect|features}
 
-The previous section detailed an imaginary cooking scenario. We identified, in
-no particular order, numerous desirable features for a knowledge representation
-system for robotics.
+We have detailed an imaginary cooking scenario where numerous desirable
+features for a knowledge representation system for robotics have been
+identified in no particular order.
 
@@
 relevant literature.
-
-\fxnote{Should we also mention the general dynamics of the knowledge
-representation system? ie, the system from the point of view of an on-going
-cognitive process. See \cite{Henderson2011}}
 
@@
 All systems allow to modify the ABox, not always possible to alter the TBox
+
+\subsubsection{Memory}
+\label{sect|memory}
 

### 2012-04-06 krs-survey.tex
Completed sections on meta-cognition and introspection

@@
 \newcommand{\eg}{{\textit{e.g.~}}}
+
+\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
+%\newcommand{\stmt}[1]{{\footnotesize $\langle$\stmttt#1\relax$\rangle$}}
+\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
+\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}
+
+\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
+\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}
 
@@
 
-Narrower, more technical dimension of the introspection.
+As stated by Josyula and Raja, meta-cognition is composed of both
+\emph{``meta-level control of cognitive activities and the introspective
+monitoring of such activities to evaluate and to explain them"}.
+
+A knowledge representation system endowed with \emph{meta-cognition} is not
+only able to manipulate knowledge but also to exhibit and manipulate the
+structure of its knowledge and the reasoning process. For instance, the ability
+to explain a logical inconsistency in a KRS is a meta-cognitive function.
+Sloman~\cite{Sloman2011}
+
+At section~\ref{sect|introspection} below, we discuss the idea of
+introspection.  Meta-cognition can be viewed as the technical facet of the
+introspection in general.
 
@@
 
+Introspection is the ability to self-describe: what are my capabilities, what
+is my state (performing some action, idling, etc.), what are my beliefs (\ie my
+knowledge), what are my intentions and my plans?
+
+Introspection must be distinguished from meta-cognition: While introspection
+may require meta-cognition (for instance to be able to expose its internal
+knowledge), it is not always mandatory. The current state of the robot can be
+represented as a simple instanciation of a specific category (for instance, if
+the robot give an object to the human, this state could be represented with the
+triplets \setstmt{robot performs action1, action1 isA Give}.
+
 \paragraph{Modelling of the robot capabilities}
+
+A particularly important aspect of introspection relates to the description of
+its own capabilities: which sensors/actuators/computation services exist and
+are currently available ?  While at a first level, these descriptions can be
+static (\eg the robot has one laser scanner and two arms), at a second level,
+the description is updated and reflect the current (and possibly past and
+future) state of the robot. Note that these description may also invlove
+geometric descriptions (a kinematic chain, the pose of a device, etc.) that may
+be deported outside of the main knowledge base. Efforts trying to formalize,
+maintain and expose the capabilities and state of a robot are not new (and
+ground themselves in work and techniques for self-descriptive remote procedure
+calls in computing science), but take a renewed importance with applications
+for high-level multi-robot cooperation. Recent work in that direction
+include~\cite{Kunze2011}.
 

### 2012-04-14 krs-survey.tex
Section on physics based reasoning

@@
 
-To the extend of our knowledge, no previous published work attempted to identify
-knowledge-bound cognitive abilities required by service robotics.
+To the extend of our knowledge, no previous published work attempt to build a
+typology of cognitive abilities related to knowledge and required by service
+robotics.
 
@@
 
-In~\cite{McCarthy2007}, McCarthy recently listed the challenges he identifies
-on the road to a \emph{human-level AI}.
+For instance, in~\cite{McCarthy2007}, McCarthy recently listed the challenges
+he identifies on the road to a \emph{human-level AI}.
 
@@
 
-We have detailed an imaginary cooking scenario where numerous desirable
+We presented an imaginary cooking scenario where numerous desirable
 features for a knowledge representation system for robotics have been
@@
 \label{sect|physics}
+
+As embodied entities, robots have to interact with physical entities.
+\emph{Naive physics reasoning} covers all the everyday reasoning the humans
+unconsciouly perform, like taking into account gravity (``if I drop a ball, it
+falls down'') or common physical properties of objects (``a glass may break if
+dropped'', etc.). Many of the interactions with our everyday environments are
+ruled by such laws that are difficult to exhaustively encode.
+
+Some systems \cite{Kunze??} rely on external dedicated physics engine to
+compute symbolic facts from on-demand physics simulation. This, however,
+requires a tight integration between the symbolic model and a geometric model
+that carries the geometries and physical properties of objects.
 
@@
 
+\begin{landscape}
+\begin{table}
+\begin{center}
+\rowcolors{2}{lightgray}{codegray}
+
+%\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
+\begin{tabular}{l|ccccccccc}
+\hiderowcolors
+
+{} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMKRF} & {\sc DY-KNOW} \\
+
+\hline
+\showrowcolors
+{\bf Expressiveness} & & & & & & & & &  \\
+Formal Logic & Prolog & OWL-DL & {\sc CycL} &  &  &  & & & \\
+OWA/CWA & CWA & OWA & & & & & & & \\
+Modeling uncertainty & +++ (add-on) & & & & & & & & \\
+Meta-cognition & ++ & ++ & & & & & & & \\
+{\bf Model} & & & & & & & & &  \\
+{\bf Reasoning} & & & & & & & & &  \\
+Standard FOL reasoning & +++ & +++ & & & & & & &  \\
+Lazy evaluation & +++ & & & & & & & &  \\
+Reasoning under uncertainty & ++ (add-on) & & & & & & & & \\
+Non-monotonic reasoning & & & & & & & & &  \\
+Presupposition accomodation & & & & & & & +++ & &  \\
+Prediction, projection, diagnosis & & & & & & & & &  \\
+Physics-based reasoning & +++ (add-on) & & & & & & & &  \\
+Planning & + & +++ (external module) & & & & & & &  \\
+Knowledge structure alteration & & ++ & & & & & & &  \\
+Learning & & & & & & & & &  \\
+{\bf Knowledge Acquisition} & & & & & & & & &  \\
+{\bf Integration} & & & & & & & & &  \\
+{\bf Knowledge Instanciation} & & & & & & & & &  \\
+\hline
+
+%\end{tabularx}
+\end{tabular}
+\end{center}
+\caption{Main domain of contribution of each surveyed systems. Rating goes from +++ (major focus) to + (secondary interest). An empty cell means that 
+the system does not consider this domain.}
+\label{table|contribution-by-systems}
+\end{table}
+\end{landscape}
+
 This section tries to summary the various approaches we surveyed in the

### 2012-04-14 krs-survey.tex
Section on memory

@@
 \label{sect|memory}
+
+Memory has been studied at lenght in the cognitive neuroscience community, and
+the ability to forget is proven to play a key role for the cognition.
+\fxwarning{Add references -> ask Claire or Gladys}.
+
+Most knowledge representation systems offers some kind of working memory as a
+pool of facts that are not forgotten by the robot until it is halted. Some
+systems may propose persistent storages that allow the robot knowledge to grow
+over time, while other may offer a larger range of memory categories, like
+short term memory (that last for a couple of seconds) or episodic memory (that
+allows the robot to selectively remember facts associated to specific events).
+\fxwarning{Add references of paper in the robotics community that have
+explicitely tackled the memory issue}.
 

### 2012-04-16 krs-survey.tex
Minor rephrasing in related work

--- 2012-04-14 krs-survey.tex
@@
 \label{sect|physics}
-
-As embodied entities, robots have to interact with physical entities.
-\emph{Naive physics reasoning} covers all the everyday reasoning the humans
-unconsciouly perform, like taking into account gravity (``if I drop a ball, it
-falls down'') or common physical properties of objects (``a glass may break if
-dropped'', etc.). Many of the interactions with our everyday environments are
-ruled by such laws that are difficult to exhaustively encode.
-
-Some systems \cite{Kunze??} rely on external dedicated physics engine to
-compute symbolic facts from on-demand physics simulation. This, however,
-requires a tight integration between the symbolic model and a geometric model
-that carries the geometries and physical properties of objects.
 

### 2012-04-17 krs-survey.tex
{Modifying->Alteration} of knowledge structures
Added ref to ACT-R/E

--- 2012-04-14 krs-survey.tex
@@
 
-This exclude \textit{de facto} classical cognitive systems like {\sc Soar} or
-{\sc ACT-R} for which we could not find recent implementations on full-scale
-service robots.
+While classical cognitive systems like {\sc Soar} or {\sc ACT-R} have
+declarative knowledge modules and have been recently used on service robots
+(see {\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are absent from this
+survey because we could not find evidence in the litterature on possible
+on-line manipulation of symbolic knowledge.
 
@@
 
-\subsubsection{Modifying the knowledge structure}
+\subsubsection{Alteration of the knowledge structure}
+
+In systems
 

### 2012-04-17 krs-survey.tex
More on logical formalism, added a section on expressive power

@@
 
+Table~\ref{table|contribution-by-systems}, at the end of the article,
+summarizes all these features with the main domain of contribution of each
+surveyed KRS.
+
 \subsection{Expressiveness: What Can be Represented?}
@@
 
-\subsubsection{Introduction: main logic formalisms}
+\subsubsection{Introduction: Main Logic Formalisms}
 
@@
 \emph{individuals}. The W3C OWL-DL standard is a widely-used language to describe
-domain with the description logic.
-
-\emph{Modal logics}, that allow for statement qualification
-like \emph{possibility} or \emph{necessity}, have been shown to be closely
-related to description logics.
+domains with the description logic.
+
+Because Description Logics have been originally created from the perspective of
+a \emph{knowledge representation language} and not a logic language, their
+terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
+\emph{individual},...) is well-suited to knowledge description and we may use
+it in the remaining of this paper outside of the strict context of  Description
+Logics.
+
+\emph{Modal logic}, that allow for statement qualification like
+\emph{possibility} or \emph{necessity}, have been shown to be closely related
+to description logics. Modal logic allows to represent conveniently parallel
+possible worlds and facts like ``the robot knows \emph{that the human knows}
+how to read a recipe''.
 
 \fxfatal{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
+
+\emph{Temporal logic} are designed to represent and manipulate assertions whose
+truth value may vary in time.
 
@@
 
+Note that most of these logic formalisms are still active research field by
+their own, and practical considerations (especially the availability of
+reasoners performant enough for on-line use on a robot) often constraint the
+choice of a logical formalism and a level of expressive power.
+
 \paragraph{Some examples}
@@
 but the robot does not know what the human actually know about it.}
+
+\subsubsection{Expressive Power}
+
+Logical formalisms each bring a certain level of expressive power. For
+instance, the following classical syllogism can not be represented in
+propositional logic because of the use of \emph{universal quantification}:
+
+\begin{quote}
+\begin{enumerate}
+    \item All men are mortal,
+    \item Socrates is a man,
+    \item Therefore, Socrates is mortal
+\end{enumerate}
+\end{quote}
+
+However, the following weak version of the syllogism can be represented in
+propositional logic:
+
+\begin{quote}
+\begin{enumerate}
+    \item If Socrates is a man, then Socrates is mortal,
+    \item Socrates is a man,
+    \item Therefore, Socrates is mortal
+\end{enumerate}
+\end{quote}
+
+Generally speaking, expressive power comes at the cost of ease of
+\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
+at section~\ref{sect|reasoning}.} tests, possibly leading to undecidable formal
+systems, \ie systems where it is proven that a proposition can not be decided
+to be true or false.
+
+The relationship between expressive power and the complexity of reasoning problem that
+follow has been partially formalized for Description Logics.
+Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
+for Description Logics that allows to conveniently explore this relationship
+and indexes most of the litterature on that subject.
 

### 2012-04-17 krs-survey.tex
Filled section on standard reasoning techniques

@@
 
+\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}
 \newcommand{\ie}{{\textit{i.e.~}}}
@@
 
-Backtracking,...
-
-\paragraph{Decidability}
-
-...
+Being based on logical languages to represent the knowledge, most of the
+systems we survey allow various forms of \emph{reasoning}.
+
+We call \emph{standard reasoning techniques} techniques based on logical
+inference, using resolution algorithms like \emph{forward chaining},
+\emph{backward chaining} or \emph{semantic tableaux}.
+
+Main reasoning problems include \emph{concept satisfiability},
+\emph{consistency checking} and \emph{instance checking}.
+
+Concept satisfiability verifies if it is possible to find a non-empty
+\emph{interpretation} of a concept (or an expression defining a concept) in the
+knowledge model. For instance, the formula \concept{Plant} $\land$
+\concept{isRed}, which defines the concept of red plants, is satisfiable in a
+model \concept{KB} iff $\exists a, $ \concept{Plant}$(a) \land$
+\concept{isRed}$(a)$, \ie if we can find at least one red plant $a$ in our
+model.
+
+Checking the consistency of a model is equivalent to checking the
+satisfiability of each of the concept defined in the knowledge model.
+
+Instance checking consists in verifying that an individual $a$ is an
+interpretation of a concept (or concept expression) $C$ in the knowledge model.
+A typical example would be that we are provided with an instance
+\concept{object1} and we want to know if this object is a kind of
+\concept{Bottle} or \concept{Glass}.
+
+Reasoners can then provide other type of inferences, like:
+
+\begin{itemize}
+    \item class subsumption (similar to inheritance)
+
+    \item reasoning on roles properties, including:
+        \begin{itemize}
+        \item entailments based on roles domain and range (for instance, if the
+        domain of the role \concept{thinksOf} is known to be
+        \concept{ThinkingAgent}, then \concept{thinksOf}$(a, b) \to
+        $\concept{ThinkingAgent}$(a)$),
+
+        \item universal, existantial and cardinality constraints (including \concept{allValue}, 
+        \concept{someValue}, \concept{hasValue}),
+
+        \item property characteristics (symmetry, transitivity, etc.)
+
+        \end{itemize}
+
+    \item complex concept expressions like: \par \footnotesize \concept{Bottle}
+    $\equiv$ \concept{Artifact} {\bf that} (\concept{hasShape} {\bf value}
+    \concept{cylinderShape})\footnote{This example uses the \emph{Manchester
+    syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}} \normalsize
+
+    \item set operations like: \par \footnotesize \concept{Color} $\equiv$ {\bf
+    unionOf}(\concept{blue}, \concept{green}, \concept{orange},
+    \concept{black}...) \normalsize
+
+\end{itemize}
+
+\paragraph{Rule Languages}
+\item generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
+        \footnotesize \concept{looksAt(?agt, ?obj)} $\land$
+        \concept{pointsAt(?agt,?obj)} \par $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
+        \normalsize
+
+\subsubsection{Alteration of the knowledge structure}
+
+In systems
+
+All systems allow to modify the ABox, not always possible to alter the TBox
 

### 2012-04-17 krs-survey.tex
ainly added section on horn clauses

@@
 robotic control.\end{inparaenum}
+
+\subsection*{A Note on Terminology}
+
+Since this article discusses at length the concept of knowledge in the context
+of robotics, it is useful to make our terminology explicit.  No general
+agreement on a definition of ``knowledge'' exists. In our context, we call
+``knowledge'' \emph{an explicit set of interrelated logical facts that are
+meaningful to the robot executive controller} (by \emph{meaningful} we mean
+that can possibly be interpreted to lead to a purposeful action).
+
+The relation of \emph{data} and \emph{information} to knowledge is a debated
+epistemology question (the interested reader is invited to refer to the
+Wikipedia page on the ``DIKW'' hierarchy). In this article, we will associate
+data to low-level material like raw sensor output, and information to
+uncontextualized symbolic facts.
+
+To give a example, image a human watching at a book and being tracked by a
+Kinect sensor: the pose of the human skeleton in the world would be the data,
+the fact \concept{looksAt(human, book)} as computed by a geometric reasoning
+module would be the information, the fact \concept{looksAt(john,
+war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
+the robot would be proper knowledge.
 
@@
 side, on the ability to solve the inference problem (called
-\emph{decidability}: is a given logical sentence true in my model?) in a
+\emph{satisfiability}: is a given logical sentence true in my model?) in a
 tractable manner.
@@
 
-The family of \emph{description logics}~\cite{Baader2008} also play an
-important role. It is a subset of the first-order logic, with some extensions
-in second-order logic. Description logics are notable because most of them are
-known to be decidable. In description logic, axioms are build from
+\emph{Horn clauses} are an important subset of FOL because the satisfiability
+of a set of such clauses is a $P$-complete problem. A Horn clause is a
+disjunction of literals (a \emph{clause}) with at most one positive literal:
+$\neg p \vee \neg q \vee \cdots \vee \neg t \vee u$, which can also be
+represented as $(p \wedge q \wedge \cdots \wedge t) \rightarrow u$.  Important
+logic programming languages like Prolog are based on Horn clauses.
+
+The family of \emph{Description Logics}~\cite{Baader2008} also play an
+important role. It is also a subset of the first-order logic, with some
+extensions in second-order logic. Description logics are notable because most
+of them are known to be decidable. In description logic, axioms are build from
 \emph{concepts}, \emph{roles} (that are unary or binary predicates) and
-\emph{individuals}. The W3C OWL-DL standard is a widely-used language to describe
-domains with the description logic.
+\emph{individuals}. The W3C OWL-DL standard is a widely-used language to
+describe domains with the description logic.
 
@@
 
-Generally speaking, expressive power comes at the cost of ease of
+Generally speaking, expressive power comes at the cost of more complex
 \emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
-at section~\ref{sect|reasoning}.} tests, possibly leading to undecidable formal
-systems, \ie systems where it is proven that a proposition can not be decided
-to be true or false.
-
-The relationship between expressive power and the complexity of reasoning problem that
-follow has been partially formalized for Description Logics.
+at section~\ref{sect|reasoning}.} computations, possibly leading to
+untractable, if not undecidable (\ie systems where it is proven that a
+proposition can not be decided to be true or false) problems.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\columnwidth]{expressive_overlap_dl_horn.pdf}
+    \caption{Expresiveness overlap of Description Logics and logic programs
+    based on Horn clauses, taken from~\cite{Grosof2003}}
+    \label{fig|overlap_dl_horn}
+\end{figure}
+
+Figure~\ref{fig|overlap_dl_horn} shows that the expresive power of description
+logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
+mention extensions to description logics based on rule systems to bring closer
+the two approaches.
+
+The relationships between expressive power and reasoning complexity that follow
+has been extensively studied for Description Logics.
 Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
-for Description Logics that allows to conveniently explore this relationship
-and indexes most of the litterature on that subject.
+that allows to conveniently explore these relationships and indexes most of the
+litterature on that subject.
 
@@
 
+\subsubsection{Memory}
+\label{sect|memory}
+
+Memory has been studied at lenght in the cognitive psychology and
+neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
+introduce the idea of \emph{short-term} and \emph{long-term} memory,
+Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
+and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organizes
+the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
+into a hierarchy. Short-term memory is refined with the concept of
+\emph{working memory} by Baddeley~\cite{Baddeley2010}
+(Figure~\ref{fig|memory_models}).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
+    \caption{Overview of the main types of memories}
+    \label{fig|memory_models}
+\end{figure}
+
+Most knowledge representation systems offers some kind of memory as a pool of
+facts that are not forgotten by the robot until it is halted (this memory is
+often refered as a \emph{working memory}, but with a meaning unrelated to
+Baddeley's definition). Some systems may propose persistent storages that allow
+the robot knowledge to grow over time, while other may offer a larger range of
+memory categories, like short term memory (that lasts for a couple of seconds)
+or episodic memory (that allows the robot to selectively remember facts
+associated to specific events).
+
 %%%%%%%%%%
@@
 \paragraph{Rule Languages}
-\item generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
+
+As mentionned earlier, knowledge models based on description logics can be extended through rule languages.
+
+Intersection of properties is an example of expression that can only be represented with rules: for instance,
+
+generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
         \footnotesize \concept{looksAt(?agt, ?obj)} $\land$
@@
 
-$Egg(a) \wedge Egg(b) \wedge MakeOmelette(a, b, c) \to \lnot Egg(a) \wedge
-\lnot Egg(b) \wedge Omelette(c)$
-
-The insertion of the proposition $MakeOmelette(a, b, c)$ leads to retraction of
-other facts. This rule requires non-monotonic reasoning to be applied.
+\concept{Egg}$(a) \wedge $ \concept{Egg} $(b) \wedge $
+\concept{MakeOmelette}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \wedge \lnot $
+\concept{Egg}$(b) \wedge $ \concept{Omelette}$(c)$
+
+The insertion of the proposition \concept{MakeOmelette}$(a, b, c)$ leads to
+retraction of other facts. This rule requires non-monotonic reasoning to be
+applied.
 
@@
 
-Some systems \cite{Kunze??} rely on external dedicated physics engine to
+Some systems \cite{Kunze2011a} rely on external dedicated physics engine to
 compute symbolic facts from on-demand physics simulation. This, however,
@@
 
-\subsubsection{Memory}
-\label{sect|memory}
-
-Memory has been studied at lenght in the cognitive neuroscience community, and
-the ability to forget is proven to play a key role for the cognition.
-\fxwarning{Add references -> ask Claire or Gladys}.
-
-Most knowledge representation systems offers some kind of working memory as a
-pool of facts that are not forgotten by the robot until it is halted. Some
-systems may propose persistent storages that allow the robot knowledge to grow
-over time, while other may offer a larger range of memory categories, like
-short term memory (that last for a couple of seconds) or episodic memory (that
-allows the robot to selectively remember facts associated to specific events).
-\fxwarning{Add references of paper in the robotics community that have
-explicitely tackled the memory issue}.
-
 \subsubsection{Learning}
@@
 \label{sect|knowledge-acquisition}
+
+In the survey inclusion criteria, we have insisted on only considering robotic
+systems that acquire knowledge by themselves, during their runtime.
+
+\emph{Acquiring knowledge} means in our context building new statements
+(usually as new logical facts) connected to our existing knowledge from
+external sources of information. We consider mainly three of them:
+proprioceptive/exteroceptive sensors, interaction with other agents, humans or
+robots, and remote databases. This process has generally at least two steps:
+the information acquisition by itself, and the \emph{transformation} of the
+information into knowledge, \emph{aligned} with the robot existing model
+(following our terminology for information and knowledge, as discussed in the
+introduction).
+
+Knowledge acquision is generally not done directly in the knowledge
+representation system. On the contrary, several, if not numerous, external
+components are usually required to convert percepts into symbolic facts and to
+ground them.
+
+While we do not review in this article all these systems, the whole process of
+knowledge acquisition is central in cognitive robotic architecture and the
+design of knowledge representation systems can influence or be influenced by
+the approach to knowledge acquisition.
+
+In particular, complex robotic systems often require multi-modal perception
+capabilities (for instance, a robot can only interpret an utterance like ``this
+is a plate'' if it is able to understand gestures, understand natural language
+and merge them in a timely manner). Multi-modal interpretation can take place
+at various levels, but in many cases (especially if the modalities are of very
+different natures, like in the example above) merging will require
+symbolic-level reasoning. The KRS has a direct impact on the feasability and
+ease of such operations.
 

### 2012-04-18 krs-survey.tex
Extended with ref to psychology + added figure to the Memory section

--- 2012-04-17 krs-survey.tex
@@
 
-Memory has been studied at lenght in the cognitive neuroscience community, and
-the ability to forget is proven to play a key role for the cognition.
-\fxwarning{Add references -> ask Claire or Gladys}.
-
-Most knowledge representation systems offers some kind of working memory as a
-pool of facts that are not forgotten by the robot until it is halted. Some
-systems may propose persistent storages that allow the robot knowledge to grow
-over time, while other may offer a larger range of memory categories, like
-short term memory (that last for a couple of seconds) or episodic memory (that
-allows the robot to selectively remember facts associated to specific events).
-\fxwarning{Add references of paper in the robotics community that have
-explicitely tackled the memory issue}.
+Memory has been studied at lenght in the cognitive psychology and
+neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
+introduce the idea of \emph{short-term} and \emph{long-term} memory,
+Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
+and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organizes
+the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
+into a hierarchy. Short-term memory is refined with the concept of
+\emph{working memory} by Baddeley~\cite{Baddeley2010}
+(Figure~\ref{fig|memory_models}).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
+    \caption{Overview of the main types of memories}
+    \label{fig|memory_models}
+\end{figure}
+
+Most knowledge representation systems offers some kind of memory as a pool of
+facts that are not forgotten by the robot until it is halted (this memory is
+often refered as a \emph{working memory}, but with a meaning unrelated to
+Baddeley's definition). Some systems may propose persistent storages that allow
+the robot knowledge to grow over time, while other may offer a larger range of
+memory categories, like short term memory (that lasts for a couple of seconds)
+or episodic memory (that allows the robot to selectively remember facts
+associated to specific events).
 

### 2012-04-19 krs-survey.tex
Moved the section on memory to the end of 3.2 How things are represented
Addded section on OMRKF + minor other changes

@@
 \emph{Default logic} is one of the formal logic that account for representing
-general truth and exceptions to it. However, due to computational complexity of
-these model (most of inferences in default logic are known to be $NP$-complete
-problem), classical logics and most of the existing reasoners do not allow
-non-monotonic reasoning. For instance, the SWRL rule language, usually
-associated to the OWL-DL ontology language, do not allow non-monotonic
-reasoning.
+general truth and exceptions to it (for instance, \emph{tomatoes are red, in
+general}). However, due to computational complexity of these model (most of
+inferences in default logic are known to be $NP$-complete problem), classical
+logics and most of the existing reasoners do not allow non-monotonic reasoning.
+For instance, the SWRL rule language, usually associated to the OWL-DL ontology
+language, do not allow non-monotonic reasoning.
 
@@
 
-\subsubsection{Presupposition accommodation}
+\subsubsection{Presupposition Accommodation}
 \label{sect|presupposition-accomodation}
@@
 Making decision based on prediction
-
-\subsubsection{Alteration of the knowledge structure}
-
-In systems
-
-All systems allow to modify the ABox, not always possible to alter the TBox
-
-\subsubsection{Memory}
-\label{sect|memory}
-
-Memory has been studied at lenght in the cognitive psychology and
-neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
-introduce the idea of \emph{short-term} and \emph{long-term} memory,
-Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
-and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organizes
-the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
-into a hierarchy. Short-term memory is refined with the concept of
-\emph{working memory} by Baddeley~\cite{Baddeley2010}
-(Figure~\ref{fig|memory_models}).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
-    \caption{Overview of the main types of memories}
-    \label{fig|memory_models}
-\end{figure}
-
-Most knowledge representation systems offers some kind of memory as a pool of
-facts that are not forgotten by the robot until it is halted (this memory is
-often refered as a \emph{working memory}, but with a meaning unrelated to
-Baddeley's definition). Some systems may propose persistent storages that allow
-the robot knowledge to grow over time, while other may offer a larger range of
-memory categories, like short term memory (that lasts for a couple of seconds)
-or episodic memory (that allows the robot to selectively remember facts
-associated to specific events).
 
@@
 \hiderowcolors
-{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model} & {\bf Reasoner} & Main reference \\
+{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & {\bf Reasoner} & Main reference \\
 \hline
@@
 GSM & & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
-OMKRF & & Suh et al. \par (Hanyang Univ.) & & & & & \cite{Suh2007} \\
+OMRKF & & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe/?mid=symbolic_reasoning} & ? & Horn Clauses & ? & \cite{Suh2007} \\
 DY-KNOW & & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
@@
 
-\subsubsection{Intrinsic features}
-\label{sect|peis-intrinsic-features}
-
-\paragraph{Expressiveness} The PEIS Knowledge representation system relies on
+\paragraph{Knowledge model} The PEIS Knowledge representation system relies on
 the {\sc ResearchCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
@@
 \fxfatal{OWA/CWA?}
-
-\subsubsection{Anchoring strategies}
-\label{sect|peis-anchoring}
 
@@
 
+The system can ask the human for help to disambiguate between concept names.
+
 \paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
@@
 
-\subsubsection{Integration in the robot architecture}
+\paragraph{Integration in the robot architecture}
 \label{sect|peis-integration}
@@
 
+\subsection{OMRKF}
+\label{sect|omrkf}
+
+The Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
+(OMKRF) is a knowledge representation system based on four inter-related
+\emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
+approach to knowledge representation that allows to integrate the grounding
+process to the knowledge representation process. OMRKF knowledge model is
+implemented with Horn clauses.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{omrkf.pdf}
+
+    \caption{OMRKF organizes knowledge into four \emph{classes}, each composed
+    of three \emph{levels}. The figure shows some examples of link between
+    knowledge classes and knowledge levels. Based on~\ref{Suh2007}.}
+
+    \label{fig|omrkf}
+\end{figure}
+
+Each level of knowledge is build as three stages of ontological realization: a
+\emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
+or ``object feature'', a taxonomy of concepts inside this level (for instance
+$cup : Object \sqsubseteq tableware : Object$) and an instanciation of the
+taxonomy ($cup1 : cup$).
+
+Environment is represented in OMRKF in the $space : Model$ knowledge level as a
+classical three layers mapping (metric, topological and semantic maps). Objects
+(in $object : Model$) are localized in $space : Model$ through Voronoi nodes.
+
+The knowledge class $Context$ proposes an explicit statement of spatial context
+(mostly geometric relations between objects), temporal context and a more
+general \emph{high-level} context, inferred from spatial and temporal contexts.
+
+Finally, the $Activity$ knowledge class store coumpound actions in a HTN-like
+structure, exploited at run-time by a planner.
+
+\paragraph{Experiments} Experiments conducted with OMRKF include finding
+kitchen objects and reporting about their state to a human.  This experiment
+also shows how OMRKF can deal with objects only partially matched by their
+descriptor by introducing a $candidate()$ function.
+
 %%%%%%%%%% Underlying knowledge model table %%%%%%%%
@@
 DY-KNOW & \\
-OMKRF & \\
+OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
 GSM &  \\
@@
 %\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
-\begin{tabular}{l|ccccccccc}
+\begin{tabular}{p{4.5cm}|ccccccccc}
 \hiderowcolors
 
-{} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMKRF} & {\sc DY-KNOW} \\
+{} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc DY-KNOW} \\
 
@@
 {\bf Expressiveness} & & & & & & & & &  \\
-Formal Logic & Prolog & OWL-DL & {\sc CycL} &  &  &  & & & \\
+Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses & \\
+Expressive power & & & & & & & & &  \\
 OWA/CWA & CWA & OWA & & & & & & & \\
-Modeling uncertainty & +++ (add-on) & & & & & & & & \\
+Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) & \\
 Meta-cognition & ++ & ++ & & & & & & & \\
 {\bf Model} & & & & & & & & &  \\
+Representation of space & & & & & & & & ++ &  \\
+Representation of temporality & & & & & & & & + &  \\
+Context & & & & & & & & ++ &  \\
+Alternative worlds & & +++ & & & & & & &  \\
+Introspection & + (SRDF) & & & & & & & &  \\
+Memory & & + & & & & & & &  \\
 {\bf Reasoning} & & & & & & & & &  \\
 Standard FOL reasoning & +++ & +++ & & & & & & &  \\
+Knowledge structure alteration & & ++ & & & & & & &  \\
 Lazy evaluation & +++ & & & & & & & &  \\
-Reasoning under uncertainty & ++ (add-on) & & & & & & & & \\
+Reasoning under uncertainty & ++ (extension) & & & & & & & & \\
 Non-monotonic reasoning & & & & & & & & &  \\
@@
 Prediction, projection, diagnosis & & & & & & & & &  \\
-Physics-based reasoning & +++ (add-on) & & & & & & & &  \\
+Physics-based reasoning & +++ (extension) & & & & & & & &  \\
 Planning & + & +++ (external module) & & & & & & &  \\
-Knowledge structure alteration & & ++ & & & & & & &  \\
 Learning & & & & & & & & &  \\
 {\bf Knowledge Acquisition} & & & & & & & & &  \\
+Acquisition techniques & & & & & & & & &  \\
+Grounding & & & & & & & & &  \\
+Instanciation & & & & & & & & &  \\
 {\bf Integration} & & & & & & & & &  \\
-{\bf Knowledge Instanciation} & & & & & & & & &  \\
+...with sensori-motor layers & & & & & & & & &  \\
+...with executive layers & & & & & & & & &  \\
+Monitoring and debugging & & & & & & & & &  \\
+Performances & & & & & & & & &  \\
+
 \hline

### 2012-04-20 krs-survey.tex
Formatting of the summary table

@@
 \usepackage{tabularx}
+\usepackage{booktabs}
 \usepackage{supertabular}
+\usepackage{multirow}
+\usepackage{varwidth}
+
+\newcommand{\turn}[3][10em]{% \turn[<width>]{<angle>}{<stuff>}
+  \rlap{\rotatebox{#2}{\begin{varwidth}[t]{#1}\bfseries#3\end{varwidth}}}%
+  }
+
 \usepackage{pdflscape} %% Used for very big table
@@
 \begin{center}
-\rowcolors{2}{lightgray}{codegray}
-
-%\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
-\begin{tabular}{p{4.5cm}|ccccccccc}
-\hiderowcolors
-
-{} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc DY-KNOW} \\
-
+
+\begin{tabular}{cp{4.5cm}ccccccccc}
+\toprule
+\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc DY-KNOW} \\
+
+\midrule
+
+\multirow{5}{0.7cm}[-0.2cm]{\turn[1.5cm]{90}{\bf Expressive power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses & \\
+ & Expressive power & & & & & & & & &  \\
+ & OWA/CWA & CWA & OWA & & & & & & & \\
+ & Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) & \\
+ & Meta-cognition & ++ & ++ & & & & & & & \\
 \hline
-\showrowcolors
-{\bf Expressiveness} & & & & & & & & &  \\
-Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses & \\
-Expressive power & & & & & & & & &  \\
-OWA/CWA & CWA & OWA & & & & & & & \\
-Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) & \\
-Meta-cognition & ++ & ++ & & & & & & & \\
-{\bf Model} & & & & & & & & &  \\
-Representation of space & & & & & & & & ++ &  \\
-Representation of temporality & & & & & & & & + &  \\
-Context & & & & & & & & ++ &  \\
-Alternative worlds & & +++ & & & & & & &  \\
-Introspection & + (SRDF) & & & & & & & &  \\
-Memory & & + & & & & & & &  \\
-{\bf Reasoning} & & & & & & & & &  \\
-Standard FOL reasoning & +++ & +++ & & & & & & &  \\
-Knowledge structure alteration & & ++ & & & & & & &  \\
-Lazy evaluation & +++ & & & & & & & &  \\
-Reasoning under uncertainty & ++ (extension) & & & & & & & & \\
-Non-monotonic reasoning & & & & & & & & &  \\
-Presupposition accomodation & & & & & & & +++ & &  \\
-Prediction, projection, diagnosis & & & & & & & & &  \\
-Physics-based reasoning & +++ (extension) & & & & & & & &  \\
-Planning & + & +++ (external module) & & & & & & &  \\
-Learning & & & & & & & & &  \\
-{\bf Knowledge Acquisition} & & & & & & & & &  \\
-Acquisition techniques & & & & & & & & &  \\
-Grounding & & & & & & & & &  \\
-Instanciation & & & & & & & & &  \\
-{\bf Integration} & & & & & & & & &  \\
-...with sensori-motor layers & & & & & & & & &  \\
-...with executive layers & & & & & & & & &  \\
-Monitoring and debugging & & & & & & & & &  \\
-Performances & & & & & & & & &  \\
-
+\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Representation of space & & & & & & & & ++ &  \\
+ & Representation of temporality & & & & & & & & + &  \\
+ & Context & & & & & & & & ++ &  \\
+ & Alternative worlds & & +++ & & & & & & &  \\
+ & Introspection & + (SRDF) & & & & & & & &  \\
+ & Memory & & + & & & & & & &  \\
 \hline
+\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & & & & &  \\
+ & Knowledge structure alteration & & ++ & & & & & & &  \\
+ & Lazy evaluation & +++ & & & & & & & &  \\
+ & Reasoning under uncertainty & ++ (extension) & & & & & & & & \\
+ & Non-monotonic reasoning & & & & & & & & &  \\
+ & Presupposition accomodation & & & & & & & +++ & &  \\
+ & Prediction, projection, diagnosis & & & & & & & & &  \\
+ & Physics-based reasoning & +++ (extension) & & & & & & & &  \\
+ & Planning & + & +++ (external module) & & & & & & &  \\
+ & Learning & & & & & & & & &  \\
+\hline
+\multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & &  \\
+ & Grounding & & & & & & & & &  \\
+ & Instanciation & & & & & & & & &  \\
+\hline
+\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & & &  \\
+ & ...with executive layers & & & & & & & & &  \\
+ & Monitoring and debugging & & & & & & & & &  \\
+ & Performances & & & & & & & & &  \\
+
+\bottomrule
 

### 2012-04-20 krs-survey.tex
Removed DyKnow from reviewed system + mentionned Heintz survey
DyKnow has been removed because no focus on 'service robot'/'human environment' + a bit too lowlevel

@@
 
+This exclude \textit{de-facto} platforms like {\sc DyKnow}~\cite{Heintz2004}
+which are focused on data fusion and knowledge grounding at lower levels.
+
 While classical cognitive systems like {\sc Soar} or {\sc ACT-R} have
@@
 
-However, several authors from fields that are connected to the robotics have
-previously listed desirable features of artificial systems aiming at cognitive
-abilities as rich as possible.
-
-For instance, in~\cite{McCarthy2007}, McCarthy recently listed the challenges
-he identifies on the road to a \emph{human-level AI}.
-
-\begin{itemize}
-
-	\item the ability to \emph{"operate successfully in the common sense
-	informatic situation"},
-
-	\item the necessity of relying on mathematical logic, as the most fruitful
-	formalism for machine intelligence,
-
-	\item the ability to deal with \emph{approximate concepts and approximate
-	theories} (that would include representing them, and reasoning with them),
-
-	\item non-monotonic reasoning,
-
-	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
-	extend \emph{on demand} the closed domain of interpretation for a
-	given assertion,
-
-	\item the ability to formalize and reason about contexts,
-
-	\item reasoning about events, and in particular, actions,
-
-	\item the capacity of introspection,
-
-	\item and finally, he points the issue of giving computer the right
-	heuristics for decision making.
-
-\end{itemize}
+In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
+\emph{knowledge processing middlewares} as systems supporting ``declarative
+specifications for flexible configuration and dynamic reconfiguration of
+context dependent processing at many different levels of abstraction''. They
+identify six characteristics: the system must be able to merge informations
+from different, possibly distributed sources; it should support quantitative as
+well as qualitative processing of information, it should offer bottom-up and
+top-down processing, it should be able to deal with uncertainty, allow for
+``flexible configuration and reconfiguration'' (which require what we call
+\emph{non-monotonicity} in this article) and eventually Heintz et al. call for
+meta-knowledge and introspective capacities (``declarative specification of the
+processing functionalities''). In this paper, we focus our analysis on similar
+system with however an emphazise on the symbolic capabilities.
 
@@
 knowledge representation strategies and are relevant to our study.
+
+At a broader scope, several authors from fields that are connected to robotics
+have previously listed desirable features of artificial systems aiming at
+cognitive abilities as rich as possible.
+
+For instance, in~\cite{McCarthy2007}, McCarthy recently listed the challenges
+he identifies on the road to a \emph{human-level AI}.
+
+\begin{itemize}
+
+	\item the ability to \emph{"operate successfully in the common sense
+	informatic situation"},
+
+	\item the necessity of relying on mathematical logic, as the most fruitful
+	formalism for machine intelligence,
+
+	\item the ability to deal with \emph{approximate concepts and approximate
+	theories} (that would include representing them, and reasoning with them),
+
+	\item non-monotonic reasoning,
+
+	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
+	extend \emph{on demand} the closed domain of interpretation for a
+	given assertion,
+
+	\item the ability to formalize and reason about contexts,
+
+	\item reasoning about events, and in particular, actions,
+
+	\item the capacity of introspection,
+
+	\item and finally, he points the issue of giving computer the right
+	heuristics for decision making.
+
+\end{itemize}
 
@@
 OMRKF & & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe/?mid=symbolic_reasoning} & ? & Horn Clauses & ? & \cite{Suh2007} \\
-DY-KNOW & & Heintz, Dowerty \par (Linköping Univ.) & & & & & \cite{Heintz2004} \\
 Ke Jia & & Chen et al. \par (Univ. of Science and Technology of China) & & & & & \cite{Chen2010} \\
@@
 \begin{center}
-\rowcolors{2}{lightgray}{codegray}
 
 \begin{tabular}{lp{4cm}}
-\hiderowcolors
+\toprule
 {\bf Project} & {\bf Common-sense \par knowledge source} \\
-\hline
-\showrowcolors
+\midrule
 {\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
@@
 PEIS Ecology & {\sc ResearchCyc} \\
-DY-KNOW & \\
 OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
@@
 ARMAR &  \\
-Hertzberg (Osnabrück) & \\
+Hertzberg & \\
 (DFKI Bremen) & \\
@@
 
-\hline
+\bottomrule
 
@@
 \toprule
-\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc DY-KNOW} \\
+\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} \\
 
@@
 
-\multirow{5}{0.7cm}[-0.2cm]{\turn[1.5cm]{90}{\bf Expressive power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses & \\
- & Expressive power & & & & & & & & &  \\
- & OWA/CWA & CWA & OWA & & & & & & & \\
- & Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) & \\
- & Meta-cognition & ++ & ++ & & & & & & & \\
+\multirow{5}{0.7cm}[-0.2cm]{\turn[1.5cm]{90}{\bf Expressive power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses \\
+ & Expressive power & & & & & & & &  \\
+ & OWA/CWA & CWA & OWA & & & & & & \\
+ & Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) \\
+ & Meta-cognition & ++ & ++ & & & & & & \\
 \hline
-\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Representation of space & & & & & & & & ++ &  \\
- & Representation of temporality & & & & & & & & + &  \\
- & Context & & & & & & & & ++ &  \\
- & Alternative worlds & & +++ & & & & & & &  \\
- & Introspection & + (SRDF) & & & & & & & &  \\
- & Memory & & + & & & & & & &  \\
+\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Representation of space & & & & & & & & ++  \\
+ & Representation of temporality & & & & & & & & +  \\
+ & Context & & & & & & & & ++  \\
+ & Alternative worlds & & +++ & & & & & &  \\
+ & Introspection & + (SRDF) & & & & & & &  \\
+ & Memory & & + & & & & & &  \\
 \hline
-\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & & & & &  \\
- & Knowledge structure alteration & & ++ & & & & & & &  \\
- & Lazy evaluation & +++ & & & & & & & &  \\
- & Reasoning under uncertainty & ++ (extension) & & & & & & & & \\
- & Non-monotonic reasoning & & & & & & & & &  \\
- & Presupposition accomodation & & & & & & & +++ & &  \\
- & Prediction, projection, diagnosis & & & & & & & & &  \\
- & Physics-based reasoning & +++ (extension) & & & & & & & &  \\
- & Planning & + & +++ (external module) & & & & & & &  \\
- & Learning & & & & & & & & &  \\
+\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & & & & +  \\
+ & Knowledge structure alteration & & ++ & & & & & &  \\
+ & Lazy evaluation & +++ & & & & & & &  \\
+ & Reasoning under uncertainty & ++ (extension) & & & & & & & \\
+ & Non-monotonic reasoning & & & & & & & &  \\
+ & Presupposition accomodation & & & & & & & +++ & \\
+ & Prediction, projection, diagnosis & & & & & & & & \\
+ & Physics-based reasoning & +++ (extension) & & & & & & &  \\
+ & Planning & + & +++ (external module) & & & & & & + \\
+ & Learning & & & & & & & &  \\
 \hline
-\multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & &  \\
- & Grounding & & & & & & & & &  \\
- & Instanciation & & & & & & & & &  \\
+\multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & \\
+ & Grounding & & & & & & & & ++ \\
+ & Instanciation & & & & & & & & \\
 \hline
-\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & & &  \\
- & ...with executive layers & & & & & & & & &  \\
- & Monitoring and debugging & & & & & & & & &  \\
- & Performances & & & & & & & & &  \\
+\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & & ++ \\
+ & ...with executive layers & & & & & & & & ++ \\
+ & Monitoring and debugging & & & & & & & & \\
+ & Performances & & & & & & & & \\
 

### 2012-04-20 krs-survey.tex
Minor update to the Soar/ACT-R paragraph

@@
 
-While classical cognitive systems like {\sc Soar} or {\sc ACT-R} have
-declarative knowledge modules and have been recently used on service robots
-(see {\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are absent from this
-survey because we could not find evidence in the litterature on possible
-on-line manipulation of symbolic knowledge.
+While classical cognitive architectures like {\sc Soar} or {\sc ACT-R} have
+declarative knowledge modules~\cite{Derbinsky2010} and have been recently used
+on service robots (see {\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are
+also absent from this survey because we did not find much references in the
+litterature on knowledge manipulation and representation applied to robotic
+scenarii in these architectures.
 

### 2012-04-20 krs-survey.tex
Added 'interaction' as subsection of knowledge acquisition
Added section on GSM

@@
 
-\paragraph{Perception}
-\paragraph{Interaction}
-\paragraph{External sources (Web, upper ontologies, ...)}
-\paragraph{Learning}
-
 \subsubsection{Grounding/anchoring strategies}
 \label{sect|grounding}
+
+\subsubsection{Interaction with other agents}
+\label{sect|interactions}
 
@@
 
+\subsection{GSM}
+\label{sect|gsm}
+
+GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
+representation system primarly built to ``facilitate cross-modal
+interoperability'',  especially in the context of verbal interaction with a
+robot.
+
+Unlike most of the previously presented systems, GSM does not rely on any
+formal language but rather on a layered datastructure (Figure~\ref{fig|gsm})
+that organizes the surrounding world into agents and relations between agents.
+Each agent (any animate or inanimate object) is attached to a physical model (made
+of \emph{body parts} that have properties like their position, color, etc.) and
+a mental model (which is a recursively embedded GSM, thus allowing a sort of
+theory of mind).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{gsm.pdf}
+
+    \caption{Simplified hierachical structure of the Grounded Situation Model,
+    based on~\cite{Mavridis2006}}
+
+    \label{fig|gsm}
+   \end{figure}
+
+Properties are represented in three layers: a stochastic representation, close
+to sensory percepts, a \emph{continous single-valued} encoding of the
+stochastic model, and a discrete, categorical model.
+
+One notable feature of GSM is the \emph{bidirectionality} of the grounding
+process: not only sensor percepts are abstracted into categories suitable for
+human conversation, but human utterance (like ``There is a red ball in the
+center of the table'') can also be turned into property descriptions. This
+basically enable the knowledge representation system of the robot to
+\emph{imagine} entities.
+
+GSM also features several strategies for managing time and events.
+\emph{Moments} are created by storing timestamped snap-shots of GSM, and
+\emph{event classifiers} allow to define and detect events.
+
+\paragraph{Experiments} GSM has mostly been tested on table-top manipulation
+and interaction tasks (a ``conversational helping hand'' as stated by the
+authors) realized on a 7-DOF arm equipped with force feedback, cameras for blob
+tracking and speech recognition (Sphinx4).
+
 \subsection{OMRKF}
@@
 PEIS Ecology & {\sc ResearchCyc} \\
-OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
-GSM &  \\
 NKLR &  None \\
 CAST Proxies &  None \\
+GSM &  Predefined categories \\
+OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
 Wrighteagle & \\
@@
  & OWA/CWA & CWA & OWA & & & & & & \\
- & Modeling uncertainty & +++ (extension) & & & & & & & + (\emph{candidate} entities) \\
+ & Modeling uncertainty & +++ (extension) & & & & & & ++ (stochastic) & + (\emph{candidate} entities) \\
  & Meta-cognition & ++ & ++ & & & & & & \\
@@
 \multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Representation of space & & & & & & & & ++  \\
- & Representation of temporality & & & & & & & & +  \\
+ & Representation of temporality & & & & & & & + (snapshots) & +  \\
  & Context & & & & & & & & ++  \\
- & Alternative worlds & & +++ & & & & & &  \\
+ & Alternative worlds & & +++ & & & & ++ & &  \\
  & Introspection & + (SRDF) & & & & & & &  \\
- & Memory & & + & & & & & &  \\
+ & Memory models & & + & & & & & &  \\
 \hline
@@
  & Lazy evaluation & +++ & & & & & & &  \\
- & Reasoning under uncertainty & ++ (extension) & & & & & & & \\
+ & Reasoning under uncertainty & ++ (extension) & & & & & + & & \\
  & Non-monotonic reasoning & & & & & & & &  \\
@@
  & Physics-based reasoning & +++ (extension) & & & & & & &  \\
- & Planning & + & +++ (external module) & & & & & & + \\
+ & Planning & + & +++ (external) & & & & & & + \\
  & Learning & & & & & & & &  \\
@@
 \multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & \\
- & Grounding & & & & & & & & ++ \\
+ & Grounding & & & & & & & ++ & ++ \\
+ & Interaction & & +++ (external) & & & & & +++ & \\
  & Instanciation & & & & & & & & \\
 \hline
-\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & & ++ \\
- & ...with executive layers & & & & & & & & ++ \\
+\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & + & + \\
+ & ...with executive layers & & & & & & & + & ++ \\
  & Monitoring and debugging & & & & & & & & \\

### 2012-04-23 krs-survey.tex
Added reference to Ke Jia project (wrighteagle)

@@
 
-One important exception if the \emph{Negation as failure} inference rule, as
+One important exception is the \emph{negation as failure} inference rule, as
 implemented by {\sc Prolog} for instance, that allows for non-monotonicity, but
@@
 \fxfatal{Give here an example of non-monotonic reasoning with Prolog}
+\fxwarning{Do we mention here Answer Set Programming?}
 
@@
 
-\subsubsection{Interaction with other agents}
-\label{sect|interactions}
-
 \subsubsection{Ability to automatically create new object instances}
@@
 
-\paragraph{Language integration}
-
 \paragraph{Events}
+
+\subsubsection{Language Processing}
+\label{sect|nlp}
 
@@
 \begin{center}
-\rowcolors{2}{lightgray}{codegray}
-
-%\begin{tabularx}{\textheight}{p{2cm}p{4cm}p{3cm}p{4cm}p{2.3cm}p{2cm}p{2cm}}
+
 \begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{2.8cm}p{1.5cm}}
-\hiderowcolors
+\toprule
 {\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & {\bf Reasoner} & Main reference \\
-\hline
-\showrowcolors
-{\sc KnowRob} & KRS & Tenorth, Beetz \par (TU Munich) & \url{http://ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
+\midrule
+{\sc KnowRob} & KRS & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
 ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
@@
 NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
-GSM & & Mavridis, Roy \par (MIT MediaLab) & & & & & \cite{Mavridis2006} \\
-OMRKF & & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe/?mid=symbolic_reasoning} & ? & Horn Clauses & ? & \cite{Suh2007} \\
-Ke Jia & & Chen et al. \par (Univ. of Science and Technology of China) & & & & & \cite{Chen2010} \\
+GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
+OMRKF & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe/?mid=symbolic_reasoning} & ? & Horn Clauses & ? & \cite{Suh2007} \\
+Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
  & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\
@@
 
-\hline
-
-%\end{tabularx}
+\bottomrule
+
 \end{tabular}
@@
 descriptor by introducing a $candidate()$ function.
+
+\subsection{Ke Jia Project}
+\label{sect|kejia}
+
+The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
+representation language with natural language processing, task planing and
+motion planing.
+
+Knowledge representation relies on \emph{Action Language C}, itself based on
+\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
+are syntactically close to Prolog, are based on \emph{stable models} of logic
+programs, and support non-monotonic reasoning. Default and non-monotonic
+reasoning has been especially researched within the Ke Jia project for symbolic
+task planing~\cite{Ji2011} and underspecified natural language processing.
+
+Amongst other features, the natural language processing capabilities of the
+system support acquisition of new logical rules at run-time.
+
+\paragraph{Experiments} The Ke Jia robot has been demonstrated in several tasks
+involving human-robot interaction with natural language. These tasks include a
+task with multiple \emph{pick \& carry} that are globally optimized, naive
+physics reasoning via teached rules or more complex scenarii with the robot
+delivering drinks, taking into account changing and mutually exclusive
+preferences of users.
 
@@
 \toprule
-\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Golog} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} \\
+\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} \\
 
@@
 
-\multirow{5}{0.7cm}[-0.2cm]{\turn[1.5cm]{90}{\bf Expressive power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  &  &  & & Horn clauses \\
- & Expressive power & & & & & & & &  \\
- & OWA/CWA & CWA & OWA & & & & & & \\
+\multirow{5}{0.7cm}[0.2cm]{\turn[1.5cm]{90}{\bf Expr. power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  & ASP &  & & Horn clauses \\
+ & OWA/CWA & CWA & OWA & & & CWA & & & \\
  & Modeling uncertainty & +++ (extension) & & & & & & ++ (stochastic) & + (\emph{candidate} entities) \\
@@
 \hline
-\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Representation of space & & & & & & & & ++  \\
- & Representation of temporality & & & & & & & + (snapshots) & +  \\
+\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Space Representation & & & & & & & & ++  \\
+ & Time representation & & & & & & & + (snapshots) & +  \\
  & Context & & & & & & & & ++  \\
@@
 \hline
-\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & & & & +  \\
- & Knowledge structure alteration & & ++ & & & & & &  \\
+\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & +++ & & & +  \\
+ & Knowledge structure alteration & & ++ & & & ++ & & &  \\
  & Lazy evaluation & +++ & & & & & & &  \\
  & Reasoning under uncertainty & ++ (extension) & & & & & + & & \\
- & Non-monotonic reasoning & & & & & & & &  \\
+ & Non-monotonic reasoning & & & & & +++ & & &  \\
  & Presupposition accomodation & & & & & & & +++ & \\
@@
  & Physics-based reasoning & +++ (extension) & & & & & & &  \\
- & Planning & + & +++ (external) & & & & & & + \\
+ & Planning & + & +++ (external) & & & ++ & & & + \\
  & Learning & & & & & & & &  \\
@@
 \multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & \\
- & Grounding & & & & & & & ++ & ++ \\
- & Interaction & & +++ (external) & & & & & +++ & \\
+ & Grounding & & ++ (external) & & & & & ++ & ++ \\
  & Instanciation & & & & & & & & \\
@@
  & ...with executive layers & & & & & & & + & ++ \\
+ & NLP & & +++ (external) & & & +++ & & +++ & \\
  & Monitoring and debugging & & & & & & & & \\
- & Performances & & & & & & & & \\
+ & Performances & & & & & + & & & \\
 

### 2012-04-23 krs-survey.tex
Filled section on symbol grounding

@@
 \label{sect|grounding}
+
+\emph{Grounding} (also called \emph{anchoring} when specifically refering to
+building links between \emph{percepts} and \emph{physical
+objects}~\cite{Coradeschi2003}) is the task consisting in building and
+maintaining a bi-directional link between sub-symbolic representations (sensors
+data, low-level actuation) and symbolic represtentations that can be
+manipulated and reasoned about~\cite{Harnad1990}.
+
+Being embodied entities with interaction with other embodied entities as a
+fundamental requirement, robots and robotic is deeply concerned by the
+grounding issue.
+
+Being actually implemented on real service robots, all the symbolic knowledge
+representation systems that we review in this study have some kind of grounding
+process. Numerous approaches exist, like amodal
+\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
+representations~\cite{Alami2011, Mavridis2006} or semantic maps
+(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{semanticmaps_hertzberg.png}
+    \caption{Example of a semantic map, taken from~\cite{Nuechter2008}.}
+    \label{fig|semanticmap}
+\end{figure}
 

### 2012-04-23 krs-survey.tex
Added a section on ARMAR

@@
 \begin{landscape}
-\begin{table}
+\begin{table}\footnotesize
 \begin{center}
@@
 CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
-Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
 NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
 GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
-OMRKF & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe/?mid=symbolic_reasoning} & ? & Horn Clauses & ? & \cite{Suh2007} \\
+OMRKF & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
 Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
+ARMAR/Tapas & KRS & Holzapfel, Waible \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
+Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
  & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\
  & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\
-ARMAR & & Schmidt-Rohr (Karlsruhe TH) \\
  & & Hertzberg (Osnabrück Univ.) \\
- & & (DFKI Bremen) \\
  (based on {\sc KnowRob} & & (JSK) \\
+ OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
 
@@
 
+\subsection{ARMAR/Tapas}
+
+{\sc Tapas} is the name of the knowledge representation system and dialogue
+manager found on the ARMARIII robot~\cite{Holzapfel2008}.
+
+Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
+knowledge. The later is split into \emph{lexical knowledge}, \emph{semantic
+knowledge} and a database of identified objects (with their properties). The
+\emph{lexical knowledge} contains lexical and grammatical informations about
+the objects. The \emph{semantic knowledge} is organized into an ontology
+relying on \emph{typed feature structures} (TFS,~\cite{Carpenter1992}, a
+formalism originating from the computational linguistics community, and a
+superset of first-order logic).
+
+{\sc Tapas} has a strong focus on natural language grounding. It proceeds by
+generating grammars from properties represented in the ontology to parse and
+understand dialogue.
+
+Another focus is put on handling unknown words and objects. {\sc Tapas}
+provides routines to recognize unknown entities, and propose and interactive
+and iterative verbal process to categorize (including adding new caterogies)
+those new concepts.
+
 %%%%%%%%%% Underlying knowledge model table %%%%%%%%
@@
 OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
-Wrighteagle & \\
+Ke Jia & None \\
+ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
 Vincze & \\
-ARMAR &  \\
-Hertzberg & \\
-(DFKI Bremen) & \\
 (JSK) & \\
@@
 \begin{landscape}
-\begin{table}
+\begin{table}\footnotesize
 \begin{center}
 
-\begin{tabular}{cp{4.5cm}ccccccccc}
+\begin{tabular}{cp{4.5cm}p{2.3cm}p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.5cm}p{2cm}p{1.5cm}p{1.5cm}}
 \toprule
-\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} \\
+\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc ARMAR} & {\sc Hertzberg} \\
 
@@
 
-\multirow{5}{0.7cm}[0.2cm]{\turn[1.5cm]{90}{\bf Expr. power}} & Logical formalism & Prolog & OWL-DL & {\sc CycL} &  & ASP &  & & Horn clauses \\
- & OWA/CWA & CWA & OWA & & & CWA & & & \\
+\multirow{5}{0.7cm}[0.2cm]{\turn[1.5cm]{90}{\bf Expr. power}} & Logical formalism & Prolog & DL (OWL) & {\sc CycL} &  & ASP &  & & Horn clauses & TFS & \\
+ & OWA/CWA & CWA & OWA & ? & ? & CWA & ? & ? & ? & ? & ? \\
  & Modeling uncertainty & +++ (extension) & & & & & & ++ (stochastic) & + (\emph{candidate} entities) \\
@@
 \multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & +++ & & & +  \\
- & Knowledge structure alteration & & ++ & & & ++ & & &  \\
+ & Knowledge structure alteration & & ++ & & & ++ & & & & ++ &  \\
  & Lazy evaluation & +++ & & & & & & &  \\
@@
  & Planning & + & +++ (external) & & & ++ & & & + \\
- & Learning & & & & & & & &  \\
+ & Learning & & + & & & & & & & ++ &  \\
 \hline
 \multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & \\
- & Grounding & & ++ (external) & & & & & ++ & ++ \\
+ & Grounding & +++ \par (semantic maps) & ++ (external amodal model) & & & & & ++ & ++ \\
  & Instanciation & & & & & & & & \\
@@
  & ...with executive layers & & & & & & & + & ++ \\
- & NLP & & +++ (external) & & & +++ & & +++ & \\
+ & NLP & & +++ (external) & & & +++ & & +++ & & +++ & \\
  & Monitoring and debugging & & & & & & & & \\

### 2012-04-23 krs-survey.tex
Reorganizing the conclusion, started to affine content

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary of Approaches}
-\label{sect|summary}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Towards the next generation of Knowledge Representation Systems for Robotics}
+\label{sect|conclusion}
+
+This section tries to summary the various approaches we surveyed in the
+previous sections and to draw some research perspectives. This section does not
+exactly follow the list of compared features presented in
+section~\ref{sect|features}. We have organized it along four axis: \emph{what
+can be represented?}, \emph{How knowledge is created and grounded?}, \emph{What
+can be done with the knowledge?} and \emph{How to use knowledge in the whole,
+larger robot architecture?}.
+
+Table~\ref{table|contribution-by-systems} proposes a summary of the main scopes
+of \emph{contribution} of the systems we have surveyed. This overview helps to
+identify the main weaknesses of current approaches of knowledge representation
+regarding the long-term goal of \emph{human-level robots}.
 
@@
 
-This section tries to summary the various approaches we surveyed in the
-previous section.  This summary does not exactly follow the list of compared
-features presented in section~\ref{sect|features}. We have organized
-it along four axis: \emph{what can be represented?}, \emph{How knowledge is
-created and grounded?}, \emph{What can be done with the knowledge?} and
-\emph{How to use knowledge in the whole, larger robot architecture?}.
-
 \subsection{Expressiveness: what can we represent in the current state-of-the-art?}
@@
 \begin{itemize}
-	\item expressiveness
+    \item common-sense knowledge
+    \begin{itemize}
+        \item \emph{default reasoning}
+        \item \emph{uncertainty representation/reasoning}
+        \item \emph{physics-based reasoning}
+    \end{itemize}
+    \item Spatio-temporal representations
+    \item Meta-cognition and introspection
+    \item Context
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Towards the next generation of Knowledge Representation Systems for Robotics}
-\label{sect|conclusion}
-
 \subsection{Current shotcomings: what is not successfully covered by current systems}

### 2012-04-24 krs-survey.tex
Mention affordance based approaches to grounding

@@
 \emph{proxies}~\cite{Jacobsson2008}, grounded amodal
-representations~\cite{Alami2011, Mavridis2006} or semantic maps
-(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}).
+representations~\cite{Alami2011, Mavridis2006}, semantic maps
+(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
+affordance-based planing and object classification~\cite{Lorken2008,
+Varadarajan2011}.
 

### 2012-04-24 krs-survey.tex
Added section on natural language processing

@@
 \label{sect|nlp}
+
+The scope of this survey is service robotic, which suggest human-robot
+interaction. While numerous modalities of communication between humans and
+robots do exist and have been extensively studied, the most challenging of
+these interactions is natural, bidirectional, unconstrained verbal
+communication. Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date
+survey of literature on situated human-robot dialogue, focusing on formal
+representation systems, bi-directionality of the interaction and context
+building.
+
+Natural Language Processing (NLP) is a large research field by itself, and
+while the robotic community may be lagging behind on many theoretical aspects,
+it brings one important aspect: the embodiment. Because the interactors, both
+the robot and the human, are establishing a communication within a shared
+physical context, the verbal communication channel is complemented by deictic
+channels, back channels and possibly shared physical experiences: a human can
+show something to a robot, saying ``Give me this''. This is simply not possible
+for a virtual agent.
+
+Several of the knowledge representation systems we present have developed
+specific built-in mechanisms or extensions to parse, ground and possibly
+rebuild natural language.
 

### 2012-04-24 krs-survey.tex
Spell checking

@@
 Those questions have in common one prerequisite to be answered: the robot must
-be able to manipulate \emph{explicitely} knowledge, and hence, need a wayto
+be able to manipulate \emph{explicitly} knowledge, and hence, need a way to
 represent the knowledge.
@@
 six broad categories: \emph{what can be represented?} or the
-\emph{expresiveness} question ; \emph{how things are represented?} ; \emph{what
+\emph{expressiveness} question ; \emph{how things are represented?} ; \emph{what
 reasoning techniques are offered?} ; \emph{how knowledge is acquired and
@@
 components of the robot?} and \emph{which actual knowledge is stored?} or the
-\emph{knowledge instanciation} question.
-
-We have identified over ten projects that are either explicitely advertised as
+\emph{knowledge instantiation} question.
+
+We have identified over ten projects that are either explicitly advertised as
 knowledge representation systems or involve explicit knowledge manipulation
@@
 assertions, be it static statements on the world or spatio-temporal events, and
-reasoning on this representation. Proeminent features of each of these systems
+reasoning on this representation. Prominent features of each of these systems
 illustrate how the various features of a knowledge representation system are
@@
 
-Every robotic system has, implicitely or not, some knowledge representation
+Every robotic system has, implicitly or not, some knowledge representation
 systems. It may range from a simple state vector to an explicit symbolic
@@
     \item  run on \emph{service robot} (that is, robots that interact with 
-    objects in a semantic-rich environment primarly designed for humans),
+    objects in a semantic-rich environment primarily designed for humans),
     \item  ground the knowledge in the physical world (physically embedded
@@
 also absent from this survey because we did not find much references in the
-litterature on knowledge manipulation and representation applied to robotic
+literature on knowledge manipulation and representation applied to robotic
 scenarii in these architectures.
@@
 processing functionalities''). In this paper, we focus our analysis on similar
-system with however an emphazise on the symbolic capabilities.
+system with however an emphasize on the symbolic capabilities.
 
@@
 the structure of the world is captured in a representation and shared), the
-relation to \emph{embodiement}, the ability to \emph{anticipate} and to
+relation to \emph{embodiment}, the ability to \emph{anticipate} and to
 \emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
@@
 Section~\ref{sect|surveyed-systems} then presents each of the systems we
-examinated, and underlines their specificities.
+examined, and underlines their specificities.
 
@@
 
-A related ability is called \emph{pre-supposition accomodation}: if one of the
+A related ability is called \emph{pre-supposition accommodation}: if one of the
 agent moves behind another one, with the brownie dough in its arm, and says
@@
 their own, and practical considerations (especially the availability of
-reasoners performant enough for on-line use on a robot) often constraint the
+reasoners efficient enough for on-line use on a robot) often constraint the
 choice of a logical formalism and a level of expressive power.
@@
     \includegraphics[width=0.6\columnwidth]{expressive_overlap_dl_horn.pdf}
-    \caption{Expresiveness overlap of Description Logics and logic programs
+    \caption{expressiveness overlap of Description Logics and logic programs
     based on Horn clauses, taken from~\cite{Grosof2003}}
@@
 
-Figure~\ref{fig|overlap_dl_horn} shows that the expresive power of description
+Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
 logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
@@
 that allows to conveniently explore these relationships and indexes most of the
-litterature on that subject.
+literature on that subject.
 
@@
 knowledge), it is not always mandatory. The current state of the robot can be
-represented as a simple instanciation of a specific category (for instance, if
+represented as a simple instantiation of a specific category (for instance, if
 the robot give an object to the human, this state could be represented with the
@@
 the description is updated and reflect the current (and possibly past and
-future) state of the robot. Note that these description may also invlove
+future) state of the robot. Note that these description may also involve
 geometric descriptions (a kinematic chain, the pose of a device, etc.) that may
@@
 
-Memory has been studied at lenght in the cognitive psychology and
+Memory has been studied at length in the cognitive psychology and
 neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
@@
 facts that are not forgotten by the robot until it is halted (this memory is
-often refered as a \emph{working memory}, but with a meaning unrelated to
+often referred as a \emph{working memory}, but with a meaning unrelated to
 Baddeley's definition). Some systems may propose persistent storages that allow
@@
 
-        \item universal, existantial and cardinality constraints (including \concept{allValue}, 
+        \item universal, existential and cardinality constraints (including \concept{allValue}, 
         \concept{someValue}, \concept{hasValue}),
@@
 
-As mentionned earlier, knowledge models based on description logics can be extended through rule languages.
+As mentioned earlier, knowledge models based on description logics can be extended through rule languages.
 
@@
 
-A monotonic system does not theoretically allow for knowledge retractation,
+A monotonic system does not theoretically allow for knowledge retraction,
 which is an important issue in the robotic context where the world model is
@@
 \subsubsection{Presupposition Accommodation}
-\label{sect|presupposition-accomodation}
+\label{sect|presupposition-accommodation}
 
@@
 
-A knowledge representation system able to copte with presupposition
+A knowledge representation system able to cope with presupposition
 accommodation would be able to take into account this (usually under-defined)
@@
 \emph{Naive physics reasoning} covers all the everyday reasoning the humans
-unconsciouly perform, like taking into account gravity (``if I drop a ball, it
+unconsciously perform, like taking into account gravity (``if I drop a ball, it
 falls down'') or common physical properties of objects (``a glass may break if
@@
 
-Knowledge acquision is generally not done directly in the knowledge
+Knowledge acquisition is generally not done directly in the knowledge
 representation system. On the contrary, several, if not numerous, external
@@
 different natures, like in the example above) merging will require
-symbolic-level reasoning. The KRS has a direct impact on the feasability and
+symbolic-level reasoning. The KRS has a direct impact on the feasibility and
 ease of such operations.
@@
 
-\emph{Grounding} (also called \emph{anchoring} when specifically refering to
+\emph{Grounding} (also called \emph{anchoring} when specifically referring to
 building links between \emph{percepts} and \emph{physical
@@
 maintaining a bi-directional link between sub-symbolic representations (sensors
-data, low-level actuation) and symbolic represtentations that can be
+data, low-level actuation) and symbolic representations that can be
 manipulated and reasoned about~\cite{Harnad1990}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Knowledge instanciation}
+\subsection{Knowledge instantiation}
 
@@
 
-Table \ref{table|surveyed-systems} presents the fifteen knowledge representation systems surveyed
-in this article.
+Table \ref{table|surveyed-systems} presents the eight knowledge representation
+systems surveyed in this article.
 
@@
 Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
- & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\
- & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\
- & & Hertzberg (Osnabrück Univ.) \\
- (based on {\sc KnowRob} & & (JSK) \\
 OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
+% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
+% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
+% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
+% (based on {\sc KnowRob} & & (JSK) \\
 
@@
 \end{center}
-\caption{List of surveyed systems.
-Projects are listed by category (\emph{KRS} for systems that are explicit knowledge representation and reasoning modules, \emph{Ubiquitous} for systems where knowledge processing is fully distributed, \emph{Language} for languages used as KRS on robots), then names.}
+
+\caption{List of surveyed systems. Projects are listed by category (\emph{KRS}
+for systems that are explicit knowledge representation and reasoning modules,
+\emph{Ubiquitous} for systems where knowledge processing is fully distributed,
+\emph{Language} for languages used as KRS on robots), then names.}
+
 \label{table|surveyed-systems}
@@
 % - object identification based on viewpoint independent SIFT features
-% - formalized anchoring system that explicitely match percieved attributes to predicates
+% - formalized anchoring system that explicitly match percieved attributes to predicates
 % - Cyc predicates
@@
 insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
-module can be seemlessly integrated into the PEIS ecosystem.
+module can be seamlessly integrated into the PEIS ecosystem.
 
@@
 \emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
-While this language is developped since a long time by Zarri~\cite{Zarri1997,
+While this language is developed since a long time by Zarri~\cite{Zarri1997,
 Zarri2008}, recent research direction include application to the robotic
@@
 CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
-comprehensive toolkit aimed at builduing cognitive architectures for robots
+comprehensive toolkit aimed at building cognitive architectures for robots
 through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
 expose a central knowledge base as seen in previous works. It instead
-represents knowledge as unrooted \emph{proxies}. Thoses proxies are formally
+represents knowledge as unrooted \emph{proxies}. Those proxies are formally
 defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
-a set of instanciated features (like $\phi^{Colour}_{red}$) and $u_p$ a
+a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
 \emph{proxies union} that form an equivalence class corresponding to one
@@
 A union of proxies forms a global amodal representation of an entity, that can
-be explicitely shared and manipulated. Being not centralized, the knowledge
-model can be qualified of \emph{ubiquitous}. Futhermore, knowledge source in
+be explicitly shared and manipulated. Being not centralized, the knowledge
+model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
 the CAS architecture is tightly bound to the on-line grounding process (be it
@@
 GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
-representation system primarly built to ``facilitate cross-modal
+representation system primarily built to ``facilitate cross-modal
 interoperability'',  especially in the context of verbal interaction with a
@@
 Unlike most of the previously presented systems, GSM does not rely on any
-formal language but rather on a layered datastructure (Figure~\ref{fig|gsm})
+formal language but rather on a layered data structure (Figure~\ref{fig|gsm})
 that organizes the surrounding world into agents and relations between agents.
@@
 
-    \caption{Simplified hierachical structure of the Grounded Situation Model,
+    \caption{Simplified hierarchical structure of the Grounded Situation Model,
     based on~\cite{Mavridis2006}}
@@
 Properties are represented in three layers: a stochastic representation, close
-to sensory percepts, a \emph{continous single-valued} encoding of the
+to sensory percepts, a \emph{continuous single-valued} encoding of the
 stochastic model, and a discrete, categorical model.
@@
 The Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
-(OMKRF) is a knowledge representation system based on four inter-related
+(OMRKF) is a knowledge representation system based on four inter-related
 \emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
@@
 or ``object feature'', a taxonomy of concepts inside this level (for instance
-$cup : Object \sqsubseteq tableware : Object$) and an instanciation of the
+$cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
 taxonomy ($cup1 : cup$).
@@
 
-Finally, the $Activity$ knowledge class store coumpound actions in a HTN-like
+Finally, the $Activity$ knowledge class store compound actions in a HTN-like
 structure, exploited at run-time by a planner.
@@
 task with multiple \emph{pick \& carry} that are globally optimized, naive
-physics reasoning via teached rules or more complex scenarii with the robot
+physics reasoning via taught rules or more complex scenarii with the robot
 delivering drinks, taking into account changing and mutually exclusive
@@
 provides routines to recognize unknown entities, and propose and interactive
-and iterative verbal process to categorize (including adding new caterogies)
+and iterative verbal process to categorize (including adding new categories)
 those new concepts.
@@
 ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
-Vincze & \\
-(JSK) & \\
 
@@
  & Non-monotonic reasoning & & & & & +++ & & &  \\
- & Presupposition accomodation & & & & & & & +++ & \\
+ & Presupposition accommodation & & & & & & & +++ & \\
  & Prediction, projection, diagnosis & & & & & & & & \\
@@
  & Grounding & +++ \par (semantic maps) & ++ (external amodal model) & & & & & ++ & ++ \\
- & Instanciation & & & & & & & & \\
+ & instantiation & & & & & & & & \\
 \hline
@@
 
-\subsection{Current shotcomings: what is not successfully covered by current systems}
+\subsection{Current shortcomings: what is not successfully covered by current systems}
 

### 2012-04-24 krs-survey.tex
Added references to Soar

@@
 
-While classical cognitive architectures like {\sc Soar} or {\sc ACT-R} have
-declarative knowledge modules~\cite{Derbinsky2010} and have been recently used
-on service robots (see {\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are
-also absent from this survey because we did not find much references in the
-literature on knowledge manipulation and representation applied to robotic
-scenarii in these architectures.
+While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006} or
+{\sc ACT-R} have declarative knowledge modules~\cite{Derbinsky2010} and have
+been recently used on service robots (see {\sc ACT-R/E}~\cite{Kennedy2009} for
+instance), they are also absent from this survey because we did not find much
+references in the literature on knowledge manipulation and representation
+applied to robotic scenarii in these architectures.
 
@@
 associated to specific events).
+
+In the larger field of cognitive architectures, the {\sc Soar}
+architecture~\cite{Lehman2006} is one of those that tries to reproduce a
+human-like memory organization.
 

### 2012-04-24 krs-survey.tex
Filled sections on debugging and performance evaluation

@@
 
+\subsubsection{Planning}
+\label{sect|planning}
+
+Making decision based on prediction
+
 \subsubsection{Physics-based reasoning}
@@
 that carries the geometries and physical properties of objects.
-
-\subsubsection{Planning}
-\label{sect|planning}
-
-Making decision based on prediction
 
@@
 
+Knowledge representation systems do not mean anything to robots if they are
+considered in isolation. This section proposes categories of features related
+to the integration of the KRS into a larger software architecture that includes
+perception routines, decision-making processes and actuation control.
+
+We also mention some practical aspects of a real-world system, like
+performances and monitoring tools that come along with the KRS.
+
 \subsubsection{Integration with sensori-motor layers}
@@
 
-Ability to ``listen'' to the robot internal structures.
+We have previously discussed (section~\ref{sect|grounding}) the principles of
+the grounding process that aims at establishing and maintaining a connection
+between percepts (and to a lesser extend, low-level actions) and symbols.
+
+While every real-world cognitive robot need some kind of grounding, the actual
+implementations lead to very different information flows.
+
+The systems can be roughly split into two classes: \emph{passive} knowledge
+repositories that process symbolic facts produced by lower-level sensori-motor
+layers (\emph{push} flow); \emph{active} knowledge managers that directly query
+(possibly by polling or on-demand) low-level layers.
+
+This macroscopic distinction is however mostly a matter of defining the
+frontiers of the KRS: some systems like KnowRob~\cite{Tenorth2009a} encompass
+geometric reasoning layers that would be considered as external by other
+systems like ORO~\cite{Lemaignan2010} that focus on the symbolic fact storage
+and rely on a ecosystem of independent modules to provide and consume symbolic
+knowledge.
+
+\fxwarning{Some archi with the ability to ``listen'' to the robot internal
+structures -> which ones?}
+
+Other systems do not fit either in such a partition between active and passive
+systems because they do not stand as independent modules but exist as diffuse,
+\emph{ubiquitous} knowledge manipulation system~\cite{Jacobsson2008}, for
+instance because they are primarily language~\cite{Ferrein2008, Sabri2011}.
 
@@
 
-\subsubsection{Is it fast enough? Scalability and responsiveness}
-\label{sect|scalability}
+It is common to have knowledge representation systems at the heart of a
+cognitive robotic architecture, and therefore KRS are easily ``burried'' in the
+system.
+
+At the same time, the symbolic model often provides a valuable synthetic view
+on the whole state of the robot, furthermore easily understandable by the human
+developer (the fact \concept{human1 issitting true} is easier to interpret than
+the suite of relative coordinates of each joints of the human skeleton, as
+provided by the human tracker, for instance).
+
+That is the reason why having at hand good tools to trace and visualize at
+run-time the evolution of the knowledge structure and contents, as well as
+post-processing tools that can be run on the trace to precisely analyze the
+cognitive behaviour of the robot, is useful.
+
+\subsubsection{Evaluation of performances}
+\label{sect|performances}
+
+Benchmarks of symbolic systems for robots are notoriously hard to conduct for
+several reasons: identifying good metrics for robotic experiments in general is
+difficult because of the complex interactions between tenth of modules running
+in parallel. Isolating the role of one specific component is thus hard.
+Furthermore, knowledge representation systems are usually tightly connected to
+numerous other modules, and the lack of standard API for knowledge services
+makes it especially hard to switch between KRS to compare them. Also, because
+service robots are design to act in rich, dynamic environments, possibly with
+humans, building repeatable experiments is challenging.
+
+Cognitive architectures and KRS in particular are consequently often
+benchmarked with synthetic datasets (which leads to other issues: how to assess
+the meaningfulness of the performance of a reasoner on an artificial
+ontology?~\cite{Bail2010}) or \emph{toy} experiments~\cite{Chong2009} that do
+not reflect real-world complexity.
+
+While evaluating the performances of knowledge representation systems
+themselves may be difficult, the cognitive performances of the robot as a whole
+can be aptly evaluated through ``classical'' tests from the psychology (like
+False-Belief experiments~\cite{Leslie2000} or the Token
+test~\cite{DiSimoni1978}).
+
+However, the performance evaluation of current tools is mostly qualitative and
+scenario-specific (``did my system do the job for my task?''), and amongst the
+surveyed KRS, only few systems proposes an evaluation of their performances
+that could be compared to other system.
 
@@
 Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
-ARMAR/Tapas & KRS & Holzapfel, Waible \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
+ARMAR/Tapas & KRS & Holzapfel, Waibel \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
 Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
@@
 authors) realized on a 7-DOF arm equipped with force feedback, cameras for blob
-tracking and speech recognition (Sphinx4).
+tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
+an in-depth analysis of the performance of GSM by the mean of a standard
+psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.
 
@@
  & Monitoring and debugging & & & & & & & & \\
- & Performances & & & & & + & & & \\
+ & Performances & & & & & + & & +++ (Token test) & & & \\
 

### 2012-04-24 krs-survey.tex
Filled section on integration with executive layers

@@
 
-\paragraph{Events}
+Conversely, the knowledge management module need a tight integration with the
+decision-making processes. As for the integration with sensori-motor
+layers, the borders of the KRS can be fuzzy and vary from one architecture to
+another: many consider symbolic task planning as an integral role of the KRS,
+while other have dedicated extensions for planning, some integrate learning as
+an on-the-flight process that is part of the KRS, others as an independent
+deliberative process, etc.
+
+The actual integration techniques vary also widely, from language extensions
+(like the integration of CRAM~\cite{Beetz2010} with KnowRob) and client-server
+architectures, to event-driven models (SHARY and ORO~\cite{Alami2011}). Choices
+at this level have notable consequences on the whole design of the upper
+control architecture of the robot, in particular regarding its modularity and
+the ease of addition of new components.
 
@@
 on the whole state of the robot, furthermore easily understandable by the human
-developer (the fact \concept{human1 issitting true} is easier to interpret than
+developer (the fact \stmt{human1 isSitting true} is easier to interpret than
 the suite of relative coordinates of each joints of the human skeleton, as

### 2012-04-25 krs-survey.tex
Cleaned headers
Proof-reading of the first half of the paper

@@
-%\documentclass[twoside,a4paper]{report}
 \documentclass[a4paper, twocolumn]{article}
@@
 \usepackage{graphicx}
-\usepackage{tabularx}
+
+%tables
 \usepackage{booktabs}
@@
   }
-
 \usepackage{pdflscape} %% Used for very big table
-\usepackage{moreverb}
-\usepackage[table]{xcolor}
-\usepackage{listings}
+
 \usepackage{paralist}
-\usepackage{fancyhdr}
+
 \usepackage[draft]{fixme}
@@
 
-\definecolor{codegray}{gray}{.95}
-\definecolor{palegray}{gray}{.65}
-\lstset{language=C,
-	numbers=left,
-	tabsize=2,
-	basicstyle=\scriptsize,
-	stringstyle=\textrm,
-	showstringspaces=false,
-	frame=none,
-	xleftmargin=3pt,
-	backgroundcolor=\color{codegray}}
+\graphicspath{{figs/}}
+
+\newcommand{\ie}{{\textit{i.e.\ }}}
+\newcommand{\cf}{{\textit{cf~}}}
+\newcommand{\eg}{{\textit{e.g.\ }}}
+
+% Macros related to concepts/statements typesetting
+\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}
+\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
+\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
+\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}
+\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
+\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}
 
@@
 \author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz}
-
-\graphicspath{{figs/}}
-
-\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}
-\newcommand{\ie}{{\textit{i.e.~}}}
-\newcommand{\cf}{{\textit{cf~}}}
-\newcommand{\eg}{{\textit{e.g.~}}}
-
-\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
-%\newcommand{\stmt}[1]{{\footnotesize $\langle$\stmttt#1\relax$\rangle$}}
-\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
-\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}
-
-\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
-\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}
 
@@
 features like knowledge acquisition and integration into pre-existent robotic
-architectures.
+architectures. We base the typology on an initial typical scenario of service
+robotics.
 
@@
 identify which knowledge-bound cognitive abilities are today well understood
-and implemented, and which would require more efforts.
+and implemented, and which would require more research efforts.
 
 We hope this article may bring a clearer picture of the challenges of knowledge
-representation and manipulation for the robotics research community.
+representation and manipulation for the robotic research community.
 
@@
     Central to this effort is to develop an understanding of the relationship
-    between the knowledge, the perception, and the action of \[...\] a robot. The
+    between the knowledge, the perception, and the action of [\ldots] a robot. The
     sorts of questions we want to be able to answer are
@@
 
-Those questions have in common one prerequisite to be answered: the robot must
+To be answered, those questions have in common one prerequisite: the robot must
 be able to manipulate \emph{explicitly} knowledge, and hence, need a way to
@@
 establishing a as-exhaustive-as-possible typology of representation needs for
-robotics, and then at painting the current landscape of approaches to the
-knowledge representation problem in the research community, in order to
-summarize the major trends and to identify the possible shortcomings.
+robotics in the specific context of service robotics, and then at painting the
+current landscape of approaches to the knowledge representation problem in the
+research community, in order to summarize the major trends and to identify the
+possible shortcomings.
 
@@
 robotic control.\end{inparaenum}
+\fxfatal{Do we really do it?}
 
@@
 of robotics, it is useful to make our terminology explicit.  No general
-agreement on a definition of ``knowledge'' exists. In our context, we call
+agreement on a definition of ``knowledge'' exists. For our needs, we call
 ``knowledge'' \emph{an explicit set of interrelated logical facts that are
-meaningful to the robot executive controller} (by \emph{meaningful} we mean
-that can possibly be interpreted to lead to a purposeful action).
+meaningful to the robot's executive controller}, where \emph{meaningful} means
+that it can possibly be interpreted to lead to a purposeful action).
 
@@
 Wikipedia page on the ``DIKW'' hierarchy). In this article, we will associate
-data to low-level material like raw sensor output, and information to
-uncontextualized symbolic facts.
-
-To give a example, image a human watching at a book and being tracked by a
+\emph{data} to low-level material like raw sensor output, and
+\emph{information} to uncontextualized symbolic facts.
+
+To give a example, imagine a human looking at a book while being tracked by a
 Kinect sensor: the pose of the human skeleton in the world would be the data,
-the fact \concept{looksAt(human, book)} as computed by a geometric reasoning
+the fact \concept{looksAt(entity1, entity2)} as computed by a geometric reasoning
 module would be the information, the fact \concept{looksAt(john,
@@
 Besides, we have decided to restraint the set of systems to those actually
-implemented on robots, and used in semantic-rich environments (\ie , dynamic,
+implemented on robots, and used in semantic-rich environments (\ie dynamic,
 partially unknown environments with a large range of different entities which
 may have interactions). The typical scenario that would involve such robots is
-a service robot in a human-friendly environment like a kitchen (as described in
-section~\ref{sect|scenario}).
+a service robot in a human-friendly environment like a kitchen (we will
+describe such a scenario in section~\ref{sect|scenario}).
 
@@
     \item  are able of on-line, dynamic knowledge acquisition and reasoning 
-    (\ie not a simple static database).
+    (\ie not simple static databases).
 \end{inparaenum}
@@
 references in the literature on knowledge manipulation and representation
-applied to robotic scenarii in these architectures.
+applied to robotic scenarii for these architectures.
 
@@
 To the extend of our knowledge, no previous published work attempt to build a
-typology of cognitive abilities related to knowledge and required by service
-robotics.
+typology of cognitive abilities strictly related to knowledge in the context of
+service robotics.
 
@@
 ``flexible configuration and reconfiguration'' (which require what we call
-\emph{non-monotonicity} in this article) and eventually Heintz et al. call for
-meta-knowledge and introspective capacities (``declarative specification of the
-processing functionalities''). In this paper, we focus our analysis on similar
-system with however an emphasize on the symbolic capabilities.
+\emph{non-monotonicity} in this article) and finally meta-knowledge and
+introspective capacities (``declarative specification of the processing
+functionalities''). In this paper, we focus our analysis on similar system with
+however an emphasize on the symbolic capabilities.
 
@@
 knowledge representation strategies and are relevant to our study.
-
-At a broader scope, several authors from fields that are connected to robotics
+\fxwarning{Detail Chong here}
+
+At an even broader scope, several authors from fields that are connected to robotics
 have previously listed desirable features of artificial systems aiming at
-cognitive abilities as rich as possible.
+rich cognitive abilities.
 
@@
 Roy and Reiter, in~\cite{Roy2005}, summarize what they see as the main
-challenges to be tackled by KRS: cross-modal representation systems, association of
-words with perceptual and action categories, modeling of context, figuring out
-the right granularity of models, integrating temporal modeling and planning,
-ability to match past (learned) experiences with the current interaction and
-ability to take into account the human perspective.
+challenges to be tackled by KRS: cross-modal representation systems,
+association of words with perceptual and action categories (\emph{grounding}),
+modeling of context, definition of the right granularity of models, integration
+of temporal modeling and planning, ability to match past (learned) experiences
+with the current interaction and ability to take into account the human
+perspective.
 
 Knowledge representation systems in robotics are concerned by most, if not all,
-of these points, and we shall reuse them later, in slightly reformulated ways.
+of these points, and we shall indeed mention them in our typology, in slightly
+reformulated ways.
 
@@
 
-The next section proposes a prototypical scenario in a rich semantic
-environment. We will not survey knowledge representation systems based on this
-scenario: we merely use it to ground the selection and illustrate the set
-features that define the dimensions of a
-knowledge representation system for robotics.
+The next section proposes a prototypical scenario of a robot acting in a rich
+semantic environment. We will not survey knowledge representation systems based
+on this scenario: we merely use it to ground the selection and illustrate the
+set of features that define the dimensions of a knowledge representation system
+for robotics.
 
@@
 
-Section~\ref{sect|surveyed-systems} then presents each of the systems we
+Section~\ref{sect|surveyed-systems} then presents each of the systems we have
 examined, and underlines their specificities.
 
-We finally summarize in section~\ref{sect|summary} the current approaches in
-relation to the previously identified set of features. This leads us to
-identify several new research directions that are not sufficiently addressed by
-the current state of the art.
+We finally summarize in section~\ref{sect|conclusion} the current approaches in
+order to identify several research directions that are not sufficiently
+addressed by the current state of the art.
 
@@
 In order to illustrate what we consider to be the required features of a
-knowledge representation system for robotics, we introduce in this section a
-demonstration scenario.
+knowledge representation system for robotics, we introduce in this section an
+(imaginary) demonstration scenario.
 
@@
 examine, we insist on the fact that our main purpose here is to ground into a
-real case the set of representational features we consider as desirable.  We
-will \emph{not} evaluate any of the systems we survey in this article by
-judging their applicability to this particular scenario.
-
-We entitle our scenario ``the Brownie scenario'' (Fig.~\ref{fig|scenario}).
-
-Robi and Roba are two service robots. They are considered to be able to freely
-move and pick objects, but they are not expected to have the exact same
-hardware and softwares architectures, and specially, they are not expected to
-have the same knowledge representation system. These robots cooperate with a
-human in a kitchen environment.
+concrete case the set of representational features we consider as desirable.
+
+We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
+Robi and Roba are two service robots, that can freely
+move and pick objects around (with possibly different hardware and softwares
+architectures, including different knowledge representation systems). They
+cooperate with a human in a kitchen environment.
 
 The main task of the scenario is the joint realization of a brownie, initiated
-by the human injunction ``Let's make a brownie for tonight!''.
-
-The scenario is successful if the task is achieved (the brownie is baked) and:
-
-\begin{itemize}
-
-	\item it took less time that it would have require for the human alone,
-
-	\item it didn't require a heavier cognitive involvement from the human that
-	what would have been required without the robots' help.
-
-\end{itemize}
+by the human: ``Let's make a brownie for tonight!''.
+
+The scenario is successful if the task is achieved (the brownie is baked) in a
+reasonable time (typically shorter than what it would have been required by the
+human alone).
 
@@
 We can further refine these categories: to prepare and bake a brownie, the
-robot first needs to make sense of the very term \emph{brownie}: what is it?
+robot first needs to make sense of the term \emph{brownie} itself: what is it?
 what is it used for? what is it made of? etc. We call this knowledge
 \emph{common-sense knowledge} and the robot must be able not only to represent
-it, but also to have some kind of access to it (for instance through a initial
+it, but also to have access to an initial source (for instance through a initial
 set of facts that are made available at startup, or via access to a Web-based
@@
 
-Bound to the action \emph{make}, this should lead the robot to build and
-represent a \emph{context}: we are in a scenario involving cooking. Actions
-related to cooking often take place in the kitchen, cooking requires ingredients, 
-utensils and a procedure that may be provided by a recipe, etc.
-
-This last sentence implies several other features for our knowledge
+Once bound to the action \emph{make}, this should lead the robot to build and
+represent a \emph{context}: we are in a scenario involving cooking. The context
+enables the robot to retrieve more common-sense knowledge, like that actions
+related to cooking often take place in the kitchen, cooking requires
+ingredients, utensils and a procedure that may be provided by a recipe.
+
+These last assertions imply several other features for our knowledge
 representation system: ``cooking often takes place in the kitchen'' implies
@@
 
-``cooking requires ingredients that may be provided by a recipe'' hints about a
-very common feature available in most knowledge representation systems:
-\emph{reasoning}. The robot \emph{infers} that cooking may require a recipe
-since a list of ingredients is a pre-requisite of the cooking action, and a
-recipe may provide such a list. If we omit the ``may'', this is a typical
-example of first-order logic reasoning. Many other reasoning techniques exist
-(including probabilistic ones -- ones able to deal with the ``may''), we shall
-illustrate some of them later in this scenario.
+``cooking requires ingredients'' hints about another very common feature
+available in most knowledge representation systems: \emph{reasoning}. The robot
+can \emph{infers} that cooking may require a recipe since a list of ingredients
+is a pre-requisite of the cooking action, and a recipe may provide such a list.
+If we omit the ``may'', this is a typical example of first-order logic
+reasoning. Many other reasoning techniques exist (including probabilistic ones
+-- ones able to deal with the ``may''), we shall illustrate some of them later
+in this scenario.
 
@@
 component: it may well be the result of the aggregation of several logical
-components working together}.
+components working together. We come back later on this aspect.}.
 
@@
 knowledge of the other agents: to think \emph{from the perspective} of another
-agent (this idea is related to the so-called Theory of Mind~\cite{Leslie2000}).
+agent (an idea that is related to the availability of a theory of
+mind~\cite{Leslie2000}).
 
 Ability to communicate is one important pre-requisite to collaboration.
-Communication in general~\cite{Jakobson1960} requires the addresser and the
-addressee to share a common interpretative framework (a shared common-sense
-knowledge -- or cultural background -- and a shared context). In our scenario,
-the agents are working in a kitchen. This element of context does not however
+Communication in general requires the addresser and the addressee to share a
+common interpretative framework (a shared common-sense knowledge -- or cultural
+background -- and a shared context)~\cite{Jakobson1960}. In our scenario, the
+agents are working in a kitchen. This element of context does not however
 suffice if, for example, an agent asks another agent to ``give {[him]} the
@@
 operation is essential, and may be tightly bound to the underlying knowledge
-representation system. Note that grounding usually denotes the {\it top-down}
-operation: from the symbol to the percept. It is however commonly bundled with
-the converse operation, that retrieve (or create) symbols from perception.
+representation system. Note that grounding is a bidirectional process: in
+covers the {\it top-down} operation (from the symbol to the percept) and the
+{\it bottom-up} converse (retrieval or creation of symbols from perception).
 
@@
 and of immediate ability) is thus often desirable. More general introspection
-(like the ability to tell ``who I am'' or ``what do I think of'' does not
+(like the ability to tell ``who I am'' or ``what do I think of'') does not
 appear to be necessary in our scenario. This may however be required in more
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Typology of Knowledge Representation Systems Requirements for Robotics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A Typology of Knowledge Representation Requirements for Robotics}
 \label{sect|features}
 
-We presented an imaginary cooking scenario where numerous desirable
-features for a knowledge representation system for robotics have been
-identified in no particular order.
+We now want to organize requirements from this imaginary cooking scenario where
+numerous desirable features for a knowledge representation system for robotics
+have been identified without any particular order.
 
@@
 
+As previously mentioned, we propose six main categories.
+\fxfatal{figure taxonomy}
+
 \subsection{Expressiveness: What Can be Represented?}
@@
 
-\subsubsection{Introduction: Main Logic Formalisms}
+\subsubsection{Main Logic Formalisms}
 
@@
 A large number of logic formalism do exist, we shall summarize below the most
-relevant for systems actually deployed in current robotic architectures.
+relevant ones for systems actually deployed in current robotic architectures.
 
@@
 \emph{Horn clauses} are an important subset of FOL because the satisfiability
-of a set of such clauses is a $P$-complete problem. A Horn clause is a
-disjunction of literals (a \emph{clause}) with at most one positive literal:
-$\neg p \vee \neg q \vee \cdots \vee \neg t \vee u$, which can also be
-represented as $(p \wedge q \wedge \cdots \wedge t) \rightarrow u$.  Important
-logic programming languages like Prolog are based on Horn clauses.
+of a set of such clauses is a $P$-complete problem (\ie practically tractable).
+A Horn clause is a disjunction of literals (a \emph{clause}) with at most one
+positive literal: $\neg p \lor \neg q \lor \cdots \lor \neg t \lor u$, which
+can also be represented as $(p \land q \land \cdots \land t) \rightarrow u$.
+Important logic programming languages like Prolog are based on Horn clauses.
 
@@
 extensions in second-order logic. Description logics are notable because most
-of them are known to be decidable. In description logic, axioms are build from
-\emph{concepts}, \emph{roles} (that are unary or binary predicates) and
-\emph{individuals}. The W3C OWL-DL standard is a widely-used language to
-describe domains with the description logic.
+of them are known to be decidable (but not always in a practically tractable
+manner). In description logic, axioms are build from \emph{concepts},
+\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
+W3C OWL-DL standard is a widely-used language to describe domains with the
+description logic.
 
@@
 terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
-\emph{individual},...) is well-suited to knowledge description and we may use
+\emph{individual},\ldots) is well-suited to knowledge description and we may use
 it in the remaining of this paper outside of the strict context of  Description
@@
 \emph{possibility} or \emph{necessity}, have been shown to be closely related
-to description logics. Modal logic allows to represent conveniently parallel
+to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
 possible worlds and facts like ``the robot knows \emph{that the human knows}
@@
 
-Note that most of these logic formalisms are still active research field by
+Note that most of these logic formalisms are still active research field on
 their own, and practical considerations (especially the availability of
@@
 choice of a logical formalism and a level of expressive power.
-
-\paragraph{Some examples}
-
-\emph{The robot knows that a blue bottle is laying on the table.}
-
-\emph{The robot knows that the human knows about the position of the bottle,
-but the robot does not know what the human actually know about it.}
 
@@
 table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
-a simple, concrete example of consequences of the CWA/OWA choice on reasoning.
+another example of consequences of the CWA/OWA choice on reasoning.
 
@@
 
-Domains constrained with the closed world assumption lead to more tractable
-inference problems, and allow for instance the use of logic languages like
-Prolog. Thus, several approaches exists to \emph{locally close} a domain (\cf
+The OWL language is specifically known to assume an open world.  Domains
+constrained with the closed world assumption lead to more tractable inference
+problems, and allow for instance the use of logic languages like Prolog. Thus,
+several approaches exists to \emph{locally close} a domain (\cf
 Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those).
@@
 
-As stated by Josyula and Raja, meta-cognition is composed of both
+As stated by Cox and Raja~\cite{Cox2007}, meta-cognition is composed of both
 \emph{``meta-level control of cognitive activities and the introspective
@@
 to explain a logical inconsistency in a KRS is a meta-cognitive function.
-Sloman~\cite{Sloman2011}
+Sloman~\cite{Sloman2011} \fxfatal{...Sloman?}
 
@@
 \subsubsection{Role Representations}
+
+Thematic roles
 
@@
 \fxfatal{Discuss time chronicles~\cite{Ghallab1996}}
+\fxfatal{Discuss Allen intervals}
 
@@
 
-\paragraph{The introspective robot}
+\paragraph{Introspection}
 
@@
 are currently available ?  While at a first level, these descriptions can be
-static (\eg the robot has one laser scanner and two arms), at a second level,
-the description is updated and reflect the current (and possibly past and
-future) state of the robot. Note that these description may also involve
+static (\eg the robot has one laser scanner and two arms), at more advanced
+levels, the description is updated and reflect the current (and possibly past
+and future) state of the robot. Note that these description may also involve
 geometric descriptions (a kinematic chain, the pose of a device, etc.) that may
@@
 
-Reasoners can then provide other type of inferences, like:
+Reasoners can often provide other type of inferences, like:
 
@@
 
-        \item universal, existential and cardinality constraints (including \concept{allValue}, 
-        \concept{someValue}, \concept{hasValue}),
-
-        \item property characteristics (symmetry, transitivity, etc.)
+        \item universal, existential and cardinality constraints,
+
+        \item property characteristics (inverse, symmetry, transitivity, etc.)
 
@@
 
-    \item set operations like: \par \footnotesize \concept{Color} $\equiv$ {\bf
-    unionOf}(\concept{blue}, \concept{green}, \concept{orange},
-    \concept{black}...) \normalsize
+    \item set operations like: \par \footnotesize \concept{Color} $\equiv \bigcup$ (\concept{blue}, \concept{green}, \concept{orange},
+    \concept{black}, \ldots) \normalsize
 
@@
 
-\concept{Egg}$(a) \wedge $ \concept{Egg} $(b) \wedge $
-\concept{MakeOmelette}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \wedge \lnot $
-\concept{Egg}$(b) \wedge $ \concept{Omelette}$(c)$
+\concept{Egg}$(a) \land $ \concept{Egg} $(b) \land $
+\concept{MakeOmelette}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \land \lnot $
+\concept{Egg}$(b) \land $ \concept{Omelette}$(c)$
 
@@
 For instance, the SWRL rule language, usually associated to the OWL-DL ontology
-language, do not allow non-monotonic reasoning.
+language, do not allow non-monotonic reasoning (only so-called DL-safe rules
+are allowed).
 
@@
 One important exception is the \emph{negation as failure} inference rule, as
-implemented by {\sc Prolog} for instance, that allows for non-monotonicity, but
-only \emph{within the closed world assumption}.
+implemented by {\sc Prolog} for instance, that allows for non-monotonicity
+within the closed world assumption.
 
@@
 
-\fxfatal{Bielefeld -> could not find much...}
+\fxfatal{Bielefeld -> could not find much\ldots}
 \fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}
@@
 % - two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
-% - uses directly Cyc (ie, 250 000 common sense concepts...), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
+% - uses directly Cyc (ie, 250 000 common sense concepts\ldots), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
 % Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
@@
 
-...seem to be mostly WIP...
+\ldots seem to be mostly WIP\ldots
 
@@
 \hline
-\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & ...with sensori-motor layers & & & & & & & + & + \\
- & ...with executive layers & & & & & & & + & ++ \\
+\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & \ldots with sensori-motor layers & & & & & & & + & + \\
+ & \ldots with executive layers & & & & & & & + & ++ \\
  & NLP & & +++ (external) & & & +++ & & +++ & & +++ & \\
@@
 \begin{itemize}
-	\item a priori evaluation/lazy evaluation and trade-offs (relations to events, scalability, detection of inconsistencies,...)
-	\item relation with planning (action representation, temporal representation...)
+	\item a priori evaluation/lazy evaluation and trade-offs (relations to events, scalability, detection of inconsistencies,\ldots)
+	\item relation with planning (action representation, temporal representation\ldots)
 	\item API/bindings/integration into executive layers

### 2012-04-25 krs-survey.tex
Added diagrams of the typology

@@
 \usepackage{graphicx}
+
+% typology diagrams
+\usepackage{tikz}
+\usetikzlibrary{shapes}
 
@@
 As previously mentioned, we propose six main categories.
-\fxfatal{figure taxonomy}
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, level distance = 3cm, level 1/.style={sibling distance=5mm}]
+    %[edge from parent fork east]
+    %every node/.style={fill=black!30,rounded corners},
+    %[parent anchor=east,child anchor=west,grow=east,
+    %edge from parent/.style={thick,draw}]
+    \node {Dimensions}
+        child {node {Instantiation}}
+        child {node {Integration}}
+        child {node {Acquisition}}
+        child {node {Reasoning}}
+        child {node {Representation}}
+        child {node {Expressiveness}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
 
@@
 \label{sect|expressiveness}
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
+    \node {\bf Expressiveness}
+            child {node {Meta-cognition}}
+            child {node {Uncertainty}}
+            child {node {Open/Closed world assumption}}
+            child {node {Expressive power}}
+            child {node {Logic formalism}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
 
@@
 
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
+    \node {\bf Representation}
+            child {node {Memory}}
+            child {node {Introspection}}
+            child {node {Possible Worlds}}
+            child {node {Context}}
+            child {node {Roles}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
 \subsubsection{Role Representations}
@@
 
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 4cm]
+    \node {\bf Reasoning}
+            child {node {Learning}}
+            child {node {Naive physics}}
+            child {node {Planning}}
+            child {node {Prediction}}
+            child {node {Presupposition}}
+            child {node {Non-monotonicity}}
+            child {node {Uncertainty}}
+            child {node {Lazy evaluation}}
+            child {node {Structural alteration}}
+            child {node {Standard reasoning}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
 \subsubsection{Standard reasoning techniques}
@@
 
-\subsubsection{Knowledge acquisition and modalities merging}
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
+    \node {\bf Acquisition}
+            child {node {Dynamic instantiation}}
+            child {node {Grounding}}
+            child {node {Acquisition and fusion}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+\subsubsection{Knowledge acquisition and modalities fusion}
 \label{sect|knowledge-acquisition}
@@
 
-\subsubsection{Ability to automatically create new object instances}
+\subsubsection{Dynamic instantiation}
 \label{sect|new-instances}
+Ability to automatically create new object instances
 
@@
 \label{sect|integration-robot}
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
+    \node {\bf Integration}
+            child {node {Performances}}
+            child {node {Monitoring}}
+            child {node {Natural Language}}
+            child {node {Executive layers}}
+            child {node {Sensori-motor}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
 

### 2012-05-08 krs-survey.tex
Added a paragraph on memory mechanisms

@@
 This survey focuses on this knowledge representation issue: we aims at first
-establishing a as-exhaustive-as-possible typology of representation needs for
+establishing a comprehensive typology of representational needs for
 robotics in the specific context of service robotics, and then at painting the
@@
 
+It worth emphasing that if memory is commonly associated to the process of
+forgetting facts after a variable amount of \emph{time}, it covers actually
+more mechanisms that are relevant to robotics, like selective remembering
+triggered by a specific context or reinforcement learning.
+
 Most knowledge representation systems offers some kind of memory as a pool of
@@
 \end{figure}
+
+\subsubsection{Intrinsic Motivation}
+
+\cite{Oudeyer2007} -> page 2 gives a detailled account of psychological grounds.
+Curiosity, Pro-active behaviour...
 
@@
 effectively representing actions or temporal information.\fxfatal{What about reasoning? can they retrieve for example 'all proxies for colorful objects'?}
+
+In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
+the context of spatial representations.
 
@@
  & Grounding & +++ \par (semantic maps) & ++ (external amodal model) & & & & & ++ & ++ \\
- & instantiation & & & & & & & & \\
+ & Intrinsic Motivation & & & & ++ & & & & \\
+ & Dynamic instantiation & & & & & & & & \\
 \hline

### 2012-05-08 krs-survey.tex
Added numberniing to the typology

@@
     \node {Dimensions}
-        child {node {Instantiation}}
-        child {node {Integration}}
-        child {node {Acquisition}}
-        child {node {Reasoning}}
-        child {node {Representation}}
-        child {node {Expressiveness}};
+        child {node {{\bf F}. Instantiation}}
+        child {node {{\bf E}. Integration}}
+        child {node {{\bf D}. Acquisition}}
+        child {node {{\bf C}. Reasoning}}
+        child {node {{\bf B}. Representation}}
+        child {node {{\bf A}. Expressiveness}};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
-    \node {\bf Expressiveness}
-            child {node {Meta-cognition}}
-            child {node {Uncertainty}}
-            child {node {Open/Closed world assumption}}
-            child {node {Expressive power}}
-            child {node {Logic formalism}};
+    \node {\bf A. Expressiveness}
+            child {node {{\bf A.5}. Meta-cognition}}
+            child {node {{\bf A.4}. Uncertainty}}
+            child {node {{\bf A.3}. OWA/CWA}}
+            child {node {{\bf A.2}. Expressive power}}
+            child {node {{\bf A.1}. Logic formalism}};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf Representation}
-            child {node {Memory}}
-            child {node {Introspection}}
-            child {node {Possible Worlds}}
-            child {node {Context}}
-            child {node {Roles}};
+    \node {\bf B. Representation}
+            child {node {{\bf B.5}. Memory}}
+            child {node {{\bf B.4}. Introspection}}
+            child {node {{\bf B.3}. Possible Worlds}}
+            child {node {{\bf B.2}. Context}}
+            child {node {{\bf B.1}. Roles}};
 \end{tikzpicture}
@@
 the robot give an object to the human, this state could be represented with the
-triplets \setstmt{robot performs action1, action1 isA Give}.
+triples \setstmt{robot performs action1, action1 isA Give}.
 
@@
 \begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 4cm]
-    \node {\bf Reasoning}
-            child {node {Learning}}
-            child {node {Naive physics}}
-            child {node {Planning}}
-            child {node {Prediction}}
-            child {node {Presupposition}}
-            child {node {Non-monotonicity}}
-            child {node {Uncertainty}}
-            child {node {Lazy evaluation}}
-            child {node {Structural alteration}}
-            child {node {Standard reasoning}};
+    \node {\bf C. Reasoning}
+            child {node {{\bf C.10}. Learning}}
+            child {node {{\bf C.9}. Naive physics}}
+            child {node {{\bf C.8}. Planning}}
+            child {node {{\bf C.7}. Prediction}}
+            child {node {{\bf C.6}. Presupposition}}
+            child {node {{\bf C.5}. Non-monotonicity}}
+            child {node {{\bf C.4}. Uncertainty}}
+            child {node {{\bf C.3}. Lazy evaluation}}
+            child {node {{\bf C.2}. Structural alteration}}
+            child {node {{\bf C.1}. Standard reasoning}};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf Acquisition}
-            child {node {Dynamic instantiation}}
-            child {node {Grounding}}
-            child {node {Acquisition and fusion}};
+    \node {\bf D. Acquisition}
+            child {node {{\bf D.4} Dynamic instantiation}}
+            child {node {{\bf D.3} Motivation}}
+            child {node {{\bf D.2} Grounding}}
+            child {node {{\bf D.1} Acquisition and fusion}};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf Integration}
-            child {node {Performances}}
-            child {node {Monitoring}}
-            child {node {Natural Language}}
-            child {node {Executive layers}}
-            child {node {Sensori-motor}};
+    \node {\bf E. Integration}
+            child {node {{\bf D.5} Performances}}
+            child {node {{\bf D.4} Monitoring}}
+            child {node {{\bf D.3} Natural Language}}
+            child {node {{\bf D.2} Executive layers}}
+            child {node {{\bf D.1} Sensori-motor}};
 \end{tikzpicture}
@@
 \subsection{Knowledge instantiation}
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
+    \node {\bf F. Knowledge instantiation}
+            child {node {{\bf F.2} Alignement with Standards}}
+            child {node {{\bf F.1} Self-Knowledge}}
+ \end{tikzpicture}
+\end{center}
+\end{scriptsize}
 

### 2012-05-08 krs-survey.tex
Added section on self knowledge

@@
 \end{itemize}
+
+\subsubsection{Self-Knowledge}
+
+``Self-knowledge'' is the term used in psychology to describe the knowledge
+that an individual has, acquires or infers about itself through its
+experiences. It answers the question ``What am I like?''.
+
+For a robot to have at its disposal self-knowledge is closely link to the
+introspective capability that we have previously mentioned
+(section~\ref{sect|introspection})
+
+\subsubsection{Alignment with Standard Upper-Ontologies}
 

### 2012-05-09 krs-survey.tex
Enriched section on knowledge instantiation

@@
 the meaningfulness of the performance of a reasoner on an artificial
-ontology?~\cite{Bail2010}) or \emph{toy} experiments~\cite{Chong2009} that do
+ontology?~\cite{Bail2010}) or toy experiments~\cite{Chong2009} that do
 not reflect real-world complexity.
@@
 themselves may be difficult, the cognitive performances of the robot as a whole
-can be aptly evaluated through ``classical'' tests from the psychology (like
+can be also evaluated through ``classical'' tests from the psychology (like
 False-Belief experiments~\cite{Leslie2000} or the Token
@@
     \node {\bf F. Knowledge instantiation}
-            child {node {{\bf F.2} Alignement with Standards}}
-            child {node {{\bf F.1} Self-Knowledge}}
+            child {node {{\bf F.4} Alignement with Standards}}
+            child {node {{\bf F.3} Self-Knowledge}}
+            child {node {{\bf F.2} Metrics}}
+            child {node {{\bf F.1} Design Strategy}};
  \end{tikzpicture}
@@
 
-How much knowledge is available? Which content? How big is the knowledge base?
+\subsubsection{Design Strategy}
+
+Ontology engineering has grown into a research field of its own\fxwarning{Give
+some idea about it? wikipedia?}.
+
+Two main approaches are common in the cognitive robotic community:
+\emph{top-down} design and \emph{bottom-up}.
+
+Several strategies ... Designing, implementing and feeding an ontological model of the beliefs of the robot
 
 \begin{itemize}
-	\item  Which underlying knowledge (\emph{common-sense}, \emph{upper knowledge}\ldots{})
-	\begin{itemize}
-		\item  top-down approach?
+    \item Dynamic, run-time instantiation
+    \item Static, a priori knowledge
+    \item Fetch from external resources (TBox and/or ABox)
 \end{itemize}
 
-\end{itemize}
+\subsubsection{Metrics}
+
+Qualitative and quantitative metrics give an oversight of the size and
+expressive complexity of ontologies.
+
+Table~\ref{table|upper_onto_metrics} gives such metrics for four major
+upper-ontologies commonly used in the semantic Web community.
+
+\fxwarning{Add metrics for SUMO, OpenCyc, ResearchCyc}
+
+\begin{table}
+\begin{center}
+
+\begin{tabular}{llllll}
+\toprule
+{\bf Project} & {\bf Expressiveness} & {\bf Axioms} & {\bf Classes} & {\bf Properties} & {\bf Instances} \\
+\midrule
+OpenCyc \\
+ResearchCyc \\
+SUMO \\
+DBPedia \\
+\bottomrule
+
+\end{tabular}
+\end{center}
+\caption{Selection of metrics for the major upper-ontologies.}
+\label{table|upper_onto_metrics}
+\end{table}
 
@@
 
-\subsubsection{Alignment with Standard Upper-Ontologies}
+\subsubsection{Alignment with Standard Upper-ontologies}
 
@@
 
+\subsubsection{Assessing performances}
+
+The systems we have presented share the property to have real implementations
+on real, physical robots. This simple fact allows to first assert that these
+tools are at least viable solutions to the question: how to endow a (real)
+robot with explicit knowledge manipulation capabilities?
+
+From a theoritical point of view, this work hopefully brings elements of
+comparison and evaluation of the existing systems.
+
+However, this do not suffice to assess and comment the suitability of such
+systems to real-world tasks. As hinted in section~\ref{sect|performances}, we
+think that first, the practical evaluation of knowledge representation systems
+for robots can not take place in isolation (links with perceptual and
+decisional layers are too important), then we should evaluate the robot
+behaviour at the scale of the whole system, basing experiments on established
+protocols, borrowed from other disciplines from cognitive sciences. We have
+already mentioned classical experiments like False-Beliefs in psychology or the
+Token test in linguistic. A systematic review of standard cognitive tests
+applicable to service robots is still to be made.
+
 \subsection{Towards a new design}

### 2012-05-10 krs-survey.tex
Shortcomings:Added assessment of performances

@@
     \node {\bf F. Knowledge instantiation}
-            child {node {{\bf F.4} Alignement with Standards}}
-            child {node {{\bf F.3} Self-Knowledge}}
-            child {node {{\bf F.2} Metrics}}
-            child {node {{\bf F.1} Design Strategy}};
+            child {node {{\bf F.2} Alignement with Standards}}
+            child {node {{\bf F.1} Self-Knowledge}}
  \end{tikzpicture}
@@
 
-\subsubsection{Design Strategy}
-
-Ontology engineering has grown into a research field of its own\fxwarning{Give
-some idea about it? wikipedia?}.
-
-Two main approaches are common in the cognitive robotic community:
-\emph{top-down} design and \emph{bottom-up}.
-
-Several strategies ... Designing, implementing and feeding an ontological model of the beliefs of the robot
+How much knowledge is available? Which content? How big is the knowledge base?
 
 \begin{itemize}
-    \item Dynamic, run-time instantiation
-    \item Static, a priori knowledge
-    \item Fetch from external resources (TBox and/or ABox)
+	\item  Which underlying knowledge (\emph{common-sense}, \emph{upper knowledge}\ldots{})
+	\begin{itemize}
+		\item  top-down approach?
 	\end{itemize}
 
-\subsubsection{Metrics}
-
-Qualitative and quantitative metrics give an oversight of the size and
-expressive complexity of ontologies.
-
-Table~\ref{table|upper_onto_metrics} gives such metrics for four major
-upper-ontologies commonly used in the semantic Web community.
-
-\fxwarning{Add metrics for SUMO, OpenCyc, ResearchCyc}
-
-\begin{table}
-\begin{center}
-
-\begin{tabular}{llllll}
-\toprule
-{\bf Project} & {\bf Expressiveness} & {\bf Axioms} & {\bf Classes} & {\bf Properties} & {\bf Instances} \\
-\midrule
-OpenCyc \\
-ResearchCyc \\
-SUMO \\
-DBPedia \\
-\bottomrule
-
-\end{tabular}
-\end{center}
-\caption{Selection of metrics for the major upper-ontologies.}
-\label{table|upper_onto_metrics}
-\end{table}
+\end{itemize}
 
@@
 
-\subsubsection{Alignment with Standard Upper-ontologies}
+\subsubsection{Alignment with Standard Upper-Ontologies}
 

### 2012-05-10 krs-survey.tex
Updated biblio + small change in knowledge instantiation

--- 2012-05-09 krs-survey.tex
@@
 Table~\ref{table|upper_onto_metrics} gives such metrics for four major
-upper-ontologies commonly used in the semantic Web community.
-
-\fxwarning{Add metrics for SUMO, OpenCyc, ResearchCyc}
+upper-ontologies commonly used in the semantic Web community (taken
+from~\cite{Mascardi2007} and the projects' respective websites).
 
@@
 \toprule
-{\bf Project} & {\bf Expressiveness} & {\bf Axioms} & {\bf Classes} & {\bf Properties} & {\bf Instances} \\
+{\bf Project} & {\bf Terms} & {\bf Assertions} \\
 \midrule
-OpenCyc \\
-ResearchCyc \\
-SUMO \\
-DBPedia \\
+Cyc & > 300 000 & > 3 000 000 & \\
+YAGO & > 10 000 000 & > 120 000 000 \\
+SUMO & 20 000 & 60 000 \\
+DBPedia (for English) & 1 840 000 & 385 000 000\\
+OpenMind Common Sense (for English) & & 1 000 000\\
 \bottomrule

### 2012-05-10 krs-survey.tex
Added subsections to 'Kn Acq and fusion' + filled part on Linked Data

@@
 ease of such operations.
+
+Let review shortly the three sub-categories of knowledge acquisition.
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
+    \node {\bf D.1 Acquisition and fusion}
+            child {node {{\bf D.1.3} Linked Resources}}
+            child {node {{\bf D.1.2} Interaction}}
+            child {node {{\bf D.1.1} Sensing}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+\paragraph{Sensing}
+
+\paragraph{Interaction}
+
+\paragraph{Linked Knowledge Resources}
+
+Robots, and in particular service robots, can access (with sometimes some
+security-related constraints) the World Wide Web and remote knowledge stores.
+
+The current shift towards the Semantic Web (\ie structured, annotated data that
+are easily machine-processable) makes increasingly easy to have robots to reuse
+autonomously this knowledge. The DBPedia project, for instance, illustrates
+well the tendency: it provides an automatically generated RDF version of the
+Wikipedia encyclopedia.
+
+In its current state, the relevance of the DBPedia project in our context is
+however limited: the triples that are extracted are mostly ``mechanical'' (like
+the categories of a term, or factual informations extracted from Wikipedia's
+InfoBox), and the vast majority of the knowledge actually contained in the
+encyclopedia pages remains out of reach of automated parsers. Efforts like {\sc
+ReVerb}~\cite{Fader2011} work however at automatic extraction of triples by
+examining subject-verb-object relations in plain text.
+
+Another notable project that seeks at providing large amount of
+machine-friendly common-sense knowledge is the MIT's OpenMind
+project~\cite{Singh2002}. The project is designed to let the general public
+easily add common-sense statements in semi-controlled natural language, that is
+then processed to a publicly available ontology.
+
+An other approach that is easier at short-term to let robots remotely access
+knowledge repositories consists is building robot-specific shared repositories,
+with declarative and/or procedural (\ie, plan library) knowledge. The
+RoboEarth~\cite{Waibel2011} project is an example of such an effort.
 

### 2012-05-10 krs-survey.tex
Filled section on task planning

@@
 
-\subsubsection{Planning}
+\subsubsection{Task Planning}
 \label{sect|planning}
@@
 Making decision based on prediction
+
+Symbolic task planning is the ability for a robot to select a sequence of
+actions in order to reach a given final state. This capability is closely
+related to the previously mentioned projection and legality tasks: for a robot
+to plan, it must be able to build hypothetical states of the world that would
+follow from the successive application of actions (prediction task), and at
+each step, select possible, legal actions based on their pre-conditions
+(legality task). As already mentioned, these tasks are highly non-monotonic.
+
+Symbolic task planning in general is a large research
+field\cite{Russell2009planning}. The so-called Classical Planning Problem,
+first, is characterised by a unique known initial state, durationless
+deterministic actions which can be taken only one at a time, and a single
+agent. STRIPS and PDDL are amongst the commonly used languages for representing
+such planning problems.
+
+Planning with nondeterministic durationless actions with probabilities can be
+represented as discrete-time \emph{Markov decision processes} (MDP), and when
+full observability is replaced by partial observability, we deal with
+\emph{partially observable Markov decision process} (POMDP).
+
+\emph{Hierarchical Task Networks} (HTN) are another common formalism for
+planning problems, where an initial set of tasks (\emph{High Level Tasks}, HLA)
+is decomposed into either primitive actions or a new set of subtasks.
+
+From the observation that the core reasoning techniques (back and forward
+chaining) are shared between planners and reasoners used in knowledge
+representation systems, task planning can be considered within the KRS.
+
+It is however common to rely on external components dedicated to symbolic task
+planning (in particular because of the non-monotonicity requirements) with a
+tight link to the KRS for domain retreival and/or resulting plan storage.
+
+%An extensive report on task modeling in OWL ontologies in appendix
 

### 2012-05-11 krs-survey.tex
Improved section on performance evaluation
Introduced several concepts and notation (relaxation time,
scalability...)

@@
 
-Benchmarks of symbolic systems for robots are notoriously hard to conduct for
+Benchmarks of symbolic systems for robots are hard to conduct for
 several reasons: identifying good metrics for robotic experiments in general is
 difficult because of the complex interactions between tenth of modules running
-in parallel. Isolating the role of one specific component is thus hard.
-Furthermore, knowledge representation systems are usually tightly connected to
-numerous other modules, and the lack of standard API for knowledge services
-makes it especially hard to switch between KRS to compare them. Also, because
+in parallel, and isolating one specific component is difficult.
+Also, knowledge representation systems are often tightly coupled to
+the other modules, and the lack of standard API for knowledge services
+makes it especially hard to switch between KRS to compare them. Finally, because
 service robots are design to act in rich, dynamic environments, possibly with
@@
 
-Cognitive architectures and KRS in particular are consequently often
-benchmarked with synthetic datasets (which leads to other issues: how to assess
-the meaningfulness of the performance of a reasoner on an artificial
-ontology?~\cite{Bail2010}) or toy experiments~\cite{Chong2009} that do
-not reflect real-world complexity.
-
-While evaluating the performances of knowledge representation systems
-themselves may be difficult, the cognitive performances of the robot as a whole
-can be also evaluated through ``classical'' tests from the psychology (like
-False-Belief experiments~\cite{Leslie2000} or the Token
-test~\cite{DiSimoni1978}).
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
+    \node {\bf D.5 Performance Evaluation}
+            child {node {{\bf D.5.2} Cognitive Performances}}
+            child {node {{\bf D.5.1} Raw Performances}};
+ \end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+\paragraph{Raw Performances} The \emph{raw performance} is evaluated on
+quantitative benchmarks. The main metric is the \emph{scalability}
+$\mathcal{S}^{KRS}$ of the system with the size of the knowledge base
+$\sigma=|\Delta^{\mathcal{I}}|$  (in term of atoms or statements).
+
+We call \emph{relaxation time} $\mathcal{R}^{\mathcal{M}}$ the (averaged) time
+required by the system after a model modification of type $\mathcal{M}$ before
+being available for further interaction, and \emph{query time}
+$\mathcal{Q}^{\mathcal{M}'}$ the (averaged) time to execute a query of
+complexity $\mathcal{M}'$ on the KRS. The type $\mathcal{M}$ of model
+alteration is either an ABox modification (addition/removal of an instance) or
+a TBox alteration (addition/removal of a class, a class restriction or a rule).
+Query complexity $\mathcal{M}'$ is...\fxwarning{find references}.
+
+\emph{Temporal scalability} is defined in term of the nature of the function
+$f^{\mathcal{M}, \mathcal{M}'}(\sigma) = \mathcal{R}^{\mathcal{M}}(\sigma) +
+\mathcal{Q}^{\mathcal{M}'}(\sigma)$ (\ie the relation of the relaxation time
+and query time to the knowledge base size). \emph{Space scalability} is the
+relation of memory consumption to the size. The scalability is tightly coupled
+to the expressiveness of the underlying knowledge model, which need to be known
+for the scalability measurement to be meaningful.
+
+Note that robotic applications may have a knowledge base size relatively small
+compared to the complexity of its TBox. In that case the relaxation and query
+times for a constant, average knowledge size $\sigma$ may be more relevant as a
+metric for KRS raw performance.
+
+Because of coupling and repeatability issues we have mentioned, raw
+performances of KRS are often benchmarked with synthetic datasets (which leads
+to issues: how to assess the meaningfulness of the performance of a reasoner on
+an artificial ontology?~\cite{Bail2010}) or toy experiments~\cite{Chong2009}
+that do not always model the whole complexity of real-world application.
+
+\paragraph{Cognitive Performances} While evaluating the raw performances of
+knowledge representation systems in a relevant manner may be difficult, the cognitive
+performances of the robot as a whole can be also evaluated through
+``classical'' tests from the psychology (like False-Belief
+experiments~\cite{Leslie2000} or the Token test~\cite{DiSimoni1978}).
 

### 2012-05-11 krs-survey.tex
Better use of tikz to style taxonomy nodes

@@
 \begin{document}
+
+\tikzstyle {taxonomy} = [grow=right, level distance = 3cm, sibling distance=7mm]
+\tikzstyle {completetaxonomy} = [grow=right]
+\tikzstyle {taxon} = [text width=3cm, text centered,inner sep=0pt, outer sep=2pt] %, draw=blue!50,thick]
 
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, level distance = 3cm, level 1/.style={sibling distance=5mm}]
+\begin{tikzpicture}[taxonomy]
     %[edge from parent fork east]
@@
     %edge from parent/.style={thick,draw}]
-    \node {Dimensions}
-        child {node {{\bf F}. Instantiation}}
-        child {node {{\bf E}. Integration}}
-        child {node {{\bf D}. Acquisition}}
-        child {node {{\bf C}. Reasoning}}
-        child {node {{\bf B}. Representation}}
-        child {node {{\bf A}. Expressiveness}};
+    \node [taxon] {Dimensions}
+        child {node [taxon] {{\bf F}. Instantiation}}
+        child {node [taxon] {{\bf E}. Integration}}
+        child {node [taxon] {{\bf D}. Acquisition}}
+        child {node [taxon] {{\bf C}. Reasoning}}
+        child {node [taxon] {{\bf B}. Representation}}
+        child {node [taxon] {{\bf A}. Expressiveness}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
-    \node {\bf A. Expressiveness}
-            child {node {{\bf A.5}. Meta-cognition}}
-            child {node {{\bf A.4}. Uncertainty}}
-            child {node {{\bf A.3}. OWA/CWA}}
-            child {node {{\bf A.2}. Expressive power}}
-            child {node {{\bf A.1}. Logic formalism}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf A. Expressiveness}
+            child {node [taxon] {{\bf A.5}. Meta-cognition}}
+            child {node [taxon] {{\bf A.4}. Uncertainty}}
+            child {node [taxon] {{\bf A.3}. OWA/CWA}}
+            child {node [taxon] {{\bf A.2}. Expressive power}}
+            child {node [taxon] {{\bf A.1}. Logic formalism}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf B. Representation}
-            child {node {{\bf B.5}. Memory}}
-            child {node {{\bf B.4}. Introspection}}
-            child {node {{\bf B.3}. Possible Worlds}}
-            child {node {{\bf B.2}. Context}}
-            child {node {{\bf B.1}. Roles}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf B. Representation}
+            child {node [taxon] {{\bf B.5}. Memory}}
+            child {node [taxon] {{\bf B.4}. Introspection}}
+            child {node [taxon] {{\bf B.3}. Possible Worlds}}
+            child {node [taxon] {{\bf B.2}. Context}}
+            child {node [taxon] {{\bf B.1}. Roles}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 4cm]
-    \node {\bf C. Reasoning}
-            child {node {{\bf C.10}. Learning}}
-            child {node {{\bf C.9}. Naive physics}}
-            child {node {{\bf C.8}. Planning}}
-            child {node {{\bf C.7}. Prediction}}
-            child {node {{\bf C.6}. Presupposition}}
-            child {node {{\bf C.5}. Non-monotonicity}}
-            child {node {{\bf C.4}. Uncertainty}}
-            child {node {{\bf C.3}. Lazy evaluation}}
-            child {node {{\bf C.2}. Structural alteration}}
-            child {node {{\bf C.1}. Standard reasoning}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf C. Reasoning}
+            child {node [taxon] {{\bf C.10}. Learning}}
+            child {node [taxon] {{\bf C.9}. Naive physics}}
+            child {node [taxon] {{\bf C.8}. Planning}}
+            child {node [taxon] {{\bf C.7}. Prediction}}
+            child {node [taxon] {{\bf C.6}. Presupposition}}
+            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
+            child {node [taxon] {{\bf C.4}. Uncertainty}}
+            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
+            child {node [taxon] {{\bf C.2}. Structural alteration}}
+            child {node [taxon] {{\bf C.1}. Standard reasoning}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf D. Acquisition}
-            child {node {{\bf D.4} Dynamic instantiation}}
-            child {node {{\bf D.3} Motivation}}
-            child {node {{\bf D.2} Grounding}}
-            child {node {{\bf D.1} Acquisition and fusion}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf D. Acquisition}
+            child {node [taxon] {{\bf D.4} Dynamic instantiation}}
+            child {node [taxon] {{\bf D.3} Motivation}}
+            child {node [taxon] {{\bf D.2} Grounding}}
+            child {node [taxon] {{\bf D.1} Acquisition and fusion}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf D.1 Acquisition and fusion}
-            child {node {{\bf D.1.3} Linked Resources}}
-            child {node {{\bf D.1.2} Interaction}}
-            child {node {{\bf D.1.1} Sensing}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf D.1 Acquisition and fusion}
+            child {node [taxon] {{\bf D.1.3} Linked Resources}}
+            child {node [taxon] {{\bf D.1.2} Interaction}}
+            child {node [taxon] {{\bf D.1.1} Sensing}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3cm]
-    \node {\bf E. Integration}
-            child {node {{\bf D.5} Performances}}
-            child {node {{\bf D.4} Monitoring}}
-            child {node {{\bf D.3} Natural Language}}
-            child {node {{\bf D.2} Executive layers}}
-            child {node {{\bf D.1} Sensori-motor}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf E. Integration}
+            child {node [taxon] {{\bf D.4} Performances}}
+            child {node [taxon] {{\bf D.3} Monitoring}}
+            child {node [taxon] {{\bf D.2} Executive layers}}
+            child {node [taxon] {{\bf D.1} Sensori-motor}};
 \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
-    \node {\bf D.5 Performance Evaluation}
-            child {node {{\bf D.5.2} Cognitive Performances}}
-            child {node {{\bf D.5.1} Raw Performances}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf D.5 Performance \\Evaluation}
+            child {node [taxon] {{\bf D.5.2} Cognitive Performances}}
+            child {node [taxon] {{\bf D.5.1} Raw Performances}};
  \end{tikzpicture}
@@
 \begin{center}
-\begin{tikzpicture}[grow=right, sibling distance=5mm, level distance = 3.5cm]
-    \node {\bf F. Knowledge instantiation}
-            child {node {{\bf F.4} Alignement with Standards}}
-            child {node {{\bf F.3} Self-Knowledge}}
-            child {node {{\bf F.2} Metrics}}
-            child {node {{\bf F.1} Design Strategy}};
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf F. Knowledge instantiation}
+            child {node [taxon] {{\bf F.4} Alignement with Standards}}
+            child {node [taxon] {{\bf F.3} Self-Knowledge}}
+            child {node [taxon] {{\bf F.2} Metrics}}
+            child {node [taxon] {{\bf F.1} Design Strategy}};
  \end{tikzpicture}

### 2012-05-11 krs-survey.tex
Refined and completed section on interaction

@@
 
-In the survey inclusion criteria, we have insisted on only considering robotic
-systems that acquire knowledge by themselves, during their runtime.
-
-\emph{Acquiring knowledge} means in our context building new statements
-(usually as new logical facts) connected to our existing knowledge from
-external sources of information. We consider mainly three of them:
-proprioceptive/exteroceptive sensors, interaction with other agents, humans or
-robots, and remote databases. This process has generally at least two steps:
-the information acquisition by itself, and the \emph{transformation} of the
-information into knowledge, \emph{aligned} with the robot existing model
-(following our terminology for information and knowledge, as discussed in the
-introduction).
-
-Knowledge acquisition is generally not done directly in the knowledge
-representation system. On the contrary, several, if not numerous, external
-components are usually required to convert percepts into symbolic facts and to
-ground them.
+In the survey inclusion criteria, we have chosen to only consider robotic
+systems that acquire knowledge by themselves, at runtime.
+
+In our context, \emph{acquiring knowledge} means to build new logical
+statements from data sources and to anchor them into the existing knowledge. We
+consider three possible sources of data: proprioceptive/exteroceptive sensing,
+interaction with other agents, humans or robots, and remote knowledge bases.
+The acquisition process has generally two steps: the information acquisition by
+itself, and the \emph{transformation} of the information into knowledge,
+\emph{aligned} with the robot existing model (following our terminology for
+information and knowledge, as discussed in the introduction).
+
+It must be observed that knowledge acquisition is generally not done directly
+in the knowledge representation system. External components (often aggregated
+into knowledge acquisition \emph{pipelines}) are usually required to convert
+percepts into symbolic facts and to ground them.
 
@@
 
+From the point of view of knowledge representation, the sensing capability can
+be split into proprioceptive sensing (\ie, sensing of the robot own internal
+state) and exteroceptive sensing (sensing of the robot environment). The
+(physical) introspection capabilities of the robot relies on the former.
+
+\fxwarning{What to say here?}
+
 \paragraph{Interaction}
+
+Interaction with other intelligent agents (humans or robots) is another
+important source of knowledge acquisition. It relies obviously on some form of
+sensing (from speech recognition to gesture recognition) but we distinguish it
+from the previous section because \emph{interaction} implies a form of
+communication. Communication is associated to specific functions
+(figure~\ref{fig|communication-model}), and in particular, it implies a shared
+\emph{context} (usually implicit) between interactors.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{jakobson_communication_model.pdf}
+
+    \caption{The Jakobson~\ref{Jakobson1960} communication model. In italics,
+    the associated \emph{communication functions}.}
+
+    \label{fig|communication-model}
+\end{figure}
+
+We distinguish between two main interaction channels: verbal and deictic.
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf D.1.2 Interaction}
+            child {node [taxon] {{\bf D.1.2.2} Deictic Interaction}}
+            child {node [taxon] {{\bf D.1.2.1} Verbal Interaction}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+\label{sect|nlp}
+
+The field of verbal interaction processing for robots spans from pattern-based,
+constrainted sentences recognition to natural, bidirectional, unconstrained
+verbal communication. Kruijff et al. provides in~\cite{Kruijff2010} an
+up-to-date survey of literature on situated human-robot dialogue, focusing on
+formal representation systems, bi-directionality of the interaction and context
+building.
+
+Natural Language Processing (NLP) is a large research field by itself, and
+while the robotic community may be lagging behind on many theoretical aspects,
+it brings one important aspect: the embodiment. Because the interactors, both
+the robot and the human, are establishing a communication within a shared
+physical context, the verbal communication channel is complemented by deictic
+channels, back channels and possibly shared physical experiences: a human can
+show something to a robot, saying ``Give me this''. This is simply not possible
+for a virtual agent.
+
+Several of the knowledge representation systems we present have developed
+specific built-in mechanisms or extensions to parse, ground and possibly
+rebuild natural language.
+
+Deictic (used in the literal meaning of ``display, demonstration, reference'')
+interaction is also an established field of research in human-robot
+interaction. Common deictic forms of communication include attentional focus
+(via face and gaze tracking) and joint attention, pointing, emotional
+expressions (based on face expressions, postures, emotional gestures).
+
+Like other knowledge acquisition modalities, the recognition and interpretation
+of deictic communication is rarely directly included in a KRS, but, as
+previously mentioned, the symbolic representation of such communication acts is
+relevant and important to achieve successful multi-modal interactions.
 
@@
 the ease of addition of new components.
-
-\subsubsection{Language Processing}
-\label{sect|nlp}
-
-The scope of this survey is service robotic, which suggest human-robot
-interaction. While numerous modalities of communication between humans and
-robots do exist and have been extensively studied, the most challenging of
-these interactions is natural, bidirectional, unconstrained verbal
-communication. Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date
-survey of literature on situated human-robot dialogue, focusing on formal
-representation systems, bi-directionality of the interaction and context
-building.
-
-Natural Language Processing (NLP) is a large research field by itself, and
-while the robotic community may be lagging behind on many theoretical aspects,
-it brings one important aspect: the embodiment. Because the interactors, both
-the robot and the human, are establishing a communication within a shared
-physical context, the verbal communication channel is complemented by deictic
-channels, back channels and possibly shared physical experiences: a human can
-show something to a robot, saying ``Give me this''. This is simply not possible
-for a virtual agent.
-
-Several of the knowledge representation systems we present have developed
-specific built-in mechanisms or extensions to parse, ground and possibly
-rebuild natural language.
 

### 2012-05-11 krs-survey.tex
Filled section on TBox alteration

@@
 
-In systems
-
-All systems allow to modify the ABox, not always possible to alter the TBox
+The content of a knowledge base is often conveniently divided into a
+\emph{structural part} that defines the conceptualisation of a domain in term
+of vocabulary and relations within the concepts, and an \emph{instantiation} of
+this structure into concrete entities.
+
+The terms TBox and ABox are commonly found to describe this two different types of
+statements in ontologies. TBox statements describe a system in terms of
+controlled vocabularies, for example, a set of classes and properties. ABox are
+TBox-compliant statements about that vocabulary.
+
+All knowledge representation systems allow to modify the ABox, which can be
+considered as the dynamic part of the knowledge base. It is however not always
+possible to alter the TBox because not all reasoning systems can dynamically
+take into account structural changes in the knowledge base.
+
+Example of TBox alteration include modification of the taxonomy (like addition
+or retractation of a \concept{subClassOf} relation), changes to the asserted
+domain or range of a predicate, addition or retractation of rules.
+
+Supporting TBox alteration has notable consequences on the learning
+capabilities of the system: teaching general facts (\ie, facts at the level of
+whole categories) like ``cars go on roads'' to a robot requires an alteration
+of the TBox.
 

### 2012-05-11 krs-survey.tex
Filled section on lazy evaluation

@@
 \label{sect|lazy-evaluation}
+
+\emph{Lazy evaluation} describes the ability for a KRS to delay active
+knowledge acquisition or resoning operations until the value is actually
+needed.
+
+For instance, a system that computes the symbolic relative placement of two
+objects only when this fact is required to answer a query would be said to
+adopt a lazy evaluation strategy, whereas a system that computes {\it a priori}
+such relations (and by consequence, carries out such computation for possibly
+all known objects) would be said to use a \emph{strict evaluation} policy.
+
+Lazy evaluation has an immediate impact on efficiency and scalability of the
+system, and some problems may even be only tractable with a lazy evaluation
+strategy (the relative placements of object is an example of combinatory
+explosion in strict evaluation approaches, although this issue is mitigated in
+real use-cases by heuristics that select a subset of objects to evaluate).
+
+One downside is that the knowledge base never contains explicitly the complete
+set of beliefs of the robot. This limits for instance the ability for the robot
+to react to logical conditions that involves fact that are lazily evaluated
+(``trigger a callback when object A is behind object B'' could be an example of
+a event triggered by a logical condition).
 

### 2012-05-11 krs-survey.tex
New proposed title + added Rachid to co-authors
Reorganized Kn Acq, introduced it

@@
 
-\title{Requirements and Current State of Knowledge Representation Systems for Service Robotics}
-\author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz}
+\title{Knowledge Representation for Service Robots: Supply and Demand}
+\author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz, Rachid Alami}
 
@@
 \paragraph{Linked Knowledge Resources}
+\label{sect|lod}
 
@@
     \node [taxon] {\bf F. Knowledge instantiation}
-            child {node [taxon] {{\bf F.4} Alignement with Standards}}
-            child {node [taxon] {{\bf F.3} Self-Knowledge}}
-            child {node [taxon] {{\bf F.2} Metrics}}
+            child {node [taxon] {{\bf F.5} Self-Knowledge}}
+            child {node [taxon] {{\bf F.4} Metrics}}
+            child {node [taxon] {{\bf F.3} Granularity}}
+            child {node [taxon] {{\bf F.2} Common-sense and Alignement}}
             child {node [taxon] {{\bf F.1} Design Strategy}};
@@
 
+This last branch of our taxonomy looks at the actual \emph{content} of the
+knowledge base: the knowledge instantiation. Here, \emph{instantiation} does
+not only refer to the instantiation of the knowledge structure (what we have
+called the ABox), but also includes the knowledge structure itself (the TBox).
+
+While we have previously mentioned features of knowledge representation system
+that enable the robot to fill its knowledge base with content and alter the
+knowledge structure, most of the systems also come with a certain amount of
+initial knowledge that often includes \emph{common-sense} knowledge (\ie facts
+that widely known to humans, and hence often implicit: ``to put a cake in the
+oven, one must first open the oven's door'').
+
+The design strategy, the choice to rely on common-sense knowledge or not, the
+reuse of standard ontologies, the quantity of {\it a priori} knowledge are many
+parameters that lead to different knowledge models.
+
 \subsubsection{Design Strategy}
 
-Ontology engineering has grown into a research field of its own\fxwarning{Give
-some idea about it? wikipedia?}.
+Ontology engineering has grown into a research field of its own, covering
+formalized methodologies for ontology creation, reuse, modularisation, etc.
+
+The creation
 
@@
 
-Several strategies ... Designing, implementing and feeding an ontological model of the beliefs of the robot
-
-\begin{itemize}
-    \item Dynamic, run-time instantiation
-    \item Static, a priori knowledge
-    \item Fetch from external resources (TBox and/or ABox)
-\end{itemize}
+\subsubsection{Common-sense and alignment with Standard Upper-ontologies}
+
+At section~\ref{sect|lod} we have presented how remote knowledge bases are a
+valuable source of knowledge, including common-sense knowledge, that robot can
+extract.
+
+\subsubsection{Granularity}
+
+The knowledge \emph{granularity} qualifies the level of details or refinement
+of the knowledge stored.
+
+...
+
+The storage of literal values is a particular case: while knowledge
+representation systems have been considered until now mostly as symbolic
+systems, they can often represent literal values (in particular, numerical
+values) as well.
+
+Depending on the representation language, literal values can be naturally
+represented and processed (common case in logic programming language) or not
+(storing numerical value, let alone matrices, in OWL is cumbersome).
 
@@
 
-Qualitative and quantitative metrics give an oversight of the size and
+Qualitative and quantitative metrics give an insight on the size and
 expressive complexity of ontologies.
@@
 (section~\ref{sect|introspection})
-
-\subsubsection{Alignment with Standard Upper-ontologies}
 

### 2012-05-12 krs-survey.tex
Filled most of the 'Role representation' section

@@
 
-This section proposes a more formal typology of desirable features for a such a
-system. For each feature, we provide a short definition along with links to
-relevant literature.
-
-Table~\ref{table|contribution-by-systems}, at the end of the article,
-summarizes all these features with the main domain of contribution of each
-surveyed KRS.
+This section proposes a more formal typology and a nomenclature of analysis
+axis for a such a system. For each feature, we provide a short definition along
+with links to relevant literature.
 
@@
 
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf B.1 Roles}
+            child {node [taxon] {{\bf B.1.3}. Actions}}
+            child {node [taxon] {{\bf B.1.2}. Time}}
+            child {node [taxon] {{\bf B.1.1}. Space}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+\paragraph{Representation of space}
+
+Symbolic representation of space is a widely studied topic. In particular, a
+large literature corpus is available on spatial ontologies.
+
+Two main classes of spatial relations are usually represented at the symbolic
+level: the topology of environments and the placement of physical entities.
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf B.1.1 Space}
+            child {node [taxon] {{\bf B.1.1.2}. Placement}}
+            child {node [taxon] {{\bf B.1.1.1}. Topology}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
+
+...topology??
+
+Symbolic representation of entities placement can be absolute or relative. The
+relation \concept{isOn}, for example, leads to absolute statements: the
+validity of the relation is independent of the nature of its subjects and
+objects, and is also independent of the observer.
+
+On the contrary, the relation \concept{nextTo} is relative, and depends on the
+relative size of the subject and the object. Two houses distant of 2 meters
+from each other can be considered as next to each other. Two ants separated by
+2 meters are not next to each other.
+
+The relation \concept{leftOf} is another example of relative spatial relation,
+this time because it depends on the observer viewpoint.
+
+\fxwarning{\concept{front}/\concept{back} -> entities have or not a concept of front/back.}
+
+Choices must be made in the knowledge representation system to adequately
+represent relative spatial relations. Options include the computation of such
+relations only on-demand, when the context is know (which viewpoints, etc.) or
+storage of these facts in different models, one per agent.
+
+\fxwarning{Find references on that subject -> classification of spatial
+relations. Wikipedia?}
+
 \paragraph{Representation of time}
@@
 timestamps to events the robot perceives -- and in the future), and
-representing \emph{passing time} (durations, timespans) like in \emph{``the
-eggs will be cooked in 10 min''}.
+representing \emph{passing time} (situations, durations, timespans) like in
+\emph{``the eggs will be cooked in 10 min''}.
+
+We call \emph{situations} all temporal concepts with a non-zero duration,
+whereas \emph{events} are durationless temporal concepts.
 
@@
 
-We call a system that do not account for time (\ie that permanently lives in
+We call a system that does not account for time (\ie that permanently lives in
 present) \emph{atemporal}.
 
-\paragraph{Representation of space}
-
-\paragraph{Representation of events and actions}
+\paragraph{Actions}
+
+Events and actions are two temporal concepts that are of particular importance to
+robotic systems, as systems that perceive, react and perform in their environment.
+
+While representation of events \fxwarning{revient à} to label a timestamp,
+representation of actions are more complex since they semantically imply
+interactions between concepts. From a taxonomy point of view, actions are a
+particular type of event that normaly leads to a situation corresponding to the
+action realization.
+
+\emph{Thematic roles}~\cite{Gruber1965} (also found as \emph{semantic roles} or
+\emph{theta roles} in the literature) allow to semantically qualify the
+parameters of an action. The recipient of the action, the performer, the object
+acted upon, the destination are some example of common thematic roles.
+
+Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
+roles proposed by \cite{Aarts1997} (see~\cite{Gutierrez2001} for a comparison
+of other sets of thematic roles present in the literature).
+
+\begin{table}
+\begin{center}
+
+\begin{tabular}{lp{5cm}}
+
+\toprule
+       Role & Meaning \\
+\midrule
+      Agent & The \emph{doer} or instigator of the action denoted by the predicate. \\
+    Patient & The \emph{undergoer} of the action or event denoted by the predicate. \\
+      Theme & The entity that is moved by the action or event denoted by the predicate. \\
+Experiencer & The living entity that experiences the action or event denoted by the predicate. \\
+       Goal & The location or entity in the direction of which something moves. \\
+Benefactive & The entity that benefits from the action or event denoted by the predicate. \\
+     Source & The location or entity from which something moves. \\
+ Instrument & The medium by which the action or event denoted by the predicate is carried out. \\
+   Locative & The specification of the place where the action or event denoted by the predicate in situated. \\
+\bottomrule
+
+\end{tabular}
+\end{center}
+\caption{Thematic Roles~\cite{Aarts1997}.}
+\label{table|theta-roles}
+\end{table}
+
+In the context of robotic, a \emph{task} represents the abstract model of an
+action. Tasks usually specify at least the conditions required to performed the
+action and the consequences of the realization of the action, and are central
+to the decisional layers of the robot, in particular for the planning,
+monitoring and execution control activities. Knowledge representation systems
+may thus have specific mechanisms to represent them (and possible, to reason
+about them, as presented in section~\ref{sect|planning}).
+
+%An extensive report on task modeling in OWL ontologies in appendix
 

### 2012-05-12 krs-survey.tex
Typos in the taxonomy nomenclature

@@
     \node [taxon] {\bf E. Integration}
-            child {node [taxon] {{\bf D.4} Performances}}
-            child {node [taxon] {{\bf D.3} Monitoring}}
-            child {node [taxon] {{\bf D.2} Executive layers}}
-            child {node [taxon] {{\bf D.1} Sensori-motor}};
+            child {node [taxon] {{\bf E.4} Performances}}
+            child {node [taxon] {{\bf E.3} Monitoring}}
+            child {node [taxon] {{\bf E.2} Executive layers}}
+            child {node [taxon] {{\bf E.1} Sensori-motor}};
 \end{tikzpicture}
@@
 \begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf D.5 Performance \\Evaluation}
-            child {node [taxon] {{\bf D.5.2} Cognitive Performances}}
-            child {node [taxon] {{\bf D.5.1} Raw Performances}};
+    \node [taxon] {\bf E.4 Performance \\Evaluation}
+            child {node [taxon] {{\bf E.4.2} Cognitive Performances}}
+            child {node [taxon] {{\bf E.4.1} Raw Performances}};
  \end{tikzpicture}

### 2012-05-12 krs-survey.tex
Introduction section Representation

@@
 
+We do not discuss in this section the general strategies to construct a
+knowledge model (they will be presented in
+section~\ref{sect|design-strategies}). We focus here on questions that involve
+representational challenges (time, space, context) or require specific
+cognitive capabilities (theory of mind, introspection, memory).
+
 \begin{scriptsize}
@@
 
-Thematic roles
-
-Spatio-Temporal Representations:
+This section discusses strategies and approaches to represent in a knowledge
+model three important \emph{roles}\fxwarning{Reference on the term 'role'?}.
 
@@
 \subsubsection{Design Strategy}
+\label{sect|design-strategies}
 

### 2012-05-13 krs-survey.tex
Mention Levesque2008 in Previous Work

@@
 
-To the extend of our knowledge, no previous published work attempt to build a
-typology of cognitive abilities strictly related to knowledge in the context of
-service robotics.
+A large part of our proposal is build as a comprehensive synthesis of
+classification and analysis found in the literatture. This synthesis is focused
+on cognitive abilities strictly related to knowledge manipulation, in the
+context of service robotics.
+
+Levesque and Lakemeyer~\cite{Levesque2008} present in a chapter on Cognitive
+Robotics several characteristics of knowledge representation systems for
+robots, stressing the need of representing the dynamics of the world.  Sensing
+is included in the knowledge representation via fluents; they introduce the
+idea of \emph{possible worlds} to represent distinct parallel mental models;
+action representation and reasoning about tasks is discussed in the context of
+situation calculus; open world vs. closed world approaches are both mentioned.
+They also discuss how robot programming and knowledge representation can be
+related. We integrate most of these items in our typology.
 

### 2012-05-13 krs-survey.tex
Mention fluents, situation calculus

@@
 
-We call \emph{situations} all temporal concepts with a non-zero duration,
-whereas \emph{events} are durationless temporal concepts.
+This introduces the disjunction in time representation between
+\emph{situations}, that is, any temporal concept with a non-zero duration, and
+\emph{events}, that is, any durationless temporal concept.
+
+Other definitions of a situation do exist, notably in the context of
+\emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque consider a
+situation to be an history of actions.
 
@@
 
+The concept of \emph{fluent} plays another important role for time
+representation. Fluents are properties (or \emph{conditions}) that change over
+time, like $sees(agent1, apple, t)$.
+
 We call a system that does not account for time (\ie that permanently lives in
@@
 
-While representation of events \fxwarning{revient à} to label a timestamp,
+While representation of events boils down to label a timestamp,
 representation of actions are more complex since they semantically imply
@@
 
-Making decision based on prediction
-
 Symbolic task planning is the ability for a robot to select a sequence of
@@
 representation systems, task planning can be considered within the KRS.
+
+Since McCarthy's \emph{Situation Calculus} in 1963, numerous knowledge
+representation formalisms dedicated to representation and reasoning about
+actions and situation have emerged: besides situation calculs, \emph{fluent
+calculus} and \emph{event calculus} are the main ones.
+Thielscher~\cite{Thielscher2011}} recently proposed a unification of these
+approach in a new \emph{action calculus}.
 

### 2012-05-13 krs-survey.tex
Merged Dynamic Instantiation with Structural Alteration

@@
             child {node [taxon] {{\bf C.3}. Lazy evaluation}}
-            child {node [taxon] {{\bf C.2}. Structural alteration}}
+            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
             child {node [taxon] {{\bf C.1}. Standard reasoning}};
@@
 
-\subsubsection{Alteration of the knowledge structure}
+\subsubsection{Dynamic Instantiation and Alteration of the Knowledge Structure}
 
@@
 The terms TBox and ABox are commonly found to describe this two different types of
-statements in ontologies. TBox statements describe a system in terms of
+statements in ontologies: TBox statements describe a system in terms of
 controlled vocabularies, for example, a set of classes and properties. ABox are
@@
 
-All knowledge representation systems allow to modify the ABox, which can be
-considered as the dynamic part of the knowledge base. It is however not always
-possible to alter the TBox because not all reasoning systems can dynamically
-take into account structural changes in the knowledge base.
+Most knowledge representation systems allow to modify the ABox, which can be
+considered as the dynamic part of the knowledge base. Alteration of the ABox
+include the addition or retractation of relations between existing instances,
+and the addition or removal of instances. We call the later \emph{dynamic
+instantiation}: the capability for a system to create new instances at
+run-time. The term dynamic instantion applies primarily to concrete entities
+(typically, a new object is discovered by the robot, a symbolic instance is
+created for it), but may also apply to abstract entities (like an instance of
+action, of a feeling, etc.).
+
+KRS may also allow to alter the TBox. This requires the underlying reasoning
+systems to be able to dynamically take into account structural changes in the
+knowledge base.
 
@@
     \node [taxon] {\bf D. Acquisition}
-            child {node [taxon] {{\bf D.4} Dynamic instantiation}}
             child {node [taxon] {{\bf D.3} Motivation}}
@@
 Curiosity, Pro-active behaviour...
-
-\subsubsection{Dynamic instantiation}
-\label{sect|new-instances}
-Ability to automatically create new object instances
 

### 2012-05-13 krs-survey.tex
Started to fill section on 'Modality' (was Possible Worlds)

@@
             child {node [taxon] {{\bf B.4}. Introspection}}
-            child {node [taxon] {{\bf B.3}. Possible Worlds}}
+            child {node [taxon] {{\bf B.3}. Modality}}
             child {node [taxon] {{\bf B.2}. Context}}
@@
 
-\subsubsection{Possible-Worlds and representing what others know}
+\subsubsection{Modality, Contingency and Theory of Mind}
 \label{sect|possible-worlds}
@@
 Linked to the context representation, but seen from another angle, knowledge
-representation systems may provide explicit ways to model other point of view
+representation systems may support logical \emph{modality}. A knowledge model
+is logically modal if is support the concept of \emph{possible worlds}, \ie,
+parallel beliefs models (or \emph{interpretations}) that can be independently
+accessed.
+
+A \emph{contingent} proposition is defined as neither always true (a tautology)
+nor always false (a contradiction) in every possible world. In knowledge
+representation systems for robotics that support logical modality,
+interpretations are often initialised with a common set of initial beliefs
+(like common-sense knowledge). This initial common knowledge is hence true in
+every possible worlds for the robot, and thus does not belong to the contingent
+knowledge.
+
+Alternative mental models may be used to represent different (possibly
+hypothetical or even imaginary) views on the world, from the robot own
+perspective, or from persepectives of other agents, as computed by the robot.
+
+provide explicit ways to model other point of view
 on the world. This ability is often referred as the \emph{perspective taking}
 ability.
-
-\cite{Levesque2008}, p. 4
 
@@
 calculus} and \emph{event calculus} are the main ones.
-Thielscher~\cite{Thielscher2011}} recently proposed a unification of these
+Thielscher~\cite{Thielscher2011} recently proposed a unification of these
 approach in a new \emph{action calculus}.

### 2012-05-14 krs-survey.tex
Completed perspective taking theroy of mind

@@
 (like common-sense knowledge). This initial common knowledge is hence true in
-every possible worlds for the robot, and thus does not belong to the contingent
+every possible worlds for the robot, and thus does not belong to its contingent
 knowledge.
 
-Alternative mental models may be used to represent different (possibly
-hypothetical or even imaginary) views on the world, from the robot own
-perspective, or from persepectives of other agents, as computed by the robot.
-
-provide explicit ways to model other point of view
-on the world. This ability is often referred as the \emph{perspective taking}
-ability.
+Logical modality allows to implement alternative mental models. They may be
+used to represent different (possibly hypothetical or even imaginary) views on
+the world, from the robot own perspective, or from other persepectives computed
+by the robot.
+
+The representation of the mental perspective of other agents has a particular
+importance in human-robot interaction. It relies first on the ability to view
+the world from a standpoint which is not \emph{egocentric}. This cognitive
+ability is referred as \emph{perspective taking} (see
+Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} for the psychological
+grounds of perspective taking, itself originated in Piaget's work on cognitive
+development, as remembered by Flavell), and has been studied by
+infants~\cite{Moll2006} and robots as well~\cite{Trafton2005, Breazeal2006,
+Ros2010}.
+
+The idea of a \emph{theory of mind}~\cite{Premack1978} emerges from the
+perspective taking ability. It can be defined as the ability for one to
+understand and acknowledge that other intelligent agents can have their own
+mental state (that includes beliefs, intents, desires, knowledge) that is
+possibly different from one's own.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.9\columnwidth]{sally_ann.pdf}
+        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken from~\cite{Leslie2000}.}
+        \label{fig|sally-ann}
+\end{figure}
+
+Perceiving and understanding other's \emph{attention} has been identified as
+essential to the development of a theory of mind by
+Baron-Cohen~\cite{Baron-Cohen1985, Leslie2000}. In
+robotics~\cite{Scassellati2002}
 

### 2012-05-14 krs-survey.tex
Completed section on theory of mind + fixed sally ann figure

@@
 \usepackage{graphicx}
+\usepackage{subfigure}
 
@@
 mental state (that includes beliefs, intents, desires, knowledge) that is
-possibly different from one's own.
+possibly different from one's own. The \emph{attention} plays a central to the
+development and recognition of a theory of mind~\cite{Baron-Cohen1985,
+Leslie2000}.
+
+A notable consequence of having a theory of mind is the representation of
+\emph{false beliefs}, \ie, facts that are believed to be true for an agent, but
+false for other ones. The ``Sally and Ann'' experiment
+(figure~\ref{fig|sally-ann}) is a classical example of a false belief
+situation. In~\ref{fig|sally-ann-final}, Sally thinks the ball is in the beige
+box because she did not see Ann moving it. An external observer asked ``Where
+will Sally look for the ball?'' would answer ``in the blue box'' without a
+theory of mind (\ie, a model of the knowledge of Sally), whereas it would
+correctly answer ``In the beige box'' with a theory of mind.
 
@@
         \centering
-        \includegraphics[width=0.9\columnwidth]{sally_ann.pdf}
-        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken from~\cite{Leslie2000}.}
+        \subfigure[Sally puts the ball in the beige box.]{
+            \includegraphics[scale=0.3]{sally_ann1.pdf}
+        } \hspace{8pt} %
+        \subfigure[She goes away.]{
+            \includegraphics[scale=0.3]{sally_ann2.pdf}
+        } \\
+        \subfigure[Ann moves the ball.]{
+            \includegraphics[scale=0.3]{sally_ann3.pdf}
+        } \hspace{8pt} %
+        \subfigure[Where will Sally look for the ball?]{
+            \includegraphics[scale=0.3]{sally_ann4.pdf}
+            \label{fig|sally-ann-final}
+        }
+        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken
+        from~\cite{Leslie2000}.}
+
         \label{fig|sally-ann}
@@
 
-Perceiving and understanding other's \emph{attention} has been identified as
-essential to the development of a theory of mind by
-Baron-Cohen~\cite{Baron-Cohen1985, Leslie2000}. In
-robotics~\cite{Scassellati2002}
+Scassellati~\cite{Scassellati2002} is amongst the researcher having studied
+theory of mind for robots.
 

### 2012-05-14 krs-survey.tex
Filled section on Explanation

@@
 \subsubsection{Open World and Close World Assumptions}
+\label{sect|owa-cwa}
 
@@
             child {node [taxon] {{\bf C.8}. Planning}}
-            child {node [taxon] {{\bf C.7}. Prediction}}
+            child {node [taxon] {{\bf C.7}. Prediction and explanation}}
             child {node [taxon] {{\bf C.6}. Presupposition}}
@@
 
-\subsubsection{Prediction, projection and diagnosis tasks}
+\subsubsection{Prediction, projection, explanation}
 \label{sect|prediction-projection}
@@
 temporal reverse of the projection task.
+
+\paragraph{Explanation} Diagnosis is also linked to the \emph{explanation} or
+\emph{justification} capabilities that may be offered by the knowledge
+representation system. The explanation of an entailment is the sequence (or
+set of sequences if several are possible) of reasoning steps that allow to
+conclude. An explanation can also conversely explain why a statement leads to a
+contradiction.
+
+The following example shows an explanation for an inconsistency in a particular
+knowledge base: by adding the statement \stmt{robot1 belongsTo human1}, we
+observe that an inconsistency is triggered. The reasoner provides the four
+following observations\footnote{This example uses the \emph{Manchester syntax},
+\url{http://www.w3.org/TR/owl2-manchester-syntax/}} to explain the
+inconsistency:
+
+\begin{quote}
+\scriptsize
+\begin{enumerate}
+    \item \concept{myself} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
+    \item \concept{belongsTo} {\bf Domain} \concept{Artifact}
+    \item \concept{myself} {\bf Type} \concept{Agent}
+    \item {\bf DisjointClasses:} \concept{Agent}, \concept{Artifact}
+\end{enumerate}
+\normalsize
+\end{quote}
+
+This explanation directly portrays the underlying structure of the knowledge
+model. A KRS may provide mechanisms to expose this kind of analysis, either
+automatically when inconsistencies survene, or on demand.
+
+Explanation of contradictions plays a particular role for robots: from a
+cognitive point of view, a logical inconsistency (\ie, a contradiction) can be
+viewed as a \emph{cognitive conflict} or \emph{cognitive dissonance} (\ie, two
+incompatible models of the world that must be dealt with). Being able to expose
+and explain such cognitive conflicts eases the control of the behaviour of the
+robot in unexpected semantic situations, and forms a first step towards an
+adquate reaction.
+
+Cognitive dissonance is also identified by developmental psychologists like
+Piaget as a motivation factor for a child to progress through the various
+stages of its cognitive development. This has been also studied in
+robotics~\cite{Oudeyer2007}.
+
+The cognitive capability of \emph{justification} is also intimately linked to
+the meta-cognition capability and participates to the overall \emph{cognitive
+observability} of the system.
 

### 2012-05-14 krs-survey.tex
Filled section on intrinsic motivation

@@
 
-\subsubsection{Intrinsic Motivation}
-
-\cite{Oudeyer2007} -> page 2 gives a detailled account of psychological grounds.
-Curiosity, Pro-active behaviour...
+\subsubsection{Intrinsic Motivation and Curiosity}
+
+The reason for a robot to acquire knowledge are various, and usually external
+to the robot itself. It is often driven by the needs of the tasks the robot has
+to execute. In this case, motivations are managed by the execution controler
+and are mostly unvisible to the knowledge representation system.
+
+However, motivation can also be intrinsic, driven by the internal state of the
+robot, without external pressure or reward. In this case, the
+
+In the context of robotics, Oudeyer notes however that \emph{the information that is
+compared \emph{[to compute a level of motivation]} has to be understood in an
+information theoretic perspective, in which what is considered is the intrinsic
+mathematical structure of the values of stimuli, independently of their
+meaning.}~\cite{Oudeyer2007}.
+
+Psychological grounds of motivation are summarized in~\cite{Oudeyer2007} while
+the main approaches to computational motivation, divided into knowledge based
+models, competence based models and morphological models, are surveyed
+in~\cite{Oudeyer2008}.
 

### 2012-05-15 krs-survey.tex
Added first pass on KnowRob review + figures for KnowRob and ORO

@@
 
+{\sc KnowRob}~\cite{Tenorth2009a} is an integrated knowledge management system
+developed at the Technical University of Munich. It is build as a set of
+modules (figure~\ref{fig|knowrob} organized around a core reasoning system
+written in Prolog. The core component interfaces through Java/Prolog or
+C/Prolog APIs with external modules.
+
+{\sc KnowRob} offers a mechanism called \emph{computables} that allow to
+evaluate certain predicates by calling external dedicated functions (for
+instance, the valuation of a proposition like \stmt{object1 isOn object2} is
+computed when required by calling a specific geometric reasonning module).
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.36]{knowrob.pdf}
+
+    \caption{Overview of the {\sc KnowRob} framework, taken
+    from~\cite{Tenorth2011}.}
+
+    \label{fig|knowrob}
+\end{figure}
+
+Extension modules plug into the system and provide specialized reasoning
+capabilities or interfaces to external data, \eg to read object detections from
+the vision system. These modules mainly operate on the level of instances
+(ABox).
+
+\paragraph{Knowledge model} {\sc KnowRob} can load OWL ontologies, and the
+\emph{KnowRob-Base} ontology is provided as a common-sense ontology, with a
+focus on household and kitchen domains. {\sc KnowRob} also store and reason on
+introspective knowledge through the \emph{Semantic Robot Description
+Language}~\cite{Kunze2011} that allow to represent symbolically the
+capabilities of the robot, and is used for planning.
+
+\paragraph{Reasoning Techniques} Amongst the notable {\sc KnowRob} extensions,
+{\sc ProbCog}~\cite{Jain2009} is an effort to provide probabilistic reasoning
+based on bayesian networks, integration with naive physics reasoning has been
+studied~\cite{Kunze2011a}, automatic parsing of Web resource in semi-natural
+language has been experimented with~\cite{Tenorth2009b}.
+
+\paragraph{Grounding} Anchoring of knowledge in the {\sc KnowRob} framework
+relies on...CoP? Semantic Maps~\cite{Blodow2011}.
+
+%\paragraph{Integration with Decisional Layers} CRAM~\cite{Beetz2010}
+
+\paragraph{Notable Experiments} {\sc KnowRob} has been deployed on several
+scenarii at the Technical University of Munich, on the PR2 robot and on a 2-arm
+custom mobile manipulator in the scope of the ``assistive
+kitchen''~\cite{Beetz2008} project. In these experiments, {\sc KnowRob} has
+been used as central knowledge hub...
+
 \subsection{ORO}
 \label{sect|oro}
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.36]{oro.pdf}
+
+    \caption{Overview of the ORO server, taken from~\cite{Lemaignan2012}.}
+
+    \label{fig|oro}
+\end{figure}
 
@@
 
-\subsubsection{Notable experiments}
+\paragraph{Notable experiments}
 \label{sect|peis-expe}
@@
 
-\subsubsection{Intrinsic language features}
+\paragraph{Intrinsic language features}
 \label{sect|nkrl-intrinsic-features}
 
-\paragraph{Expressiveness}
-
-\subsubsection{Integration with physical world and in the robot architecture}
+Expressiveness...
+
+\paragraph{Integration with physical world and in the robot architecture}
 \label{sect|nkrl-integration}

### 2012-05-16 krs-survey.tex
Partially updated OMKRF to OUR-K

@@
 GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
-OMRKF & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
+OUR-K & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
 Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
@@
 In the presented implementation, the CAST knowledge model does no allow for
-effectively representing actions or temporal information.\fxfatal{What about reasoning? can they retrieve for example 'all proxies for colorful objects'?}
+effectively representing actions or temporal information.\fxfatal{What about
+reasoning? can they retrieve for example 'all proxies for colorful objects'?}
 
@@
 
-\subsection{OMRKF}
+\subsection{OUR-K and OMRKF}
 \label{sect|omrkf}
 
-The Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
-(OMRKF) is a knowledge representation system based on four inter-related
+The Ontology-based Unified Robot
+Knowledge~\cite{Lim2011} (OUR-K) framework, successor of the Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
+(OMRKF), is a knowledge representation system based on five inter-related
 \emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
 approach to knowledge representation that allows to integrate the grounding
-process to the knowledge representation process. OMRKF knowledge model is
-implemented with Horn clauses.
+process to the knowledge representation process. OUR-K knowledge model is
+implemented with a mix of Description Logics for the concept hierarchies and Horn clauses.
 
@@
     \centering
-    \includegraphics[width=0.8\columnwidth]{omrkf.pdf}
-
-    \caption{OMRKF organizes knowledge into four \emph{classes}, each composed
-    of three \emph{levels}. The figure shows some examples of link between
-    knowledge classes and knowledge levels. Based on~\ref{Suh2007}.}
+    \includegraphics[width=0.8\columnwidth]{ourk.pdf}
+
+    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
+    of \emph{levels}. Figure based on~\ref{Lim2011}.}
 
@@
 \emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
-or ``object feature'', a taxonomy of concepts inside this level (for instance
+or ``object feature''), a taxonomy of concepts inside this level (for instance
 $cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
@@
 
-Environment is represented in OMRKF in the $space : Model$ knowledge level as a
-classical three layers mapping (metric, topological and semantic maps). Objects
-(in $object : Model$) are localized in $space : Model$ through Voronoi nodes.
+\paragraph{Representation} The environment is represented in OUR-K in the
+$space : Model$ knowledge level as a classical three layers mapping (metric,
+topological and semantic maps). Objects (in $object : Model$) are localized in
+$space : Model$ through Voronoi nodes.
 
@@
 
-\paragraph{Experiments} Experiments conducted with OMRKF include finding
-kitchen objects and reporting about their state to a human.  This experiment
-also shows how OMRKF can deal with objects only partially matched by their
-descriptor by introducing a $candidate()$ function.
+\paragraph{Experiments} Experiments conducted with OUR-K and OMRKF include
+finding kitchen objects and reporting about their state to a human.  This
+experiment also shows how OUR-K can deal with objects only partially matched by
+their descriptor by introducing a $candidate()$ function.
 
@@
 GSM &  Predefined categories \\
-OMRKF & {\it A priori} knowledge structure and axioms, custom set of instances\\
+OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
 Ke Jia & None \\
@@
 \toprule
-\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OMRKF} & {\sc ARMAR} & {\sc Hertzberg} \\
+\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OUR-K} & {\sc ARMAR} & {\sc Hertzberg} \\
 

### 2012-05-16 krs-survey.tex
Add mentions of GOLOG and derivatives

@@
 which are focused on data fusion and knowledge grounding at lower levels.
+
+We have also chosen not to include the GOLOG language and its
+derivatives~\cite{Levesque1997, Ferrein2008, Gspandl2011} in this survey. While
+several implementations on robots, including service robots, do exist, the
+focus of this language is on representation and reasoning about actions and
+situations, and the link with symbolic, abstract knowledge is not explicit.
 
@@
 in case of failure to provide possible explanation. This can be seen as the
-temporal reverse of the projection task.
+temporal reverse of the projection task. Because of their modelisation of
+situations as an history of actions, derivatives of the GOLOG logic programming
+languages are a good example of diagnosis tasks integrated to the knowledge
+representation system~\cite{Gspandl2011}.
 
@@
 approach in a new \emph{action calculus}.
+
+The GOLOG~\cite{Levesque1997} language and its derivatives (like {\sc
+ReadyLOG}~\cite{Ferrein2008} and {\sc IndiLOG}~\ref{Gspandl2011}) propose
+implementations of the situation calculus focused on robotic applications.
 

### 2012-05-16 krs-survey.tex
Reorder reviews in alphabetical order

@@
 \midrule
+ARMAR/Tapas & KRS & Holzapfel, Waibel \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
+CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
+GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
+Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
 {\sc KnowRob} & KRS & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
+NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
+%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
+OUR-K & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
 ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
 PEIS KR\&R & KRS & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
-CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
-NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
-GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
-OUR-K & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
-Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
-ARMAR/Tapas & KRS & Holzapfel, Waibel \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
-Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
-OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
+%Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
 % & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
@@
 
-\fxfatal{Bielefeld -> could not find much\ldots}
-\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}
+%\fxfatal{Bielefeld -> could not find much\ldots}
+%\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}
+
+\subsection{ARMAR/Tapas}
+
+{\sc Tapas} is the name of the knowledge representation system and dialogue
+manager found on the ARMARIII robot~\cite{Holzapfel2008}.
+
+Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
+knowledge. The later is split into \emph{lexical knowledge}, \emph{semantic
+knowledge} and a database of identified objects (with their properties). The
+\emph{lexical knowledge} contains lexical and grammatical informations about
+the objects. The \emph{semantic knowledge} is organized into an ontology
+relying on \emph{typed feature structures} (TFS,~\cite{Carpenter1992}, a
+formalism originating from the computational linguistics community, and a
+superset of first-order logic).
+
+{\sc Tapas} has a strong focus on natural language grounding. It proceeds by
+generating grammars from properties represented in the ontology to parse and
+understand dialogue.
+
+Another focus is put on handling unknown words and objects. {\sc Tapas}
+provides routines to recognize unknown entities, and propose and interactive
+and iterative verbal process to categorize (including adding new categories)
+those new concepts.
+
+\subsection{CAST Knowledge model}
+\label{sect|cast}
+
+CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
+comprehensive toolkit aimed at building cognitive architectures for robots
+through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
+expose a central knowledge base as seen in previous works. It instead
+represents knowledge as unrooted \emph{proxies}. Those proxies are formally
+defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
+a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
+\emph{proxies union} that form an equivalence class corresponding to one
+entity.
+
+A union of proxies forms a global amodal representation of an entity, that can
+be explicitly shared and manipulated. Being not centralized, the knowledge
+model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
+the CAS architecture is tightly bound to the on-line grounding process (be it
+grounded in perception or in dialogue). While nothing seems to prevent it, no
+{\it a priori} knowledge (including common-sense knowledge) is used.
+
+Knowledge sharing is ensured by the event mechanism of CAST: modules can
+monitor proxies for alteration by other modules. Jacobsson et al. mention how
+this can apply to reinforcement learning: the vision module creates a proxy for
+an orange object. This proxy get monitored by a learning module. In parallel,
+the proxy is bound to an union by the natural language understanding module
+that add new a feature like \emph{"this object is a fruit"}. The learning
+module is called back, and can add this new information to its model.
+
+In the presented implementation, the CAST knowledge model does no allow for
+effectively representing actions or temporal information.\fxfatal{What about
+reasoning? can they retrieve for example 'all proxies for colorful objects'?}
+
+In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
+the context of spatial representations.
+
+\subsection{GSM}
+\label{sect|gsm}
+
+GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
+representation system primarily built to ``facilitate cross-modal
+interoperability'',  especially in the context of verbal interaction with a
+robot.
+
+Unlike most of the previously presented systems, GSM does not rely on any
+formal language but rather on a layered data structure (Figure~\ref{fig|gsm})
+that organizes the surrounding world into agents and relations between agents.
+Each agent (any animate or inanimate object) is attached to a physical model (made
+of \emph{body parts} that have properties like their position, color, etc.) and
+a mental model (which is a recursively embedded GSM, thus allowing a sort of
+theory of mind).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{gsm.pdf}
+
+    \caption{Simplified hierarchical structure of the Grounded Situation Model,
+    based on~\cite{Mavridis2006}}
+
+    \label{fig|gsm}
+\end{figure}
+
+Properties are represented in three layers: a stochastic representation, close
+to sensory percepts, a \emph{continuous single-valued} encoding of the
+stochastic model, and a discrete, categorical model.
+
+One notable feature of GSM is the \emph{bidirectionality} of the grounding
+process: not only sensor percepts are abstracted into categories suitable for
+human conversation, but human utterance (like ``There is a red ball in the
+center of the table'') can also be turned into property descriptions. This
+basically enable the knowledge representation system of the robot to
+\emph{imagine} entities.
+
+GSM also features several strategies for managing time and events.
+\emph{Moments} are created by storing timestamped snap-shots of GSM, and
+\emph{event classifiers} allow to define and detect events.
+
+\paragraph{Experiments} GSM has mostly been tested on table-top manipulation
+and interaction tasks (a ``conversational helping hand'' as stated by the
+authors) realized on a 7-DOF arm equipped with force feedback, cameras for blob
+tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
+an in-depth analysis of the performance of GSM by the mean of a standard
+psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.
+
+\subsection{GOLOG}
+\label{sect|kejia}
+
+\subsection{Ke Jia Project}
+\label{sect|kejia}
+
+The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
+representation language with natural language processing, task planing and
+motion planing.
+
+Knowledge representation relies on \emph{Action Language C}, itself based on
+\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
+are syntactically close to Prolog, are based on \emph{stable models} of logic
+programs, and support non-monotonic reasoning. Default and non-monotonic
+reasoning has been especially researched within the Ke Jia project for symbolic
+task planing~\cite{Ji2011} and underspecified natural language processing.
+
+Amongst other features, the natural language processing capabilities of the
+system support acquisition of new logical rules at run-time.
+
+\paragraph{Experiments} The Ke Jia robot has been demonstrated in several tasks
+involving human-robot interaction with natural language. These tasks include a
+task with multiple \emph{pick \& carry} that are globally optimized, naive
+physics reasoning via taught rules or more complex scenarii with the robot
+delivering drinks, taking into account changing and mutually exclusive
+preferences of users.
 
@@
 been used as central knowledge hub...
+
+\subsection{NKRL}
+\label{sect|nkrl}
+
+\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
+While this language is developed since a long time by Zarri~\cite{Zarri1997,
+Zarri2008}, recent research direction include application to the robotic
+field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
+system, as it is primarily a language. However, it is used as the
+representation and reasoning mechanism for robots by Sabri et al.
+
+\paragraph{Intrinsic language features}
+\label{sect|nkrl-intrinsic-features}
+
+Expressiveness...
+
+\paragraph{Integration with physical world and in the robot architecture}
+\label{sect|nkrl-integration}
+
+\ldots seem to be mostly WIP\ldots
+
+\subsection{OUR-K and OMRKF}
+\label{sect|omrkf}
+
+The Ontology-based Unified Robot
+Knowledge~\cite{Lim2011} (OUR-K) framework, successor of the Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
+(OMRKF), is a knowledge representation system based on five inter-related
+\emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
+approach to knowledge representation that allows to integrate the grounding
+process to the knowledge representation process. OUR-K knowledge model is
+implemented with a mix of Description Logics for the concept hierarchies and Horn clauses.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{ourk.pdf}
+
+    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
+    of \emph{levels}. Figure based on~\cite{Lim2011}.}
+
+    \label{fig|omrkf}
+\end{figure}
+
+Each level of knowledge is build as three stages of ontological realization: a
+\emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
+or ``object feature''), a taxonomy of concepts inside this level (for instance
+$cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
+taxonomy ($cup1 : cup$).
+
+\paragraph{Representation} The environment is represented in OUR-K in the
+$space : Model$ knowledge level as a classical three layers mapping (metric,
+topological and semantic maps). Objects (in $object : Model$) are localized in
+$space : Model$ through Voronoi nodes.
+
+The knowledge class $Context$ proposes an explicit statement of spatial context
+(mostly geometric relations between objects), temporal context and a more
+general \emph{high-level} context, inferred from spatial and temporal contexts.
+
+Finally, the $Activity$ knowledge class store compound actions in a HTN-like
+structure, exploited at run-time by a planner.
+
+\paragraph{Experiments} Experiments conducted with OUR-K and OMRKF include
+finding kitchen objects and reporting about their state to a human.  This
+experiment also shows how OUR-K can deal with objects only partially matched by
+their descriptor by introducing a $candidate()$ function.
 
@@
 \label{sect|peis-expe}
-
-\subsection{NKRL}
-\label{sect|nkrl}
-
-\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
-While this language is developed since a long time by Zarri~\cite{Zarri1997,
-Zarri2008}, recent research direction include application to the robotic
-field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
-system, as it is primarily a language. However, it is used as the
-representation and reasoning mechanism for robots by Sabri et al.
-
-\paragraph{Intrinsic language features}
-\label{sect|nkrl-intrinsic-features}
-
-Expressiveness...
-
-\paragraph{Integration with physical world and in the robot architecture}
-\label{sect|nkrl-integration}
-
-\ldots seem to be mostly WIP\ldots
-
-\subsection{CAST Knowledge model}
-\label{sect|cast}
-
-CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
-comprehensive toolkit aimed at building cognitive architectures for robots
-through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
-expose a central knowledge base as seen in previous works. It instead
-represents knowledge as unrooted \emph{proxies}. Those proxies are formally
-defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
-a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
-\emph{proxies union} that form an equivalence class corresponding to one
-entity.
-
-A union of proxies forms a global amodal representation of an entity, that can
-be explicitly shared and manipulated. Being not centralized, the knowledge
-model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
-the CAS architecture is tightly bound to the on-line grounding process (be it
-grounded in perception or in dialogue). While nothing seems to prevent it, no
-{\it a priori} knowledge (including common-sense knowledge) is used.
-
-Knowledge sharing is ensured by the event mechanism of CAST: modules can
-monitor proxies for alteration by other modules. Jacobsson et al. mention how
-this can apply to reinforcement learning: the vision module creates a proxy for
-an orange object. This proxy get monitored by a learning module. In parallel,
-the proxy is bound to an union by the natural language understanding module
-that add new a feature like \emph{"this object is a fruit"}. The learning
-module is called back, and can add this new information to its model.
-
-In the presented implementation, the CAST knowledge model does no allow for
-effectively representing actions or temporal information.\fxfatal{What about
-reasoning? can they retrieve for example 'all proxies for colorful objects'?}
-
-In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
-the context of spatial representations.
-
-\subsection{GSM}
-\label{sect|gsm}
-
-GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
-representation system primarily built to ``facilitate cross-modal
-interoperability'',  especially in the context of verbal interaction with a
-robot.
-
-Unlike most of the previously presented systems, GSM does not rely on any
-formal language but rather on a layered data structure (Figure~\ref{fig|gsm})
-that organizes the surrounding world into agents and relations between agents.
-Each agent (any animate or inanimate object) is attached to a physical model (made
-of \emph{body parts} that have properties like their position, color, etc.) and
-a mental model (which is a recursively embedded GSM, thus allowing a sort of
-theory of mind).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{gsm.pdf}
-
-    \caption{Simplified hierarchical structure of the Grounded Situation Model,
-    based on~\cite{Mavridis2006}}
-
-    \label{fig|gsm}
-\end{figure}
-
-Properties are represented in three layers: a stochastic representation, close
-to sensory percepts, a \emph{continuous single-valued} encoding of the
-stochastic model, and a discrete, categorical model.
-
-One notable feature of GSM is the \emph{bidirectionality} of the grounding
-process: not only sensor percepts are abstracted into categories suitable for
-human conversation, but human utterance (like ``There is a red ball in the
-center of the table'') can also be turned into property descriptions. This
-basically enable the knowledge representation system of the robot to
-\emph{imagine} entities.
-
-GSM also features several strategies for managing time and events.
-\emph{Moments} are created by storing timestamped snap-shots of GSM, and
-\emph{event classifiers} allow to define and detect events.
-
-\paragraph{Experiments} GSM has mostly been tested on table-top manipulation
-and interaction tasks (a ``conversational helping hand'' as stated by the
-authors) realized on a 7-DOF arm equipped with force feedback, cameras for blob
-tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
-an in-depth analysis of the performance of GSM by the mean of a standard
-psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.
-
-\subsection{OUR-K and OMRKF}
-\label{sect|omrkf}
-
-The Ontology-based Unified Robot
-Knowledge~\cite{Lim2011} (OUR-K) framework, successor of the Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
-(OMRKF), is a knowledge representation system based on five inter-related
-\emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
-approach to knowledge representation that allows to integrate the grounding
-process to the knowledge representation process. OUR-K knowledge model is
-implemented with a mix of Description Logics for the concept hierarchies and Horn clauses.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{ourk.pdf}
-
-    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
-    of \emph{levels}. Figure based on~\ref{Lim2011}.}
-
-    \label{fig|omrkf}
-\end{figure}
-
-Each level of knowledge is build as three stages of ontological realization: a
-\emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
-or ``object feature''), a taxonomy of concepts inside this level (for instance
-$cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
-taxonomy ($cup1 : cup$).
-
-\paragraph{Representation} The environment is represented in OUR-K in the
-$space : Model$ knowledge level as a classical three layers mapping (metric,
-topological and semantic maps). Objects (in $object : Model$) are localized in
-$space : Model$ through Voronoi nodes.
-
-The knowledge class $Context$ proposes an explicit statement of spatial context
-(mostly geometric relations between objects), temporal context and a more
-general \emph{high-level} context, inferred from spatial and temporal contexts.
-
-Finally, the $Activity$ knowledge class store compound actions in a HTN-like
-structure, exploited at run-time by a planner.
-
-\paragraph{Experiments} Experiments conducted with OUR-K and OMRKF include
-finding kitchen objects and reporting about their state to a human.  This
-experiment also shows how OUR-K can deal with objects only partially matched by
-their descriptor by introducing a $candidate()$ function.
-
-\subsection{Ke Jia Project}
-\label{sect|kejia}
-
-The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
-representation language with natural language processing, task planing and
-motion planing.
-
-Knowledge representation relies on \emph{Action Language C}, itself based on
-\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
-are syntactically close to Prolog, are based on \emph{stable models} of logic
-programs, and support non-monotonic reasoning. Default and non-monotonic
-reasoning has been especially researched within the Ke Jia project for symbolic
-task planing~\cite{Ji2011} and underspecified natural language processing.
-
-Amongst other features, the natural language processing capabilities of the
-system support acquisition of new logical rules at run-time.
-
-\paragraph{Experiments} The Ke Jia robot has been demonstrated in several tasks
-involving human-robot interaction with natural language. These tasks include a
-task with multiple \emph{pick \& carry} that are globally optimized, naive
-physics reasoning via taught rules or more complex scenarii with the robot
-delivering drinks, taking into account changing and mutually exclusive
-preferences of users.
-
-\subsection{ARMAR/Tapas}
-
-{\sc Tapas} is the name of the knowledge representation system and dialogue
-manager found on the ARMARIII robot~\cite{Holzapfel2008}.
-
-Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
-knowledge. The later is split into \emph{lexical knowledge}, \emph{semantic
-knowledge} and a database of identified objects (with their properties). The
-\emph{lexical knowledge} contains lexical and grammatical informations about
-the objects. The \emph{semantic knowledge} is organized into an ontology
-relying on \emph{typed feature structures} (TFS,~\cite{Carpenter1992}, a
-formalism originating from the computational linguistics community, and a
-superset of first-order logic).
-
-{\sc Tapas} has a strong focus on natural language grounding. It proceeds by
-generating grammars from properties represented in the ontology to parse and
-understand dialogue.
-
-Another focus is put on handling unknown words and objects. {\sc Tapas}
-provides routines to recognize unknown entities, and propose and interactive
-and iterative verbal process to categorize (including adding new categories)
-those new concepts.
 

### 2012-10-08 krs-survey.tex
Merged from dissertation
-> many changes in many places

@@
 
+\fxnote{Support material: \emph{What is a knowledge representation} by Davis,
+Shrobe and Szolovits,
+\url{http://groups.csail.mit.edu/medg/ftp/psz/k-rep.html}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
-In a chapter on that subject in \emph{Fondations of Artifical
+In a chapter on that subject in \emph{Foundations of Artificial
 Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
@@
         \item to execute a program, what information does a robot need to have
-        at the outset vs. the information that it can acquire \emph{en route}
+        at the outset versus the information that it can acquire \emph{en route}
         by perceptual means?
 
-        \item what does the robot need to know about its environment vs. what
+        \item what does the robot need to know about its environment versus what
         need only be known by the designer?
@@
 
-\subsection*{A Note on Terminology}
-
-Since this article discusses at length the concept of knowledge in the context
-of robotics, it is useful to make our terminology explicit.  No general
-agreement on a definition of ``knowledge'' exists. For our needs, we call
-``knowledge'' \emph{an explicit set of interrelated logical facts that are
-meaningful to the robot's executive controller}, where \emph{meaningful} means
-that it can possibly be interpreted to lead to a purposeful action).
+\paragraph{What do we call ``knowledge''?}
+\label{sect|on-knowledge}
+
+Since we will discuss at length the concept of knowledge in the context of
+robotics in the coming pages, it is useful to make our terminology explicit.
+
+Be it in philosophy, cognitive sciences or computer sciences, reaching an
+agreement on a definition of ``knowledge'' seems difficult.
+
+Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
+starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
+and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
+with representation, the knowledge level deals with language, semantics;
+whereas the symbol level deals with inference, the knowledge level deals with
+entailment. We will, at the conclusion of the thesis, give a second look to
+this distinction.
+
+In our robotic context, we define knowledge as a narrower concept, while
+keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
+interrelated logical facts that are meaningful to the robot executive
+controller}. By \emph{meaningful} we mean that can possibly be interpreted to
+lead to a purposeful action. We will see that our main challenge while
+designing a cognitive architecture is furthermore to make this knowledge as
+\emph{explicit} as possible.
 
 The relation of \emph{data} and \emph{information} to knowledge is a debated
-epistemology question (the interested reader is invited to refer to the
-Wikipedia page on the ``DIKW'' hierarchy). In this article, we will associate
-\emph{data} to low-level material like raw sensor output, and
-\emph{information} to uncontextualized symbolic facts.
-
-To give a example, imagine a human looking at a book while being tracked by a
-Kinect sensor: the pose of the human skeleton in the world would be the data,
-the fact \concept{looksAt(entity1, entity2)} as computed by a geometric reasoning
+epistemology question known as the ``DIKW'' hierarchy question. In this thesis,
+we will associate data to low-level material like raw sensor output, and
+information to uncontextualised symbolic facts.
+
+To give a example, we can imagine a human reading a book while being tracked by
+a Kinect sensor: the pose of the human skeleton in the world would be the data,
+the fact \concept{looksAt(human, book)} as interpreted by a geometric reasoning
 module would be the information, the fact \concept{looksAt(john,
@@
 
-This exclude \textit{de-facto} platforms like {\sc DyKnow}~\cite{Heintz2004}
+These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
 which are focused on data fusion and knowledge grounding at lower levels.
@@
 
-While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006} or
-{\sc ACT-R} have declarative knowledge modules~\cite{Derbinsky2010} and have
-been recently used on service robots (see {\sc ACT-R/E}~\cite{Kennedy2009} for
-instance), they are also absent from this survey because we did not find much
-references in the literature on knowledge manipulation and representation
-applied to robotic scenarii for these architectures.
-
-\subsection{Previous Work}
+While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
+GLAIR~\cite{Shapiro2009} or {\sc ACT-R} have declarative knowledge
+modules~\cite{Derbinsky2010} and have been recently used on service robots (see
+{\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are also absent from this
+survey because we did not find much references in the literature on knowledge
+manipulation and representation applied to real-world robotic scenarii for
+these architectures.
+
+A comprehensive reference on (bio-inspired) cognitive architectures is also
+available from the BICA Society~\cite{BicaCogArch2011}.
+
+\subsection{Previous work}
 \label{sect|evaluation-literature}
 
-A large part of our proposal is build as a comprehensive synthesis of
-classification and analysis found in the literatture. This synthesis is focused
-on cognitive abilities strictly related to knowledge manipulation, in the
+The typology we propose is in part based on a comprehensive synthesis of
+classifications and analysis found in the literature. This synthesis is focused
+on cognitive abilities strictly related to knowledge manipulation in the
 context of service robotics.
 
-Levesque and Lakemeyer~\cite{Levesque2008} present in a chapter on Cognitive
-Robotics several characteristics of knowledge representation systems for
-robots, stressing the need of representing the dynamics of the world.  Sensing
-is included in the knowledge representation via fluents; they introduce the
-idea of \emph{possible worlds} to represent distinct parallel mental models;
-action representation and reasoning about tasks is discussed in the context of
-situation calculus; open world vs. closed world approaches are both mentioned.
-They also discuss how robot programming and knowledge representation can be
-related. We integrate most of these items in our typology.
+Levesque and Lakemeyer~\cite{Levesque2008} present in their chapter on
+Cognitive Robotics several characteristics of knowledge representation systems
+for robots, stressing the need of representing the \emph{dynamics of the
+world}.  Sensing is included in the knowledge representation via
+\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
+distinct parallel mental models; action representation and reasoning about
+tasks is discussed in the context of \emph{situation calculus}; \emph{open
+world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
+robot programming and knowledge representation can be related. We integrate
+most of these items in our typology.
 
 In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
-\emph{knowledge processing middlewares} as systems supporting ``declarative
+\emph{knowledge processing middleware} as systems supporting ``declarative
 specifications for flexible configuration and dynamic reconfiguration of
 context dependent processing at many different levels of abstraction''. They
-identify six characteristics: the system must be able to merge informations
-from different, possibly distributed sources; it should support quantitative as
-well as qualitative processing of information, it should offer bottom-up and
-top-down processing, it should be able to deal with uncertainty, allow for
-``flexible configuration and reconfiguration'' (which require what we call
-\emph{non-monotonicity} in this article) and finally meta-knowledge and
-introspective capacities (``declarative specification of the processing
-functionalities''). In this paper, we focus our analysis on similar system with
-however an emphasize on the symbolic capabilities.
-
-Several surveys compare global cognitive architectures (\cite{Vernon2007, Chong2009}).
+identify six characteristics: the system must be able to \emph{merge
+informations} from different, possibly distributed sources; it should support
+quantitative as well as qualitative processing of information, it should offer
+\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
+\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
+(which require what we call here \emph{non-monotonicity}) and finally
+\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
+specification of the processing functionalities'').
+
+Several surveys compare global cognitive architectures \cite{Langley2006,
+Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
+distinguish nine capabilities: recognition and categorisation, decision making,
+perception and situation assessment, prediction and monitoring, planning,
+reasoning, execution control, interaction and
+learning/remembering/introspection. They also separately identify four
+\emph{properties} of a cognitive architecture, that categorise how knowledge is
+handled by the architecture: representation of knowledge, organisation of
+knowledge, utilisation of knowledge and acquisition and refinement of
+knowledge. This categorisation had a notable influence on our typology, and
+many of these categories are also present in our proposal.
+
 Vernon et al.~\cite{Vernon2007} split these architectures into two broad
@@
 al. propose twelve \emph{characteristics of cognitive system} to compare
-architectures. Amongst them, they mention the \emph{inter-agent epistemology} (how
-the structure of the world is captured in a representation and shared), the
-relation to \emph{embodiment}, the ability to \emph{anticipate} and to
+architectures. Amongst them, they mention the \emph{inter-agent epistemology}
+(how the structure of the world is captured in a representation and shared),
+the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
 \emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
@@
 knowledge representation strategies and are relevant to our study.
-\fxwarning{Detail Chong here}
-
-At an even broader scope, several authors from fields that are connected to robotics
-have previously listed desirable features of artificial systems aiming at
-rich cognitive abilities.
-
-For instance, in~\cite{McCarthy2007}, McCarthy recently listed the challenges
-he identifies on the road to a \emph{human-level AI}.
+
+Chong et al.~\cite{Chong2009} also provide a recent review of the main cognitive
+architectures, with a focus on eight functions: perception, memory, goals
+management, problem solving capabilities, planning, reasoning, learning and
+links to neurobiology.
+
+At an even broader scope, several authors from fields that are connected to
+robotics have previously listed desirable features of artificial systems aiming
+at rich cognitive abilities.
+
+For instance McCarthy recently listed in~\cite{McCarthy2007} the challenges he
+identifies on the road to a \emph{human-level AI}.
 
@@
 
-	\item the ability to formalize and reason about contexts,
+	\item the ability to formalise and reason about contexts,
 
@@
 Coming from the perspective of natural language processing in situated context,
-Roy and Reiter, in~\cite{Roy2005}, summarize what they see as the main
-challenges to be tackled by KRS: cross-modal representation systems,
-association of words with perceptual and action categories (\emph{grounding}),
-modeling of context, definition of the right granularity of models, integration
-of temporal modeling and planning, ability to match past (learned) experiences
-with the current interaction and ability to take into account the human
-perspective.
-
-Knowledge representation systems in robotics are concerned by most, if not all,
-of these points, and we shall indeed mention them in our typology, in slightly
-reformulated ways.
+Roy and Reiter summarise in~\cite{Roy2005} what they see as the main challenges
+to be tackled by knowledge representation systems: \emph{cross-modal
+representation} systems, association of words with perceptual and action
+categories (\emph{grounding}), modeling of \emph{context}, definition of the
+right \emph{granularity} of models, integration of \emph{temporal modeling and
+planning}, ability to \emph{match past (learned) experiences} with the current
+interaction and ability to take into account the \emph{human perspective}.
+
+Knowledge representation systems in robotics are directly affected by these
+points, and we indeed integrate them in our typology, in slightly reformulated
+ways.
 
@@
 	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
-	\caption{A possible representation of the scenario, in the MORSE simulator}
+	\caption{The {\em Brownie} scenario}
 	\label{fig|scenario}
@@
 
-In order to illustrate what we consider to be the required features of a
-knowledge representation system for robotics, we introduce in this section an
-(imaginary) demonstration scenario.
-
-While devising a scenario inevitably constraints the breadth of abilities we
-examine, we insist on the fact that our main purpose here is to ground into a
-concrete case the set of representational features we consider as desirable.
+The aim of this imaginary scenario (that has not been implemented, neither in
+simulation nor on a robot) is to materialise early in this paper the context
+and challenges of knowledge representation and manipulation for service, social
+and interactive robotics. It underlines the place, the role and the need of
+knowledge in a near-future, everyday situation where several robots and humans
+co-exist and cooperate. This scenario will also be a source of support examples
+for later sections of the thesis. Also, we have emphasised several keywords in
+this description: we will come back to them in the next section to
+explain them formally and relate them to each other.
 
 We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
-Robi and Roba are two service robots, that can freely
-move and pick objects around (with possibly different hardware and softwares
-architectures, including different knowledge representation systems). They
-cooperate with a human in a kitchen environment.
+Robi and Roba are two service robots, that can freely move and pick objects
+around (with possibly different hardware and software architectures, including
+different knowledge representation systems: typically, two robots built by two
+different companies). They cooperate with a human in a kitchen environment.
 
 The main task of the scenario is the joint realization of a brownie, initiated
-by the human: ``Let's make a brownie for tonight!''.
-
-The scenario is successful if the task is achieved (the brownie is baked) in a
-reasonable time (typically shorter than what it would have been required by the
-human alone).
+by Tom, the human: ``Let's make a brownie for tonight!''.
+
+The scenario is successful {\it 1)} if the task is achieved (the brownie is baked),
+{\it 2)} in a reasonable time (typically shorter than what it would have been
+required by the human alone), {\it 3)} and the human is happy by the help it received
+(he/she did not feel useless, uncomfortable, unsafe).
 
 We voluntarily do not detail the subtasks of the scenario, neither we define
-how they are shared amongst agents. In our analysis we focus on the general,
-\textit{a priori} representational needs of the scenario.
+how they are shared amongst agents: our focus is on knowledge needs and flows.
 
 A ``first-order'' analysis of this task leads to a rough partition of the
-required representation abilities:
+required \emph{representation} abilities:
 
@@
 \emph{common-sense knowledge} and the robot must be able not only to represent
-it, but also to have access to an initial source (for instance through a initial
-set of facts that are made available at startup, or via access to a Web-based
-knowledge base like Wikipedia, etc.)
+it, but also to have access to an initial source (for instance through a
+initial set of facts that are made available at startup, or via access to a
+Web-based knowledge base like Wikipedia)
 
@@
 
-These last assertions imply several other features for our knowledge
-representation system: ``cooking often takes place in the kitchen'' implies
-that representation of both uncertainty and likelihood is desirable. The fact
-that cooking is associated to a place further implies that the system models
-locations and is able to attach \emph{thematic relations} to concept
-(here, the likely location of the cooking action).
-
-``cooking requires ingredients'' hints about another very common feature
-available in most knowledge representation systems: \emph{reasoning}. The robot
-can \emph{infers} that cooking may require a recipe since a list of ingredients
-is a pre-requisite of the cooking action, and a recipe may provide such a list.
-If we omit the ``may'', this is a typical example of first-order logic
+These last assertions imply several other capabilities: ``cooking often takes
+place in the kitchen'' implies that representation of both uncertainty and
+likelihood is desirable. The fact that cooking is associated to a place further
+implies that the system models locations and is able to attach \emph{thematic
+relations} to concepts (here, the likely location of the cooking action).
+
+``cooking requires a procedure and ingredients'' hints about another important
+feature closely tied on knowledge manipulation: \emph{reasoning}. The robot can
+\emph{infer} that cooking may require a recipe since a procedure and a list of
+ingredients are pre-requisites of the cooking action that may be provided by a
+recipe.  If we omit the ``may'', this is a typical example of first-order logic
 reasoning.  Many other reasoning techniques exist (including probabilistic ones
@@
 We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
-robot should be able to store this plan in a way that allow later execution.
+robot should be able to store this plan in a way that allows later execution.
 The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
 the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
-must as well be appropriately represented.
-
-To make decision, a robot may also want to \emph{predict} the state of the world
-after some action (``if I leave the cake 2h in the oven, it will burn'').
+must be as well appropriately handled.
+
+To make decision, a robot may also want to \emph{predict} the state of the
+world after some action (``if I leave the cake 2h in the oven, it will burn'').
 Such ability to project itself in future or, generally speaking, in other
 possible state of the world is related to several cognitive ability and
-reasoning techniques: \emph{planning}, \emph{representation of possible worlds}
-and \emph{non-monotonic reasoning}, in addition to common-sense knowledge and
-\emph{physics-based} reasoning (that allows to predict that an egg is likely to
-break if dropped).
-
-Procedures are in addition often \emph{underspecified}: we can expect the recipe
-to provide a cooking duration, but we usually do not expect the recipe to tell
-us to first open the oven door, and then put the cake into it, since it is
-self-evident that the door must first be opened to put the cake in the oven.
-Such underspecification should be detectable, representable, and ideally
-completable by the knowledge representation system\footnote{Note that we do not
-assume here the {\it knowledge representation system} to be a single software
-component: it may well be the result of the aggregation of several logical
-components working together. We come back later on this aspect.}.
+reasoning techniques: \emph{planning}, \emph{projection}, \emph{representation
+of possible worlds} and \emph{non-monotonic reasoning}, in addition to
+common-sense knowledge and \emph{physics-based} reasoning (that allows for
+instance to predict that an egg is likely to break if dropped).
+
+Procedures are in addition often \emph{underspecified}: we can expect the
+recipe to provide a cooking duration, but we usually do not expect the recipe
+to tell us to first open the oven door, and then put the cake into it, since it
+is self-evident that the door must first be opened to put the cake in the oven.
+Our cognitive robot should ideally be able to detect and possibly complete such
+underspecification.
 
@@
 knowledge of the other agents: to think \emph{from the perspective} of another
-agent (an idea that is related to the availability of a theory of
-mind~\cite{Leslie2000}).
+agent (an idea that is related to the availability of a theory of mind, we will
+come back to it later on).
 
@@
 common interpretative framework (a shared common-sense knowledge -- or cultural
-background -- and a shared context)~\cite{Jakobson1960}. In our scenario, the
-agents are working in a kitchen. This element of context does not however
-suffice if, for example, an agent asks another agent to ``give {[him]} the
-bowl''. Behind the symbol ``bowl'', which physical entity are we actually
-talking about? If we want to act on the world, this so-called \emph{grounding}
-operation is essential, and may be tightly bound to the underlying knowledge
-representation system. Note that grounding is a bidirectional process: in
-covers the {\it top-down} operation (from the symbol to the percept) and the
-{\it bottom-up} converse (retrieval or creation of symbols from perception).
+background -- and a shared context). In our scenario, the agents are working in
+a kitchen. This element of context does not however suffice if, for example,the 
+human asks a robot to ``give {[him]} the bowl''. Behind the symbol
+``bowl'', which physical entity are we actually talking about? If we want to
+talk and act on the world, this so-called \emph{grounding} operation is
+essential. It is a bidirectional process: in covers the {\it top-down}
+operation (from the symbol to the percept) and the {\it bottom-up} converse
+(retrieval or creation of symbols from perception).
 
@@
 agent want to move, it must take into account the new obstacle) something that
-was not directly perceived. A knowledge representation system may be able to
-provide structures to represent and reason about such cases.
-
-Also central to cooperation are the notions of \emph{joint
-intentions} and \emph{joint goals}~\cite{Tomasello2005, Bratman2009}: to help
-the human during the cooking session, the robots need to track how far
-they are into the recipe, what is the next step the human is likely to go for,
-how task are currently split between agents, what action is currently
-blocking the procedure, etc. This knowledge should let the robot identify the
-intentions of other agents and create accordingly joint goals. Hence, a
-knowledge representation system aiming at dealing with cooperative behaviours
-is likely to have goal management structures taking explicitly into account
-other agents' actions and goals.
+is not directly perceived.
+
+Also central to cooperation are the notions of \emph{joint intentions} and
+\emph{joint goals}: to help the human during the cooking session, the robots
+need to track how far they are into the recipe, what is the next step the human
+is likely to go for, how tasks are currently split between agents, what action
+is currently blocking the procedure, etc. This knowledge should let the robot
+identify the intentions of other agents and create accordingly joint goals.
+Hence, a knowledge representation system aiming at dealing with cooperative
+behaviours is likely to have goal management structures taking explicitly into
+account other agents' actions and goals.
 
@@
 capable of: \emph{capability introspection} (both in term of general capability
-and of immediate ability) is thus often desirable. More general introspection
-(like the ability to tell ``who I am'' or ``what do I think of'') does not
-appear to be necessary in our scenario. This may however be required in more
-general human-robot interaction.
+and of immediate ability) is thus often desirable. It can be extended to
+general introspection (like the ability to tell ``who I am'' or ``what do I
+think of'') that may be required for the interaction.
 
 Last but not least, our scenario assumes implicitly \emph{natural interaction}
-between humans and robots (as showed by the casual style of the
-order ``Let's make a brownie!''). While natural language processing {\it
-per-se} is usually out of the scope of a knowledge representation system, we
-may want to be sure these systems may successfully interoperate, \ie that the
-knowledge representation system provides efficient support to the natural
-language processing module (for instance by adopting models and vocabulary that
-are both well suited for machine processing and remain as close as possible to
-the humans own structures and vocabulary).
+between humans and robots (as shown by the casual style of the order ``Let's
+make a brownie!''), and we want to ensure that the
+knowledge available to the robot provides efficient support to the natural
+language understanding (for instance by adopting models and vocabulary that are
+both well suited for machine processing and remain as close as possible to the
+humans own structures and vocabulary), and also to \emph{non-verbal forms of
+communication}, like gestures.
 
@@
 
-\subsubsection{Main Logic Formalisms}
+This first axis of analysis is its intrinsic expressive power. It answers the
+question: what can be possibly represented. When it explicitly exists, the
+\emph{language} of representation plays here an obvious role.
+
+\subsubsection{Main logic formalisms}
 
 The main role of a knowledge representation system is to provide an adequate
-representation system to store facts and concepts that can be informally
+representation system to formally store facts and concepts that could be informally
 described in natural language.
@@
 
-Most (but not all) of the systems we survey rely on a particular logic
+Most (but not all) of the systems we have surveyed rely on a particular logic
 formalism. The choice of the formalism has a strong impact, on one side, on the
@@
 expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
-side, on the ability to solve the inference problem (called
+side, on the ability to solve the fundamental inference problem (called
 \emph{satisfiability}: is a given logical sentence true in my model?) in a
@@
 
-A large number of logic formalism do exist, we shall summarize below the most
-relevant ones for systems actually deployed in current robotic architectures.
+A large number of logic formalisms do exist and we briefly present below the most
+relevant ones for systems actually deployed in robotic architectures.
 
@@
 
-The family of \emph{Description Logics}~\cite{Baader2008} also play an
+The family of \emph{Description Logics}~\cite{Baader2008} also plays an
 important role. It is also a subset of the first-order logic, with some
@@
 
-Because Description Logics have been originally created from the perspective of
+Because description logics have been originally created from the perspective of
 a \emph{knowledge representation language} and not a logic language, their
 terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
-\emph{individual},\ldots) is well-suited to knowledge description and we may use
-it in the remaining of this paper outside of the strict context of  Description
-Logics.
-
-\emph{Modal logic}, that allow for statement qualification like
+\emph{individual},\ldots) is well-suited to knowledge description.
+
+\emph{Modal logic}, that allows for statement qualification like
 \emph{possibility} or \emph{necessity}, have been shown to be closely related
@@
 
-\fxfatal{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
+\fxwarning{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
 
 \emph{Temporal logic} are designed to represent and manipulate assertions whose
-truth value may vary in time.
+truth value may vary in time. We introduce one of its key idea (the
+\emph{fluents}) later in the typology.
 
@@
 their own, and practical considerations (especially the availability of
-reasoners efficient enough for on-line use on a robot) often constraint the
+reasoners efficient enough for on-line use on a robot) often constrain the
 choice of a logical formalism and a level of expressive power.
 
-\subsubsection{Expressive Power}
+\subsubsection{Expressive power}
 
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{expressive_overlap_dl_horn.pdf}
+    \includegraphics[width=0.3\columnwidth]{expressive_overlap_dl_horn.pdf}
     \caption{expressiveness overlap of Description Logics and logic programs
@@
 logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
-mention extensions to description logics based on rule systems to bring closer
+mention extensions to description logics based on rule systems that bring closer
 the two approaches.
@@
 
-\subsubsection{Open World and Close World Assumptions}
+It can be noted that the relation between expressiveness and reasoning
+complexity is fragile: for instance, adding the following axiom \stmt{farFrom
+disjointProperty near} (that states that two individuals can not be at the same
+time near and far from each other) changes the expressiveness power of the ORO
+Common-Sense ontology (presented at chapter~\ref{chapt|oroserver}) from
+$\mathcal{SHOIQ(D)}$ to $\mathcal{SROIQ(D)}$\footnote{See
+appendix~\ref{chapt|dl} for a brief explanation of this notation}: this
+seemingly innocuous assertion change the complexity class of the whole
+ontology, and the concept satisfiability reasoning problem switches from a {\it
+NExpTime-complete} problem to a {\it NExpTime-hard} problem (\ie, \emph{at
+least} as hard as the hardest problem in {\it NExpTime}).
+
+This ``instability'' has practical consequences on run-time performances on
+the robot because a light alteration of the knowledge structure can lead to
+very noticeable performance drops.
+
+\subsubsection{Open world and close world assumptions}
 \label{sect|owa-cwa}
 
-The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumption names a
+The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumptions name a
 modelling choice on the \emph{completeness} of a knowledge domain. In the close
@@
 
-This distinction is important in robotics were the robot may have to manipulate
-concepts with only partial knowledge on them. For instance, let imagine a robot
+This distinction is important in robotics where the robot may have to manipulate
+concepts with only partial knowledge on them. For instance, let us imagine a robot
 that sees a bottle on a table, whose bottom is hidden by another object. The
@@
 several approaches exists to \emph{locally close} a domain (\cf
-Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those).
+Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those ones).
 
@@
 
+\fxfatal{Representation of uncertainty: To be completed!}
+
 \subsubsection{Meta-cognition: knowledge on the knowledge}
@@
 
+Sloman proposes in~\cite{Sloman2011} a detailed analysis of meta-cognition and
+its different aspects in both natural (human) and artificial systems.
+
 A knowledge representation system endowed with \emph{meta-cognition} is not
@@
 structure of its knowledge and the reasoning process. For instance, the ability
-to explain a logical inconsistency in a KRS is a meta-cognitive function.
-Sloman~\cite{Sloman2011} \fxfatal{...Sloman?}
+to explain a logical inconsistency in a KRS is a meta-cognitive function, as is
+the ability to expose and alter the knowledge structure (these two reasoning
+techniques have their own entries in the taxonomy, at
+section~\ref{sect|reasoning}.
 
@@
 
-\subsubsection{Role Representations}
+\subsubsection{Role representations}
 
 This section discusses strategies and approaches to represent in a knowledge
-model three important \emph{roles}\fxwarning{Reference on the term 'role'?}.
+model three important \emph{roles}: spatial relations, time and representation
+of actions.
 
@@
 Symbolic representation of space is a widely studied topic. In particular, a
-large literature corpus is available on spatial ontologies.
+large literature corpus is available on spatial ontologies\fxwarning{Which one?}.
 
@@
 
-...topology??
+Topological maps are abstractions of an environment as graphs where nodes
+represent places and edges represent connections between places. Because they
+are symbolic representation, topological maps allow higher-level reasoning
+(such as containment, connectivity, regions) than metric maps.
+
+One important contribution to the building of a coherent representational stack
+for space representation is the Spatial Semantic Hierarchy~\cite{Kuipers2000}
+introduced by Kuipers. It consists in multiple interacting representations of
+space, both qualitative and quantitative, that span from sensor-level
+representation to ontologies of places and regions.
 
@@
 represent relative spatial relations. Options include the computation of such
-relations only on-demand, when the context is know (which viewpoints, etc.) or
+relations only on-demand, when the context is known (which viewpoints, etc.) or
 storage of these facts in different models, one per agent.
@@
 \paragraph{Representation of time}
+\label{sect|time-representation}
 
 As an agent acting at human-like time scale and dealing with temporal concepts
-(like actions), a robot may want to represent, and possibly to reason, about
+(like actions), a robot needs to represent and reason about
 time. Time representation is split into two distinct abilities: representing
@@
 timestamps to events the robot perceives -- and in the future), and
-representing \emph{passing time} (situations, durations, timespans) like in
+representing passing time (situations, durations, timespans) like in
 \emph{``the eggs will be cooked in 10 min''}.
 
-This introduces the disjunction in time representation between
-\emph{situations}, that is, any temporal concept with a non-zero duration, and
-\emph{events}, that is, any durationless temporal concept.
-
-Other definitions of a situation do exist, notably in the context of
-\emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque consider a
-situation to be an history of actions.
-
-\fxfatal{Discuss time chronicles~\cite{Ghallab1996}}
-\fxfatal{Discuss Allen intervals}
-
-The concept of \emph{fluent} plays another important role for time
-representation. Fluents are properties (or \emph{conditions}) that change over
-time, like $sees(agent1, apple, t)$.
-
-We call a system that does not account for time (\ie that permanently lives in
-present) \emph{atemporal}.
+This is usually formalised as a disjunction in time representation between
+\emph{events}, that is, any durationless temporal concept, and
+\emph{situations}, that is, any temporal concept with a non-zero
+duration\footnote{Other definitions of a situation do exist, notably in the
+context of \emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque
+consider a situation to be an history of actions.}.
+
+Numerous techniques to represent and reason about time have been devised.
+Amongst the most significant ones, we can mention Allen's interval
+algebra~\cite{Allen1984} and Ghallab's chronicles~\cite{Ghallab1996}.
+\fxwarning{We could detail more chronicles and allen algebra...}
+The concept of \emph{fluent} also play an important role for time
+representation: fluents are properties (or \emph{conditions}) that change over
+time, like $sees(agent1, apple, t)$.\fxwarning{A bit short... citation?}
+
+We call a system that does not account for time (\ie that mentally permanently
+lives in present) \emph{atemporal}.
 
@@
 
-While representation of events boils down to label a timestamp,
-representation of actions are more complex since they semantically imply
-interactions between concepts. From a taxonomy point of view, actions are a
-particular type of event that normaly leads to a situation corresponding to the
-action realization.
+While representation of events boils down to label a timestamp, representation
+of actions are more complex since not only they deal with durations, but they
+also imply semantic interactions between concepts. From a taxonomy point of
+view, actions are a particular type of event that normally leads to a situation
+corresponding to the action realisation.
 
@@
 acted upon, the destination are some example of common thematic roles.
-
 Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
@@
 \end{center}
-\caption{Thematic Roles~\cite{Aarts1997}.}
+
+\caption{A list of thematic roles, as proposed by Aarts~\cite{Aarts1997}.
+Depending on the action, only certain roles are meaningful.}
+
 \label{table|theta-roles}
@@
 
-In the context of robotic, a \emph{task} represents the abstract model of an
-action. Tasks usually specify at least the conditions required to performed the
-action and the consequences of the realization of the action, and are central
-to the decisional layers of the robot, in particular for the planning,
-monitoring and execution control activities. Knowledge representation systems
-may thus have specific mechanisms to represent them (and possible, to reason
-about them, as presented in section~\ref{sect|planning}).
-
-%An extensive report on task modeling in OWL ontologies in appendix
+In the context of robotic, \emph{task} often represents an aggregate of
+(atomic) actions. We use the term here as the representation of the abstract
+model of an action. Tasks usually specify at least the conditions required to
+performed the action and the consequences of the realization of the action, and
+are central to the decisional layers of the robot, in particular for the
+planning, monitoring and execution control activities. Knowledge representation
+systems may thus have specific mechanisms to represent them (and possible, to
+reason about them, as presented in section~\ref{sect|planning}).
+
+Plans representation is closely related to action and task representation. It
+is specifically discussed at section~\ref{sect|planning}.
 
@@
 
-\emph{Knowledge is contextualized information}\fxfatal{Find someone respectable
-how said that :-)}: it is essential for the robot to associate the facts it
-represents to a \emph{context}. The context carries the keys for the
-interpretation of the information. It covers the \emph{domain of validity} of
-the facts, the \emph{common-sense} knowledge required to fill the gaps in the
-representation\fxfatal{give an example}, \fxfatal{What more?}.
-
-\subsubsection{Modality, Contingency and Theory of Mind}
+Knowledge is contextualized information: it is essential for the robot to
+associate the facts it represents to a \emph{context}. The context carries the
+keys for the interpretation of the information and set a common ground for
+interaction (and, in particular, communication~\cite{Jakobson1960}). It
+implicitely defines the domain of validity of the facts and carries the
+\emph{common-sense} knowledge required to fill the gaps of the knowledge
+explicitely shared between the agents.
+
+Context is generally difficult to recognise, represent and reuse because it is
+multiform and largely implicit nature. It is also never unique: at a given
+moment, several contexts, at different temporal, spatial, social scales,
+overlap.
+
+In the current literature in robotics and cognitive architectures, the term
+\emph{context} usually simply refers to a set of beliefs that initiate a
+representation (and reasoning) frame: in~\cite{Lemaignan2011a}, the robot
+creates a context of interaction with a specific human by storing in a separate
+model the beliefs of this human and using this knowledge when dialoging with
+the human, the reasoning network is reinitialised in the GLAIR architecture
+when the hypotheses that defined the current situation are not believed
+anymore~\cite{Shapiro2009}.
+
+This acception of \emph{context} is simplistic, and omit the overlapping and
+multi-scale aspects of context modeling.
+
+We see context representation as one of the main challenge of knowledge
+representatio in general, and we will further discuss the importance and issues
+brought by context modeling in the conclusion of the thesis.
+
+\subsubsection{Modality, contingency and theory of mind}
 \label{sect|possible-worlds}
@@
 A \emph{contingent} proposition is defined as neither always true (a tautology)
-nor always false (a contradiction) in every possible world. In knowledge
-representation systems for robotics that support logical modality,
-interpretations are often initialised with a common set of initial beliefs
-(like common-sense knowledge). This initial common knowledge is hence true in
-every possible worlds for the robot, and thus does not belong to its contingent
-knowledge.
-
-Logical modality allows to implement alternative mental models. They may be
+nor always false (a contradiction) in every possible world: its truth value
+depends on the context. In knowledge representation systems for robotics that
+support logical modality, interpretations are often initialised with a common
+set of initial beliefs (like common-sense knowledge). This initial common
+knowledge is hence true in every possible worlds for the robot, and thus does
+not belong to its contingent knowledge.
+
+On the contrary, alternative mental models with contingent knowledge may be
 used to represent different (possibly hypothetical or even imaginary) views on
-the world, from the robot own perspective, or from other persepectives computed
-by the robot.
+the world, from the robot own perspective or context, or from other
+perspectives computed by the robot.
 
 The representation of the mental perspective of other agents has a particular
-importance in human-robot interaction. It relies first on the ability to view
-the world from a standpoint which is not \emph{egocentric}. This cognitive
-ability is referred as \emph{perspective taking} (see
-Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} for the psychological
-grounds of perspective taking, itself originated in Piaget's work on cognitive
-development, as remembered by Flavell), and has been studied by
-infants~\cite{Moll2006} and robots as well~\cite{Trafton2005, Breazeal2006,
-Ros2010}.
+importance in human-robot interaction. It relies first on the ability to
+literally \emph{view} the world from a standpoint which is not
+\emph{egocentric}. This cognitive ability is referred as \emph{perspective
+taking}. Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} define the
+psychological grounds of perspective taking, that are themself originated in
+Piaget's work on cognitive development (recent studies on infants include
+Moll~\cite{Moll2006} for instance). Perspective taking begins to be studied on
+robots as well~\cite{Trafton2005, Breazeal2006, Ros2010}.
 
@@
 \emph{false beliefs}, \ie, facts that are believed to be true for an agent, but
-false for other ones. The ``Sally and Ann'' experiment
+false for other ones. The \emph{Sally and Ann} experiment~\cite{Leslie2000}
 (figure~\ref{fig|sally-ann}) is a classical example of a false belief
@@
 
-Scassellati~\cite{Scassellati2002} is amongst the researcher having studied
-theory of mind for robots.
-
-\subsubsection{Introspection: Who am I? What can I do?}
+Scassellati~\cite{Scassellati2002} is one of the first to have implemented a
+theory of mind on a humanoid robot.
+
+\subsubsection{Self-knowledge: Who am I? What can I do?}
 \label{sect|introspection}
 
-\paragraph{Introspection}
-
-Introspection is the ability to self-describe: what are my capabilities, what
-is my state (performing some action, idling, etc.), what are my beliefs (\ie my
-knowledge), what are my intentions and my plans?
+\paragraph{Self-knowledge}
+
+\emph{Self-knowledge} is the term used in philosophy and psychology to describe
+the knowledge that an individual has, acquires or infers about itself through
+its experiences. It answers the question ``What am I like?''.
+
+Introspection is the technical meaning to access to self-knowledge by the
+ability to self describe: what are my capabilities, what is my state
+(performing some action, idling, etc.), what are my beliefs, what are my
+intentions and my plans?
 
@@
 represented as a simple instantiation of a specific category (for instance, if
-the robot give an object to the human, this state could be represented with the
-triples \setstmt{robot performs action1, action1 isA Give}.
-
-\paragraph{Modelling of the robot capabilities}
-
-A particularly important aspect of introspection relates to the description of
-its own capabilities: which sensors/actuators/computation services exist and
-are currently available ?  While at a first level, these descriptions can be
-static (\eg the robot has one laser scanner and two arms), at more advanced
-levels, the description is updated and reflect the current (and possibly past
-and future) state of the robot. Note that these description may also involve
-geometric descriptions (a kinematic chain, the pose of a device, etc.) that may
-be deported outside of the main knowledge base. Efforts trying to formalize,
-maintain and expose the capabilities and state of a robot are not new (and
-ground themselves in work and techniques for self-descriptive remote procedure
-calls in computing science), but take a renewed importance with applications
-for high-level multi-robot cooperation. Recent work in that direction
-include~\cite{Kunze2011}.
+the robot gives an object to the human, this state could be represented with the
+triples \setstmt{robot performs action1,action1 isA Give}.
+
+\paragraph{Modeling of the robot capabilities}
+
+A particularly important aspect of self-knowledge for robots relates to the
+description of its own capabilities: which sensors/actuators/computation
+services exist and are currently available ?  While at a first level, these
+descriptions can be static (\eg the robot has one laser scanner and two arms),
+at more advanced levels, the description is updated and reflect the current
+(and possibly past and future) state of the robot. Note that these descriptions
+may also involve geometric descriptions (a kinematic chain, the pose of a
+device, etc.) that may be deported outside of the main knowledge base. Efforts
+trying to formalize, maintain and expose the capabilities and state of a robot
+are not new (and ground themselves in work and techniques for self-descriptive
+remote procedure calls in computing science), but take a renewed importance
+with applications for high-level multi-robot cooperation.
+
+Recent works by Kunze et al.~\cite{Kunze2011} seeks at defining a formal
+language to represent the capabilities of a robot.
 
@@
 Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
-and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organizes
+and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organises
 the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
@@
     \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
-    \caption{Overview of the main types of memories}
+    \caption{Overview of the main types of memories, based
+    on~\cite{Atkinson1968, Anderson1976, Tulving1985, Baddeley2010}}
+
     \label{fig|memory_models}
@@
 
-It worth emphasing that if memory is commonly associated to the process of
-forgetting facts after a variable amount of \emph{time}, it covers actually
+It is worth emphasising that if memory is commonly associated to the process of
+forgetting facts after a variable amount of \emph{time}, it actually covers
 more mechanisms that are relevant to robotics, like selective remembering
@@
 Baddeley's definition). Some systems may propose persistent storages that allow
-the robot knowledge to grow over time, while other may offer a larger range of
+the robot knowledge to grow over time, while others may offer a larger range of
 memory categories, like short term memory (that lasts for a couple of seconds)
@@
 architecture~\cite{Lehman2006} is one of those that tries to reproduce a
-human-like memory organization.
+human-like memory organisation. The GLAIR cognitive architecture also have a
+concept of long term/short term and episodic/semantic memories.
 
 %%%%%%%%%%
-\subsection{Reasoning Techniques}
+\subsection{Reasoning techniques}
 \label{sect|reasoning}
@@
 \subsubsection{Standard reasoning techniques}
-
-Being based on logical languages to represent the knowledge, most of the
-systems we survey allow various forms of \emph{reasoning}.
+\label{sect|std-reasoning}
 
@@
 
-Reasoners can often provide other type of inferences, like:
+Inferences can also be drawn from other constructs, whose availability depends
+on the representation language. OWL, for instance, has constructs for:
 
 \begin{itemize}
-    \item class subsumption (similar to inheritance)
+    \item class subsumption (to represent inheritance relations)
 
@@
         domain of the role \concept{thinksOf} is known to be
-        \concept{ThinkingAgent}, then \concept{thinksOf}$(a, b) \to
+        \concept{ThinkingAgent}, then \par \concept{thinksOf}$(a, b) \to
         $\concept{ThinkingAgent}$(a)$),
@@
 
-        \item property characteristics (inverse, symmetry, transitivity, etc.)
+        \item several second-order predicates (inverse, symmetry, transitivity, etc.)
 
@@
 
-    \item complex concept expressions like: \par \footnotesize \concept{Bottle}
-    $\equiv$ \concept{Artifact} {\bf that} (\concept{hasShape} {\bf value}
-    \concept{cylinderShape})\footnote{This example uses the \emph{Manchester
-    syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}} \normalsize
-
-    \item set operations like: \par \footnotesize \concept{Color} $\equiv \bigcup$ (\concept{blue}, \concept{green}, \concept{orange},
-    \concept{black}, \ldots) \normalsize
+    \item \emph{class restrictions} like: \par 
+	\footnotesize 
+	\concept{Bottle} $\equiv$ \concept{Artifact} {\bf that} (\concept{hasShape}
+	{\bf value} \concept{cylinderShape})\footnote{This example uses the \emph{Manchester
+    syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}}
+	\normalsize
+
+    \item set operations like: \par 
+    \footnotesize 
+    \concept{Color} $\equiv \bigcup$ (\concept{blue}, \concept{green},
+    \concept{orange}, \concept{black}, \ldots) 
+    \normalsize
 
@@
 
-\paragraph{Rule Languages}
-
-As mentioned earlier, knowledge models based on description logics can be extended through rule languages.
-
-Intersection of properties is an example of expression that can only be represented with rules: for instance,
-
-generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
-        \footnotesize \concept{looksAt(?agt, ?obj)} $\land$
-        \concept{pointsAt(?agt,?obj)} \par $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
+\paragraph{Rule Languages} As mentioned earlier, knowledge models based on
+description logics can be extended through rule languages (typically for OWL,
+the SWRL language).
+
+An intersection of properties is an example of expression that can only be
+represented with rules. For instance:
+
+\footnotesize
+\concept{looksAt(?agt, ?obj)} $\land$ \concept{pointsAt(?agt,?obj)}
+$\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
 \normalsize
 
-\subsubsection{Dynamic Instantiation and Alteration of the Knowledge Structure}
+\subsubsection{Dynamic instantiation and alteration of the knowledge structure}
 
@@
 \emph{structural part} that defines the conceptualisation of a domain in term
-of vocabulary and relations within the concepts, and an \emph{instantiation} of
+of vocabulary and relations between the concepts, and an \emph{instantiation} of
 this structure into concrete entities.
 
-The terms TBox and ABox are commonly found to describe this two different types of
-statements in ontologies: TBox statements describe a system in terms of
-controlled vocabularies, for example, a set of classes and properties. ABox are
-TBox-compliant statements about that vocabulary.
-
-Most knowledge representation systems allow to modify the ABox, which can be
+The terms TBox and ABox are commonly found to describe these two different types
+of statements in ontologies: TBox statements describe a system structure (made,
+for example, of a set of classes and properties) whereas the ABox contains
+TBox-compliant statements that are asserted in this structure.
+
+Knowledge representation systems allow to modify the ABox, which can be
 considered as the dynamic part of the knowledge base. Alteration of the ABox
-include the addition or retractation of relations between existing instances,
+include the addition or retraction of relations between existing instances,
 and the addition or removal of instances. We call the later \emph{dynamic
 instantiation}: the capability for a system to create new instances at
-run-time. The term dynamic instantion applies primarily to concrete entities
+run-time. The term dynamic instantiation applies primarily to concrete entities
 (typically, a new object is discovered by the robot, a symbolic instance is
@@
 
-KRS may also allow to alter the TBox. This requires the underlying reasoning
-systems to be able to dynamically take into account structural changes in the
-knowledge base.
+The knowledge representation system may also allow to alter the TBox. This
+requires the underlying reasoning systems to be able to dynamically take into
+account structural changes in the knowledge base.
 
 Example of TBox alteration include modification of the taxonomy (like addition
-or retractation of a \concept{subClassOf} relation), changes to the asserted
-domain or range of a predicate, addition or retractation of rules.
+or retraction of a \concept{subClassOf} relation), changes to the asserted
+domain or range of a predicate, addition or retraction of rules.
 
@@
 \emph{Lazy evaluation} describes the ability for a KRS to delay active
-knowledge acquisition or resoning operations until the value is actually
+knowledge acquisition or reasoning operations until the value is actually
 needed.
@@
 strategy (the relative placements of object is an example of combinatory
-explosion in strict evaluation approaches, although this issue is mitigated in
-real use-cases by heuristics that select a subset of objects to evaluate).
+explosion in strict evaluation approaches, although this issue could be
+mitigated in real use-cases by heuristics that would select a subset of objects
+to evaluate).
 
@@
 set of beliefs of the robot. This limits for instance the ability for the robot
-to react to logical conditions that involves fact that are lazily evaluated
-(``trigger a callback when object A is behind object B'' could be an example of
-a event triggered by a logical condition).
+to react to logical conditions that involve facts that are lazily evaluated
+(``trigger a callback when object A is behind object B'' would not be triggered
+with a purely lazy evaluation strategy for spatial relations).
 
@@
 
-\subsubsection{(Non) Monotonic Reasoning}
+\subsubsection{(Non) monotonic reasoning}
 
@@
 birds, but clearly we can think up as many additional exceptions like birds
-with their feet encased in concrete as we like. Formalized non-monotonic
+with their feet encased in concrete as we like. Formalised non-monotonic
 reasoning provides a way of saying that a bird can fly unless there
@@
 
-Another important application of non-monotonic reasoning is representation of
-change: for example, to make an omelette, you need to crack eggs and wipe them.
-The eggs disappear and are replaced by an omelette:
-
-\concept{Egg}$(a) \land $ \concept{Egg} $(b) \land $
-\concept{MakeOmelette}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \land \lnot $
-\concept{Egg}$(b) \land $ \concept{Omelette}$(c)$
-
-The insertion of the proposition \concept{MakeOmelette}$(a, b, c)$ leads to
+An important application of non-monotonic reasoning is the representation of
+change: for example, to make a brownie, one needs to crack eggs and mix them to the chocolate.
+The eggs disappear and are replaced by a dough:
+
+\concept{Egg}$(a) \land $ \concept{Chocolate} $(b) \land $
+\concept{MakeDough}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \land \lnot $
+\concept{Chocolate}$(b) \land $ \concept{Dough}$(c)$
+
+The insertion of the proposition \concept{MakeDough}$(a, b, c)$ leads to
 retraction of other facts. This rule requires non-monotonic reasoning to be
@@
 
-\emph{Default logic} is one of the formal logic that account for representing
+\emph{Default logic} is one of the formal logics that account for representing
 general truth and exceptions to it (for instance, \emph{tomatoes are red, in
-general}). However, due to computational complexity of these model (most of
+general}). However, due to computational complexity of these models (most of
 inferences in default logic are known to be $NP$-complete problem), classical
@@
 For instance, the SWRL rule language, usually associated to the OWL-DL ontology
-language, do not allow non-monotonic reasoning (only so-called DL-safe rules
+language, does not allow non-monotonic reasoning (only so-called DL-safe rules
 are allowed).
 
-\fxfatal{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
+\fxwarning{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
 languages? reasoner?}
@@
 
-\fxfatal{Give here an example of non-monotonic reasoning with Prolog}
-\fxwarning{Do we mention here Answer Set Programming?}
-
-A monotonic system does not theoretically allow for knowledge retraction,
-which is an important issue in the robotic context where the world model is
-likely to be often altered.  However it is a practical issue only if the
-reasoning process is \emph{continuous} during the whole robot's activity
-lifespan. It is often possible to stop the reasoner, alter the knowledge, and
-restart the inference process on a new domain.
-\fxfatal{Rephrase to emphasize that when new evidences appear, it is anyway often a
-good idea to restart the reasoner.}
-
-\fxfatal{Mention that the 'change of world' issue can also be dealt with appropriate
-time representation.}
-\fxfatal{Mention that probabilistic reasoning lead to implicit non-monotonic reasoning}
-
-\subsubsection{Presupposition Accommodation}
+\fxwarning{Give here an example of non-monotonic reasoning with Prolog}
+\fxnote{Do we mention here Answer Set Programming?}
+
+\subsubsection{Presupposition accommodation}
 \label{sect|presupposition-accommodation}
 
-\emph{Presupposition accommodation} is the ability for the system to
-automatically create a context allowing to make sense of a proposition.
-
-For instance, we can imagine a human telling a robot \emph{Please get me the
-bottle that is behind you}. If the robot has not yet see what is behind it, it
+\emph{Presupposition accommodation}~\cite{VonFintel2008} is the ability for a
+system to automatically create a context allowing to make sense of a
+proposition.
+
+Applied to robotics, we can imagine a human telling a robot ``Please get me the
+bottle that is behind you''. If the robot has not yet see what is behind it, it
 needs to assume (and represents in its knowledge model) that a undefined bottle
@@
 
-Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection
-task} and the \emph{legality task}.
+Levesque~\cite{Levesque2008} distinguishes two main tasks related to reasoning
+on actions and consequences of actions, the \emph{projection task} and the
+\emph{legality task}.
+
+We call \emph{diagnosis} the converse of the projection task: the ability to
+track back the origin of a decision, and \emph{explanation} the more general
+ability to explicit a reasoning or a decision.
+
+\begin{scriptsize}
+\begin{center}
+\begin{tikzpicture}[taxonomy]
+    \node [taxon] {\bf C.6 Prediction and Explaination}
+            child {node [taxon] {{\bf D.4} Explaination}}
+            child {node [taxon] {{\bf D.3} Diagnosis}}
+            child {node [taxon] {{\bf D.2} Legality}}
+            child {node [taxon] {{\bf D.1} Projection}};
+\end{tikzpicture}
+\end{center}
+\end{scriptsize}
 
 \paragraph{Projection task}: determining whether or not some condition while
-hold after a sequence of actions.
+hold after a sequence of actions. The projection task is a typical
+non-monotonic reasoning task, since at each step, the system must add but also
+retract beliefs, as defined in the tasks post-conditions.
 
@@
 in case of failure to provide possible explanation. This can be seen as the
-temporal reverse of the projection task. Because of their modelisation of
+temporal reverse of the projection task. Because of their modelling of
 situations as an history of actions, derivatives of the GOLOG logic programming
-languages are a good example of diagnosis tasks integrated to the knowledge
+languages are a good example of the diagnosis task integrated to the knowledge
 representation system~\cite{Gspandl2011}.
@@
 \emph{justification} capabilities that may be offered by the knowledge
-representation system. The explanation of an entailment is the sequence (or
-set of sequences if several are possible) of reasoning steps that allow to
-conclude. An explanation can also conversely explain why a statement leads to a
-contradiction.
+representation system. The explanation of an entailment is the sequence (or set
+of sequences if several are possible) of reasoning steps that allow to reach a
+conclusion. An explanation can also conversely explain why a statement leads to
+a contradiction.
 
@@
 observe that an inconsistency is triggered. The reasoner provides the four
-following observations\footnote{This example uses the \emph{Manchester syntax},
-\url{http://www.w3.org/TR/owl2-manchester-syntax/}} to explain the
-inconsistency:
+following observations to explain the inconsistency:
 
@@
 \begin{enumerate}
-    \item \concept{myself} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
+    \item \concept{robot1} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
     \item \concept{belongsTo} {\bf Domain} \concept{Artifact}
-    \item \concept{myself} {\bf Type} \concept{Agent}
+    \item \concept{robot1} {\bf Type} \concept{Agent}
     \item {\bf DisjointClasses:} \concept{Agent}, \concept{Artifact}
@@
 This explanation directly portrays the underlying structure of the knowledge
-model. A KRS may provide mechanisms to expose this kind of analysis, either
-automatically when inconsistencies survene, or on demand.
+model: we can understand that a robot is modelled here as an agent, that agents
+and artifacts are disjoint classes and that only artifacts can belong to
+someone, hence the inconsistency. A KRS may provide mechanisms to expose this
+kind of analysis, either automatically when inconsistencies occur, or on
+demand.
 
@@
 robot in unexpected semantic situations, and forms a first step towards an
-adquate reaction.
+adequate reaction.
 
@@
 
-\subsubsection{Task Planning}
+\subsubsection{Task planning}
 \label{sect|planning}
@@
 Symbolic task planning in general is a large research
-field\cite{Russell2009planning}. The so-called Classical Planning Problem,
+field~\cite{Russell2009planning}. The so-called Classical Planning Problem,
 first, is characterised by a unique known initial state, durationless
@@
 representation formalisms dedicated to representation and reasoning about
-actions and situation have emerged: besides situation calculs, \emph{fluent
+actions and situation have emerged: besides situation calculus, \emph{fluent
 calculus} and \emph{event calculus} are the main ones.
@@
 The GOLOG~\cite{Levesque1997} language and its derivatives (like {\sc
-ReadyLOG}~\cite{Ferrein2008} and {\sc IndiLOG}~\ref{Gspandl2011}) propose
+ReadyLOG}~\cite{Ferrein2008} and {\sc IndiLOG}~\cite{Gspandl2011}) propose
 implementations of the situation calculus focused on robotic applications.
 
-It is however common to rely on external components dedicated to symbolic task
+It is however also common to rely on external components dedicated to symbolic task
 planning (in particular because of the non-monotonicity requirements) with a
-tight link to the KRS for domain retreival and/or resulting plan storage.
+tight link to the KRS for domain retrieval and/or resulting plan storage.
+
+Finally, in the context of interaction between several agents, the management
+of \emph{joint intentions} and \emph{joint goals}~\cite{Tomasello2005,
+Dominey2011} are additional aspects that have to be represented and
+appropriately handled by the planning subsystem.
 
@@
 Some systems \cite{Kunze2011a} rely on external dedicated physics engine to
-compute symbolic facts from on-demand physics simulation. This, however,
-requires a tight integration between the symbolic model and a geometric model
-that carries the geometries and physical properties of objects.
+compute symbolic facts from on-demand physics simulation. This requires a tight
+integration between the symbolic model and a geometric model that carries the
+geometries and physical properties of objects.
 
@@
 
+\fxfatal{Learning section to write}
+
 %%%%%%%%%%%%%%%%%
-\subsection{Acquiring Knowledge}
+\subsection{Acquiring knowledge}
 
@@
 
-\fxwarning{What to say here?}
+Exteroceptive sensing is the most obvious and largely studied mean of knowledge
+acquisition. Traditional sensing devices (IR, cameras, laser-scanners), while
+still present and widely used (navigation based on 2D localisation and obstacle
+avoidance is today the standard), are being step-by-step replaced by
+\emph{synthetic sensors}.
+
+These synthetic sensors include post-processing to provide higher-level
+percepts that ease the grounding. The prototypical example of such a device is
+the Kinect sensor. At a first level, it replaces traditional stereo vision
+algorithms by providing a fast, robust depth map. At a second level, it
+provides accurate, real-time tracking recognition and tracking of whole body
+poses of human.
+
+Other examples of such synthetic sensors exist: face recognition, off-the-shelf
+performant SLAM solutions, automatic cluster segmentation in point clouds (with
+the PCL library), etc.
+
+While the progresses of these sensing technologies are remarkable, one field of
+perception remain a very difficult challenge: accurate and generic object
+recognition. Most of the current approaches to object recognition rely on a mix
+of point cloud segmentation and fitting with visual feature recognition
+(SIFT-like algorithms), but it remains a slow and fragile task.
 
@@
 communication. Communication is associated to specific functions
-(figure~\ref{fig|communication-model}), and in particular, it implies a shared
+(as shown on Jakobson's diagram, figure~\ref{fig|communication-model}), and in particular, it implies a shared
 \emph{context} (usually implicit) between interactors.
@@
 The field of verbal interaction processing for robots spans from pattern-based,
-constrainted sentences recognition to natural, bidirectional, unconstrained
-verbal communication. Kruijff et al. provides in~\cite{Kruijff2010} an
-up-to-date survey of literature on situated human-robot dialogue, focusing on
-formal representation systems, bi-directionality of the interaction and context
-building.
-
-Natural Language Processing (NLP) is a large research field by itself, and
-while the robotic community may be lagging behind on many theoretical aspects,
-it brings one important aspect: the embodiment. Because the interactors, both
-the robot and the human, are establishing a communication within a shared
-physical context, the verbal communication channel is complemented by deictic
-channels, back channels and possibly shared physical experiences: a human can
-show something to a robot, saying ``Give me this''. This is simply not possible
-for a virtual agent.
+constrained sentences recognition to natural, bidirectional, unconstrained
+verbal communication. A large literature corpus exists on Natural Language
+Processing (NLP) which is presented at section~\ref{sect|dialogs-related-work},
+page~\pageref{sect|dialogs-related-work}.
+
+NLP is an established research field by itself, and while the robotic community
+is still lagging behind on many theoretical aspects, it brings one important
+aspect: the embodiment. Because the interactors, both the robot and the human,
+are establishing a communication within a shared physical context, the verbal
+communication channel is complemented by deictic channels, back channels and
+possibly shared physical experiences: a human can show something to a robot,
+saying ``Give me this''. This is not possible for a virtual agent.
 
@@
 interaction is also an established field of research in human-robot
-interaction. Common deictic forms of communication include attentional focus
-(via face and gaze tracking) and joint attention, pointing, emotional
-expressions (based on face expressions, postures, emotional gestures).
+interaction. Common deictic forms of communication~\cite{Li2012} include
+attentional focus (via face and gaze tracking) and joint attention, pointing,
+emotional expressions (based on face expressions, postures, emotional
+gestures).
 
@@
 
-Robots, and in particular service robots, can access (with sometimes some
-security-related constraints) the World Wide Web and remote knowledge stores.
+Robots, and in particular service robots, have usually an access (with possibly
+security-related constraints) to the World Wide Web and remote knowledge stores.
 
@@
 InfoBox), and the vast majority of the knowledge actually contained in the
-encyclopedia pages remains out of reach of automated parsers. Efforts like {\sc
-ReVerb}~\cite{Fader2011} work however at automatic extraction of triples by
-examining subject-verb-object relations in plain text.
+encyclopedia pages remains out of reach of automated parsers. Efforts in that
+direction however exist~\cite{Nyga2009, Fader2011}.
 
@@
 project~\cite{Singh2002}. The project is designed to let the general public
-easily add common-sense statements in semi-controlled natural language, that is
+easily add common-sense statements in semi-controlled natural language, which is
 then processed to a publicly available ontology.
 
-An other approach that is easier at short-term to let robots remotely access
+Another approach that is easier at short-term to let robots remotely access
 knowledge repositories consists is building robot-specific shared repositories,
@@
 
-\emph{Grounding} (also called \emph{anchoring} when specifically referring to
-building links between \emph{percepts} and \emph{physical
+\emph{Grounding} (also called \emph{anchoring} when specifically referring to the
+building of links between \emph{percepts} and \emph{physical
 objects}~\cite{Coradeschi2003}) is the task consisting in building and
@@
 
-\subsubsection{Intrinsic Motivation and Curiosity}
-
-The reason for a robot to acquire knowledge are various, and usually external
-to the robot itself. It is often driven by the needs of the tasks the robot has
-to execute. In this case, motivations are managed by the execution controler
-and are mostly unvisible to the knowledge representation system.
+\subsubsection{Intrinsic motivation and curiosity}
+
+The reasons for a robot to acquire knowledge are diverse, and usually external
+to the robot itself. It is often driven by the requirements of tasks that the
+robot has to execute. In this case, motivations are managed by the execution
+controller and are mostly invisible to the knowledge representation system.
 
 However, motivation can also be intrinsic, driven by the internal state of the
-robot, without external pressure or reward. In this case, the
+robot's beliefs, without external pressure or reward. In this case, the \fxwarning{to be completed}
 
@@
 %%%%%%%%%%%%%%%%%
-\subsection{Practical Integration in Robotic Architectures}
+\subsection{Practical integration in robotic architectures}
 \label{sect|integration-robot}
@@
 systems because they do not stand as independent modules but exist as diffuse,
-\emph{ubiquitous} knowledge manipulation system~\cite{Jacobsson2008}, for
-instance because they are primarily language~\cite{Ferrein2008, Sabri2011}.
+\emph{ubiquitous} knowledge manipulation system (case of the CAST knowledge
+model~\cite{Jacobsson2008}), for instance because they are primarily
+language~\cite{Ferrein2008, Sabri2011}.
 
@@
 It is common to have knowledge representation systems at the heart of a
-cognitive robotic architecture, and therefore KRS are easily ``burried'' in the
+cognitive robotic architecture, and therefore KRS are easily ``buried'' in the
 system.
@@
 
-That is the reason why having at hand good tools to trace and visualize at
-run-time the evolution of the knowledge structure and contents, as well as
-post-processing tools that can be run on the trace to precisely analyze the
-cognitive behaviour of the robot, is useful.
+Tools to trace and visualise at run-time the evolution of the knowledge
+structure and contents may be available with the KRS, as well as
+post-processing tools that run on the trace to analyze {\it a posteriori} the
+cognitive behaviour of the robot.
 
@@
 
-Benchmarks of symbolic systems for robots are hard to conduct for
-several reasons: identifying good metrics for robotic experiments in general is
+Benchmarks of symbolic systems for robots are hard to conduct for several
+reasons: identifying good metrics for robotic experiments in general is
 difficult because of the complex interactions between tenth of modules running
-in parallel, and isolating one specific component is difficult.
-Also, knowledge representation systems are often tightly coupled to
-the other modules, and the lack of standard API for knowledge services
-makes it especially hard to switch between KRS to compare them. Finally, because
-service robots are design to act in rich, dynamic environments, possibly with
-humans, building repeatable experiments is challenging.
+in parallel, and isolating one specific component is difficult.  Also,
+knowledge representation systems are often tightly coupled to the other
+modules. To quote Langley~\cite{Langley2006}:
+
+\begin{quote}
+The conventional wisdom of software engineering is that one should
+develop independent modules that have minimal interaction. In contrast, a
+cognitive architecture offers a \emph{unified} theory of cognition with tightly
+interleaved modules that support synergistic effects.
+\end{quote}
+
+The lack of standard API for knowledge services makes it also hard to switch
+between KRS to compare them.
+
+Finally, because service robots are designed to act in rich, dynamic
+environments, possibly with humans, building repeatable experiments is
+challenging, and quantitative measurements are often not the right
+metric~\cite{Langley2006}.
+
+We will however present here some quantitative metrics (related to scalability,
+for instance), followed by qualitative evaluation approaches, grouped under the
+term \emph{Cognitive Performances}.
 
@@
 $\mathcal{Q}^{\mathcal{M}'}$ the (averaged) time to execute a query of
-complexity $\mathcal{M}'$ on the KRS. The type $\mathcal{M}$ of model
+complexity $\mathcal{C}$ on the KRS. The type $\mathcal{M}$ of model
 alteration is either an ABox modification (addition/removal of an instance) or
 a TBox alteration (addition/removal of a class, a class restriction or a rule).
-Query complexity $\mathcal{M}'$ is...\fxwarning{find references}.
+%Query complexity $\mathcal{C}$ is...\fxwarning{find references}.
 
 \emph{Temporal scalability} is defined in term of the nature of the function
-$f^{\mathcal{M}, \mathcal{M}'}(\sigma) = \mathcal{R}^{\mathcal{M}}(\sigma) +
-\mathcal{Q}^{\mathcal{M}'}(\sigma)$ (\ie the relation of the relaxation time
+$f^{\mathcal{M}, \mathcal{C}}(\sigma) = \mathcal{R}^{\mathcal{M}}(\sigma) +
+\mathcal{Q}^{\mathcal{C}}(\sigma)$ (\ie the relation of the relaxation time
 and query time to the knowledge base size). \emph{Space scalability} is the
-relation of memory consumption to the size. The scalability is tightly coupled
-to the expressiveness of the underlying knowledge model, which need to be known
-for the scalability measurement to be meaningful.
-
-Note that robotic applications may have a knowledge base size relatively small
-compared to the complexity of its TBox. In that case the relaxation and query
-times for a constant, average knowledge size $\sigma$ may be more relevant as a
-metric for KRS raw performance.
+relation of memory consumption to the size $\sigma$. The scalability is tightly
+coupled to the expressiveness of the underlying knowledge model, which need to
+be known for the scalability measurement to be meaningful.
 
@@
 performances of KRS are often benchmarked with synthetic datasets (which leads
-to issues: how to assess the meaningfulness of the performance of a reasoner on
-an artificial ontology?~\cite{Bail2010}) or toy experiments~\cite{Chong2009}
-that do not always model the whole complexity of real-world application.
+to another issues: how to assess the meaningfulness of the performance of a
+reasoner on an artificial ontology?~\cite{Bail2010}) or ``toy'' experiments
+that do not always model the whole complexity of real-world
+application~\cite{Chong2009}.
 
@@
 knowledge representation systems in a relevant manner may be difficult, the cognitive
-performances of the robot as a whole can be also evaluated through
-``classical'' tests from the psychology (like False-Belief
-experiments~\cite{Leslie2000} or the Token test~\cite{DiSimoni1978}).
-
-However, the performance evaluation of current tools is mostly qualitative and
-scenario-specific (``did my system do the job for my task?''), and amongst the
-surveyed KRS, only few systems proposes an evaluation of their performances
-that could be compared to other system.
-
+performances of the robot as a whole can be also evaluated.
+
+Langley et al.~\cite{Langley2006} propose five such dimensions of evaluation:
+the \emph{generality} of the system (can it adapt easily to new tasks?), the
+\emph{rationality} or relevant of the inference/reasoning/decisions the system
+take, the \emph{reactivity} and \emph{persistence} that evaluates if the
+behaviour of a cognitive system is appropriate under unpredicted changes, the
+\emph{improvability} of the system as a function of the knowledge added to it,
+and finally, the resulting \emph{autonomy} of the system.
+
+Cognitive performance can also be evaluated with the support of tools developed
+in cognitive psychology. Several standard tests (like False-Belief
+experiments~\cite{Leslie2000} or the Token test~\cite{DiSimoni1978}) have been
+used to judge the cognitive abilities of robots~\cite{Mavridis2006,
+Breazeal2006}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \node [taxon] {\bf F. Knowledge instantiation}
-            child {node [taxon] {{\bf F.5} Self-Knowledge}}
-            child {node [taxon] {{\bf F.4} Metrics}}
-            child {node [taxon] {{\bf F.3} Granularity}}
+            child {node [taxon] {{\bf F.4} Granularity}}
+            child {node [taxon] {{\bf F.3} Metrics}}
             child {node [taxon] {{\bf F.2} Common-sense and Alignement}}
@@
 
-While we have previously mentioned features of knowledge representation system
+While we have previously mentioned features of knowledge representation systems
 that enable the robot to fill its knowledge base with content and alter the
@@
 
-\subsubsection{Design Strategy}
+\subsubsection{Design strategy}
 \label{sect|design-strategies}
 
-Ontology engineering has grown into a research field of its own, covering
-formalized methodologies for ontology creation, reuse, modularisation, etc.
-
-The creation
-
-Two main approaches are common in the cognitive robotic community:
-\emph{top-down} design and \emph{bottom-up}.
-
-\subsubsection{Common-sense and alignment with Standard Upper-ontologies}
+The main challenge of knowledge representation can be summarised as \emph{How
+to model the real-world state and interactions in a symbolic way, processable
+by the robot to make decisions}. We have already introduced the term
+\emph{grounding} to describe the (bi-directional) process of binding percepts
+to symbols.
+
+We have also seen that the instantiation of the knowledge (\ie, the actual,
+practical knowledge available to the robot) comes either from some variant of
+perception plus grounding, or from knowledge that the developer considers as
+already meaningful for the robot: remote semantic databases or initial
+(common-sense or situation specific) knowledge.
+
+This translates into two main strategies to drive knowledge instantiation: a
+\emph{top-down} design or a \emph{bottom-up} design.
+
+A bottom-up approach to knowledge instantiation takes the output of sensors as
+the primary source of knowledge: instances of objects, agents, and their
+relations directly result from what is perceived. No abstract, ``from the more
+generic to the more specific'' process takes place.
+
+Steels~\cite{Steels2007} considers this approach to be a solved issue, and
+according to Sloman~\cite{Sloman2007} (and his stance against the
+\emph{``Symbol Grounding meme''}), since bottom-up grounding boils down to grounding
+of \emph{somatic} concepts (\ie roughly, the sensori-motor relationships that
+the robot learns from its interaction with the world), it constrains in an
+unacceptable way the range of concepts accessible to the robot.
+
+Knowledge instantiation can also be approached as a \emph{top-down} activity:
+in natural language grounding, for instance, the robot needs to automatically
+bind an abstract representation (a group of words uttered by a human) to an
+unambiguous, context-dependent, internal concept. This concept may (or may not)
+be \textit{a priori} available to the robot as a pre-loaded ontology (what we
+previously called the cultural background of the robot). In any case, the robot
+must conduct a cognitive process that leads from an abstract concept to a
+concrete entity.
+
+The bottom-up and top-down strategies also reflect how the knowledge structure
+itself is constructed: either by successive classification and refinement of
+percepts, or from generic categories, typically extracted from standard
+upper-ontologies.
+
+\subsubsection{Common-sense and alignment with standard upper-ontologies}
 
@@
 
-\subsubsection{Granularity}
-
-The knowledge \emph{granularity} qualifies the level of details or refinement
-of the knowledge stored.
-
-...
-
-The storage of literal values is a particular case: while knowledge
-representation systems have been considered until now mostly as symbolic
-systems, they can often represent literal values (in particular, numerical
-values) as well.
-
-Depending on the representation language, literal values can be naturally
-represented and processed (common case in logic programming language) or not
-(storing numerical value, let alone matrices, in OWL is cumbersome).
-
-\subsubsection{Metrics}
-
-Qualitative and quantitative metrics give an insight on the size and
-expressive complexity of ontologies.
-
-Table~\ref{table|upper_onto_metrics} gives such metrics for four major
+Building {\it a priori} knowledge with a top-down strategy leads to populate
+the upper part of the taxonomy with abstract concepts like \emph{Thing},
+\emph{Time}, \emph{Person}, etc. The organisation of the whole knowledge
+depends on the design of this abstract part of the common-sense knowledge.
+
+Different knowledge structures at this level can lead to serious
+misunderstanding: for instance, if one robot considers the concept of a
+\emph{person} as representing some intelligent, possibly disembodied, entity,
+and another robot represents a \emph{person} as a subclass of mammals, their
+models of the world are likely to be conflicting.
+
+To be able to successfully exchange knowledge with other systems (robots,
+databases, natural language parsers, etc.), the common-sense knowledge of
+robots is thus often aligned with a standard upper-ontology.
+
+Many such upper-ontologies exist (table~\ref{table|upper_onto_metrics} lists
+some of them). While {\sc Cyc} and SUMO are current the main two, many efforts
+take place to make these ontologies compatible with each other (in particular
+by using the {\sc WordNet} thesaurus as intermediate unambiguous source of
+semantics).
+
+\subsubsection{Metrics and quality criteria}
+
+Qualitative and quantitative metrics give an insight on the size, complexity
+and effectiveness of ontologies used with the robots.
+
+Table~\ref{table|upper_onto_metrics} gives such metrics for five major
 upper-ontologies commonly used in the semantic Web community (taken
-from~\cite{Mascardi2007} and the projects' respective websites).
+from~\cite{Mascardi2007} and the projects' respective websites). It is
+interesting to note the large variations between projects like {\sc DBpedia}
+whose content is automatically generated from Wikipedia, {\sc Cyc} or {\sc
+SUMO}, which are both hand written, but do not adopt the same strategy to
+select the knowledge to represent.
+
+These metrics do not reflect adequately the \emph{expressive complexity},
+though: in most of these ontologies with a large amount of terms and
+assertions, taxonomic relations (\concept{isA}) or technical predicates (URI,
+translations, etc.) account for a large part of the assertions, at the expense
+of real semantic relations.
 
@@
 \toprule
-{\bf Project} & {\bf Terms} & {\bf Assertions} \\
+{\bf Project} & {\bf Terms} & {\bf Assertions} (triples) \\
 \midrule
-Cyc & > 300 000 & > 3 000 000 & \\
+{\sc Cyc} & > 300 000 & > 3 000 000 & \\
 YAGO & > 10 000 000 & > 120 000 000 \\
 SUMO & 20 000 & 60 000 \\
-DBPedia (for English) & 1 840 000 & 385 000 000\\
-OpenMind Common Sense (for English) & & 1 000 000\\
+{\sc DBpedia} (for English) & 1 840 000 & 385 000 000\\
+{\sc OpenMind} Common Sense (for English) & & 1 000 000\\
 \bottomrule
@@
 \end{center}
-\caption{Selection of metrics for the major upper-ontologies.}
+\caption{Raw size of major upper-ontologies.}
 \label{table|upper_onto_metrics}
@@
 
-\subsubsection{Self-Knowledge}
-
-``Self-knowledge'' is the term used in psychology to describe the knowledge
-that an individual has, acquires or infers about itself through its
-experiences. It answers the question ``What am I like?''.
-
-For a robot to have at its disposal self-knowledge is closely link to the
-introspective capability that we have previously mentioned
-(section~\ref{sect|introspection})
+Other metrics that include the type of predicates that are used, or the
+computed DL expressiveness, can be more significant (some are presented in
+table~\ref{table|onto-stats}, page~\pageref{table|onto-stats}).
+
+Qualitative evaluation of ontologies is also a well studied field. In Staab's
+{\it Handbook on Ontologies}, Vrande{\v{c}}i{\'c}~\cite{Vrandevcic2009}
+provides a synthesis of qualitative criteria for ontology assessment.  He lists
+eight of them: \emph{accuracy, adaptability, clarity, completeness,
+computational efficiency, conciseness, consistency,} and
+\emph{organizational fitness}. Details and relevant literature can be found
+in Vrande{\v{c}}i{\'c}'s chapter.
+
+\subsubsection{Granularity}
+
+Amongst the characteristics of ontologies, knowledge \emph{granularity}
+qualifies the level of details or refinement of the knowledge stored. The level
+of granularity of a robot's ontology hints on the place of the symbolic layer
+in the whole robotic architecture: some systems (like OMRKF~\cite{Suh2007}) go
+as down as storing SIFT features (\ie a large volume of numerical values) in
+the ontology. The storage of literal values is indeed a relevant case for
+robotics. Depending on the representation language, literal values can be
+naturally represented and processed (common case in logic programming language)
+or not (storing numerical value, let alone matrices, in OWL is cumbersome and
+not efficient due to the serialisation to XML).
+
+The issue of the granularity of models can also be partially addressed by
+splitting the knowledge representation into a geometric level (where all
+numerical values are stored) and a purely symbolic level. The communication
+between the two layers is however a complex question.
+
+It must finally be noted that complex robotic systems are likely to use
+different level of the knowledge depending on the task to achieve, and the
+granularity of the knowledge should probably be considered as a dynamic
+property.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Surveyed systems}
+\section{Existing systems for knowledge representation in service robotics}
 \label{sect|surveyed-systems}
 
-Table \ref{table|surveyed-systems} presents the eight knowledge representation
-systems surveyed in this article.
+Table \ref{table|surveyed-systems} lists the knowledge representation
+systems that we have surveyed.
 
@@
 \begin{landscape}
-\begin{table}\footnotesize
+\begin{table}\scriptsize
 \begin{center}
 
-\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{2.8cm}p{1.5cm}}
+\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{1.5cm}}
 \toprule
-{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & {\bf Reasoner} & Main reference \\
+{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
 \midrule
-ARMAR/Tapas & KRS & Holzapfel, Waibel \par (Karlsruhe TH) & & ? & TFS (Typed Feature Structures) & & \cite{Holzapfel2008}\\
-CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & & \cite{Jacobsson2008} \\
-GSM & KRS & Mavridis, Roy \par (MIT MediaLab) & & ? & & & \cite{Mavridis2006} \\
-Ke Jia Project & KRS & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP & ASP & ASP & \cite{Chen2010} \\
-{\sc KnowRob} & KRS & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL & Custom \par ({\sc Prolog}) & \cite{Tenorth2009a} \\
-NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & & \cite{Sabri2011} \\
+ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
+CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & \cite{Jacobsson2008} \\
+GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & & \cite{Mavridis2006} \\
+Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
+{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
+NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & \cite{Sabri2011} \\
 %OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
-OUR-K & KRS & Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & Horn Clauses & ? & \cite{Suh2007} \\
-ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
-PEIS KR\&R & KRS & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & & \cite{Daoutis2009} \\
+OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
+%ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
+PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
 %Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
@@
 
-\caption{List of surveyed systems. Projects are listed by category (\emph{KRS}
-for systems that are explicit knowledge representation and reasoning modules,
-\emph{Ubiquitous} for systems where knowledge processing is fully distributed,
-\emph{Language} for languages used as KRS on robots), then names.}
+\caption{List of surveyed systems. Categories are \emph{Formal} for systems
+that have a formal underlying knowledge representation, \emph{Ubiquitous} for
+systems where knowledge is fully distributed, \emph{Language} for languages
+used as KRS on robots or \emph{Structural} for KRS where knowledge is
+represented as special data structures.}
 
@@
 {\sc Tapas} is the name of the knowledge representation system and dialogue
-manager found on the ARMARIII robot~\cite{Holzapfel2008}.
+manager found on the ARMAR~III robot~\cite{Holzapfel2008} for the Karlsruhe
+Institute of Technology.
 
@@
 \emph{lexical knowledge} contains lexical and grammatical informations about
-the objects. The \emph{semantic knowledge} is organized into an ontology
-relying on \emph{typed feature structures} (TFS,~\cite{Carpenter1992}, a
+the objects. The \emph{semantic knowledge} is organised into an ontology
+relying on \emph{typed feature structures} (TFS~\cite{Carpenter1992}, a
 formalism originating from the computational linguistics community, and a
@@
 Another focus is put on handling unknown words and objects. {\sc Tapas}
-provides routines to recognize unknown entities, and propose and interactive
-and iterative verbal process to categorize (including adding new categories)
+provides routines to recognise unknown entities, and propose and interactive
+and iterative verbal process to categorise (including adding new categories)
 those new concepts.
 
-\subsection{CAST Knowledge model}
+\paragraph{Experiments} {\sc Tapas} has been used experimentally in a kitchen
+environment where naive users had to ask the robot for an object and get
+information about another object.
+
+\subsection{CAST knowledge model}
 \label{sect|cast}
@@
 comprehensive toolkit aimed at building cognitive architectures for robots
-through a set of interconnected SAs (\emph{subarchitectures}). The CAS does not
-expose a central knowledge base as seen in previous works. It instead
-represents knowledge as unrooted \emph{proxies}. Those proxies are formally
+through a set of interconnected \emph{SA} (\emph{subarchitectures}). The CAS
+does not expose a central knowledge base as seen in other works. It represents
+instead knowledge as unrooted \emph{proxies}. Those proxies are formally
 defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
@@
 A union of proxies forms a global amodal representation of an entity, that can
-be explicitly shared and manipulated. Being not centralized, the knowledge
+be explicitly shared and manipulated. Being not centralised, the knowledge
 model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
@@
 In the presented implementation, the CAST knowledge model does no allow for
-effectively representing actions or temporal information.\fxfatal{What about
+effectively representing actions or temporal information.\fxwarning{What about
 reasoning? can they retrieve for example 'all proxies for colorful objects'?}
 
+\paragraph{Knowledge Acquisition} Several techniques for knowledge acquisition
+have been explored within the CAST framework. Cross-modal knowledge
+fusion~\cite{Hawes2007a} is well studied, and the interaction with natural
+language processing~\cite{Kruijff2010, Kruijff2010a} is a particular emphasise
+of the project.
+
 In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
-the context of spatial representations.
+the context of spatial representations with the robot \emph{Dora}.
+
+\paragraph{Experiments} CAST has been used in several experiments, including
+table-top manipulation (with a focus on language understanding) and more
+recently on the Dora robot~\cite{Hawes2011} for indoor exploration.
 
@@
 
-Unlike most of the previously presented systems, GSM does not rely on any
-formal language but rather on a layered data structure (Figure~\ref{fig|gsm})
-that organizes the surrounding world into agents and relations between agents.
-Each agent (any animate or inanimate object) is attached to a physical model (made
-of \emph{body parts} that have properties like their position, color, etc.) and
-a mental model (which is a recursively embedded GSM, thus allowing a sort of
-theory of mind).
+GSM does not rely on any formal language but rather on a layered data structure
+(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
+relations between agents.  Each agent (any animate or inanimate object) is
+attached to a physical model (made of \emph{body parts} that have properties
+like their position, color, etc.) and a mental model (which is a recursively
+embedded GSM, thus allowing a sort of theory of mind).
 
@@
     \caption{Simplified hierarchical structure of the Grounded Situation Model,
-    based on~\cite{Mavridis2006}}
+    based on~\cite{Mavridis2006}.}
 
@@
 and interaction tasks (a ``conversational helping hand'' as stated by the
-authors) realized on a 7-DOF arm equipped with force feedback, cameras for blob
+authors) implemented on a 7-DOF arm equipped with force feedback, cameras for blob
 tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
@@
 psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.
-
-\subsection{GOLOG}
-\label{sect|kejia}
 
@@
 involving human-robot interaction with natural language. These tasks include a
-task with multiple \emph{pick \& carry} that are globally optimized, naive
+task with multiple \emph{pick \& carry} that are globally optimised, naive
 physics reasoning via taught rules or more complex scenarii with the robot
@@
 developed at the Technical University of Munich. It is build as a set of
-modules (figure~\ref{fig|knowrob} organized around a core reasoning system
-written in Prolog. The core component interfaces through Java/Prolog or
+modules (figure~\ref{fig|knowrob} organised around a core reasoning system
+written in Prolog. This core module interfaces through Java/Prolog or
 C/Prolog APIs with external modules.
-
-{\sc KnowRob} offers a mechanism called \emph{computables} that allow to
-evaluate certain predicates by calling external dedicated functions (for
-instance, the valuation of a proposition like \stmt{object1 isOn object2} is
-computed when required by calling a specific geometric reasonning module).
 
@@
     \centering
-    \includegraphics[scale=0.36]{knowrob.pdf}
+    \includegraphics[width=0.7\columnwidth]{knowrob.pdf}
 
@@
 
-Extension modules plug into the system and provide specialized reasoning
-capabilities or interfaces to external data, \eg to read object detections from
-the vision system. These modules mainly operate on the level of instances
-(ABox).
+Extension modules can plug into the system to provide specialised reasoning
+capabilities or interfaces to external data sources, \eg to read object
+detections from the vision system. These modules operate on the level of
+instances (ABox).
 
@@
 based on bayesian networks, integration with naive physics reasoning has been
-studied~\cite{Kunze2011a}, automatic parsing of Web resource in semi-natural
-language has been experimented with~\cite{Tenorth2009b}.
-
-\paragraph{Grounding} Anchoring of knowledge in the {\sc KnowRob} framework
-relies on...CoP? Semantic Maps~\cite{Blodow2011}.
+studied~\cite{Kunze2011a}, automatic parsing of Web resources in semi-natural
+language has been also experimented with~\cite{Nyga2009}.
+
+\paragraph{Grounding} {\sc KnowRob} offers a mechanism called \emph{computables} that allow to
+evaluate certain predicates by calling external dedicated functions (for
+instance, the valuation of a proposition like \stmt{object1 isOn object2} is
+computed when required by calling a specific geometric reasoning module). In
+combination with Prolog's lazy evaluation strategy, this supports a good
+scalability.
+
+Computables rely on various subsystem to evaluate. In particular, it relies on
+the CoP framework~\cite{Klank2009} and \emph{semantic maps}~\cite{Blodow2011}
+for the recognition of objects and environment.
 
@@
 
-\paragraph{Notable Experiments} {\sc KnowRob} has been deployed on several
-scenarii at the Technical University of Munich, on the PR2 robot and on a 2-arm
-custom mobile manipulator in the scope of the ``assistive
-kitchen''~\cite{Beetz2008} project. In these experiments, {\sc KnowRob} has
-been used as central knowledge hub...
+\paragraph{Experiments} {\sc KnowRob} has been deployed on several scenarii at
+the Technical University of Munich, on the PR2 robot and on a 2-arm custom
+mobile manipulator in the scope of the ``assistive kitchen''~\cite{Beetz2008}
+project. These experiments include retrieval and automated parsing of recipies
+from the Web, retrieval and manipulation of various kitchen tools, cooperation
+between two robots.
 
@@
 
-\paragraph{Intrinsic language features}
-\label{sect|nkrl-intrinsic-features}
-
-Expressiveness...
-
-\paragraph{Integration with physical world and in the robot architecture}
-\label{sect|nkrl-integration}
-
-\ldots seem to be mostly WIP\ldots
+\paragraph{Knowledge representation} The NKRL language semantics are stored in
+two ontologies: an ontology of concepts $\Omega$ and an ontology of events
+$\Psi$. The ontology of events is made of action or situation templates.
+Templates are a set of predicates (MOVE, PRODUCE, RECEIVE, EXPERIENCE, BEHAVE,
+OWN and EXIST) associated to thematic roles. Grounding and reasoning with NKRL
+is based on template matching.
+
+\paragraph{Experiments} The main scenario of development for NKRL-based robots
+is the Smart Home and monitoring of eldery people. Knowledge acquisition
+partially rely on ambient intelligence (RFID, pressure sensors in the chairs,
+etc.). The scenario is still being implementated.
 
@@
 
-The Ontology-based Unified Robot
-Knowledge~\cite{Lim2011} (OUR-K) framework, successor of the Ontology-based Multi-layered Robot Knowledge Framework~\cite{Suh2007}
-(OMRKF), is a knowledge representation system based on five inter-related
-\emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It proposes a layered
-approach to knowledge representation that allows to integrate the grounding
-process to the knowledge representation process. OUR-K knowledge model is
-implemented with a mix of Description Logics for the concept hierarchies and Horn clauses.
+The Ontology-based Unified Robot Knowledge~\cite{Lim2011} (OUR-K) framework,
+successor of the Ontology-based Multi-layered Robot Knowledge
+Framework~\cite{Suh2007} (OMRKF), is a knowledge representation system based on
+five inter-related \emph{classes} of knowledge (figure~\ref{fig|omrkf}). It
+proposes a layered approach to knowledge representation that allows to
+integrate the grounding process to the knowledge representation process. OUR-K
+knowledge model is implemented with a mix of Description Logics for the concept
+hierarchies and Horn clauses.
 
@@
 
-    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
+    \caption{OUR-K organises knowledge into fives \emph{classes}, each composed
     of \emph{levels}. Figure based on~\cite{Lim2011}.}
@@
 \paragraph{Representation} The environment is represented in OUR-K in the
-$space : Model$ knowledge level as a classical three layers mapping (metric,
-topological and semantic maps). Objects (in $object : Model$) are localized in
-$space : Model$ through Voronoi nodes.
+$spaces : Model$ knowledge level as a classical three layers mapping (metric,
+topological and semantic maps). Objects (in $objects : Model$) are localised in
+$spaces : Model$ through Voronoi nodes.
 
@@
 structure, exploited at run-time by a planner.
+
+\fxwarning{re-read the paper to improve the section...}
 
@@
 
-{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim to binds autonomous
-robotics with ambient intelligence (network of sensors). \emph{PEIS} stands for
-\emph{Physically Embedded Intelligent System}: every robots or intelligent
-device in the environment is abstracted as a PEIS.
-
-Each PEIS physical component is running a \emph{PEIS Kernel} instance. Communication
-between instance relies on a custom P2P communication protocol.
-
-The PEIS architecture allows for adding new abilities through software components sharing the common \emph{tuple space}.
-
-We survey here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS KR\&R}, that includes symbolic representation and reasoning.
+{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim
+to binds autonomous robotics with ambient intelligence (network of sensors).
+\emph{PEIS} stands for \emph{Physically Embedded Intelligent System}: every
+robots or intelligent device in the environment is abstracted as a PEIS.
+
+Each PEIS physical component is running a \emph{PEIS Kernel} instance.
+Communication between instance relies on a custom P2P communication protocol.
+
+The PEIS architecture allows for adding new abilities through software
+components sharing the common \emph{tuple space}.
+
+We consider here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS
+KR\&R}, that includes symbolic representation and reasoning.
+
+\fxwarning{have a look to latest Daoutis paper!}
 
@@
 % - object identification based on viewpoint independent SIFT features
-% - formalized anchoring system that explicitly match percieved attributes to predicates
+% - formalised anchoring system that explicitly match percieved attributes to predicates
 % - Cyc predicates
@@
 
-\fxfatal{Is modal logics and higher order logics actually used in PEIS?}
+\fxwarning{Is modal logics and higher order logics actually used in PEIS?}
 
@@
 
-\fxfatal{OWA/CWA?}
+\fxwarning{OWA/CWA?}
 
@@
 \paragraph{Knowledge Acquisition} The primary source for knowledge acquisition
-is perception.  The PEIS ecosystem provides a SIFT-based object recognizer used
-in conjunction with ceiling cameras for object localization.  Other perceptual
+is perception.  The PEIS ecosystem provides a SIFT-based object recogniser used
+in conjunction with ceiling cameras for object localisation.  Other perceptual
 modalities are available (like human tracking, ambient environment monitoring).
@@
 
-\paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
+\paragraph{Anchoring} Daoutis et al. formalise the issue of anchoring as
 finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
@@
 
-\paragraph{Notable experiments}
-\label{sect|peis-expe}
+\paragraph{Experiments} Experiments involving PEIS take place in a Smart Home
+environment (\emph{PEIS Home}). The implemented case studies explore
+dialogue-based interaction with the robot about known objects.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Table~\ref{table|contribution-by-systems} proposes a summary of the main scopes
-of \emph{contribution} of the systems we have surveyed. This overview helps to
-identify the main weaknesses of current approaches of knowledge representation
-regarding the long-term goal of \emph{human-level robots}.
+The table~\ref{table|contribution-by-systems} sketches the landscape of current
+research. The main fields of \emph{contribution} for the systems we have
+surveyed are sorted along the different axis, dimensions of knowledge
+representation that were identified in previous section.
 
 \begin{landscape}
-\begin{table}\footnotesize
+\begin{table}\tiny
 \begin{center}
-
-\begin{tabular}{cp{4.5cm}p{2.3cm}p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.5cm}p{2cm}p{1.5cm}p{1.5cm}}
+\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}|p{2cm}}
 \toprule
-\multicolumn{2}{c}{\bf Category} & {\sc KnowRob} & {ORO} & {\sc PEIS} & {\sc CAST} & {\sc Ke Jia} & {\sc NKRL} & {\sc GSM} & {\sc OUR-K} & {\sc ARMAR} & {\sc Hertzberg} \\
+\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}       & ORO \cite{Lemaignan2010}                      \\
 
@@
 
-\multirow{5}{0.7cm}[0.2cm]{\turn[1.5cm]{90}{\bf Expr. power}} & Logical formalism & Prolog & DL (OWL) & {\sc CycL} &  & ASP &  & & Horn clauses & TFS & \\
- & OWA/CWA & CWA & OWA & ? & ? & CWA & ? & ? & ? & ? & ? \\
- & Modeling uncertainty & +++ (extension) & & & & & & ++ (stochastic) & + (\emph{candidate} entities) \\
- & Meta-cognition & ++ & ++ & & & & & & \\
+\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL + Horn clauses             & {\sc CycL}                    & DL (OWL)                                      \\
+                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 &                               &                               & OWA                                           \\
+                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         & + (\emph{candidate} entities) &                               &                                               \\
+                                                                    & Meta-cognition &                           &                             &                             &                             & ++                                 & + (transformation rules)                        &                               &                               & ++ (reification, taxonomy walking)            \\
 \hline
-\multirow{6}{0.2cm}{\turn{90}{\bf Model}} & Space Representation & & & & & & & & ++  \\
- & Time representation & & & & & & & + (snapshots) & +  \\
- & Context & & & & & & & & ++  \\
- & Alternative worlds & & +++ & & & & ++ & &  \\
- & Introspection & + (SRDF) & & & & & & &  \\
- & Memory models & & + & & & & & &  \\
+\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++                            &                               & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})\\
+                                                               & Time representation &                           &                             & ++ (snapshots)              &                             &                                    & ++ (event oriented)                             & +                             &                               &                                               \\
+                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            & +++                           &                               &                                               \\
+                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          & ++                            & ++ (microtheories)            &                                               \\
+                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 &                               &                               & ++ (Theory of Mind \cite{Warnier2012a})       \\
+                                                                    & Self-knowledge &                           &                             &                             &                             & + (SRDF \cite{Kunze2011})          &                                                 &                               &                               &                                               \\
+                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 &                               & +                             & +                                             \\
 \hline
-\multirow{10}{0.2cm}{\turn{90}{\bf Reasoning}} & Standard FOL reasoning & +++ & +++ & & & +++ & & & +  \\
- & Knowledge structure alteration & & ++ & & & ++ & & & & ++ &  \\
- & Lazy evaluation & +++ & & & & & & &  \\
- & Reasoning under uncertainty & ++ (extension) & & & & & + & & \\
- & Non-monotonic reasoning & & & & & +++ & & &  \\
- & Presupposition accommodation & & & & & & & +++ & \\
- & Prediction, projection, diagnosis & & & & & & & & \\
- & Physics-based reasoning & +++ (extension) & & & & & & &  \\
- & Planning & + & +++ (external) & & & ++ & & & + \\
- & Learning & & + & & & & & & & ++ &  \\
+\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & +                             & ++                            & ++                                            \\
+                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 &                               &                               & ++ (TBox alteration)                          \\
+                                                                   & Lazy evaluation &                           &                             &                             &                             & +++ (Prolog, computables)          &                                                 &                               &                               &                                               \\
+                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           &                                    &                                                 &                               &                               &                                               \\
+                                                      & Presupposition accommodation &                           &                             & +++                         &                             &                                    &                                                 &                               &                               &                                               \\
+                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             &                                    & ++                                              &                               &                               & + (explanation)                               \\
+                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +                             & {\it++}                       & +++ ({\em HATP})                              \\
+                                                           & Physics-based reasoning &                           &                             &                             & +                           & {\it+++} ({\em naive physics \cite{Kunze2011a}}) &                                   &                               &                               &                                               \\
+%                                                                         & Learning & ++                        & ++ \cite{Jacobsson2008}     &                             &                             &                                    &                                                 &                               &                               &                                               \\
 \hline
-\multirow{3}{0.7cm}{\turn[1cm]{90}{\bf Knw. acq.}} & Acquisition techniques & & & & & & & & \\
- & Grounding & +++ \par (semantic maps) & ++ (external amodal model) & & & & & ++ & ++ \\
- & Intrinsic Motivation & & & & ++ & & & & \\
- & Dynamic instantiation & & & & & & & & \\
+\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 &                               & ++ ({\em ambient intelligence})& ++                                           \\
+                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 &                               & + (template-based)            & +++ ({\em Dialogs} \cite{Lemaignan2011a})     \\
+                                                                     & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{Nyga2009}})&                             &                               &                               &                                               \\
+                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & +++ (bottom-up)               & +++ \cite{Loutfi2008}         & ++ ({\em amodal model} \cite{Lemaignan2012})  \\
+                                                              & Intrinsic Motivation &                           & ++ \cite{Hawes2011}         &                             &                             &                                    &                                                 &                               &                               &                                               \\
 \hline
-\multirow{4}{0.2cm}{\turn{90}{\bf Integ.}} & \ldots with sensori-motor layers & & & & & & & + & + \\
- & \ldots with executive layers & & & & & & & + & ++ \\
- & NLP & & +++ (external) & & & +++ & & +++ & & +++ & \\
- & Monitoring and debugging & & & & & & & & \\
- & Performances & & & & & + & & +++ (Token test) & & & \\
+\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          & +                             & ++ (tuple space)              &                                               \\
+                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++                            & ++ (tuple space)              & ++ (semantic events)                          \\
+                                                          & Monitoring and debugging &                           &                             &                             &                             &                                    &                                                 &                               &                               & + ({\em remote visualisation})                \\
+                                                           & Performances evaluation &                           & ++ \cite{Hawes2008}         & ++ (Token test)             & +                           & + \cite{Tenorth2011}               &                                                 &                               &                               &                                               \\
 
@@
 \end{center}
-\caption{Main domain of contribution of each surveyed systems. Rating goes from +++ (major focus) to + (secondary interest). An empty cell means that 
-the system does not consider this domain.}
+
+\caption{Main domains of contribution of current KRS. Italics mean that the
+feature is implemented as an external module. Main references are given in the
+table header. When relevant, feature-specific publications have been provided.
+An empty cell means that either the system has no specific focus on this domain
+or we could not find relevant literature.}
+
 \label{table|contribution-by-systems}
@@
 
-\subsection{Expressiveness: what can we represent in the current state-of-the-art?}
-\label{sect|summary-expressiveness}
-
-\begin{itemize}
-    \item common-sense knowledge
-    \begin{itemize}
-        \item \emph{default reasoning}
-        \item \emph{uncertainty representation/reasoning}
-        \item \emph{physics-based reasoning}
-    \end{itemize}
-    \item Spatio-temporal representations
-    \item Meta-cognition and introspection
-    \item Context
-\end{itemize}
-
-\subsection{Creating and grounding knowledge in the physical world}
-\label{sect|summary-grounding}
-
-\fxnote{Mention symmetry and research of invariants \cite{Henderson2011} as a process to create knowledge?}
-
-\begin{itemize}
-	\item which knowledge modalities can be merged
-	\item which grounding strategies
-\end{itemize}
-
-\subsection{Exploiting knowledge}
-\label{sect|summary-knowledge-sources-reasoning}
-
-\begin{itemize}
-	\item which common-sense knowledge is used
-	\item what kind of reasoning is made
-	\item introspection
-\end{itemize}
-
-\subsection{Using a knowledge representation system in a larger robotic cognitive architecture}
-\label{sect|summary-integration}
-
-\begin{itemize}
-	\item a priori evaluation/lazy evaluation and trade-offs (relations to events, scalability, detection of inconsistencies,\ldots)
-	\item relation with planning (action representation, temporal representation\ldots)
-	\item API/bindings/integration into executive layers
-\end{itemize}
-
-\subsection{Current shortcomings: what is not successfully covered by current systems}
-
-\subsubsection{Assessing performances}
-
-The systems we have presented share the property to have real implementations
-on real, physical robots. This simple fact allows to first assert that these
-tools are at least viable solutions to the question: how to endow a (real)
-robot with explicit knowledge manipulation capabilities?
-
-From a theoritical point of view, this work hopefully brings elements of
-comparison and evaluation of the existing systems.
-
-However, this do not suffice to assess and comment the suitability of such
-systems to real-world tasks. As hinted in section~\ref{sect|performances}, we
-think that first, the practical evaluation of knowledge representation systems
-for robots can not take place in isolation (links with perceptual and
-decisional layers are too important), then we should evaluate the robot
-behaviour at the scale of the whole system, basing experiments on established
-protocols, borrowed from other disciplines from cognitive sciences. We have
-already mentioned classical experiments like False-Beliefs in psychology or the
-Token test in linguistic. A systematic review of standard cognitive tests
-applicable to service robots is still to be made.
+To comment this table, we propose select nine topic based on some of the
+challenges that McCarthy and Roy have identified (there are listed at
+section~\ref{sect|evaluation-literature}) on the road to natural interaction
+and, to slightly paraphrase McCarthy, ``human-level robots''.
+
+\subsection{Logic formalisms, continuous world}
+
+McCarthy underlines the necessity of relying on mathematical logic, as the most
+fruitful formalism for machine intelligence. This is largely the case. Almost
+all the systems we have surveyed rely on logical formalisms, in several cases
+as a mix of declarative languages (often based on Decription Logics) and
+logical programming languages (Prolog, of course, but not only).
+
+The two exceptions (the CAST knowledge model and GSM) are however also
+interesting: CAST proposes a pervasive model of knowledge that is seducing from
+the grounding perspective (but likely suboptimal for deliberative tasks), and
+GSM encodes knowledge in an amodal (continuous, geometric) model while
+preserving features that are usually specific of symbolic models (like theory
+of mind or categorical knowledge).
+
+The interleaving of discreet symbolic models with continuous geometric models
+remains a major challenge, and techniques vary a lot: besides the CAST proxies
+and the GSM amodal model, most systems directly extract physical attributes of
+the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
+OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
+adopt a top-down approach where local geometric models are set up on-demand to
+compute symbolic properties like relative locations.
+
+In our approach, ORO has tight (but unidirectional) links with SPARK, the
+component in charge of geometric modeling and reasoning. Similarly to the GSM
+approach, SPARK is an amodal model of the environment we can manipulate (3D
+motion planing, for instance takes place in MHP, the twin brother of SPARK).
+
+\subsection{Management of uncertainty and approximation}
+
+McCarthy also mentions the necessity to deal with \emph{approximate concepts
+and approximate theories} (that includes representing them and reasoning with
+them).
+
+The most notable attempt at modeling uncertainty down to the knowledge
+representation is {\sc ProgCog}, an extension of {\sc KnowRob}. The theoretical
+and practical performances (including decidability) remain however difficult to
+overcome.
+
+The challenge of representing and reasoning under uncertainty has not been
+tackled in ORO server, at least not \emph{within} the current ORO server
+knowledge model. Two reasons explain this absence: the decisional architecture
+developed at LAAS for the robots has not consistent approach to uncertainty
+representation and management, neither at the \emph{symbol production} level
+(\ie SPARK or {\sc Dialogs}) or at the symbol consumption level (execution
+control). Thus, no strong incentive pushed us in this direction. Then, the
+current tools available from the Semantic Web domain to manipulate and reason
+about knowledge do not provide mature support for uncertainty management. Some
+effort do exist (like the {\sc Pronto} reasoner~\cite{Klinov2008}), but this
+does not appear to be a major focus in the currently available tools.
+
+\subsection{Non-monotonic reasoning}
+
+McCarthy gives a particular importance to non-monotonic reasoning. Systems like
+Ke Jia explicitly tackle this issue.
+
+A monotonic system does not theoretically allow for knowledge retraction during
+the reasoning, which is an important issue in the robotic context where the
+world model is likely to be often altered.  However it is a practical issue
+only if the reasoning process has to be \emph{continuous} during the whole
+robot's activity lifespan, which is rarely the case. It is often possible to
+stop the reasoner, alter the knowledge, and restart the inference process on a
+new domain.
+
+Non-monotonicity can also be partially dealt with with appropriate time
+representation and reasoning (the reasoner then only takes into account
+statements that are set as valid for a given moment).
+
+Finally, probabilistic reasoning also implicitly leads to non-monotonic
+reasoning, by relying on a continuous description of the state of the
+world.
+
+Non-monotonicity is however of broader significance to the knowledge
+representation field, in particular for the representation of common-sense
+knowledge where \emph{default} representation is currently sorely lacking.
+
+It is also related to what McCarthy calls \emph{elaboration tolerance}: the
+ability to extend \emph{on demand} the closed domain of interpretation for a
+given assertion, to take into account new contextual knowledge.
+
+\subsection{Modeling of contexts}
+
+Both Roy and McCarthy mention the importance of formalizing and reasoning about
+contexts. Many of the KRS we have surveyed mention at some point the context
+modeling, but no consistent interpretation, let alone theory, of context
+management has clearly emerged.
+
+Several approaches for building contextualized knowledge have been presented in
+this thesis: symbolic environment interpretation, perspective taking and
+independent mental state for each agent, grounded natural language resolution,
+self-awareness of its own activity. Much remains to be done for a robot to
+actually identify its current context as well as contexts that may be referred
+to.
+
+We will further develop this important topic in the conclusion of the thesis.
+
+\subsection{Reasoning about time, events and actions}
+
+While representation of spatial roles (topology, placement, with or without
+perspective-taking) is well studied and KRS usually integrate with spatial
+perception and reasoning components, time representation and integration in
+knowledge systems is less uniform.
+
+The two main tasks that require time reasoning are task planing and sequence
+recognition (in particular, action recognition). Systems that directly
+integrate task planing (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
+fluents in {\sc KnowRob}) to represent time.
+
+Other approaches include storing snapshots of the knowledge state (like GSM),
+that is used to move back to past mental states (of the robot or of another
+agent).
+
+Integration of these techniques with memory mechanisms (forgetting and
+reinforcement learning) remains to explore.
+
+Temporal modeling is currently absent of ORO, and symbolic and geometric
+planning are accomplished outside of the main knowledge representation
+layer. As a matter of fact, knowledge is mostly atemporal in the ORO server: the
+set of triples stored in the server represents the beliefs of the robot
+\emph{at the current time}. The robot represents knowledge about the \emph{here
+and now}. Dealing with time reasoning is delegated to dedicated tools.
+
+However, ORO does provide a simple implementation of a memory mechanism that
+relies on time: when a statement is attached to a memory profile, the statement
+is reified and the date of creation is stored. This allows to remove the
+statement after a period of time that depends on the memory profile.
+
+Statement reification to store timestamps of creation for all statements is
+technically very easy to do in ORO server, but has not been actually enabled
+because the performance hit (each reified statement produces four triples
+instead of one) was not justified by any current use case of ORO server.
+
+Regarding events, several systems (GSM, NKRL, ORO) have adopted an {\em
+event-oriented} architecture where conditions (or templates) are used to
+trigger decisional and execution processes.
+
+A side note about how thematic roles and action models are initialized in ORO:
+The current implementation relies on a small, predefined set of action verbs
+that can be recognised from natural language
+(section~\ref{processing_of_actions}).  This constraint does not come from the
+resolution algorithm itself, but rather from the difficulty to automatically
+extract the thematic roles associated to a verb. An interesting yet easy to
+implement extension would consist in binding the dialogue processor with the
+symbolic task planner. The task planner could dynamically provide the list of
+actions that the robot can process, \ie actions for which the robot can produce
+a plan. In certain case, it could also validate the understanding of a desire
+by attempting to plan it.
+
+Also, we could exploit on-line resources like
+\textsc{VerbNet}~\cite{Kipper2008}, which provides a large machine-processable
+lexicon of English verbs along with their thematic roles.
+
+\subsection{Grounding in a multi-modal environment}
+
+Symbol grounding remains a focus for most of the developers of cognitive
+architectures for robots, and all the systems we have surveyed implement
+grounding strategies (...it was an inclusion criteria). Most of the systems
+adapt their grounding strategies to the sensing modality, so that grounding can
+not be considered as a single, well delimited process.
+
+Some systems (ORO, GSM) introduce an intermediate step in the grounding process
+by the mean of an amodal model of the environment that aggregates the
+perceptions (or suppositions) in a single place. This enables geometric
+reasoning that takes into account all the perceptual modalities (typically
+required for spatial perspective taking: we need to know where the humans are
+looking at, and also where are the objects), and also improve the observability
+of the system.
+
+We note in addition that the development of what we have called ``synthetic
+sensors'' is simplifying the grounding task. The most obvious example is the
+human tracking system provided by the low-cost Kinect device, and now used
+pervasively in the robotics lab. Not only this system segments and computes the
+pose of humans that enter its field of view, but it also tracks them (including
+in case they are partially or completely occluded). This kind of high level
+sensing device does not completely remove the need of grounding one symbolic
+instance of a human with the physical human, but the task is ways simpler than
+it used to be.
+
+Many systems also tackle the difficult question of natural language processing.
+The CAST middleware, in particular, has been used in the European CoSy and CogX
+projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
+the huge interest for the grounding of verbal interaction within the cognitive
+robotics community, and we are likely to reach important milestones in this
+field in the coming years.
+
+Merging more modalities (especially, back-channel communication, deictic
+gestures and social gazes) also sparks a lot of attention, and is becoming more
+and more present as symbolic knowledge available to the control layers.
+
+\subsection{Common-sense}
+
+McCarthy actually starts his list by affirming that intelligent systems must be
+able to \emph{``operate successfully in the common sense informatic
+situation''}. This question of the \emph{common-sense} is probably one of the
+toughest because what \emph{common-sense} mean is not very clear at first place
+(since, by definition, common-sense is, well, common-sense...). We feel,
+however, that it is related to a diffuse cultural background, and one may even
+claim that, for a system to acquire common-sense reasoning is equivalent to
+solve the strong AI challenge (this is at least more or less McCarthy's opinion).
+
+Our knowledge representation systems have a pragmatic and potentially very
+powerful approach to common-sense: reuse knowledge stored on the Web. While not
+so many of the systems we have surveyed directly tackle this question ({\sc
+KnowRob} is the only one explicitly working on common-sense reasoning, through
+physics-based reasoning, integration with Web databases or parsing of
+semi-structured Web documents), several rely however on Web standards (OWL,
+OpenCyc) to represent their knowledge. With the development of initiatives like
+OpenMind that encode with the same standards common-sense facts and rules,
+there is a strong potential for our robot to gain common-sense knowledge from
+well-structured online resources in the coming years.
+
+%%%%%%%%%% Underlying knowledge model table %%%%%%%%
+\begin{table}
+\begin{center}
+
+\begin{tabular}{ll}
+\toprule
+{\bf Project} & {\bf Common-sense knowledge source} \\
+\midrule
+ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
+CAST Proxies &  None \\
+GSM &  Predefined categories \\
+Ke Jia & None \\
+{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology, physics simulation \\
+NKLR &  None \\
+OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
+ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
+PEIS Ecology & {\sc ResearchCyc} \\
+
+\bottomrule
+
+\end{tabular}
+\end{center}
+\caption{Underlying common-sense knowledge sources for each project.}
+\label{table|knowledge-sources}
+\end{table}
+
+Table~\ref{table|knowledge-sources} lists for the systems we have surveyed the
+current sources of common-sense knowledge.
+
+\subsection{Learning, representation of experience, introspection}
+
+Learning has not been addressed in this work. This is a complex, still emerging
+field, whose exact meaning and perimeter varies a lot. It was difficult to
+synthesise in the KRS taxonomy, and we decided to omit it. We did not either
+really tackled this question in ORO, except in term of knowledge structure
+alteration based on verbal interaction (this is presented in one case-study, in
+the next section).
+
+Learning at the level of the knowledge representation system has been explored
+in conjunction with visual perception systems (mentioned for instance, in the
+CAST project, in~\cite{Jakobsson2008}). {\sc KnowRob}, by filling its pool of
+facts from informations automatically extracted from the Web~\ref{Nyga2009},
+can also be considered as a \emph{learning} system (it autonomously acquire
+knowledge). We lack however a formal study of learning strategies and
+techniques at the knowledge representation level.
+
+Representation and matching of past experience is a related topic. This ability
+is a key step for general action recognition, and is of particular importance
+for the robot to assess the state of the interaction with the human.
+
+While we already mentioned that several systems are able to reason on past
+states, we are not aware of existing implementation of algorithms to reflect on
+past experiences.
+
+This is itself related to introspection and meta-cognition: the shift towards
+explicit knowledge representation exemplified in the nine systems we have
+presented, has a major impact on the meta-cognitive capabilities of our robots.
+They can exhibit, manipulate and reason on their internal belief state (what we
+will call in the conclusion the \emph{cognitive observability}). We have still
+to discover all the possibilities that are open by this important cognitive
+ability.
+
+\subsection{Perspective-awareness}
+
+Finally, Roy mentions \emph{the ability to take into account the human
+perspective}: this cognitive ability, that relates to the representation of
+different modalities, is present in ORO and GSM. ORO explicits the perspectives
+in different symbolic mental states while GSM recursively stores models for
+each agent.
+
+We have already explained how perspective awareness enables advanced cognitive
+capabilities like a theory of mind, and this is probably the main contribution
+of this thesis regarding techniques for knowledge representation.
+
+% \subsection{Justification and resolution of cognitive conflicts}
+% 
+% Logical inconsistencies in one ORO's model (or, from a cognitive point of view,
+% cognitive conflicts) could be better dealt with: while (partially) provided by
+% the Pellet reasoner, and output to the developers for debugging purposes, the
+% logical justification of the inconsistency is not really exploited.
+
+% \subsection{Performances analysis}
+% 
+% \subsubsection{Relaxation time}
+% 
+% The relaxation time (as defined at section~\ref{sect|performances}) of ORO for
+% the average workloads we have encountered during the experiments is in the
+% range of 100ms.  The temporal resolution that follows for the symbolic layer,
+% while acceptable in most cases, is not satisfactory. In particular, multi-modal
+% interaction requires fine grained temporal alignement between
+% modalities~\cite{Li2012} which are not possible with the current temporal
+% performances of ORO.
+% 
+% \subsubsection{Cognitive evaluation}
+% 
+% The performance evaluation of current tools is still mostly system- and
+% scenario-specific (``did my system do the job for my task?''), and amongst the
+% surveyed KRS, only few systems proposes an evaluation of their performances
+% that could be compared to other system.
+% 
+% 
+% \fxwarning{Mention reasoning and expressiveness, cf section oro-server}
+%
 

### 2012-10-08 krs-survey.tex
Integration of Moritz's comments

@@
 
-    Central to this effort is to develop an understanding of the relationship
+    ``Central to this effort is to develop an understanding of the relationship
     between the knowledge, the perception, and the action of [\ldots] a robot. The
@@
     generating actions in the world that are appropriate as a function of some
-    current set of beliefs and desires.
+    current set of beliefs and desires.''
 
@@
 whereas the symbol level deals with inference, the knowledge level deals with
-entailment. We will, at the conclusion of the thesis, give a second look to
+entailment. We will, at the conclusion of the article, give a second look to
 this distinction.
@@
 The relation of \emph{data} and \emph{information} to knowledge is a debated
-epistemology question known as the ``DIKW'' hierarchy question. In this thesis,
+epistemology question known as the ``DIKW'' hierarchy question. In this article,
 we will associate data to low-level material like raw sensor output, and
@@
 a Kinect sensor: the pose of the human skeleton in the world would be the data,
-the fact \concept{looksAt(human, book)} as interpreted by a geometric reasoning
+the fact \concept{looksAt(human, coords)} as interpreted by a geometric reasoning
 module would be the information, the fact \concept{looksAt(john,
@@
 co-exist and cooperate. This scenario will also be a source of support examples
-for later sections of the thesis. Also, we have emphasised several keywords in
+for later sections of the article. Also, we have emphasised several keywords in
 this description: we will come back to them in the next section to
@@
 
-A ``first-order'' analysis of this task leads to a rough partition of the
+A \textit{first order} analysis of this task leads to a rough partition of the
 required \emph{representation} abilities:
@@
 
-These last assertions imply several other capabilities: ``cooking often takes
+These last assertions imply several other capabilities: ``cooking \emph{often} takes
 place in the kitchen'' implies that representation of both uncertainty and
@@
 the first agent knows that the recipient agent both needs the information but
-does not know it. This in turn requires the robot to be able to model the
-knowledge of the other agents: to think \emph{from the perspective} of another
-agent (an idea that is related to the availability of a theory of mind, we will
-come back to it later on).
+does not know it. This in turn requires the robot to be able to access its own
+knowledge (\emph{meta-cognition}), as well as to model the knowledge of the
+other agents (to think \emph{from the perspective} of another agent, an idea
+that is related to the availability of a theory of mind, we will come back to
+it later on).
 
@@
 complexity is fragile: for instance, adding the following axiom \stmt{farFrom
-disjointProperty near} (that states that two individuals can not be at the same
-time near and far from each other) changes the expressiveness power of the ORO
-Common-Sense ontology (presented at chapter~\ref{chapt|oroserver}) from
-$\mathcal{SHOIQ(D)}$ to $\mathcal{SROIQ(D)}$\footnote{See
-appendix~\ref{chapt|dl} for a brief explanation of this notation}: this
-seemingly innocuous assertion change the complexity class of the whole
-ontology, and the concept satisfiability reasoning problem switches from a {\it
-NExpTime-complete} problem to a {\it NExpTime-hard} problem (\ie, \emph{at
-least} as hard as the hardest problem in {\it NExpTime}).
-
-This ``instability'' has practical consequences on run-time performances on
-the robot because a light alteration of the knowledge structure can lead to
-very noticeable performance drops.
+disjointProperty near} that states that two individuals can not be at the same
+time near and far from each other, while looking semantically seemingly
+innocuous, actually introduces what is called a \emph{complex role} that leads
+to a much harder reasoning problem: concept satisfiability, for instance,
+switches from a {\it NExpTime-complete} problem to a {\it NExpTime-hard}
+problem (\ie, \emph{at least} as hard as the hardest problem in {\it
+NExpTime}).
+
+This kind of behaviours have practical consequences on run-time performances on
+the robot, and require careful design of the knowledge structure to be mitigated.
 
@@
 
-Symbolic representation of entities placement can be absolute or relative. The
-relation \concept{isOn}, for example, leads to absolute statements: the
-validity of the relation is independent of the nature of its subjects and
-objects, and is also independent of the observer.
-
-On the contrary, the relation \concept{nextTo} is relative, and depends on the
-relative size of the subject and the object. Two houses distant of 2 meters
-from each other can be considered as next to each other. Two ants separated by
-2 meters are not next to each other.
+Symbolic representations of entities placement are either topological or
+directional. The topological relation \concept{isOn}, for example, leads to
+absolute statements: the validity of the relation is independent of the nature
+of its subjects and objects, and is also independent of the observer. These
+relations are also sometimes qualified as \emph{allocentric}.
+
+On the contrary, the directional relation \concept{nextTo} is relative, and
+depends on the relative size of the subject and the object. Two houses distant
+of 2 meters from each other can be considered as next to each other. Two ants
+separated by 2 meters are not next to each other.
 
 The relation \concept{leftOf} is another example of relative spatial relation,
-this time because it depends on the observer viewpoint.
+this time because it depends on the observer viewpoint. These relations are
+also called \emph{egocentric}.
 
@@
 representatio in general, and we will further discuss the importance and issues
-brought by context modeling in the conclusion of the thesis.
+brought by context modeling in the conclusion of the paper.
 
@@
 set of beliefs of the robot. This limits for instance the ability for the robot
-to react to logical conditions that involve facts that are lazily evaluated
-(``trigger a callback when object A is behind object B'' would not be triggered
-with a purely lazy evaluation strategy for spatial relations).
+to react to logical conditions that involve facts that are lazily evaluated:
+``trigger a callback when object A is behind object B'' would not be triggered
+with a purely lazy evaluation strategy for spatial relations. This can be
+mitigated by selecting (possibly dynamically) a subset of relations that are to
+be strictly evaluated.
 
@@
 
-\paragraph{Projection task}: determining whether or not some condition while
+\paragraph{Projection task}: determining whether or not some condition
 hold after a sequence of actions. The projection task is a typical
@@
 (Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
-affordance-based planing and object classification~\cite{Lorken2008,
+affordance-based planning and object classification~\cite{Lorken2008,
 Varadarajan2011}.
@@
 
+A last qualitative facet relates to the possible bias of an ontology
+regarding the domain it pretends to cover. For instance, Wikipedia (and
+consequently, DBPedia and YAGO) is known to have a bias towards contents
+related to people, geography or historic events. {\sc OpenCYC} also suffer of
+non-homogeneous concept coverage.
+
 \subsubsection{Granularity}
@@
 %OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
+ORO & Formal & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) + {\sc Pellet} & \cite{Lemaignan2010} \\
 OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
-%ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
 PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
@@
 The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
-representation language with natural language processing, task planing and
-motion planing.
+representation language with natural language processing, task planning and
+motion planning.
 
@@
 reasoning has been especially researched within the Ke Jia project for symbolic
-task planing~\cite{Ji2011} and underspecified natural language processing.
+task planning~\cite{Ji2011} and underspecified natural language processing.
 
@@
 \begin{center}
-\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}|p{2cm}}
+\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}}
 \toprule
-\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}       & ORO \cite{Lemaignan2010}                      \\
+\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & ORO \cite{Lemaignan2010}                      & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}        \\
 
@@
 
-\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL + Horn clauses             & {\sc CycL}                    & DL (OWL)                                      \\
-                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 &                               &                               & OWA                                           \\
-                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         & + (\emph{candidate} entities) &                               &                                               \\
-                                                                    & Meta-cognition &                           &                             &                             &                             & ++                                 & + (transformation rules)                        &                               &                               & ++ (reification, taxonomy walking)            \\
+\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL (OWL)                                      & DL + Horn clauses             & {\sc CycL}                     \\
+                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 & OWA                                           &                               &                                \\
+                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         &                                               & + (\emph{candidate} entities) &                                \\
+                                                                    & Meta-cognition &                           &                             &                             &                             & +                                  & + (transformation rules)                        & ++ (reification, taxonomy walking)            &                               &                                \\
 \hline
-\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++                            &                               & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})\\
-                                                               & Time representation &                           &                             & ++ (snapshots)              &                             &                                    & ++ (event oriented)                             & +                             &                               &                                               \\
-                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            & +++                           &                               &                                               \\
-                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          & ++                            & ++ (microtheories)            &                                               \\
-                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 &                               &                               & ++ (Theory of Mind \cite{Warnier2012a})       \\
+\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})& ++         &                               \\
+                                                               & Time representation &                           &                             & ++ (snapshots)              &                             & + (Allen's intervals, via \emph{computables} & ++ (event oriented)                   &                                               & +                             &                                \\
+                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            &                                               & +++                           &                                \\
+                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          &                                               & ++                            & ++ (microtheories)             \\
+                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 & ++ (Theory of Mind \cite{Warnier2012a})       &                               &                                \\
                                                                     & Self-knowledge &                           &                             &                             &                             & + (SRDF \cite{Kunze2011})          &                                                 &                                               &                               &                                \\
-                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 &                               & +                             & +                                             \\
+                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 & +                                             &                               & +                              \\
 \hline
-\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & +                             & ++                            & ++                                            \\
-                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 &                               &                               & ++ (TBox alteration)                          \\
+\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & ++                                            & +                             & ++                             \\
+                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 & ++ (TBox alteration)                          &                               &                                \\
                                                                    & Lazy evaluation &                           &                             &                             &                             & +++ (Prolog, computables)          &                                                 &                                               &                               &                                \\
-                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           &                                    &                                                 &                               &                               &                                               \\
+                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           & + \cite{Tenorth2012}               &                                                 &                                               &                               &                                \\
                                                       & Presupposition accommodation &                           &                             & +++                         &                             &                                    &                                                 &                                               &                               &                                \\
-                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             &                                    & ++                                              &                               &                               & + (explanation)                               \\
-                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +                             & {\it++}                       & +++ ({\em HATP})                              \\
+                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             & + \cite{Tenorth2012}                                   & ++                                              & + (explanation)                               &                               &                                \\
+                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +++ ({\em HATP})                              & +                             & {\it++}                        \\
                                                            & Physics-based reasoning &                           &                             &                             & +                           & {\it+++} ({\em naive physics \cite{Kunze2011a}}) &                                   &                                               &                               &                                \\
@@
 \hline
-\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 &                               & ++ ({\em ambient intelligence})& ++                                           \\
-                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 &                               & + (template-based)            & +++ ({\em Dialogs} \cite{Lemaignan2011a})     \\
+\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 & ++                                            &                               & ++ ({\em ambient intelligence})\\
+                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 & +++ ({\em Dialogs} \cite{Lemaignan2011a})     &                               & + (template-based)             \\
                                                                      & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{Nyga2009}})&                             &                                               &                               &                                \\
-                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & +++ (bottom-up)               & +++ \cite{Loutfi2008}         & ++ ({\em amodal model} \cite{Lemaignan2012})  \\
+                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & ++ ({\em amodal model} \cite{Lemaignan2012})  & +++ (bottom-up)               & +++ \cite{Loutfi2008}          \\
                                                               & Intrinsic Motivation &                           & ++ \cite{Hawes2011}         &                             &                             &                                    &                                                 &                                               &                               &                                \\
 \hline
-\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          & +                             & ++ (tuple space)              &                                               \\
-                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++                            & ++ (tuple space)              & ++ (semantic events)                          \\
-                                                          & Monitoring and debugging &                           &                             &                             &                             &                                    &                                                 &                               &                               & + ({\em remote visualisation})                \\
+\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          &                                               & +                             & ++ (tuple space)               \\
+                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++ (semantic events)                          & ++                            & ++ (tuple space)               \\
+                                                          & Monitoring and debugging &                           &                             &                             &                             & +(Prolog tracer, visualisations...)&                                                 & + ({\em remote visualisation})                &                               &                                \\
                                                            & Performances evaluation &                           & ++ \cite{Hawes2008}         & ++ (Token test)             & +                           & + \cite{Tenorth2011}               &                                                 &                                               &                               &                                \\
@@
 
-The interleaving of discreet symbolic models with continuous geometric models
+The interleaving of discrete symbolic models with continuous geometric models
 remains a major challenge, and techniques vary a lot: besides the CAST proxies
@@
 approach, SPARK is an amodal model of the environment we can manipulate (3D
-motion planing, for instance takes place in MHP, the twin brother of SPARK).
+motion planning, for instance takes place in MHP, the twin brother of SPARK).
 
@@
 Several approaches for building contextualized knowledge have been presented in
-this thesis: symbolic environment interpretation, perspective taking and
+this article: symbolic environment interpretation, perspective taking and
 independent mental state for each agent, grounded natural language resolution,
@@
 
-We will further develop this important topic in the conclusion of the thesis.
+We will further develop this important topic in the conclusion of the paper.
 
@@
 
-The two main tasks that require time reasoning are task planing and sequence
+The two main tasks that require time reasoning are task planning and sequence
 recognition (in particular, action recognition). Systems that directly
-integrate task planing (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
+integrate task planning (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
 fluents in {\sc KnowRob}) to represent time.
@@
 capabilities like a theory of mind, and this is probably the main contribution
-of this thesis regarding techniques for knowledge representation.
+of this paper regarding techniques for knowledge representation.
 

### 2014-04-11 letter-to-editor.tex
Draft of the 'letter to editor in chief' of TRO

\documentclass{letter}
\usepackage{letterbib}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{booktabs}

\signature{Séverin Lemaignan, Moritz Tenorth}
\begin{document}

\begin{letter}{Editor-in-Chief \\ Frank Park \\ Seoul National University}
\opening{Dear Editor-in-Chief,}

As recommended on the website of the \emph{Transactions on Robotics}, we contact
you to bring to your attention a survey proposal on Knowledge Representation \&
Reasoning systems for cognitive robotics.

Those systems or frameworks have rapidly grown in number over the last few
years: Brachman and Levesque's book~\cite{brachman2004knowledge} in 2004, dozens
of systems have emerged to fulfil the need for explicit, symbolic (and
sub-symbolic) \emph{representation} and \emph{reasoning} on knowledge in
cognitive architectures for robots.

Because these systems have explored many different paths, from building upon purely
logic languages, to proposing distributed, diffuse knowledge representation
schemes; because more and more research groups that are looking for increased
robot autonomy, look towards implementing their own framework for knowledge
representation; because, also, no previous survey of these tools has ever been
published to our best knowledge, we propose to write such a review of the
existing literature, starting with brief typology of main features to be found
in those systems, followed by a systematic review of the existing tools.

This survey would be a joint work by Moritz Tenorth, from the University of
Bremen, Germany (Michael Beetz's group) and Séverin Lemaignan, currently working
at EPFL, Switzerland and previously with Rachid Alami's group, LAAS-CNRS, France.

Moritz Tenorth is the author and main developer of the well-known {\sc KnowRob}
suite for knowledge management for service robots. This framework has been
deployed on many different robots, ranging from the Willow Garage's PR2 to
[...?]. His different previous publications on this project have been cited over
500 times, and he is regarded as an expert in this field.

Séverin Lemaignan is the author of the {\sc ORO} knowledge base, that has
received special attention in the human-robot interaction community for being
one of the first knowledge representation and reasoning system taking explicitly
into account the cognitive models of the humans interacting with the robot. His
platform has also been deployed on several robots, in at least five different
laboratories. Séverin Lemaignan is regularly invited for talks on cognitive
architectures.

For this survey, we plan to consider the following systems:

\begin{tabular}{p{2.8cm}p{8cm}p{1.5cm}}
\toprule
{\bf Project} & {\bf Authors (Institution)} & Main references \\
\midrule
ARMAR/Tapas &  Holzapfel, Waibel \par (Karlsruhe TH) &  \cite{Holzapfel2008}\\
CAST Proxies &  Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) &  \cite{Jacobsson2008} \\
GOLOG & Levesque (Toronto Univ.) & \cite{levesque1997golog} \\
GSM &  Mavridis, Roy \par (MIT MediaLab) & \cite{Mavridis2006} \\
Ke Jia Project & Chen et al. \par (Univ. of Science and Technology of China) &  \cite{Chen2010} \\
{\sc KnowRob} &  Tenorth, Beetz \par (TU Munich) &   \cite{Tenorth2009a} \\
NKRL &  Zarri et al. \par (Paris Est Créteil Univ.) &  \cite{Sabri2011} \\
%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
ORO &  Lemaignan, Alami \par (LAAS-CNRS) &  \cite{Lemaignan2010} \\
OUR-K/OMRKF &  Lim, Suh et al. \par (Hanyang Univ.) &   \cite{Lim2011, Suh2007} \\
PEIS KR\&R &  Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) &  \cite{Daoutis2009} \\
% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
% (based on {\sc KnowRob} & & (JSK) \\

\bottomrule

\end{tabular}

We hope that you may consider our proposal as interesting and relevant for a future issue of
\emph{Transaction on Robotics}.

\closing{Yours Sincerly,}

\ps{Séverin Lemaignan is with EPFL, CHILI Lab, Lausanne, Switzerland; Moritz Tenorth is with Institute for Artificial Intelligence, Universität Bremen, Germany.}

%\url{http://www.gnu.org/copyleft/fdl.html}.}
%\encl{Copyright permission form}

\end{letter}

\bibliographystyle{ieeetr}
\bibliography{biblio}

\end{document}

### 2014-04-16 krs-survey.tex
TRO template (IEEEtran.cls)
Removed Prototypical scenario part

@@
-\documentclass[a4paper, twocolumn]{article}
+%\documentclass[a4paper, twocolumn]{article}
+\documentclass{IEEEtran}
 
@@
 addressed by the current state of the art.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Prototypical Scenario}
-\label{sect|scenario}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
-	\caption{The {\em Brownie} scenario}
-	\label{fig|scenario}
-\end{figure*}
-
-The aim of this imaginary scenario (that has not been implemented, neither in
-simulation nor on a robot) is to materialise early in this paper the context
-and challenges of knowledge representation and manipulation for service, social
-and interactive robotics. It underlines the place, the role and the need of
-knowledge in a near-future, everyday situation where several robots and humans
-co-exist and cooperate. This scenario will also be a source of support examples
-for later sections of the article. Also, we have emphasised several keywords in
-this description: we will come back to them in the next section to
-explain them formally and relate them to each other.
-
-We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
-Robi and Roba are two service robots, that can freely move and pick objects
-around (with possibly different hardware and software architectures, including
-different knowledge representation systems: typically, two robots built by two
-different companies). They cooperate with a human in a kitchen environment.
-
-The main task of the scenario is the joint realization of a brownie, initiated
-by Tom, the human: ``Let's make a brownie for tonight!''.
-
-The scenario is successful {\it 1)} if the task is achieved (the brownie is baked),
-{\it 2)} in a reasonable time (typically shorter than what it would have been
-required by the human alone), {\it 3)} and the human is happy by the help it received
-(he/she did not feel useless, uncomfortable, unsafe).
-
-We voluntarily do not detail the subtasks of the scenario, neither we define
-how they are shared amongst agents: our focus is on knowledge needs and flows.
-
-A \textit{first order} analysis of this task leads to a rough partition of the
-required \emph{representation} abilities:
-
-\begin{enumerate}
-
-	\item Representation abilities related to the execution of a complex
-	spatio-temporal task,
-
-	\item Representation abilities related to cooperation with other agents.
-
-\end{enumerate}
-
-% Representation of a complex task
-We can further refine these categories: to prepare and bake a brownie, the
-robot first needs to make sense of the term \emph{brownie} itself: what is it?
-what is it used for? what is it made of? etc. We call this knowledge
-\emph{common-sense knowledge} and the robot must be able not only to represent
-it, but also to have access to an initial source (for instance through a
-initial set of facts that are made available at startup, or via access to a
-Web-based knowledge base like Wikipedia)
-
-Once bound to the action \emph{make}, this should lead the robot to build and
-represent a \emph{context}: we are in a scenario involving cooking. The context
-enables the robot to retrieve more common-sense knowledge, like that actions
-related to cooking often take place in the kitchen, cooking requires
-ingredients, utensils and a procedure that may be provided by a recipe.
-
-These last assertions imply several other capabilities: ``cooking \emph{often} takes
-place in the kitchen'' implies that representation of both uncertainty and
-likelihood is desirable. The fact that cooking is associated to a place further
-implies that the system models locations and is able to attach \emph{thematic
-relations} to concepts (here, the likely location of the cooking action).
-
-``cooking requires a procedure and ingredients'' hints about another important
-feature closely tied on knowledge manipulation: \emph{reasoning}. The robot can
-\emph{infer} that cooking may require a recipe since a procedure and a list of
-ingredients are pre-requisites of the cooking action that may be provided by a
-recipe.  If we omit the ``may'', this is a typical example of first-order logic
-reasoning.  Many other reasoning techniques exist (including probabilistic ones
--- ones able to deal with the ``may''), we shall illustrate some of them later
-in this scenario.
-
-We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
-robot should be able to store this plan in a way that allows later execution.
-The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
-the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
-must be as well appropriately handled.
-
-To make decision, a robot may also want to \emph{predict} the state of the
-world after some action (``if I leave the cake 2h in the oven, it will burn'').
-Such ability to project itself in future or, generally speaking, in other
-possible state of the world is related to several cognitive ability and
-reasoning techniques: \emph{planning}, \emph{projection}, \emph{representation
-of possible worlds} and \emph{non-monotonic reasoning}, in addition to
-common-sense knowledge and \emph{physics-based} reasoning (that allows for
-instance to predict that an egg is likely to break if dropped).
-
-Procedures are in addition often \emph{underspecified}: we can expect the
-recipe to provide a cooking duration, but we usually do not expect the recipe
-to tell us to first open the oven door, and then put the cake into it, since it
-is self-evident that the door must first be opened to put the cake in the oven.
-Our cognitive robot should ideally be able to detect and possibly complete such
-underspecification.
-
-% Representation feature that enable cooperation
-Then, we want our three agents to cooperate. This, in turn, leads to another
-set of cognitive abilities.
-
-Cooperation in our scenario can intervene at many places. For instance, an
-agent may want to inform another one about the number of eggs that are
-necessary for the brownie. This \emph{helping} behaviour makes sense only if
-the first agent knows that the recipient agent both needs the information but
-does not know it. This in turn requires the robot to be able to access its own
-knowledge (\emph{meta-cognition}), as well as to model the knowledge of the
-other agents (to think \emph{from the perspective} of another agent, an idea
-that is related to the availability of a theory of mind, we will come back to
-it later on).
-
-Ability to communicate is one important pre-requisite to collaboration.
-Communication in general requires the addresser and the addressee to share a
-common interpretative framework (a shared common-sense knowledge -- or cultural
-background -- and a shared context). In our scenario, the agents are working in
-a kitchen. This element of context does not however suffice if, for example,the 
-human asks a robot to ``give {[him]} the bowl''. Behind the symbol
-``bowl'', which physical entity are we actually talking about? If we want to
-talk and act on the world, this so-called \emph{grounding} operation is
-essential. It is a bidirectional process: in covers the {\it top-down}
-operation (from the symbol to the percept) and the {\it bottom-up} converse
-(retrieval or creation of symbols from perception).
-
-A related ability is called \emph{pre-supposition accommodation}: if one of the
-agent moves behind another one, with the brownie dough in its arm, and says
-``be careful, I'm behind you!'', we want the first agent to be able to
-represent both symbolically and geometrically (because, for instance, if the
-agent want to move, it must take into account the new obstacle) something that
-is not directly perceived.
-
-Also central to cooperation are the notions of \emph{joint intentions} and
-\emph{joint goals}: to help the human during the cooking session, the robots
-need to track how far they are into the recipe, what is the next step the human
-is likely to go for, how tasks are currently split between agents, what action
-is currently blocking the procedure, etc. This knowledge should let the robot
-identify the intentions of other agents and create accordingly joint goals.
-Hence, a knowledge representation system aiming at dealing with cooperative
-behaviours is likely to have goal management structures taking explicitly into
-account other agents' actions and goals.
-
-In order to effectively share tasks, the robot must also know what it is
-capable of: \emph{capability introspection} (both in term of general capability
-and of immediate ability) is thus often desirable. It can be extended to
-general introspection (like the ability to tell ``who I am'' or ``what do I
-think of'') that may be required for the interaction.
-
-Last but not least, our scenario assumes implicitly \emph{natural interaction}
-between humans and robots (as shown by the casual style of the order ``Let's
-make a brownie!''), and we want to ensure that the
-knowledge available to the robot provides efficient support to the natural
-language understanding (for instance by adopting models and vocabulary that are
-both well suited for machine processing and remain as close as possible to the
-humans own structures and vocabulary), and also to \emph{non-verbal forms of
-communication}, like gestures.
 

### 2014-04-16 krs-survey.tex
Summarized the 'features' section in 2 pages

@@
 \newcommand{\eg}{{\textit{e.g.\ }}}
+
+\newcommand{\taxon}[2]{%
+    \textbf{[#1] \emph{#2}}
+}
 
@@
 
-\tikzstyle {taxonomy} = [grow=right, level distance = 3cm, sibling distance=7mm]
+\tikzstyle {taxonomy} = [grow=right, level distance = 3cm, sibling
+distance=7mm]
 \tikzstyle {completetaxonomy} = [grow=right]
-\tikzstyle {taxon} = [text width=3cm, text centered,inner sep=0pt, outer sep=2pt] %, draw=blue!50,thick]
+\tikzstyle {taxon} = [text width=3cm, text centered,inner sep=0pt, outer
+sep=2pt] %, draw=blue!50,thick]
 
@@
 
-The paper is organized in four main sections.
-
-The next section proposes a prototypical scenario of a robot acting in a rich
-semantic environment. We will not survey knowledge representation systems based
-on this scenario: we merely use it to ground the selection and illustrate the
-set of features that define the dimensions of a knowledge representation system
-for robotics.
-
-Section~\ref{sect|features} presents these features in detail, and organize
-them into a typology.
+The paper is organized in three main sections.
+
+Section~\ref{sect|features} introduces a typology of requirements and features
+of knowledge representation systems for robotics, that we use as foundations to
+compare and evaluate the systems we examine in this survey.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Typology of Knowledge Representation Requirements for Robotics}
+\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
 \label{sect|features}
 
-We now want to organize requirements from this imaginary cooking scenario where
+We propose to organize requirements from this imaginary cooking scenario where
 numerous desirable features for a knowledge representation system for robotics
@@
 
-As previously mentioned, we propose six main categories.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    %[edge from parent fork east]
-    %every node/.style={fill=black!30,rounded corners},
-    %[parent anchor=east,child anchor=west,grow=east,
-    %edge from parent/.style={thick,draw}]
-    \node [taxon] {Dimensions}
-        child {node [taxon] {{\bf F}. Instantiation}}
-        child {node [taxon] {{\bf E}. Integration}}
-        child {node [taxon] {{\bf D}. Acquisition}}
-        child {node [taxon] {{\bf C}. Reasoning}}
-        child {node [taxon] {{\bf B}. Representation}}
-        child {node [taxon] {{\bf A}. Expressiveness}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-Table~\ref{table|contribution-by-systems}, at the end of the article,
-summarizes all these features with the main domain of contribution of each
-surveyed KRS.
+As previously mentioned, we propose six main categories, that cover both basic
+and practical aspects: the formal
+\taxon{A}{Expressiveness} of the system, the breadth and nature
+of the \taxon{B}{Representations} allowed by the system, the
+\taxon{C}{Reasoning} capabilities, how knowledge it acquired
+\taxon{D}{Acquisition}, how the system integrates with the other components of
+the robot sotware architecture \taxon{E}{Integration} and finally what we call
+the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
+base.
+
+The next paragraph details only briefly the exact meaning of these categories
+and introduce accordingly sub-categories. We kindly refer the interested reader
+to~\cite{blabla} for in-depth discussion of each of these dimensions of
+knowledge representation systems, as well as related literature.
+
+Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
+the main contributions of each of the surveyed KRS with respect to these
+dimensions.
 
@@
 \label{sect|expressiveness}
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf A. Expressiveness}
-            child {node [taxon] {{\bf A.5}. Meta-cognition}}
-            child {node [taxon] {{\bf A.4}. Uncertainty}}
-            child {node [taxon] {{\bf A.3}. OWA/CWA}}
-            child {node [taxon] {{\bf A.2}. Expressive power}}
-            child {node [taxon] {{\bf A.1}. Logic formalism}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
 
@@
 
-\subsubsection{Main logic formalisms}
-
-The main role of a knowledge representation system is to provide an adequate
-representation system to formally store facts and concepts that could be informally
-described in natural language.
-
-Formal logic aims at providing such a representation system with the added
-value of providing a tractable support for inference and reasoning.
-
-Most (but not all) of the systems we have surveyed rely on a particular logic
-formalism. The choice of the formalism has a strong impact, on one side, on the
-range of ideas that can be expressed conveniently (\emph{practical
-expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
-side, on the ability to solve the fundamental inference problem (called
-\emph{satisfiability}: is a given logical sentence true in my model?) in a
-tractable manner.
-
-A large number of logic formalisms do exist and we briefly present below the most
-relevant ones for systems actually deployed in robotic architectures.
-
-\emph{Predicate logic} is the family of logic formalisms the most commonly
-found in knowledge representation. It distinguishes itself from the simpler
-\emph{propositional logic} by the use of quantification to increase generality.
-\emph{First-order logic} (FOL) is the subpart of \emph{predicate logic} where the
-objects of \emph{predicates} (or \emph{formulae}) are simple \emph{terms},
-while in \emph{higher-order logics}, predicates can be themselves objects of
-other predicates.
-
-\emph{Horn clauses} are an important subset of FOL because the satisfiability
-of a set of such clauses is a $P$-complete problem (\ie practically tractable).
-A Horn clause is a disjunction of literals (a \emph{clause}) with at most one
-positive literal: $\neg p \lor \neg q \lor \cdots \lor \neg t \lor u$, which
-can also be represented as $(p \land q \land \cdots \land t) \rightarrow u$.
-Important logic programming languages like Prolog are based on Horn clauses.
-
-The family of \emph{Description Logics}~\cite{Baader2008} also plays an
-important role. It is also a subset of the first-order logic, with some
-extensions in second-order logic. Description logics are notable because most
-of them are known to be decidable (but not always in a practically tractable
-manner). In description logic, axioms are build from \emph{concepts},
-\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
-W3C OWL-DL standard is a widely-used language to describe domains with the
-description logic.
-
-Because description logics have been originally created from the perspective of
-a \emph{knowledge representation language} and not a logic language, their
-terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
-\emph{individual},\ldots) is well-suited to knowledge description.
-
-\emph{Modal logic}, that allows for statement qualification like
-\emph{possibility} or \emph{necessity}, have been shown to be closely related
-to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
-possible worlds and facts like ``the robot knows \emph{that the human knows}
-how to read a recipe''.
-
-\fxwarning{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}
-
-\emph{Temporal logic} are designed to represent and manipulate assertions whose
-truth value may vary in time. We introduce one of its key idea (the
-\emph{fluents}) later in the typology.
-
-One last class of logics that is of particular relevance for robotic
-applications is the \emph{probabilistic logics} or \emph{Bayesian logics}.
-These logics provide a formal framework to reason on propositions whose truth
-or falsity is uncertain. We elaborate below on the representation of uncertainty.
-
-Note that most of these logic formalisms are still active research field on
-their own, and practical considerations (especially the availability of
-reasoners efficient enough for on-line use on a robot) often constrain the
-choice of a logical formalism and a level of expressive power.
-
-\subsubsection{Expressive power}
-
-Logical formalisms each bring a certain level of expressive power. For
-instance, the following classical syllogism can not be represented in
-propositional logic because of the use of \emph{universal quantification}:
-
-\begin{quote}
-\begin{enumerate}
-    \item All men are mortal,
-    \item Socrates is a man,
-    \item Therefore, Socrates is mortal
-\end{enumerate}
-\end{quote}
-
-However, the following weak version of the syllogism can be represented in
-propositional logic:
-
-\begin{quote}
-\begin{enumerate}
-    \item If Socrates is a man, then Socrates is mortal,
-    \item Socrates is a man,
-    \item Therefore, Socrates is mortal
-\end{enumerate}
-\end{quote}
-
-Generally speaking, expressive power comes at the cost of more complex
-\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
-at section~\ref{sect|reasoning}.} computations, possibly leading to
-untractable, if not undecidable (\ie systems where it is proven that a
-proposition can not be decided to be true or false) problems.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.3\columnwidth]{expressive_overlap_dl_horn.pdf}
-    \caption{expressiveness overlap of Description Logics and logic programs
-    based on Horn clauses, taken from~\cite{Grosof2003}}
-    \label{fig|overlap_dl_horn}
-\end{figure}
-
-Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
-logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
-mention extensions to description logics based on rule systems that bring closer
-the two approaches.
-
-The relationships between expressive power and reasoning complexity that follow
-has been extensively studied for Description Logics.
-Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
-that allows to conveniently explore these relationships and indexes most of the
-literature on that subject.
-
-It can be noted that the relation between expressiveness and reasoning
-complexity is fragile: for instance, adding the following axiom \stmt{farFrom
-disjointProperty near} that states that two individuals can not be at the same
-time near and far from each other, while looking semantically seemingly
-innocuous, actually introduces what is called a \emph{complex role} that leads
-to a much harder reasoning problem: concept satisfiability, for instance,
-switches from a {\it NExpTime-complete} problem to a {\it NExpTime-hard}
-problem (\ie, \emph{at least} as hard as the hardest problem in {\it
-NExpTime}).
-
-This kind of behaviours have practical consequences on run-time performances on
-the robot, and require careful design of the knowledge structure to be mitigated.
-
-\subsubsection{Open world and close world assumptions}
-\label{sect|owa-cwa}
-
-The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumptions name a
-modelling choice on the \emph{completeness} of a knowledge domain. In the close
-world assumption, a proposition that can not be proven true is assumed to be
-false (\emph{negation by failure}), while in the open world assumption, a
-proposition may be considered either true, false or unknown.
-
-This distinction is important in robotics where the robot may have to manipulate
-concepts with only partial knowledge on them. For instance, let us imagine a robot
-that sees a bottle on a table, whose bottom is hidden by another object. The
-robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
-representation system relying on the closed world assumption would then assume
-the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
-whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
-table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
-another example of consequences of the CWA/OWA choice on reasoning.
-
-\begin{table}
-    \begin{center}
-    \begin{tabular}{ll}
-    {\bf Action} & {\bf Part involved} \\
-    \hline
-    {\tt PickSoftly} & hand \\
-    {\tt PickAndPlace} & arm, hand \\
-    {\tt MoveArm} & arm \\
-    \hline
-    \end{tabular}
-    \end{center}
-    \caption{Assuming the question is: \emph{select actions that do not require
-    to move the arm}, a CWA reasoner would return {\tt PickSoftly} whereas an
-    OWA reasoner would not return anything if the {\tt PickSoftly} action is
-    not explicitly said not to involve the arm.}
-    \label{table|cwa-owa-example}
-\end{table}
-
-The OWL language is specifically known to assume an open world.  Domains
-constrained with the closed world assumption lead to more tractable inference
-problems, and allow for instance the use of logic languages like Prolog. Thus,
-several approaches exists to \emph{locally close} a domain (\cf
-Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those ones).
-
-\subsubsection{Representation of uncertainty and likelihood}
-
-Sources of uncertainty for a robot are two-fold: uncertainty \emph{intrinsic}
-to facts (like \emph{``It may rain tomorrow''}), uncertainty caused by
-imperfect perception of the world (\emph{``Is the bottle really on the
-table?''}). Most logics do not account explicitly for uncertainty. It must be
-either relied on specific logics (like Bayesian logics) or on extensions of
-classical logics.
-
-\fxfatal{Representation of uncertainty: To be completed!}
-
-\subsubsection{Meta-cognition: knowledge on the knowledge}
-
-As stated by Cox and Raja~\cite{Cox2007}, meta-cognition is composed of both
-\emph{``meta-level control of cognitive activities and the introspective
-monitoring of such activities to evaluate and to explain them"}.
-
-Sloman proposes in~\cite{Sloman2011} a detailed analysis of meta-cognition and
-its different aspects in both natural (human) and artificial systems.
-
-A knowledge representation system endowed with \emph{meta-cognition} is not
-only able to manipulate knowledge but also to exhibit and manipulate the
-structure of its knowledge and the reasoning process. For instance, the ability
-to explain a logical inconsistency in a KRS is a meta-cognitive function, as is
-the ability to expose and alter the knowledge structure (these two reasoning
-techniques have their own entries in the taxonomy, at
-section~\ref{sect|reasoning}.
-
-At section~\ref{sect|introspection} below, we discuss the idea of
-introspection.  Meta-cognition can be viewed as the technical facet of the
-introspection in general.
+We subdivide the category \taxon{A}{Expressiveness} into five subcategories:
+\taxon{A.1}{Logic formalism} classifies the underlying logical formalism of the
+KRS (Horn clauses like for Prolog-based systems, Description
+Logics~\cite{Baader2008}, Bayesian logics, etc.). \taxon{A.2}{Expressive Power}
+quantifies the design choices made with respect to the actual expressive power
+of the logical formalism: tradeoffs have to be found to allow both for high
+expressiveness and tractable and decidable representations, and KRS authors
+often add constraints on their representation logics to this end.
+\taxon{A.3}{OWA/CWA} reflects the design choice between the \emph{open world}
+(OWA) versus the \emph{closed world} (CWA) assumptions, which related to the
+possibility of representing \emph{unknown} facts. \taxon{A.4}{Uncertainty}
+relates to the ability for the system to represent both the \emph{intrinsic}
+uncertainty of given facts, and the \emph{extrinsic} uncertainty due to
+imperfect perception by the robot. \taxon{A.4}{} relates only to the possibility
+of \emph{representation}. Reasoning on uncertain facts is discussed in
+\taxon{C.4}{}. The last features of a KRS related to representation is
+\taxon{A.5}{Meta-cognition} that conveys the capability for a system to
+represent and exhibit it own internal structures and representation mechanisms.
 
@@
 
-We do not discuss in this section the general strategies to construct a
-knowledge model (they will be presented in
-section~\ref{sect|design-strategies}). We focus here on questions that involve
+This category focuses on questions that involve
 representational challenges (time, space, context) or require specific
 cognitive capabilities (theory of mind, introspection, memory).
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf B. Representation}
-            child {node [taxon] {{\bf B.5}. Memory}}
-            child {node [taxon] {{\bf B.4}. Introspection}}
-            child {node [taxon] {{\bf B.3}. Modality}}
-            child {node [taxon] {{\bf B.2}. Context}}
-            child {node [taxon] {{\bf B.1}. Roles}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\subsubsection{Role representations}
-
-This section discusses strategies and approaches to represent in a knowledge
-model three important \emph{roles}: spatial relations, time and representation
-of actions.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf B.1 Roles}
-            child {node [taxon] {{\bf B.1.3}. Actions}}
-            child {node [taxon] {{\bf B.1.2}. Time}}
-            child {node [taxon] {{\bf B.1.1}. Space}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\paragraph{Representation of space}
-
-Symbolic representation of space is a widely studied topic. In particular, a
-large literature corpus is available on spatial ontologies\fxwarning{Which one?}.
-
-Two main classes of spatial relations are usually represented at the symbolic
-level: the topology of environments and the placement of physical entities.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf B.1.1 Space}
-            child {node [taxon] {{\bf B.1.1.2}. Placement}}
-            child {node [taxon] {{\bf B.1.1.1}. Topology}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-Topological maps are abstractions of an environment as graphs where nodes
-represent places and edges represent connections between places. Because they
-are symbolic representation, topological maps allow higher-level reasoning
-(such as containment, connectivity, regions) than metric maps.
-
-One important contribution to the building of a coherent representational stack
-for space representation is the Spatial Semantic Hierarchy~\cite{Kuipers2000}
-introduced by Kuipers. It consists in multiple interacting representations of
-space, both qualitative and quantitative, that span from sensor-level
-representation to ontologies of places and regions.
-
-Symbolic representations of entities placement are either topological or
-directional. The topological relation \concept{isOn}, for example, leads to
-absolute statements: the validity of the relation is independent of the nature
-of its subjects and objects, and is also independent of the observer. These
-relations are also sometimes qualified as \emph{allocentric}.
-
-On the contrary, the directional relation \concept{nextTo} is relative, and
-depends on the relative size of the subject and the object. Two houses distant
-of 2 meters from each other can be considered as next to each other. Two ants
-separated by 2 meters are not next to each other.
-
-The relation \concept{leftOf} is another example of relative spatial relation,
-this time because it depends on the observer viewpoint. These relations are
-also called \emph{egocentric}.
-
-\fxwarning{\concept{front}/\concept{back} -> entities have or not a concept of front/back.}
-
-Choices must be made in the knowledge representation system to adequately
-represent relative spatial relations. Options include the computation of such
-relations only on-demand, when the context is known (which viewpoints, etc.) or
-storage of these facts in different models, one per agent.
-
-\fxwarning{Find references on that subject -> classification of spatial
-relations. Wikipedia?}
-
-\paragraph{Representation of time}
-\label{sect|time-representation}
-
-As an agent acting at human-like time scale and dealing with temporal concepts
-(like actions), a robot needs to represent and reason about
-time. Time representation is split into two distinct abilities: representing
-time points (both in the past -- which is roughly equivalent to assignment of
-timestamps to events the robot perceives -- and in the future), and
-representing passing time (situations, durations, timespans) like in
-\emph{``the eggs will be cooked in 10 min''}.
-
-This is usually formalised as a disjunction in time representation between
-\emph{events}, that is, any durationless temporal concept, and
-\emph{situations}, that is, any temporal concept with a non-zero
-duration\footnote{Other definitions of a situation do exist, notably in the
-context of \emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque
-consider a situation to be an history of actions.}.
-
-Numerous techniques to represent and reason about time have been devised.
-Amongst the most significant ones, we can mention Allen's interval
-algebra~\cite{Allen1984} and Ghallab's chronicles~\cite{Ghallab1996}.
-\fxwarning{We could detail more chronicles and allen algebra...}
-The concept of \emph{fluent} also play an important role for time
-representation: fluents are properties (or \emph{conditions}) that change over
-time, like $sees(agent1, apple, t)$.\fxwarning{A bit short... citation?}
-
-We call a system that does not account for time (\ie that mentally permanently
-lives in present) \emph{atemporal}.
-
-\paragraph{Actions}
-
-Events and actions are two temporal concepts that are of particular importance to
-robotic systems, as systems that perceive, react and perform in their environment.
-
-While representation of events boils down to label a timestamp, representation
-of actions are more complex since not only they deal with durations, but they
-also imply semantic interactions between concepts. From a taxonomy point of
-view, actions are a particular type of event that normally leads to a situation
-corresponding to the action realisation.
-
-\emph{Thematic roles}~\cite{Gruber1965} (also found as \emph{semantic roles} or
-\emph{theta roles} in the literature) allow to semantically qualify the
-parameters of an action. The recipient of the action, the performer, the object
-acted upon, the destination are some example of common thematic roles.
-Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
-roles proposed by \cite{Aarts1997} (see~\cite{Gutierrez2001} for a comparison
-of other sets of thematic roles present in the literature).
-
-\begin{table}
-\begin{center}
-
-\begin{tabular}{lp{5cm}}
-
-\toprule
-       Role & Meaning \\
-\midrule
-      Agent & The \emph{doer} or instigator of the action denoted by the predicate. \\
-    Patient & The \emph{undergoer} of the action or event denoted by the predicate. \\
-      Theme & The entity that is moved by the action or event denoted by the predicate. \\
-Experiencer & The living entity that experiences the action or event denoted by the predicate. \\
-       Goal & The location or entity in the direction of which something moves. \\
-Benefactive & The entity that benefits from the action or event denoted by the predicate. \\
-     Source & The location or entity from which something moves. \\
- Instrument & The medium by which the action or event denoted by the predicate is carried out. \\
-   Locative & The specification of the place where the action or event denoted by the predicate in situated. \\
-\bottomrule
-
-\end{tabular}
-\end{center}
-
-\caption{A list of thematic roles, as proposed by Aarts~\cite{Aarts1997}.
-Depending on the action, only certain roles are meaningful.}
-
-\label{table|theta-roles}
-\end{table}
-
-In the context of robotic, \emph{task} often represents an aggregate of
-(atomic) actions. We use the term here as the representation of the abstract
-model of an action. Tasks usually specify at least the conditions required to
-performed the action and the consequences of the realization of the action, and
-are central to the decisional layers of the robot, in particular for the
-planning, monitoring and execution control activities. Knowledge representation
-systems may thus have specific mechanisms to represent them (and possible, to
-reason about them, as presented in section~\ref{sect|planning}).
-
-Plans representation is closely related to action and task representation. It
-is specifically discussed at section~\ref{sect|planning}.
-
-\subsubsection{Context modeling}
-
-Knowledge is contextualized information: it is essential for the robot to
-associate the facts it represents to a \emph{context}. The context carries the
-keys for the interpretation of the information and set a common ground for
-interaction (and, in particular, communication~\cite{Jakobson1960}). It
-implicitely defines the domain of validity of the facts and carries the
-\emph{common-sense} knowledge required to fill the gaps of the knowledge
-explicitely shared between the agents.
-
-Context is generally difficult to recognise, represent and reuse because it is
-multiform and largely implicit nature. It is also never unique: at a given
-moment, several contexts, at different temporal, spatial, social scales,
-overlap.
-
-In the current literature in robotics and cognitive architectures, the term
-\emph{context} usually simply refers to a set of beliefs that initiate a
-representation (and reasoning) frame: in~\cite{Lemaignan2011a}, the robot
-creates a context of interaction with a specific human by storing in a separate
-model the beliefs of this human and using this knowledge when dialoging with
-the human, the reasoning network is reinitialised in the GLAIR architecture
-when the hypotheses that defined the current situation are not believed
-anymore~\cite{Shapiro2009}.
-
-This acception of \emph{context} is simplistic, and omit the overlapping and
-multi-scale aspects of context modeling.
-
-We see context representation as one of the main challenge of knowledge
-representatio in general, and we will further discuss the importance and issues
-brought by context modeling in the conclusion of the paper.
-
-\subsubsection{Modality, contingency and theory of mind}
-\label{sect|possible-worlds}
-
-Linked to the context representation, but seen from another angle, knowledge
-representation systems may support logical \emph{modality}. A knowledge model
-is logically modal if is support the concept of \emph{possible worlds}, \ie,
-parallel beliefs models (or \emph{interpretations}) that can be independently
-accessed.
-
-A \emph{contingent} proposition is defined as neither always true (a tautology)
-nor always false (a contradiction) in every possible world: its truth value
-depends on the context. In knowledge representation systems for robotics that
-support logical modality, interpretations are often initialised with a common
-set of initial beliefs (like common-sense knowledge). This initial common
-knowledge is hence true in every possible worlds for the robot, and thus does
-not belong to its contingent knowledge.
-
-On the contrary, alternative mental models with contingent knowledge may be
-used to represent different (possibly hypothetical or even imaginary) views on
-the world, from the robot own perspective or context, or from other
-perspectives computed by the robot.
-
-The representation of the mental perspective of other agents has a particular
-importance in human-robot interaction. It relies first on the ability to
-literally \emph{view} the world from a standpoint which is not
-\emph{egocentric}. This cognitive ability is referred as \emph{perspective
-taking}. Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} define the
-psychological grounds of perspective taking, that are themself originated in
-Piaget's work on cognitive development (recent studies on infants include
-Moll~\cite{Moll2006} for instance). Perspective taking begins to be studied on
-robots as well~\cite{Trafton2005, Breazeal2006, Ros2010}.
-
-The idea of a \emph{theory of mind}~\cite{Premack1978} emerges from the
-perspective taking ability. It can be defined as the ability for one to
-understand and acknowledge that other intelligent agents can have their own
-mental state (that includes beliefs, intents, desires, knowledge) that is
-possibly different from one's own. The \emph{attention} plays a central to the
-development and recognition of a theory of mind~\cite{Baron-Cohen1985,
-Leslie2000}.
-
-A notable consequence of having a theory of mind is the representation of
-\emph{false beliefs}, \ie, facts that are believed to be true for an agent, but
-false for other ones. The \emph{Sally and Ann} experiment~\cite{Leslie2000}
-(figure~\ref{fig|sally-ann}) is a classical example of a false belief
-situation. In~\ref{fig|sally-ann-final}, Sally thinks the ball is in the beige
-box because she did not see Ann moving it. An external observer asked ``Where
-will Sally look for the ball?'' would answer ``in the blue box'' without a
-theory of mind (\ie, a model of the knowledge of Sally), whereas it would
-correctly answer ``In the beige box'' with a theory of mind.
-
-\begin{figure}
-        \centering
-        \subfigure[Sally puts the ball in the beige box.]{
-            \includegraphics[scale=0.3]{sally_ann1.pdf}
-        } \hspace{8pt} %
-        \subfigure[She goes away.]{
-            \includegraphics[scale=0.3]{sally_ann2.pdf}
-        } \\
-        \subfigure[Ann moves the ball.]{
-            \includegraphics[scale=0.3]{sally_ann3.pdf}
-        } \hspace{8pt} %
-        \subfigure[Where will Sally look for the ball?]{
-            \includegraphics[scale=0.3]{sally_ann4.pdf}
-            \label{fig|sally-ann-final}
-        }
-        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken
-        from~\cite{Leslie2000}.}
-
-        \label{fig|sally-ann}
-\end{figure}
-
-Scassellati~\cite{Scassellati2002} is one of the first to have implemented a
-theory of mind on a humanoid robot.
-
-\subsubsection{Self-knowledge: Who am I? What can I do?}
-\label{sect|introspection}
-
-\paragraph{Self-knowledge}
-
-\emph{Self-knowledge} is the term used in philosophy and psychology to describe
-the knowledge that an individual has, acquires or infers about itself through
-its experiences. It answers the question ``What am I like?''.
-
-Introspection is the technical meaning to access to self-knowledge by the
-ability to self describe: what are my capabilities, what is my state
-(performing some action, idling, etc.), what are my beliefs, what are my
-intentions and my plans?
-
-Introspection must be distinguished from meta-cognition: While introspection
-may require meta-cognition (for instance to be able to expose its internal
-knowledge), it is not always mandatory. The current state of the robot can be
-represented as a simple instantiation of a specific category (for instance, if
-the robot gives an object to the human, this state could be represented with the
-triples \setstmt{robot performs action1,action1 isA Give}.
-
-\paragraph{Modeling of the robot capabilities}
-
-A particularly important aspect of self-knowledge for robots relates to the
-description of its own capabilities: which sensors/actuators/computation
-services exist and are currently available ?  While at a first level, these
-descriptions can be static (\eg the robot has one laser scanner and two arms),
-at more advanced levels, the description is updated and reflect the current
-(and possibly past and future) state of the robot. Note that these descriptions
-may also involve geometric descriptions (a kinematic chain, the pose of a
-device, etc.) that may be deported outside of the main knowledge base. Efforts
-trying to formalize, maintain and expose the capabilities and state of a robot
-are not new (and ground themselves in work and techniques for self-descriptive
-remote procedure calls in computing science), but take a renewed importance
-with applications for high-level multi-robot cooperation.
-
-Recent works by Kunze et al.~\cite{Kunze2011} seeks at defining a formal
-language to represent the capabilities of a robot.
-
-\subsubsection{Memory}
-\label{sect|memory}
-
-Memory has been studied at length in the cognitive psychology and
-neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
-introduce the idea of \emph{short-term} and \emph{long-term} memory,
-Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
-and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organises
-the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
-into a hierarchy. Short-term memory is refined with the concept of
-\emph{working memory} by Baddeley~\cite{Baddeley2010}
-(Figure~\ref{fig|memory_models}).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
-    \caption{Overview of the main types of memories, based
-    on~\cite{Atkinson1968, Anderson1976, Tulving1985, Baddeley2010}}
-
-    \label{fig|memory_models}
-\end{figure}
-
-It is worth emphasising that if memory is commonly associated to the process of
-forgetting facts after a variable amount of \emph{time}, it actually covers
-more mechanisms that are relevant to robotics, like selective remembering
-triggered by a specific context or reinforcement learning.
-
-Most knowledge representation systems offers some kind of memory as a pool of
-facts that are not forgotten by the robot until it is halted (this memory is
-often referred as a \emph{working memory}, but with a meaning unrelated to
-Baddeley's definition). Some systems may propose persistent storages that allow
-the robot knowledge to grow over time, while others may offer a larger range of
-memory categories, like short term memory (that lasts for a couple of seconds)
-or episodic memory (that allows the robot to selectively remember facts
-associated to specific events).
-
-In the larger field of cognitive architectures, the {\sc Soar}
-architecture~\cite{Lehman2006} is one of those that tries to reproduce a
-human-like memory organisation. The GLAIR cognitive architecture also have a
-concept of long term/short term and episodic/semantic memories.
+Note that the general strategies to actually \emph{construct} a
+knowledge model are treated in section~\ref{sect|design-strategies}.
+
+The dimension \taxon{B.1}{Roles}, with its sub-categories \taxon{B.1.1}{Space},
+\taxon{B.1.2}{Time} and \taxon{B.1.3}{Actions}, deals with the approaches that
+the KRS suggests to represent these three important \emph{roles} in robotics
+(spatial relations, time and representation of actions). To give a few examples
+(in no way exhaustive), \emph{Spatial relations} covers techniques like
+topological maps, allocentric and egocentric frame projections. \emph{Time
+representation} includes Allen's interval algebra~\cite{Allen1984}, Ghallab's
+chronicles~\cite{Ghallab1996} or the concept of \emph{fluent}. \emph{Actions}
+concern themselves with representation of \emph{pre-, post-conditions},
+\emph{thematic roles} or \emph{plans}.
+
+\taxon{B.2}{Context} describes if and how a given KRS represents the
+\emph{interpretation frame} of the situation and its own knowledge. This is
+often related to the initial \emph{common-sense knowledge} available to the
+robot. \taxon{B.3}{Modality} refers to the ability, for the KRS, to represent
+possible alternate worlds or belief states. A typical application of a
+\emph{modal} knowledge model is the implementation of a \emph{theory of mind},
+\ie an independent belief model for the other agents the robot interact with.
+\taxon{B.4}{Introspection} covers both \emph{self-knowledge} (what do I know
+about myself?) and a weaker version focused on the explicit representation and
+reasoning on the \emph{robot's own skills} (can I execute a given task?).
+Finally, \taxon{B.5}{Memory} refers to the implementation of
+different kind of bio-inspired memory models like \emph{declarative},
+\emph{procedural} or \emph{episodic}, and the attached processes
+(\emph{forgetting}, \emph{selectively remembering}, etc.).
 
@@
 
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf C. Reasoning}
-            child {node [taxon] {{\bf C.10}. Learning}}
-            child {node [taxon] {{\bf C.9}. Naive physics}}
-            child {node [taxon] {{\bf C.8}. Planning}}
-            child {node [taxon] {{\bf C.7}. Prediction and explanation}}
-            child {node [taxon] {{\bf C.6}. Presupposition}}
-            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
-            child {node [taxon] {{\bf C.4}. Uncertainty}}
-            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
-            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
-            child {node [taxon] {{\bf C.1}. Standard reasoning}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\subsubsection{Standard reasoning techniques}
-\label{sect|std-reasoning}
-
-We call \emph{standard reasoning techniques} techniques based on logical
-inference, using resolution algorithms like \emph{forward chaining},
-\emph{backward chaining} or \emph{semantic tableaux}.
-
-Main reasoning problems include \emph{concept satisfiability},
-\emph{consistency checking} and \emph{instance checking}.
-
-Concept satisfiability verifies if it is possible to find a non-empty
-\emph{interpretation} of a concept (or an expression defining a concept) in the
-knowledge model. For instance, the formula \concept{Plant} $\land$
-\concept{isRed}, which defines the concept of red plants, is satisfiable in a
-model \concept{KB} iff $\exists a, $ \concept{Plant}$(a) \land$
-\concept{isRed}$(a)$, \ie if we can find at least one red plant $a$ in our
-model.
-
-Checking the consistency of a model is equivalent to checking the
-satisfiability of each of the concept defined in the knowledge model.
-
-Instance checking consists in verifying that an individual $a$ is an
-interpretation of a concept (or concept expression) $C$ in the knowledge model.
-A typical example would be that we are provided with an instance
-\concept{object1} and we want to know if this object is a kind of
-\concept{Bottle} or \concept{Glass}.
-
-Inferences can also be drawn from other constructs, whose availability depends
-on the representation language. OWL, for instance, has constructs for:
-
-\begin{itemize}
-    \item class subsumption (to represent inheritance relations)
-
-    \item reasoning on roles properties, including:
-        \begin{itemize}
-        \item entailments based on roles domain and range (for instance, if the
-        domain of the role \concept{thinksOf} is known to be
-        \concept{ThinkingAgent}, then \par \concept{thinksOf}$(a, b) \to
-        $\concept{ThinkingAgent}$(a)$),
-
-        \item universal, existential and cardinality constraints,
-
-        \item several second-order predicates (inverse, symmetry, transitivity, etc.)
-
-        \end{itemize}
-
-    \item \emph{class restrictions} like: \par 
-	\footnotesize 
-	\concept{Bottle} $\equiv$ \concept{Artifact} {\bf that} (\concept{hasShape}
-	{\bf value} \concept{cylinderShape})\footnote{This example uses the \emph{Manchester
-    syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}}
-	\normalsize
-
-    \item set operations like: \par 
-    \footnotesize 
-    \concept{Color} $\equiv \bigcup$ (\concept{blue}, \concept{green},
-    \concept{orange}, \concept{black}, \ldots) 
-    \normalsize
-
-\end{itemize}
-
-\paragraph{Rule Languages} As mentioned earlier, knowledge models based on
-description logics can be extended through rule languages (typically for OWL,
-the SWRL language).
-
-An intersection of properties is an example of expression that can only be
-represented with rules. For instance:
-
-\footnotesize
-\concept{looksAt(?agt, ?obj)} $\land$ \concept{pointsAt(?agt,?obj)}
-$\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
-\normalsize
-
-\subsubsection{Dynamic instantiation and alteration of the knowledge structure}
-
-The content of a knowledge base is often conveniently divided into a
-\emph{structural part} that defines the conceptualisation of a domain in term
-of vocabulary and relations between the concepts, and an \emph{instantiation} of
-this structure into concrete entities.
-
-The terms TBox and ABox are commonly found to describe these two different types
-of statements in ontologies: TBox statements describe a system structure (made,
-for example, of a set of classes and properties) whereas the ABox contains
-TBox-compliant statements that are asserted in this structure.
-
-Knowledge representation systems allow to modify the ABox, which can be
-considered as the dynamic part of the knowledge base. Alteration of the ABox
-include the addition or retraction of relations between existing instances,
-and the addition or removal of instances. We call the later \emph{dynamic
-instantiation}: the capability for a system to create new instances at
-run-time. The term dynamic instantiation applies primarily to concrete entities
-(typically, a new object is discovered by the robot, a symbolic instance is
-created for it), but may also apply to abstract entities (like an instance of
-action, of a feeling, etc.).
-
-The knowledge representation system may also allow to alter the TBox. This
-requires the underlying reasoning systems to be able to dynamically take into
-account structural changes in the knowledge base.
-
-Example of TBox alteration include modification of the taxonomy (like addition
-or retraction of a \concept{subClassOf} relation), changes to the asserted
-domain or range of a predicate, addition or retraction of rules.
-
-Supporting TBox alteration has notable consequences on the learning
-capabilities of the system: teaching general facts (\ie, facts at the level of
-whole categories) like ``cars go on roads'' to a robot requires an alteration
-of the TBox.
-
-\subsubsection{Lazy evaluation}
-\label{sect|lazy-evaluation}
-
-\emph{Lazy evaluation} describes the ability for a KRS to delay active
+As one can expect, many reasoning techniques born in artificial intelligence
+have found their place in robotics over the years. We propose to consider ten
+subcategories of \taxon{C}{Reasoning} that reflects in our opinion important
+features in regards to the general requirements of robotics (like decision
+making being tractable in pseudo-real time, being able to deal with partially
+unknown or uncertain domain or specific reasoning related to interaction with
+the physical world).
+
+Again, the descriptions of these ten subcategories hereafter are very
+superficial. Readers can refer to~\cite{blabla} for the details.
+
+\taxon{C.1}{Standard reasoning} covers conventional reasoning techniques based
+on logical inferences (for instance, \emph{forward/backward chaining} or
+\emph{semantic tableaux}). The main problems that these techniques address are
+\emph{concept satisfiability}, \emph{consistency checking} and \emph{instance
+checking}. \taxon{C.2}{Instantiation and structural alteration} describes how
+dynamic a system is with respect to its internal knowledge structure (allowing
+for modification of the knowledge structure, often called \emph{TBox}, may
+impact the capability for the system to \emph{learn} for instance).
+\taxon{C.3}{Lazy evaluation} describes the ability for a KRS to delay active
 knowledge acquisition or reasoning operations until the value is actually
-needed.
-
-For instance, a system that computes the symbolic relative placement of two
-objects only when this fact is required to answer a query would be said to
-adopt a lazy evaluation strategy, whereas a system that computes {\it a priori}
-such relations (and by consequence, carries out such computation for possibly
-all known objects) would be said to use a \emph{strict evaluation} policy.
-
-Lazy evaluation has an immediate impact on efficiency and scalability of the
-system, and some problems may even be only tractable with a lazy evaluation
-strategy (the relative placements of object is an example of combinatory
-explosion in strict evaluation approaches, although this issue could be
-mitigated in real use-cases by heuristics that would select a subset of objects
-to evaluate).
-
-One downside is that the knowledge base never contains explicitly the complete
-set of beliefs of the robot. This limits for instance the ability for the robot
-to react to logical conditions that involve facts that are lazily evaluated:
-``trigger a callback when object A is behind object B'' would not be triggered
-with a purely lazy evaluation strategy for spatial relations. This can be
-mitigated by selecting (possibly dynamically) a subset of relations that are to
-be strictly evaluated.
-
-\subsubsection{Reasoning with uncertainty}
-
-\subsubsection{(Non) monotonic reasoning}
-
-\emph{Monotonic reasoning} means that addition of new assertions to a knowledge base
-can only extend the set of assertions that can be inferred, while a
-\emph{non-monotonic} reasoning scheme may lead to retraction of facts.
-McCarthy coined a famous example to illustrate the need of non-monotonic reasoning:
-
-\begin{quotation}
-Consider putting an axiom in a common sense database asserting that birds can
-fly. Clearly the axiom must be qualified in some way since penguins, dead birds
-and birds whose feet are encased in concrete can't fly. A careful construction
-of the axiom might succeed in including the exceptions of penguins and dead
-birds, but clearly we can think up as many additional exceptions like birds
-with their feet encased in concrete as we like. Formalised non-monotonic
-reasoning provides a way of saying that a bird can fly unless there
-is an abnormal circumstance and reasoning that only the abnormal circumstances
-whose existence follows from the facts being taken into account will be
-considered.
-\end{quotation}
-
-An important application of non-monotonic reasoning is the representation of
-change: for example, to make a brownie, one needs to crack eggs and mix them to the chocolate.
-The eggs disappear and are replaced by a dough:
-
-\concept{Egg}$(a) \land $ \concept{Chocolate} $(b) \land $
-\concept{MakeDough}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \land \lnot $
-\concept{Chocolate}$(b) \land $ \concept{Dough}$(c)$
-
-The insertion of the proposition \concept{MakeDough}$(a, b, c)$ leads to
-retraction of other facts. This rule requires non-monotonic reasoning to be
-applied.
-
-\emph{Default logic} is one of the formal logics that account for representing
-general truth and exceptions to it (for instance, \emph{tomatoes are red, in
-general}). However, due to computational complexity of these models (most of
-inferences in default logic are known to be $NP$-complete problem), classical
-logics and most of the existing reasoners do not allow non-monotonic reasoning.
-For instance, the SWRL rule language, usually associated to the OWL-DL ontology
-language, does not allow non-monotonic reasoning (only so-called DL-safe rules
-are allowed).
-
-\fxwarning{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
-languages? reasoner?}
-
-One important exception is the \emph{negation as failure} inference rule, as
-implemented by {\sc Prolog} for instance, that allows for non-monotonicity
-within the closed world assumption.
-
-\fxwarning{Give here an example of non-monotonic reasoning with Prolog}
-\fxnote{Do we mention here Answer Set Programming?}
-
-\subsubsection{Presupposition accommodation}
-\label{sect|presupposition-accommodation}
-
-\emph{Presupposition accommodation}~\cite{VonFintel2008} is the ability for a
-system to automatically create a context allowing to make sense of a
-proposition.
-
-Applied to robotics, we can imagine a human telling a robot ``Please get me the
-bottle that is behind you''. If the robot has not yet see what is behind it, it
-needs to assume (and represents in its knowledge model) that a undefined bottle
-can be found somewhere in the half of space behind it.
-
-A knowledge representation system able to cope with presupposition
-accommodation would be able to take into account this (usually under-defined)
-information that is not grounded into perception for later inferences.
-
-This ability to imagine a physically state of the world that is not actually
-perceived can be seen as the converse of the grounding ability.
-
-Note also that presupposition accommodation implies a bidirectional link of the
-symbolic knowledge model with a geometric (or physical) model of the
-environment. This article focuses on symbolic knowledge representation systems,
-but we shall mention when a KRS explicitly provides support for presupposition
-accommodation.
-
-\subsubsection{Prediction, projection, explanation}
-\label{sect|prediction-projection}
-
-Levesque~\cite{Levesque2008} distinguishes two main tasks related to reasoning
-on actions and consequences of actions, the \emph{projection task} and the
-\emph{legality task}.
-
-We call \emph{diagnosis} the converse of the projection task: the ability to
-track back the origin of a decision, and \emph{explanation} the more general
-ability to explicit a reasoning or a decision.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf C.6 Prediction and Explaination}
-            child {node [taxon] {{\bf D.4} Explaination}}
-            child {node [taxon] {{\bf D.3} Diagnosis}}
-            child {node [taxon] {{\bf D.2} Legality}}
-            child {node [taxon] {{\bf D.1} Projection}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\paragraph{Projection task}: determining whether or not some condition
-hold after a sequence of actions. The projection task is a typical
-non-monotonic reasoning task, since at each step, the system must add but also
-retract beliefs, as defined in the tasks post-conditions.
-
-\paragraph{Legality task}: determining whether a sequence of action can be
-performed starting in some initial state.
-
-\paragraph{Diagnosis}: this corresponds to the ability to rewind on past events
-in case of failure to provide possible explanation. This can be seen as the
-temporal reverse of the projection task. Because of their modelling of
-situations as an history of actions, derivatives of the GOLOG logic programming
-languages are a good example of the diagnosis task integrated to the knowledge
-representation system~\cite{Gspandl2011}.
-
-\paragraph{Explanation} Diagnosis is also linked to the \emph{explanation} or
-\emph{justification} capabilities that may be offered by the knowledge
-representation system. The explanation of an entailment is the sequence (or set
-of sequences if several are possible) of reasoning steps that allow to reach a
-conclusion. An explanation can also conversely explain why a statement leads to
-a contradiction.
-
-The following example shows an explanation for an inconsistency in a particular
-knowledge base: by adding the statement \stmt{robot1 belongsTo human1}, we
-observe that an inconsistency is triggered. The reasoner provides the four
-following observations to explain the inconsistency:
-
-\begin{quote}
-\scriptsize
-\begin{enumerate}
-    \item \concept{robot1} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
-    \item \concept{belongsTo} {\bf Domain} \concept{Artifact}
-    \item \concept{robot1} {\bf Type} \concept{Agent}
-    \item {\bf DisjointClasses:} \concept{Agent}, \concept{Artifact}
-\end{enumerate}
-\normalsize
-\end{quote}
-
-This explanation directly portrays the underlying structure of the knowledge
-model: we can understand that a robot is modelled here as an agent, that agents
-and artifacts are disjoint classes and that only artifacts can belong to
-someone, hence the inconsistency. A KRS may provide mechanisms to expose this
-kind of analysis, either automatically when inconsistencies occur, or on
-demand.
-
-Explanation of contradictions plays a particular role for robots: from a
-cognitive point of view, a logical inconsistency (\ie, a contradiction) can be
-viewed as a \emph{cognitive conflict} or \emph{cognitive dissonance} (\ie, two
-incompatible models of the world that must be dealt with). Being able to expose
-and explain such cognitive conflicts eases the control of the behaviour of the
-robot in unexpected semantic situations, and forms a first step towards an
-adequate reaction.
-
-Cognitive dissonance is also identified by developmental psychologists like
-Piaget as a motivation factor for a child to progress through the various
-stages of its cognitive development. This has been also studied in
-robotics~\cite{Oudeyer2007}.
-
-The cognitive capability of \emph{justification} is also intimately linked to
-the meta-cognition capability and participates to the overall \emph{cognitive
-observability} of the system.
-
-\subsubsection{Task planning}
-\label{sect|planning}
-
-Symbolic task planning is the ability for a robot to select a sequence of
-actions in order to reach a given final state. This capability is closely
-related to the previously mentioned projection and legality tasks: for a robot
-to plan, it must be able to build hypothetical states of the world that would
-follow from the successive application of actions (prediction task), and at
-each step, select possible, legal actions based on their pre-conditions
-(legality task). As already mentioned, these tasks are highly non-monotonic.
-
-Symbolic task planning in general is a large research
-field~\cite{Russell2009planning}. The so-called Classical Planning Problem,
-first, is characterised by a unique known initial state, durationless
-deterministic actions which can be taken only one at a time, and a single
-agent. STRIPS and PDDL are amongst the commonly used languages for representing
-such planning problems.
-
-Planning with nondeterministic durationless actions with probabilities can be
-represented as discrete-time \emph{Markov decision processes} (MDP), and when
-full observability is replaced by partial observability, we deal with
-\emph{partially observable Markov decision process} (POMDP).
-
-\emph{Hierarchical Task Networks} (HTN) are another common formalism for
-planning problems, where an initial set of tasks (\emph{High Level Tasks}, HLA)
-is decomposed into either primitive actions or a new set of subtasks.
-
-From the observation that the core reasoning techniques (back and forward
-chaining) are shared between planners and reasoners used in knowledge
-representation systems, task planning can be considered within the KRS.
-
-Since McCarthy's \emph{Situation Calculus} in 1963, numerous knowledge
-representation formalisms dedicated to representation and reasoning about
-actions and situation have emerged: besides situation calculus, \emph{fluent
-calculus} and \emph{event calculus} are the main ones.
-Thielscher~\cite{Thielscher2011} recently proposed a unification of these
-approach in a new \emph{action calculus}.
-
-The GOLOG~\cite{Levesque1997} language and its derivatives (like {\sc
-ReadyLOG}~\cite{Ferrein2008} and {\sc IndiLOG}~\cite{Gspandl2011}) propose
-implementations of the situation calculus focused on robotic applications.
-
-It is however also common to rely on external components dedicated to symbolic task
-planning (in particular because of the non-monotonicity requirements) with a
-tight link to the KRS for domain retrieval and/or resulting plan storage.
-
-Finally, in the context of interaction between several agents, the management
-of \emph{joint intentions} and \emph{joint goals}~\cite{Tomasello2005,
-Dominey2011} are additional aspects that have to be represented and
-appropriately handled by the planning subsystem.
-
-%An extensive report on task modeling in OWL ontologies in appendix
-
-\subsubsection{Physics-based reasoning}
-\label{sect|physics}
-
-As embodied entities, robots have to interact with physical entities.
-\emph{Naive physics reasoning} covers all the everyday reasoning the humans
-unconsciously perform, like taking into account gravity (``if I drop a ball, it
-falls down'') or common physical properties of objects (``a glass may break if
-dropped'', etc.). Many of the interactions with our everyday environments are
-ruled by such laws that are difficult to exhaustively encode.
-
-Some systems \cite{Kunze2011a} rely on external dedicated physics engine to
-compute symbolic facts from on-demand physics simulation. This requires a tight
-integration between the symbolic model and a geometric model that carries the
-geometries and physical properties of objects.
-
-\subsubsection{Learning}
-\label{sect|learning}
-
-\fxfatal{Learning section to write}
+needed, and may have a strong impact on the overall scalability of the system.
+\taxon{C.4}{Uncertainty} groups all the features related to reasoning over
+uncertain knowledge (like probabilistic inference).
+\taxon{C.5}{Non-monotonicity} discusses how systems deal with non-monotonic
+inferences, \ie exceptions to a rule or inferences that lead to the retractation
+of previously inferred facts. Techniques to deal with non-monotonic reasoning
+include for instance \emph{Default Logics}.  \taxon{C.6}{Presupposition
+accomodation} is the ability for a system to automatically create a context
+allowing to make sense of a proposition. This is tighly linked to the capability
+for a system to deal with \emph{under-specification}. \taxon{C.7}{Prediction and
+explanation} (with subcategories \taxon{C.7.1}{Projection},
+\taxon{C.7.2}{Legality}, \taxon{C.7.3}{Diagnosis} and
+\taxon{C.7.4}{Explanation}) relates to the capabilites for a system to predict
+and project the outcomes of a sequence of actions onto the logical system, and
+conversely, to justify and explain the sequence of entailments that lead to a
+conclusion. Related to this category, \taxon{C.8}{Planning} discusses to which
+extend a KRS can be used for \emph{symbolic task planning}, \ie the ability for
+a robot to select a sequence of actions in order to reach a given final state.
+\taxon{C.9}{Physics-based reasoning} describes the capability for a symbolic
+knowledge representation system to rely on physical reasoning (which may include
+geometry, dynamics, collision checking, etc.) to derive symbolic inferences.
+Finally, \taxon{C.10}{Learning} groups the various capabilities that lead to
+learning behaviours.
 
@@
 
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf D. Acquisition}
-            child {node [taxon] {{\bf D.3} Motivation}}
-            child {node [taxon] {{\bf D.2} Grounding}}
-            child {node [taxon] {{\bf D.1} Acquisition and fusion}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\subsubsection{Knowledge acquisition and modalities fusion}
-\label{sect|knowledge-acquisition}
-
 In the survey inclusion criteria, we have chosen to only consider robotic
-systems that acquire knowledge by themselves, at runtime.
-
-In our context, \emph{acquiring knowledge} means to build new logical
-statements from data sources and to anchor them into the existing knowledge. We
-consider three possible sources of data: proprioceptive/exteroceptive sensing,
-interaction with other agents, humans or robots, and remote knowledge bases.
-The acquisition process has generally two steps: the information acquisition by
-itself, and the \emph{transformation} of the information into knowledge,
-\emph{aligned} with the robot existing model (following our terminology for
-information and knowledge, as discussed in the introduction).
-
-It must be observed that knowledge acquisition is generally not done directly
-in the knowledge representation system. External components (often aggregated
-into knowledge acquisition \emph{pipelines}) are usually required to convert
-percepts into symbolic facts and to ground them.
-
-While we do not review in this article all these systems, the whole process of
-knowledge acquisition is central in cognitive robotic architecture and the
-design of knowledge representation systems can influence or be influenced by
-the approach to knowledge acquisition.
-
-In particular, complex robotic systems often require multi-modal perception
-capabilities (for instance, a robot can only interpret an utterance like ``this
-is a plate'' if it is able to understand gestures, understand natural language
-and merge them in a timely manner). Multi-modal interpretation can take place
-at various levels, but in many cases (especially if the modalities are of very
-different natures, like in the example above) merging will require
-symbolic-level reasoning. The KRS has a direct impact on the feasibility and
-ease of such operations.
-
-Let review shortly the three sub-categories of knowledge acquisition.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf D.1 Acquisition and fusion}
-            child {node [taxon] {{\bf D.1.3} Linked Resources}}
-            child {node [taxon] {{\bf D.1.2} Interaction}}
-            child {node [taxon] {{\bf D.1.1} Sensing}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\paragraph{Sensing}
-
-From the point of view of knowledge representation, the sensing capability can
-be split into proprioceptive sensing (\ie, sensing of the robot own internal
-state) and exteroceptive sensing (sensing of the robot environment). The
-(physical) introspection capabilities of the robot relies on the former.
-
-Exteroceptive sensing is the most obvious and largely studied mean of knowledge
-acquisition. Traditional sensing devices (IR, cameras, laser-scanners), while
-still present and widely used (navigation based on 2D localisation and obstacle
-avoidance is today the standard), are being step-by-step replaced by
-\emph{synthetic sensors}.
-
-These synthetic sensors include post-processing to provide higher-level
-percepts that ease the grounding. The prototypical example of such a device is
-the Kinect sensor. At a first level, it replaces traditional stereo vision
-algorithms by providing a fast, robust depth map. At a second level, it
-provides accurate, real-time tracking recognition and tracking of whole body
-poses of human.
-
-Other examples of such synthetic sensors exist: face recognition, off-the-shelf
-performant SLAM solutions, automatic cluster segmentation in point clouds (with
-the PCL library), etc.
-
-While the progresses of these sensing technologies are remarkable, one field of
-perception remain a very difficult challenge: accurate and generic object
-recognition. Most of the current approaches to object recognition rely on a mix
-of point cloud segmentation and fitting with visual feature recognition
-(SIFT-like algorithms), but it remains a slow and fragile task.
-
-\paragraph{Interaction}
-
-Interaction with other intelligent agents (humans or robots) is another
-important source of knowledge acquisition. It relies obviously on some form of
-sensing (from speech recognition to gesture recognition) but we distinguish it
-from the previous section because \emph{interaction} implies a form of
-communication. Communication is associated to specific functions
-(as shown on Jakobson's diagram, figure~\ref{fig|communication-model}), and in particular, it implies a shared
-\emph{context} (usually implicit) between interactors.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{jakobson_communication_model.pdf}
-
-    \caption{The Jakobson~\ref{Jakobson1960} communication model. In italics,
-    the associated \emph{communication functions}.}
-
-    \label{fig|communication-model}
-\end{figure}
-
-We distinguish between two main interaction channels: verbal and deictic.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf D.1.2 Interaction}
-            child {node [taxon] {{\bf D.1.2.2} Deictic Interaction}}
-            child {node [taxon] {{\bf D.1.2.1} Verbal Interaction}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\label{sect|nlp}
-
-The field of verbal interaction processing for robots spans from pattern-based,
-constrained sentences recognition to natural, bidirectional, unconstrained
-verbal communication. A large literature corpus exists on Natural Language
-Processing (NLP) which is presented at section~\ref{sect|dialogs-related-work},
-page~\pageref{sect|dialogs-related-work}.
-
-NLP is an established research field by itself, and while the robotic community
-is still lagging behind on many theoretical aspects, it brings one important
-aspect: the embodiment. Because the interactors, both the robot and the human,
-are establishing a communication within a shared physical context, the verbal
-communication channel is complemented by deictic channels, back channels and
-possibly shared physical experiences: a human can show something to a robot,
-saying ``Give me this''. This is not possible for a virtual agent.
-
-Several of the knowledge representation systems we present have developed
-specific built-in mechanisms or extensions to parse, ground and possibly
-rebuild natural language.
-
-Deictic (used in the literal meaning of ``display, demonstration, reference'')
-interaction is also an established field of research in human-robot
-interaction. Common deictic forms of communication~\cite{Li2012} include
-attentional focus (via face and gaze tracking) and joint attention, pointing,
-emotional expressions (based on face expressions, postures, emotional
-gestures).
-
-Like other knowledge acquisition modalities, the recognition and interpretation
-of deictic communication is rarely directly included in a KRS, but, as
-previously mentioned, the symbolic representation of such communication acts is
-relevant and important to achieve successful multi-modal interactions.
-
-\paragraph{Linked Knowledge Resources}
-\label{sect|lod}
-
-Robots, and in particular service robots, have usually an access (with possibly
-security-related constraints) to the World Wide Web and remote knowledge stores.
-
-The current shift towards the Semantic Web (\ie structured, annotated data that
-are easily machine-processable) makes increasingly easy to have robots to reuse
-autonomously this knowledge. The DBPedia project, for instance, illustrates
-well the tendency: it provides an automatically generated RDF version of the
-Wikipedia encyclopedia.
-
-In its current state, the relevance of the DBPedia project in our context is
-however limited: the triples that are extracted are mostly ``mechanical'' (like
-the categories of a term, or factual informations extracted from Wikipedia's
-InfoBox), and the vast majority of the knowledge actually contained in the
-encyclopedia pages remains out of reach of automated parsers. Efforts in that
-direction however exist~\cite{Nyga2009, Fader2011}.
-
-Another notable project that seeks at providing large amount of
-machine-friendly common-sense knowledge is the MIT's OpenMind
-project~\cite{Singh2002}. The project is designed to let the general public
-easily add common-sense statements in semi-controlled natural language, which is
-then processed to a publicly available ontology.
-
-Another approach that is easier at short-term to let robots remotely access
-knowledge repositories consists is building robot-specific shared repositories,
-with declarative and/or procedural (\ie, plan library) knowledge. The
-RoboEarth~\cite{Waibel2011} project is an example of such an effort.
-
-\subsubsection{Grounding/anchoring strategies}
-\label{sect|grounding}
-
-\emph{Grounding} (also called \emph{anchoring} when specifically referring to the
-building of links between \emph{percepts} and \emph{physical
-objects}~\cite{Coradeschi2003}) is the task consisting in building and
-maintaining a bi-directional link between sub-symbolic representations (sensors
-data, low-level actuation) and symbolic representations that can be
-manipulated and reasoned about~\cite{Harnad1990}.
-
-Being embodied entities with interaction with other embodied entities as a
-fundamental requirement, robots and robotic is deeply concerned by the
-grounding issue.
-
-Being actually implemented on real service robots, all the symbolic knowledge
-representation systems that we review in this study have some kind of grounding
-process. Numerous approaches exist, like amodal
-\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
-representations~\cite{Alami2011, Mavridis2006}, semantic maps
-(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
-affordance-based planning and object classification~\cite{Lorken2008,
-Varadarajan2011}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\columnwidth]{semanticmaps_hertzberg.png}
-    \caption{Example of a semantic map, taken from~\cite{Nuechter2008}.}
-    \label{fig|semanticmap}
-\end{figure}
-
-\subsubsection{Intrinsic motivation and curiosity}
-
-The reasons for a robot to acquire knowledge are diverse, and usually external
-to the robot itself. It is often driven by the requirements of tasks that the
-robot has to execute. In this case, motivations are managed by the execution
-controller and are mostly invisible to the knowledge representation system.
-
-However, motivation can also be intrinsic, driven by the internal state of the
-robot's beliefs, without external pressure or reward. In this case, the \fxwarning{to be completed}
-
-In the context of robotics, Oudeyer notes however that \emph{the information that is
-compared \emph{[to compute a level of motivation]} has to be understood in an
-information theoretic perspective, in which what is considered is the intrinsic
-mathematical structure of the values of stimuli, independently of their
-meaning.}~\cite{Oudeyer2007}.
-
-Psychological grounds of motivation are summarized in~\cite{Oudeyer2007} while
-the main approaches to computational motivation, divided into knowledge based
-models, competence based models and morphological models, are surveyed
-in~\cite{Oudeyer2008}.
+systems that acquire knowledge by themselves, at runtime. The category
+\taxon{D}{Acquisition} reviews the different approaches to knowledge acquisition
+and semantics extraction. \taxon{D.1}{Acquisition and fusion} examines the
+techniques to first acquire new logical statements and then \emph{align} them to
+the corpus already available to the robot. Depending on the source of knowledge,
+we distinguish three further subcategories: \taxon{D.1.1}{Sensing} covers the
+techniques that translate perceptions into symbolic knowledge,
+\taxon{D.1.2}{Interaction} focuses on knowledge specifically stemming from
+interaction (typically, language processing), and \taxon{D.1.3}{Linked
+Resources} relates to the connections and alignement issues with external
+(often Web-based) knowledge bases.
+
+\taxon{D.2}{Grounding} discusses the KRS capabilities in regards to attaching a
+practical meaning to the symbol it manipulates. In the context of robotics, it
+mainly consists in building and maintaining a bi-directional link between
+sub-symbolic representations (sensors data, low-level actuation) and symbolic
+representations.
+
+Finally, \taxon{D.3}{Instrinsic Motivation} describes the possible
+\emph{internal} mechanisms by which the robot seeks by itself to acquire more
+knowledge.
 
@@
 
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf E. Integration}
-            child {node [taxon] {{\bf E.4} Performances}}
-            child {node [taxon] {{\bf E.3} Monitoring}}
-            child {node [taxon] {{\bf E.2} Executive layers}}
-            child {node [taxon] {{\bf E.1} Sensori-motor}};
-\end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-Knowledge representation systems do not mean anything to robots if they are
-considered in isolation. This section proposes categories of features related
-to the integration of the KRS into a larger software architecture that includes
+Knowledge representation systems have limited usefulness for robotics if they are
+considered in isolation, and it seems important to also evaluate KRS from the
+perspective their integration into larger control architectures that includes
 perception routines, decision-making processes and actuation control.
-
-We also mention some practical aspects of a real-world system, like
-performances and monitoring tools that come along with the KRS.
-
-\subsubsection{Integration with sensori-motor layers}
-\label{sect|integration-sensorimotor}
-
-We have previously discussed (section~\ref{sect|grounding}) the principles of
-the grounding process that aims at establishing and maintaining a connection
-between percepts (and to a lesser extend, low-level actions) and symbols.
-
-While every real-world cognitive robot need some kind of grounding, the actual
-implementations lead to very different information flows.
-
-The systems can be roughly split into two classes: \emph{passive} knowledge
-repositories that process symbolic facts produced by lower-level sensori-motor
-layers (\emph{push} flow); \emph{active} knowledge managers that directly query
-(possibly by polling or on-demand) low-level layers.
-
-This macroscopic distinction is however mostly a matter of defining the
-frontiers of the KRS: some systems like KnowRob~\cite{Tenorth2009a} encompass
-geometric reasoning layers that would be considered as external by other
-systems like ORO~\cite{Lemaignan2010} that focus on the symbolic fact storage
-and rely on a ecosystem of independent modules to provide and consume symbolic
-knowledge.
-
-\fxwarning{Some archi with the ability to ``listen'' to the robot internal
-structures -> which ones?}
-
-Other systems do not fit either in such a partition between active and passive
-systems because they do not stand as independent modules but exist as diffuse,
-\emph{ubiquitous} knowledge manipulation system (case of the CAST knowledge
-model~\cite{Jacobsson2008}), for instance because they are primarily
-language~\cite{Ferrein2008, Sabri2011}.
-
-\subsubsection{Integration with executive layers}
-\label{sect|integration-executive-layers}
-
-Conversely, the knowledge management module need a tight integration with the
-decision-making processes. As for the integration with sensori-motor
-layers, the borders of the KRS can be fuzzy and vary from one architecture to
-another: many consider symbolic task planning as an integral role of the KRS,
-while other have dedicated extensions for planning, some integrate learning as
-an on-the-flight process that is part of the KRS, others as an independent
-deliberative process, etc.
-
-The actual integration techniques vary also widely, from language extensions
-(like the integration of CRAM~\cite{Beetz2010} with KnowRob) and client-server
-architectures, to event-driven models (SHARY and ORO~\cite{Alami2011}). Choices
-at this level have notable consequences on the whole design of the upper
-control architecture of the robot, in particular regarding its modularity and
-the ease of addition of new components.
-
-\subsubsection{Monitoring and debugging}
-\label{sect|debugging}
-
-It is common to have knowledge representation systems at the heart of a
-cognitive robotic architecture, and therefore KRS are easily ``buried'' in the
-system.
-
-At the same time, the symbolic model often provides a valuable synthetic view
-on the whole state of the robot, furthermore easily understandable by the human
-developer (the fact \stmt{human1 isSitting true} is easier to interpret than
-the suite of relative coordinates of each joints of the human skeleton, as
-provided by the human tracker, for instance).
-
-Tools to trace and visualise at run-time the evolution of the knowledge
-structure and contents may be available with the KRS, as well as
-post-processing tools that run on the trace to analyze {\it a posteriori} the
-cognitive behaviour of the robot.
-
-\subsubsection{Evaluation of performances}
-\label{sect|performances}
-
-Benchmarks of symbolic systems for robots are hard to conduct for several
-reasons: identifying good metrics for robotic experiments in general is
-difficult because of the complex interactions between tenth of modules running
-in parallel, and isolating one specific component is difficult.  Also,
-knowledge representation systems are often tightly coupled to the other
-modules. To quote Langley~\cite{Langley2006}:
-
-\begin{quote}
-The conventional wisdom of software engineering is that one should
-develop independent modules that have minimal interaction. In contrast, a
-cognitive architecture offers a \emph{unified} theory of cognition with tightly
-interleaved modules that support synergistic effects.
-\end{quote}
-
-The lack of standard API for knowledge services makes it also hard to switch
-between KRS to compare them.
-
-Finally, because service robots are designed to act in rich, dynamic
-environments, possibly with humans, building repeatable experiments is
-challenging, and quantitative measurements are often not the right
-metric~\cite{Langley2006}.
-
-We will however present here some quantitative metrics (related to scalability,
-for instance), followed by qualitative evaluation approaches, grouped under the
-term \emph{Cognitive Performances}.
-
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf E.4 Performance \\Evaluation}
-            child {node [taxon] {{\bf E.4.2} Cognitive Performances}}
-            child {node [taxon] {{\bf E.4.1} Raw Performances}};
- \end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-\paragraph{Raw Performances} The \emph{raw performance} is evaluated on
-quantitative benchmarks. The main metric is the \emph{scalability}
-$\mathcal{S}^{KRS}$ of the system with the size of the knowledge base
-$\sigma=|\Delta^{\mathcal{I}}|$  (in term of atoms or statements).
-
-We call \emph{relaxation time} $\mathcal{R}^{\mathcal{M}}$ the (averaged) time
-required by the system after a model modification of type $\mathcal{M}$ before
-being available for further interaction, and \emph{query time}
-$\mathcal{Q}^{\mathcal{M}'}$ the (averaged) time to execute a query of
-complexity $\mathcal{C}$ on the KRS. The type $\mathcal{M}$ of model
-alteration is either an ABox modification (addition/removal of an instance) or
-a TBox alteration (addition/removal of a class, a class restriction or a rule).
-%Query complexity $\mathcal{C}$ is...\fxwarning{find references}.
-
-\emph{Temporal scalability} is defined in term of the nature of the function
-$f^{\mathcal{M}, \mathcal{C}}(\sigma) = \mathcal{R}^{\mathcal{M}}(\sigma) +
-\mathcal{Q}^{\mathcal{C}}(\sigma)$ (\ie the relation of the relaxation time
-and query time to the knowledge base size). \emph{Space scalability} is the
-relation of memory consumption to the size $\sigma$. The scalability is tightly
-coupled to the expressiveness of the underlying knowledge model, which need to
-be known for the scalability measurement to be meaningful.
-
-Because of coupling and repeatability issues we have mentioned, raw
-performances of KRS are often benchmarked with synthetic datasets (which leads
-to another issues: how to assess the meaningfulness of the performance of a
-reasoner on an artificial ontology?~\cite{Bail2010}) or ``toy'' experiments
-that do not always model the whole complexity of real-world
-application~\cite{Chong2009}.
-
-\paragraph{Cognitive Performances} While evaluating the raw performances of
-knowledge representation systems in a relevant manner may be difficult, the cognitive
-performances of the robot as a whole can be also evaluated.
-
-Langley et al.~\cite{Langley2006} propose five such dimensions of evaluation:
-the \emph{generality} of the system (can it adapt easily to new tasks?), the
-\emph{rationality} or relevant of the inference/reasoning/decisions the system
-take, the \emph{reactivity} and \emph{persistence} that evaluates if the
-behaviour of a cognitive system is appropriate under unpredicted changes, the
-\emph{improvability} of the system as a function of the knowledge added to it,
-and finally, the resulting \emph{autonomy} of the system.
-
-Cognitive performance can also be evaluated with the support of tools developed
-in cognitive psychology. Several standard tests (like False-Belief
-experiments~\cite{Leslie2000} or the Token test~\cite{DiSimoni1978}) have been
-used to judge the cognitive abilities of robots~\cite{Mavridis2006,
-Breazeal2006}.
+\emph{Practical} integration also suggests to take into account real-world
+constraints, like performances and monitoring tools that come along with the KRS.
+
+\taxon{E}{Integration} is hence subdivided: The integration of the KRS with the
+\taxon{E.1}{Sensori-motor} layers may adopt various paradigms (from passive
+\emph{semantic blackboards} to active KRS that directly query the sub-symbolic
+layers. \taxon{E.2}{Executive layers} examines conversely how the KRS is used
+from the perspective of the robot controller. Here also, many different
+approaches have been investigated, from classical server-client architecture, to
+DSL-like language extensions. \taxon{E.3}{Monitoring} details the support tools
+provided by the system for the developpers to monitor and introspect their
+system. This is especially important for knowledge representation systems
+because they often lie at the heart of the architecture, and act as hubs for
+many other components. \taxon{E.4}{Performances}, finally, evaluates the
+performances of the KRS, both in term of scalability and classification speed
+\taxon{E.4.1}{Raw performances}, and in term of cognitive abilities provided by
+the system \taxon{E.4.2}{Cognitive performances}. Note that performances of KRS
+are notoriously difficult to assess since on one hand, cognitive architectures
+for robotics are often build upon tightly interleaved modules which makes it
+difficult to test a KRS in isolation, and on the other hand, we lack standard
+benchmarks and scales for cognitive capabilities.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{scriptsize}
-\begin{center}
-\begin{tikzpicture}[taxonomy]
-    \node [taxon] {\bf F. Knowledge instantiation}
-            child {node [taxon] {{\bf F.4} Granularity}}
-            child {node [taxon] {{\bf F.3} Metrics}}
-            child {node [taxon] {{\bf F.2} Common-sense and Alignement}}
-            child {node [taxon] {{\bf F.1} Design Strategy}};
- \end{tikzpicture}
-\end{center}
-\end{scriptsize}
-
-This last branch of our taxonomy looks at the actual \emph{content} of the
+The last category of our taxonomy looks at the actual \emph{content} of the
 knowledge base: the knowledge instantiation. Here, \emph{instantiation} does
@@
 
-While we have previously mentioned features of knowledge representation systems
-that enable the robot to fill its knowledge base with content and alter the
-knowledge structure, most of the systems also come with a certain amount of
-initial knowledge that often includes \emph{common-sense} knowledge (\ie facts
-that widely known to humans, and hence often implicit: ``to put a cake in the
-oven, one must first open the oven's door'').
-
-The design strategy, the choice to rely on common-sense knowledge or not, the
-reuse of standard ontologies, the quantity of {\it a priori} knowledge are many
-parameters that lead to different knowledge models.
-
-\subsubsection{Design strategy}
-\label{sect|design-strategies}
-
-The main challenge of knowledge representation can be summarised as \emph{How
-to model the real-world state and interactions in a symbolic way, processable
-by the robot to make decisions}. We have already introduced the term
-\emph{grounding} to describe the (bi-directional) process of binding percepts
-to symbols.
-
-We have also seen that the instantiation of the knowledge (\ie, the actual,
-practical knowledge available to the robot) comes either from some variant of
-perception plus grounding, or from knowledge that the developer considers as
-already meaningful for the robot: remote semantic databases or initial
-(common-sense or situation specific) knowledge.
-
-This translates into two main strategies to drive knowledge instantiation: a
-\emph{top-down} design or a \emph{bottom-up} design.
-
-A bottom-up approach to knowledge instantiation takes the output of sensors as
-the primary source of knowledge: instances of objects, agents, and their
-relations directly result from what is perceived. No abstract, ``from the more
-generic to the more specific'' process takes place.
-
-Steels~\cite{Steels2007} considers this approach to be a solved issue, and
-according to Sloman~\cite{Sloman2007} (and his stance against the
-\emph{``Symbol Grounding meme''}), since bottom-up grounding boils down to grounding
-of \emph{somatic} concepts (\ie roughly, the sensori-motor relationships that
-the robot learns from its interaction with the world), it constrains in an
-unacceptable way the range of concepts accessible to the robot.
-
-Knowledge instantiation can also be approached as a \emph{top-down} activity:
-in natural language grounding, for instance, the robot needs to automatically
-bind an abstract representation (a group of words uttered by a human) to an
-unambiguous, context-dependent, internal concept. This concept may (or may not)
-be \textit{a priori} available to the robot as a pre-loaded ontology (what we
-previously called the cultural background of the robot). In any case, the robot
-must conduct a cognitive process that leads from an abstract concept to a
-concrete entity.
-
-The bottom-up and top-down strategies also reflect how the knowledge structure
-itself is constructed: either by successive classification and refinement of
-percepts, or from generic categories, typically extracted from standard
-upper-ontologies.
-
-\subsubsection{Common-sense and alignment with standard upper-ontologies}
-
-At section~\ref{sect|lod} we have presented how remote knowledge bases are a
-valuable source of knowledge, including common-sense knowledge, that robot can
-extract.
-
-Building {\it a priori} knowledge with a top-down strategy leads to populate
-the upper part of the taxonomy with abstract concepts like \emph{Thing},
-\emph{Time}, \emph{Person}, etc. The organisation of the whole knowledge
-depends on the design of this abstract part of the common-sense knowledge.
-
-Different knowledge structures at this level can lead to serious
-misunderstanding: for instance, if one robot considers the concept of a
-\emph{person} as representing some intelligent, possibly disembodied, entity,
-and another robot represents a \emph{person} as a subclass of mammals, their
-models of the world are likely to be conflicting.
-
-To be able to successfully exchange knowledge with other systems (robots,
-databases, natural language parsers, etc.), the common-sense knowledge of
-robots is thus often aligned with a standard upper-ontology.
-
-Many such upper-ontologies exist (table~\ref{table|upper_onto_metrics} lists
-some of them). While {\sc Cyc} and SUMO are current the main two, many efforts
-take place to make these ontologies compatible with each other (in particular
-by using the {\sc WordNet} thesaurus as intermediate unambiguous source of
-semantics).
-
-\subsubsection{Metrics and quality criteria}
-
-Qualitative and quantitative metrics give an insight on the size, complexity
-and effectiveness of ontologies used with the robots.
-
-Table~\ref{table|upper_onto_metrics} gives such metrics for five major
-upper-ontologies commonly used in the semantic Web community (taken
-from~\cite{Mascardi2007} and the projects' respective websites). It is
-interesting to note the large variations between projects like {\sc DBpedia}
-whose content is automatically generated from Wikipedia, {\sc Cyc} or {\sc
-SUMO}, which are both hand written, but do not adopt the same strategy to
-select the knowledge to represent.
-
-These metrics do not reflect adequately the \emph{expressive complexity},
-though: in most of these ontologies with a large amount of terms and
-assertions, taxonomic relations (\concept{isA}) or technical predicates (URI,
-translations, etc.) account for a large part of the assertions, at the expense
-of real semantic relations.
-
-\begin{table}
-\begin{center}
-
-\begin{tabular}{llllll}
-\toprule
-{\bf Project} & {\bf Terms} & {\bf Assertions} (triples) \\
-\midrule
-{\sc Cyc} & > 300 000 & > 3 000 000 & \\
-YAGO & > 10 000 000 & > 120 000 000 \\
-SUMO & 20 000 & 60 000 \\
-{\sc DBpedia} (for English) & 1 840 000 & 385 000 000\\
-{\sc OpenMind} Common Sense (for English) & & 1 000 000\\
-\bottomrule
-
-\end{tabular}
-\end{center}
-\caption{Raw size of major upper-ontologies.}
-\label{table|upper_onto_metrics}
-\end{table}
-
-Other metrics that include the type of predicates that are used, or the
-computed DL expressiveness, can be more significant (some are presented in
-table~\ref{table|onto-stats}, page~\pageref{table|onto-stats}).
-
-Qualitative evaluation of ontologies is also a well studied field. In Staab's
-{\it Handbook on Ontologies}, Vrande{\v{c}}i{\'c}~\cite{Vrandevcic2009}
-provides a synthesis of qualitative criteria for ontology assessment.  He lists
-eight of them: \emph{accuracy, adaptability, clarity, completeness,
-computational efficiency, conciseness, consistency,} and
-\emph{organizational fitness}. Details and relevant literature can be found
-in Vrande{\v{c}}i{\'c}'s chapter.
-
-A last qualitative facet relates to the possible bias of an ontology
-regarding the domain it pretends to cover. For instance, Wikipedia (and
-consequently, DBPedia and YAGO) is known to have a bias towards contents
-related to people, geography or historic events. {\sc OpenCYC} also suffer of
-non-homogeneous concept coverage.
-
-\subsubsection{Granularity}
-
-Amongst the characteristics of ontologies, knowledge \emph{granularity}
-qualifies the level of details or refinement of the knowledge stored. The level
-of granularity of a robot's ontology hints on the place of the symbolic layer
-in the whole robotic architecture: some systems (like OMRKF~\cite{Suh2007}) go
-as down as storing SIFT features (\ie a large volume of numerical values) in
-the ontology. The storage of literal values is indeed a relevant case for
-robotics. Depending on the representation language, literal values can be
-naturally represented and processed (common case in logic programming language)
-or not (storing numerical value, let alone matrices, in OWL is cumbersome and
-not efficient due to the serialisation to XML).
-
-The issue of the granularity of models can also be partially addressed by
-splitting the knowledge representation into a geometric level (where all
-numerical values are stored) and a purely symbolic level. The communication
-between the two layers is however a complex question.
-
-It must finally be noted that complex robotic systems are likely to use
-different level of the knowledge depending on the task to achieve, and the
-granularity of the knowledge should probably be considered as a dynamic
-property.
+The \taxon{F.1}{Design strategy} of the KRS can be \emph{top-down} (from an
+initial symbolic model down to the percepts), \emph{bottom-up} or  a mix of
+both. The bottom-up and top-down strategies reflect how the knowledge structure
+itself is constructed: either from generic categories, typically extracted from
+standard upper-ontologies or by successive classification and refinement of
+percepts.
+
+\taxon{F.2}{Common-sense and alignment} discusses if and how the KRS reuse
+existing pools of knowledge like \emph{upper-ontologies}, and how the question
+of knowledge \emph{alignment} is handled. \taxon{F.3}{Metrics and quality
+criteria} includes qualitative and quantitative metrics that give an insight on
+the size, complexity and effectiveness of knowledge base available to the robot.
+These metrics include for instance the number of terms and assertions, but also
+the type of predicates that are used, or the computed Description Logics
+expressiveness (when applicable).  Qualitative evaluation of ontologies is a
+research field in its own right, and introduce criteria like \emph{accuracy,
+adaptability, clarity, completeness, computational efficiency, conciseness,
+consistency}. The last category is \taxon{F.4}{Granularity}, that accounts for
+the level of details considered in the knowledge representation system, that can
+go down to storing SIFT features~\cite{Suh2007} directly in the KRS.
 

### 2014-04-16 krs-survey.tex
Table of systems fit on a double column

@@
 
-\begin{landscape}
-\begin{table}\scriptsize
+\begin{table*}\scriptsize
 \begin{center}
 
-\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{1.5cm}}
+\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}p{2.4cm}p{3.4cm}p{1.5cm}}
 \toprule
-{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
+{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
 \midrule
-ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
-CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & \cite{Jacobsson2008} \\
-GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & & \cite{Mavridis2006} \\
-Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
-{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
-NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & \cite{Sabri2011} \\
+ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
+CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ.,
+DFKI Saarbrücken) & & Amodal proxies & \cite{Jacobsson2008} \\
+GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & \cite{Mavridis2006} \\
+Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
+{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
+NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & NKRL & & \cite{Sabri2011} \\
 %OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
-ORO & Formal & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) + {\sc Pellet} & \cite{Lemaignan2010} \\
-OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
-PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
+ORO & Formal & Lemaignan, Alami \par (LAAS-CNRS) & {\sc Java} & OWL-DL ({\sc Jena}) + {\sc Pellet} & \cite{Lemaignan2010} \\
+OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
+PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
 %Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
@@
 \label{table|surveyed-systems}
-\end{table}
-\end{landscape}
+\end{table*}
 

### 2014-04-16 krs-survey.tex
Fixed cite, redundant table,...

@@
 
-\title{Knowledge Representation for Service Robots: Supply and Demand}
-\author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz, Rachid Alami}
+\title{Knowledge Representation for Service Robots: A Survey}
+\author{Séverin Lemaignan, Moritz Tenorth, Rachid Alami, Michael Beetz}
 
@@
 
-\tikzstyle {taxonomy} = [grow=right, level distance = 3cm, sibling
-distance=7mm]
-\tikzstyle {completetaxonomy} = [grow=right]
-\tikzstyle {taxon} = [text width=3cm, text centered,inner sep=0pt, outer
-sep=2pt] %, draw=blue!50,thick]
-
 \maketitle
-\tableofcontents
 
@@
 may have interactions). The typical scenario that would involve such robots is
-a service robot in a human-friendly environment like a kitchen (we will
-describe such a scenario in section~\ref{sect|scenario}).
+a service robot in a human-friendly environment like a kitchen.
 
@@
 We have also chosen not to include the GOLOG language and its
-derivatives~\cite{Levesque1997, Ferrein2008, Gspandl2011} in this survey. While
+derivatives~\cite{levesque1997golog, Ferrein2008, Gspandl2011} in this survey. While
 several implementations on robots, including service robots, do exist, the
@@
 and introduce accordingly sub-categories. We kindly refer the interested reader
-to~\cite{blabla} for in-depth discussion of each of these dimensions of
-knowledge representation systems, as well as related literature.
+to~\cite{lemaignan2012symbolic} for in-depth discussion of each of these
+dimensions of knowledge representation systems, as well as related literature.
 
@@
 cognitive capabilities (theory of mind, introspection, memory).
-Note that the general strategies to actually \emph{construct} a
-knowledge model are treated in section~\ref{sect|design-strategies}.
 
@@
 Again, the descriptions of these ten subcategories hereafter are very
-superficial. Readers can refer to~\cite{blabla} for the details.
+superficial. Readers can refer to~\cite{lemaignan2012symbolic} for the details.
 
@@
 CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ.,
-DFKI Saarbrücken) & & Amodal proxies & \cite{Jacobsson2008} \\
+DFKI Saarbrücken) & & Amodal proxies & \cite{jacobsson2008crossmodal} \\
 GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & \cite{Mavridis2006} \\
@@
 instead knowledge as unrooted \emph{proxies}. Those proxies are formally
-defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
+defined in \cite{jacobsson2008crossmodal} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
 a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%% Underlying knowledge model table %%%%%%%%
-\begin{table}
-\begin{center}
-
-\begin{tabular}{lp{4cm}}
-\toprule
-{\bf Project} & {\bf Common-sense \par knowledge source} \\
-\midrule
-{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
-ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
-PEIS Ecology & {\sc ResearchCyc} \\
-NKLR &  None \\
-CAST Proxies &  None \\
-GSM &  Predefined categories \\
-OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
-Ke Jia & None \\
-ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
-
-\bottomrule
-
-\end{tabular}
-\end{center}
-\caption{Underlying knowledge sources for each project}
-\label{table|knowledge-sources}
-\end{table}
 
@@
 
-\begin{tabular}{ll}
+\begin{tabular}{lp{4cm}}
 \toprule
-{\bf Project} & {\bf Common-sense knowledge source} \\
+{\bf Project} & {\bf Common-sense \par knowledge source} \\
 \midrule
-ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
+{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
+ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
+PEIS Ecology & {\sc ResearchCyc} \\
+NKLR &  None \\
 CAST Proxies &  None \\
 GSM &  Predefined categories \\
+OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
 Ke Jia & None \\
-{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology, physics simulation \\
-NKLR &  None \\
-OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
-ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
-PEIS Ecology & {\sc ResearchCyc} \\
+ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
 
@@
 \end{center}
-\caption{Underlying common-sense knowledge sources for each project.}
+\caption{Underlying knowledge sources for each project}
 \label{table|knowledge-sources}
@@
 in conjunction with visual perception systems (mentioned for instance, in the
-CAST project, in~\cite{Jakobsson2008}). {\sc KnowRob}, by filling its pool of
-facts from informations automatically extracted from the Web~\ref{Nyga2009},
+CAST project, in~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
+facts from informations automatically extracted from the Web~\cite{Nyga2009},
 can also be considered as a \emph{learning} system (it autonomously acquire

### 2014-04-16 krs-survey.tex
Fixed ORO specifics and proof-read discussion

@@
 
-Section~\ref{sect|features} introduces a typology of requirements and features
-of knowledge representation systems for robotics, that we use as foundations to
-compare and evaluate the systems we examine in this survey.
-
-Section~\ref{sect|surveyed-systems} then presents each of the systems we have
-examined, and underlines their specificities.
-
-We finally summarize in section~\ref{sect|conclusion} the current approaches in
-order to identify several research directions that are not sufficiently
+Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
+representation in robotics. It introduces briefly a range of different
+approaches that do not focus on a specific field of robotics, and instead
+provide insights on the wealth of knowledge representation techniques available
+to researchers.
+
+Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
+systems deployed on \emph{service robots}: Section~\ref{sect|features}
+introduces a extensive typology of requirements and features of knowledge
+representation systems, that we use as foundation to compare and evaluate the
+systems we examine, and section~\ref{sect|surveyed-systems} then presents each
+of the eight systems, and underlines their specificities.
+
+We finally summarize in section~\ref{sect|conclusion} the current approaches and
+we attempt to identify several research directions that are not sufficiently
 addressed by the current state of the art.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Knowledge Representation in Robotics: an Overview}
+\label{sect|krsoverview}
+
+TDB
+
 \section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Towards the next generation of Knowledge Representation Systems for Robotics}
+\section{Synthesis: How Current Systems Tackle the Knowledge Representation
+Challenge?}
 \label{sect|conclusion}
 
-This section tries to summary the various approaches we surveyed in the
-previous sections and to draw some research perspectives. This section does not
-exactly follow the list of compared features presented in
-section~\ref{sect|features}. We have organized it along four axis: \emph{what
-can be represented?}, \emph{How knowledge is created and grounded?}, \emph{What
-can be done with the knowledge?} and \emph{How to use knowledge in the whole,
-larger robot architecture?}.
-
-The table~\ref{table|contribution-by-systems} sketches the landscape of current
-research. The main fields of \emph{contribution} for the systems we have
-surveyed are sorted along the different axis, dimensions of knowledge
-representation that were identified in previous section.
+This section summarizes the different approaches to knowledge representation and
+reasoning that are surveyed in the previous section, and attempt to identify
+research perspectives.
+
+First, the Table~\ref{table|contribution-by-systems} presents, system by system
+and feature by feature, a landscape of current research, underlining the main
+design choices and scientific \emph{contribution} of each of the KRS that we
+have examined.
+
+Then, we go over again the nine ``challenges'' proposed by McCarthy and Roy (see
+section~\ref{sect|evaluation-literature}) to evaluate how far we are on the road
+to natural interaction and, to slightly paraphrase McCarthy, ``human-level
+robots''.
+
+\subsection{Logic formalisms, continuous world}
+
+McCarthy underlines the necessity of relying on mathematical logic, as the most
+fruitful formalism for machine intelligence. This is largely the case. Almost
+all the systems we have surveyed rely on logical formalisms, in several cases
+as a mix of declarative languages (often based on Decription Logics) and
+logical programming languages (Prolog, expectedly, but not only).
+
+The two exceptions (the CAST knowledge model and GSM) are however 
+interesting: CAST proposes a pervasive model of knowledge that is convincing from
+the grounding perspective (but likely suboptimal for deliberative tasks), and
+GSM encodes knowledge in an amodal (continuous, geometric) model while
+preserving features that are usually specific of symbolic models (like theory
+of mind or categorical knowledge).
+
+The interleaving of discrete symbolic models with the continuous physics of the
+world remains a major challenge, and techniques vary a lot: besides the CAST
+proxies and the GSM amodal model, most systems directly extract physical
+attributes of the environment to insert them in the symbolic model ({\sc Tapas},
+Ke Jia, OUR-K, PEIS), thus skipping altogether an intermediate geometric model.
+ORO has tight (but unidirectional) links with an intermediate geometric
+representation (called {\sc Spark}), and {\sc KnowRob} adopt a top-down approach
+where local geometric models are set up on-demand to compute symbolic properties
+like relative locations.
+
+\subsection{Management of uncertainty and approximation}
+
+McCarthy also mentions the necessity to deal with \emph{approximate concepts
+and approximate theories} (that includes representing them and reasoning with
+them).
+
+The most notable attempt at modeling uncertainty down to the knowledge
+representation is {\sc ProgCog}, an extension of {\sc KnowRob}. Theoretical
+and practical performances issues (including decidability) remain however
+difficult to overcome.
+
+It can be also noted that some effort do exist to provided probabilistic
+reasoning to Description Logics based systems (like the {\sc Pronto}
+reasoner~\cite{Klinov2008}), but this does not appear to be a major focus in the
+currently available tools.
 
@@
 
-\caption{Main domains of contribution of current KRS. Italics mean that the
-feature is implemented as an external module. Main references are given in the
-table header. When relevant, feature-specific publications have been provided.
-An empty cell means that either the system has no specific focus on this domain
-or we could not find relevant literature.}
+\caption{Design choices and main domains of contribution of current KRS (+
+indicates \emph{moderate contribution}, +++ indicates \emph{major
+contribution}). Italics mean that the feature is implemented as an external
+module. Main references are given in the table header. When relevant,
+feature-specific publications have been provided.  An empty cell means that
+either the system has no specific focus on this domain or we could not find
+relevant literature.}
 
@@
 \end{landscape}
-
-To comment this table, we propose select nine topic based on some of the
-challenges that McCarthy and Roy have identified (there are listed at
-section~\ref{sect|evaluation-literature}) on the road to natural interaction
-and, to slightly paraphrase McCarthy, ``human-level robots''.
-
-\subsection{Logic formalisms, continuous world}
-
-McCarthy underlines the necessity of relying on mathematical logic, as the most
-fruitful formalism for machine intelligence. This is largely the case. Almost
-all the systems we have surveyed rely on logical formalisms, in several cases
-as a mix of declarative languages (often based on Decription Logics) and
-logical programming languages (Prolog, of course, but not only).
-
-The two exceptions (the CAST knowledge model and GSM) are however also
-interesting: CAST proposes a pervasive model of knowledge that is seducing from
-the grounding perspective (but likely suboptimal for deliberative tasks), and
-GSM encodes knowledge in an amodal (continuous, geometric) model while
-preserving features that are usually specific of symbolic models (like theory
-of mind or categorical knowledge).
-
-The interleaving of discrete symbolic models with continuous geometric models
-remains a major challenge, and techniques vary a lot: besides the CAST proxies
-and the GSM amodal model, most systems directly extract physical attributes of
-the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
-OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
-adopt a top-down approach where local geometric models are set up on-demand to
-compute symbolic properties like relative locations.
-
-In our approach, ORO has tight (but unidirectional) links with SPARK, the
-component in charge of geometric modeling and reasoning. Similarly to the GSM
-approach, SPARK is an amodal model of the environment we can manipulate (3D
-motion planning, for instance takes place in MHP, the twin brother of SPARK).
-
-\subsection{Management of uncertainty and approximation}
-
-McCarthy also mentions the necessity to deal with \emph{approximate concepts
-and approximate theories} (that includes representing them and reasoning with
-them).
-
-The most notable attempt at modeling uncertainty down to the knowledge
-representation is {\sc ProgCog}, an extension of {\sc KnowRob}. The theoretical
-and practical performances (including decidability) remain however difficult to
-overcome.
-
-The challenge of representing and reasoning under uncertainty has not been
-tackled in ORO server, at least not \emph{within} the current ORO server
-knowledge model. Two reasons explain this absence: the decisional architecture
-developed at LAAS for the robots has not consistent approach to uncertainty
-representation and management, neither at the \emph{symbol production} level
-(\ie SPARK or {\sc Dialogs}) or at the symbol consumption level (execution
-control). Thus, no strong incentive pushed us in this direction. Then, the
-current tools available from the Semantic Web domain to manipulate and reason
-about knowledge do not provide mature support for uncertainty management. Some
-effort do exist (like the {\sc Pronto} reasoner~\cite{Klinov2008}), but this
-does not appear to be a major focus in the currently available tools.
 
@@
 representation and reasoning (the reasoner then only takes into account
-statements that are set as valid for a given moment).
+statements that are set as valid for a given moment) or with dedicated reasoners
+for tasks requiring non-monotonic reasoning like symbolic task planning
+\fxwarning{mention ORO + HATP}.
 
@@
 
-Several approaches for building contextualized knowledge have been presented in
-this article: symbolic environment interpretation, perspective taking and
-independent mental state for each agent, grounded natural language resolution,
-self-awareness of its own activity. Much remains to be done for a robot to
-actually identify its current context as well as contexts that may be referred
-to.
-
-We will further develop this important topic in the conclusion of the paper.
+Several approaches for building contextualized knowledge have been however
+hinted in section~\ref{sect|features}: symbolic environment interpretation,
+perspective taking and independent mental state for each agent, grounded natural
+language resolution, self-awareness of its own activity. Much remains to be done
+for a robot to actually identify its current context as well as contexts that
+may be referred to.
 
@@
 Integration of these techniques with memory mechanisms (forgetting and
-reinforcement learning) remains to explore.
-
-Temporal modeling is currently absent of ORO, and symbolic and geometric
-planning are accomplished outside of the main knowledge representation
-layer. As a matter of fact, knowledge is mostly atemporal in the ORO server: the
-set of triples stored in the server represents the beliefs of the robot
-\emph{at the current time}. The robot represents knowledge about the \emph{here
-and now}. Dealing with time reasoning is delegated to dedicated tools.
-
-However, ORO does provide a simple implementation of a memory mechanism that
-relies on time: when a statement is attached to a memory profile, the statement
-is reified and the date of creation is stored. This allows to remove the
-statement after a period of time that depends on the memory profile.
-
-Statement reification to store timestamps of creation for all statements is
-technically very easy to do in ORO server, but has not been actually enabled
-because the performance hit (each reified statement produces four triples
-instead of one) was not justified by any current use case of ORO server.
+reinforcement learning) remains mostly to be explored.
 
@@
 
-A side note about how thematic roles and action models are initialized in ORO:
-The current implementation relies on a small, predefined set of action verbs
-that can be recognised from natural language
-(section~\ref{processing_of_actions}).  This constraint does not come from the
-resolution algorithm itself, but rather from the difficulty to automatically
-extract the thematic roles associated to a verb. An interesting yet easy to
-implement extension would consist in binding the dialogue processor with the
-symbolic task planner. The task planner could dynamically provide the list of
-actions that the robot can process, \ie actions for which the robot can produce
-a plan. In certain case, it could also validate the understanding of a desire
-by attempting to plan it.
-
-Also, we could exploit on-line resources like
-\textsc{VerbNet}~\cite{Kipper2008}, which provides a large machine-processable
-lexicon of English verbs along with their thematic roles.
-
 \subsection{Grounding in a multi-modal environment}
@@
 architectures for robots, and all the systems we have surveyed implement
-grounding strategies (...it was an inclusion criteria). Most of the systems
+grounding strategies. Most of the systems
 adapt their grounding strategies to the sensing modality, so that grounding can
@@
 
-We note in addition that the development of what we have called ``synthetic
-sensors'' is simplifying the grounding task. The most obvious example is the
-human tracking system provided by the low-cost Kinect device, and now used
+We note in addition that the development of high-level sensors is simplifying
+the grounding task. The most obvious example is the human tracking system
+provided with the low-cost Kinect family of devices, and now used
 pervasively in the robotics lab. Not only this system segments and computes the
@@
 sensing device does not completely remove the need of grounding one symbolic
-instance of a human with the physical human, but the task is ways simpler than
+instance of a human with the physical human, but the task is much simpler than
 it used to be.
@@
 projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
-the huge interest for the grounding of verbal interaction within the cognitive
+the interest for the grounding of verbal interaction within the cognitive
 robotics community, and we are likely to reach important milestones in this
@@
 
-Our knowledge representation systems have a pragmatic and potentially very
-powerful approach to common-sense: reuse knowledge stored on the Web. While not
-so many of the systems we have surveyed directly tackle this question ({\sc
-KnowRob} is the only one explicitly working on common-sense reasoning, through
+Knowledge representation systems have usually a pragmatic approach to
+common-sense: they reuse knowledge bases stored on the Web. While {\sc KnowRob}
+is the only one explicitly working on common-sense reasoning, through
 physics-based reasoning, integration with Web databases or parsing of
-semi-structured Web documents), several rely however on Web standards (OWL,
-OpenCyc) to represent their knowledge. With the development of initiatives like
-OpenMind that encode with the same standards common-sense facts and rules,
-there is a strong potential for our robot to gain common-sense knowledge from
-well-structured online resources in the coming years.
+semi-structured Web documents), several other KRS rely on Web standards (OWL,
+OpenCyc) to represent their knowledge (Table~\ref{table|knowledge-sources} lists
+the sources of common-sense knowledge found in the surveyed systems). With the
+development of initiatives like OpenMind that encode with the same standards
+common-sense facts and rules, there is a strong potential for our robot to gain
+common-sense knowledge from well-structured online resources in the coming
+years.
 
@@
 
-Table~\ref{table|knowledge-sources} lists for the systems we have surveyed the
-current sources of common-sense knowledge.
-
 \subsection{Learning, representation of experience, introspection}
 
-Learning has not been addressed in this work. This is a complex, still emerging
-field, whose exact meaning and perimeter varies a lot. It was difficult to
-synthesise in the KRS taxonomy, and we decided to omit it. We did not either
-really tackled this question in ORO, except in term of knowledge structure
-alteration based on verbal interaction (this is presented in one case-study, in
-the next section).
-
 Learning at the level of the knowledge representation system has been explored
-in conjunction with visual perception systems (mentioned for instance, in the
-CAST project, in~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
+in conjunction with visual perception systems (mentioned for instance for the
+CAST project~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
 facts from informations automatically extracted from the Web~\cite{Nyga2009},
@@
 perspective}: this cognitive ability, that relates to the representation of
-different modalities, is present in ORO and GSM. ORO explicits the perspectives
-in different symbolic mental states while GSM recursively stores models for
-each agent.
-
-We have already explained how perspective awareness enables advanced cognitive
-capabilities like a theory of mind, and this is probably the main contribution
-of this paper regarding techniques for knowledge representation.
-
-% \subsection{Justification and resolution of cognitive conflicts}
-% 
-% Logical inconsistencies in one ORO's model (or, from a cognitive point of view,
-% cognitive conflicts) could be better dealt with: while (partially) provided by
-% the Pellet reasoner, and output to the developers for debugging purposes, the
-% logical justification of the inconsistency is not really exploited.
-
-% \subsection{Performances analysis}
-% 
-% \subsubsection{Relaxation time}
-% 
-% The relaxation time (as defined at section~\ref{sect|performances}) of ORO for
-% the average workloads we have encountered during the experiments is in the
-% range of 100ms.  The temporal resolution that follows for the symbolic layer,
-% while acceptable in most cases, is not satisfactory. In particular, multi-modal
-% interaction requires fine grained temporal alignement between
-% modalities~\cite{Li2012} which are not possible with the current temporal
-% performances of ORO.
-% 
-% \subsubsection{Cognitive evaluation}
-% 
-% The performance evaluation of current tools is still mostly system- and
-% scenario-specific (``did my system do the job for my task?''), and amongst the
-% surveyed KRS, only few systems proposes an evaluation of their performances
-% that could be compared to other system.
-% 
-% 
-% \fxwarning{Mention reasoning and expressiveness, cf section oro-server}
-%
-
-\subsection{Towards a new design}
+different modalities, is explicitly advertised in ORO and GSM only: ORO
+explicits the perspectives in different symbolic mental states while GSM
+recursively stores models for each agent.  We think however that several of the
+other systems could easily gain support for alternative belief models, leading
+the way to becoming perspective-aware KRS.
+
+\subsection{Conclusion: the Future Directions to Explore}
+
+TDB
 

### 2014-04-17 krs-survey.tex
resolved conflict
added references and some text, to be adjusted and set in context...

--- 2014-04-16 krs-survey.tex
@@
-%\documentclass[a4paper, twocolumn]{article}
-\documentclass{IEEEtran}
-
+\documentclass[journal]{IEEEtran}
 \usepackage[utf8]{inputenc}
@@
 }
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}}
 
@@
 
+\todo{change wording, copied from other papers so far:}
+\paragraph{Semantic maps}
+
+Recent advances in object recognition however lead to more research on semantic
+maps and thereby to a wider use of higher-level semantic knowledge in robotic
+applications, though the ``level of semantics'' ranges from a  mere  classification
+into different parts and object types~\cite{Rusu08RAS2} or the co-occurrence of
+objects~\cite{vasudevan08} to environment representations in description
+logic~\cite{zender2008conceptual}, statistical relational environment
+models~\cite{limketkai05relational} and an embedding of spatial information
+into an encyclopedic and common-sense knowledge base~\cite{tenorth10envmodel}.
+The use of semantic maps as knowledge source for task planning has been
+investigated by~\cite{galindo08taskplanning}. Other AI techniques that have
+been used in robotics include the grounding of natural language in perception
+and action \cite{Mavridis2006,kollar10natural}, planning of actions and
+motions~\cite{wolfe10combined,kaelbling11planning}, and the generation of
+robot controllers from logical constraints~\cite{kressgazit09planning}.
+
+\paragraph{natural language understanding for direction following}
+\cite{matuszek12commands}
+\cite{duvallet13imitation}
+
+\paragraph{(epsiodic) memories in robotics}
+
+RACE \cite{rockel13ontology}
+STRANDS?
+CRAM$_m$ \cite{winkler13memory}
+
+episodic memories in SOAR~\cite{nuxoll12episodic}
+
+\paragraph{misc}
+mining knowledge about object locations from the Web
+\cite{zhou12webmining}
+
+\todo{add something on the IEEE standard draft for a robotics and automation ontology}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}

### 2014-04-21 krs-survey.tex
reworked text, added references

--- 2014-04-16 krs-survey.tex
@@
 }
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:} #1}}
 
@@
 
-The aim of this contribution is two-fold: we first propose a typology of
+The aim of this article is two-fold: We first propose a typology of
 knowledge representation systems for service robotics that covers intrinsic
-features like expressiveness or reasoning techniques as well as extrinsic
-features like knowledge acquisition and integration into pre-existent robotic
-architectures. We base the typology on an initial typical scenario of service
-robotics.
-
-We then survey the current knowledge representation system in the community to
+features, such as expressiveness or reasoning techniques, as well as extrinsic
+features, such as knowledge acquisition and integration into pre-existent 
+robotic architectures. We base the typology on an initial typical scenario of 
+service robotics.
+
+We then survey current knowledge representation systems in the community to
 identify which knowledge-bound cognitive abilities are today well understood
-and implemented, and which would require more research efforts.
+and implemented, and which ones require more research efforts.
 
 We hope this article may bring a clearer picture of the challenges of knowledge
-representation and manipulation for the robotic research community.
+representation and manipulation to the robotic research community.
 
@@
 
+\todo{MT: For a T-RO paper, we should maybe start a bit differently and 
+      emphasize the relevance for robotics and the timeliness of such an 
+      overview more?}
+
+\todo{Other ideas:}
+
+\begin{itemize}
+  \item Check if/which systems are available as open-source software?
+  \item KnowRob: add more recent developments like reasoning about 3D CAD models
+  \item 
+\end{itemize}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{Knowledge as the Root of Cognitive Abilities}
@@
 In a chapter on that subject in \emph{Foundations of Artificial 
-Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
-wrote together in 1998:
+Intelligence}~\cite{Levesque2008}, Levesque reminds the reader of the manifesto 
+they wrote together in 1998:
 
@@
     \end{itemize}
-
+    \noindent
     and so on. With respect to robotics, our goal (like that of many in AI) is
@@
 
-To be answered, those questions have in common one prerequisite: the robot must
-be able to manipulate \emph{explicitly} knowledge, and hence, need a way to
+To be answered, those questions have one prerequisite in common: The robot must
+be able to \emph{explicitly} manipulate knowledge, and hence, needs a way to
 represent the knowledge.
-
-This survey focuses on this knowledge representation issue: we aims at first
+% 
+This survey focuses on this knowledge representation issue: We first aim at
 establishing a comprehensive typology of representational needs for
 robotics in the specific context of service robotics, and then at painting the
-current landscape of approaches to the knowledge representation problem in the
-research community, in order to summarize the major trends and to identify the
-possible shortcomings.
+current landscape of approaches in the research community in order to summarize 
+the major trends and to identify possible shortcomings.
 
 We have classified the features of knowledge representation systems (KRS) into
-six broad categories: \emph{what can be represented?} or the
-\emph{expressiveness} question ; \emph{how things are represented?} ; \emph{what
-reasoning techniques are offered?} ; \emph{how knowledge is acquired and
-anchored?} ; \emph{how the knowledge is made available and used by the other
-components of the robot?} and \emph{which actual knowledge is stored?} or the
+six broad categories: \emph{What can be represented?} or the
+\emph{expressiveness} question ; \emph{How are things represented?} ; \emph{Which
+reasoning techniques are offered?} ; \emph{How is knowledge acquired and
+anchored?} ; \emph{How is the knowledge made available and used by other
+components of the robot?} and \emph{Which actual knowledge is stored?} or the
 \emph{knowledge instantiation} question.
@@
 We have also tried, for each project, to make clear \begin{inparaenum} \item
-how informations from perception or interaction are turned into knowledge,
-\item how, in return, symbolic concept are \emph{anchored} into the robot
+how information from perception or interaction is turned into knowledge,
+\item how, in return, symbolic concept are \emph{anchored} into the robot's
 sensori-motor space, \item how the knowledge representation system integrates
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \paragraph{What do we call ``knowledge''?}
@@
 
-Since we will discuss at length the concept of knowledge in the context of
-robotics in the coming pages, it is useful to make our terminology explicit.
-
+Since we will discuss the concept of knowledge in the context of robotics in 
+depth over the course of the the following pages, it is useful to make our 
+terminology explicit first.
+% 
 Be it in philosophy, cognitive sciences or computer sciences, reaching an
@@
 starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
-and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
-with representation, the knowledge level deals with language, semantics;
+and \emph{goals}, \emph{actions}, and \emph{bodies}. Whereas the symbol level deals
+with representation, the knowledge level deals with language and semantics;
 whereas the symbol level deals with inference, the knowledge level deals with
@@
 interrelated logical facts that are meaningful to the robot executive
-controller}. By \emph{meaningful} we mean that can possibly be interpreted to
-lead to a purposeful action. We will see that our main challenge while
+controller}. By \emph{meaningful} we mean that they can possibly be interpreted 
+to lead to a purposeful action. We will see that our main challenge while
 designing a cognitive architecture is furthermore to make this knowledge as
@@
 The relation of \emph{data} and \emph{information} to knowledge is a debated
-epistemology question known as the ``DIKW'' hierarchy question. In this article,
-we will associate data to low-level material like raw sensor output, and
-information to uncontextualised symbolic facts.
-
+epistemology question known as the ``DIKW'' hierarchy question (with D,I,K 
+and W standing for ``data'', ``information'', ``knowledge'' and ``wisdom'', 
+respectively). In this article, we will associate ``data'' with low-level 
+material like raw sensor output, and ``information'' with uncontextualised 
+symbolic facts.
 To give a example, we can imagine a human reading a book while being tracked by
@@
 war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
-the robot would be proper knowledge.
-
+the robot, would be proper knowledge.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection*{Survey Inclusion Criteria}
@@
 
-Every robotic system has, implicitly or not, some knowledge representation
-systems. It may range from a simple state vector to an explicit symbolic
-knowledge base.  This survey focuses on the right end of this spectrum:
-symbolic systems, suited for abstract reasoning.
-
-Besides, we have decided to restraint the set of systems to those actually
-implemented on robots, and used in semantic-rich environments (\ie dynamic,
+Every robotic system includes, explicitly or not, some kinds of knowledge 
+representation systems. They may range from a simple state vector to an 
+explicit symbolic knowledge base. This survey focuses on the right end of 
+this spectrum: symbolic systems, suited for abstract reasoning.
+% 
+Besides, we have decided to restrain the set of systems to those actually
+implemented on robots, and used in semantically rich environments (\ie dynamic,
 partially unknown environments with a large range of different entities which
-may have interactions). The typical scenario that would involve such robots is
+may have interactions). A typical scenario that would involve such robots is
 a service robot in a human-friendly environment like a kitchen.
-
-We have limited ourselves to systems that
+% 
+We have further limited ourselves to systems that
 \begin{inparaenum} 
-    \item  run on \emph{service robot} (that is, robots that interact with 
-    objects in a semantic-rich environment primarily designed for humans),
+  \item  run on a \emph{service robot} (that is, robots that interact with 
+  objects in a semantically rich environment primarily designed for humans),
   \item  ground the knowledge in the physical world (physically embedded
@@
 
+\todo{MT: I think we should rethink this decision. DyKnow is a knowledge 
+      processing system running on physical robots, and there's also some
+      work by Lakemeyer etc in Aachen on using Golog-like things on household 
+      robots}
 These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
 which are focused on data fusion and knowledge grounding at lower levels.
-
+% 
 We have also chosen not to include the GOLOG language and its
@@
 While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
-GLAIR~\cite{Shapiro2009} or {\sc ACT-R} have declarative knowledge
-modules~\cite{Derbinsky2010} and have been recently used on service robots (see
-{\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are also absent from this
-survey because we did not find much references in the literature on knowledge
+GLAIR~\cite{Shapiro2009} or {\sc ACT-R} do have declarative knowledge
+modules~\cite{Derbinsky2010} and have recently been used on service robots (see
+ for instance {\sc ACT-R/E}~\cite{Kennedy2009}), they are also absent from this
+survey because we did not find many references in the literature on knowledge
 manipulation and representation applied to real-world robotic scenarii for
 these architectures.
-
-A comprehensive reference on (bio-inspired) cognitive architectures is also
+% 
+A comprehensive reference on (bio-inspired) cognitive architectures is
 available from the BICA Society~\cite{BicaCogArch2011}.
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Previous work}
@@
 
-The typology we propose is in part based on a comprehensive synthesis of
+The typology we propose is partly based on a comprehensive synthesis of
 classifications and analysis found in the literature. This synthesis is focused
@@
 
-Levesque and Lakemeyer~\cite{Levesque2008} present in their chapter on
-Cognitive Robotics several characteristics of knowledge representation systems
+In their chapter on Cognitive Robotics, Levesque and Lakemeyer~\cite{Levesque2008} 
+present several characteristics of knowledge representation systems
 for robots, stressing the need of representing the \emph{dynamics of the
@@
 
-In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
-\emph{knowledge processing middleware} as systems supporting ``declarative
+In a slightly broader context, Heintz et al.~\cite{Heintz2008} define a
+\emph{knowledge processing middleware} as a system supporting ``declarative
 specifications for flexible configuration and dynamic reconfiguration of
-context dependent processing at many different levels of abstraction''. They
-identify six characteristics: the system must be able to \emph{merge
-informations} from different, possibly distributed sources; it should support
+context-dependent processing at many different levels of abstraction''. They
+identify six characteristics: The system must be able to \emph{merge
+information} from different, possibly distributed sources, it should support
 quantitative as well as qualitative processing of information, it should offer
@@
 \emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
-(which require what we call here \emph{non-monotonicity}) and finally
+(which requires what we here call \emph{non-monotonicity}), and finally support
 \emph{meta-knowledge} and \emph{introspective capacities} (``declarative
@@
 
-Several surveys compare global cognitive architectures \cite{Langley2006,
+Several surveys compare general cognitive architectures \cite{Langley2006,
 Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
@@
 learning/remembering/introspection. They also separately identify four
-\emph{properties} of a cognitive architecture, that categorise how knowledge is
-handled by the architecture: representation of knowledge, organisation of
-knowledge, utilisation of knowledge and acquisition and refinement of
-knowledge. This categorisation had a notable influence on our typology, and
+\emph{properties} of a cognitive architecture that categorize how knowledge is
+handled by the architecture: representation of knowledge, organization of
+knowledge, utilization of knowledge, and acquisition and refinement of
+knowledge. This categorization had a notable influence on our typology, and
 many of these categories are also present in our proposal.
@@
 interaction of the system with its environment). The approaches presented in
-this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
+this paper are, with a few exceptions, prototypical \emph{cognitivist} approaches
 that aim at making knowledge explicit within the robot architecture. Vernon et
@@
 robotics have previously listed desirable features of artificial systems aiming
-at rich cognitive abilities.
-
-For instance McCarthy recently listed in~\cite{McCarthy2007} the challenges he
-identifies on the road to a \emph{human-level AI}.
+at rich cognitive abilities. For instance McCarthy recently listed the 
+challenges he identified on the road to a \emph{human-level AI}~\cite{McCarthy2007}.
 
@@
 
-	\item the ability to \emph{"operate successfully in the common sense
+	\item The ability to \emph{"operate successfully in the common sense
 	informatic situation"},
 
-	\item the necessity of relying on mathematical logic, as the most fruitful
+	\item The necessity of relying on mathematical logic, as the most fruitful
 	formalism for machine intelligence,
 
-	\item the ability to deal with \emph{approximate concepts and approximate
+	\item The ability to deal with \emph{approximate concepts and approximate
 	theories} (that would include representing them, and reasoning with them),
 
-	\item non-monotonic reasoning,
-
-	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
+	\item Non-monotonic reasoning,
+
+	\item What McCarthy calls \emph{Elaboration Tolerance}: the ability to
 	extend \emph{on demand} the closed domain of interpretation for a
@@
 
-	\item the ability to formalise and reason about contexts,
-
-	\item reasoning about events, and in particular, actions,
-
-	\item the capacity of introspection,
-
-	\item and finally, he points the issue of giving computer the right
+	\item The ability to formalize and reason about contexts,
+
+	\item Reasoning about events, and in particular, actions,
+
+	\item The capacity of introspection,
+
+	\item and finally, he points at the issue of giving computer the right
 	heuristics for decision making.
@@
 Coming from the perspective of natural language processing in situated context,
-Roy and Reiter summarise in~\cite{Roy2005} what they see as the main challenges
-to be tackled by knowledge representation systems: \emph{cross-modal
+Roy and Reiter summarize in~\cite{Roy2005} what they see as the main challenges
+to be tackled by knowledge representation systems: \emph{Cross-modal
 representation} systems, association of words with perceptual and action
@@
 interaction and ability to take into account the \emph{human perspective}.
-
+% 
 Knowledge representation systems in robotics are directly affected by these
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{Article overview}
@@
 The paper is organized in three main sections.
-
+% 
 Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
@@
 to researchers.
-
+% 
 Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
@@
 of the eight systems, and underlines their specificities.
-
+% 
 We finally summarize in section~\ref{sect|conclusion} the current approaches and
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \section{Knowledge Representation in Robotics: an Overview}
@@
 
-TDB
-
+Representations for a robot's knowledge have been investigated for decades, 
+starting with the seminal work on the Shakey~\cite{nilsson84shakey} robot --
+which already had an internal world model based on predicate logics. Shakey
+however lived in a very simple artificial world consisting of a few blocks
+that could rather easily be described in logical form. Scaling towards more
+complex settings required research in both AI and robotics, and in the 
+following decades, these two communities operated without much interaction.
+Today's robots are now able to operate in much more complex environments 
+compared to Shakey's blocks world, have perception techniques that can handle 
+the complexity of real scenes, manipulation capabilities that allow them to 
+interact with objects, and massively more computational resources. On the one 
+hand, this makes robots an interesting application for AI research again. On
+the other hand, roboticists strive for more flexibility, adaptability and in 
+general more semantic representation and therefore consider AI methods again. 
+Besides the large-scale knowledge representation systems that will be discussed 
+in the following sections, there have been many approaches to using knowledge
+in parts of the robotic system.
+
+One field that has seen much progress in the past years is the creation and 
+representation of environment models, often termed ``semantic maps''. Advances 
+in object recognition allowed to identify single objects in a map, which 
+suggests to also store higher-level semantic information about them. The 
+term ``semantic'' is however used for a range of things -- from a mere 
+classification into different parts and object types~\cite{Rusu08RAS2}or the 
+co-occurrence of objects in scenes~\cite{vasudevan08} up to environment 
+representations in description logic~\cite{zender2008conceptual}, statistical 
+relational environment models~\cite{limketkai05relational} and an embedding 
+of spatial information into an encyclopedic and common-sense knowledge 
+base~\cite{tenorth10envmodel}. The use of semantic maps as knowledge source 
+for task planning has been investigated by~\cite{galindo08taskplanning}. 
+% 
+Another area of research that often includes some knowledge representation
+aspects is natural-language interaction. For understanding spoken dialog
+about things in its environment\cite{Mavridis2006} and for following spoken
+directions \cite{kollar10natural,matuszek12commands,duvallet13imitation}, a 
+robot needs to be able to ground the symbols (i.e. words) in percepts and 
+actions. The natural-language text may be the result of direct interaction 
+with a human, but can also be obtained from resources on the Web, for example
+for mining object locality knowledge~\cite{zhou12webmining} or extracting 
+task instructions from Websites~\cite{tenorth10webinstructions}.
+% 
+Recently, learning from experiences and ``life-long learning'' have
+become popular topics in robotics. The goal is to record rich episodic 
+memories during task execution~\cite{niemueller2012iros}, to reason about 
+them later to extract more generic rules~\cite{winkler13memory} and to learn
+generalizes task models from the collected experiences~\cite{rockel13ontology}.
+
+\todo{add something on the IEEE standard draft for a robotics and automation ontology}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
@@
 
-We propose to organize requirements from this imaginary cooking scenario where
-numerous desirable features for a knowledge representation system for robotics
-have been identified without any particular order.
-
-This section proposes a more formal typology and a nomenclature of analysis
-axis for a such a system. For each feature, we provide a short definition along
-with links to relevant literature.
-
-As previously mentioned, we propose six main categories, that cover both basic
-and practical aspects: the formal
-\taxon{A}{Expressiveness} of the system, the breadth and nature
-of the \taxon{B}{Representations} allowed by the system, the
-\taxon{C}{Reasoning} capabilities, how knowledge it acquired
-\taxon{D}{Acquisition}, how the system integrates with the other components of
-the robot sotware architecture \taxon{E}{Integration} and finally what we call
+% We propose to organize requirements from an imaginary cooking scenario where
+% numerous desirable features for a knowledge representation system for robotics
+% have been identified without any particular order.
+% 
+This section proposes a formal typology and a nomenclature of analysis 
+dimensions for a knowledge representation system. For each feature, we provide 
+a short definition along with links to relevant literature. As mentioned 
+earlier, we propose six main categories that cover both basic and practical 
+aspects: The formal \taxon{A}{Expressiveness} of the system, the breadth and 
+nature of the \taxon{B}{Representations} allowed by the system, the
+\taxon{C}{Reasoning} capabilities, its methods for knowledge 
+\taxon{D}{Acquisition}, the \taxon{E}{Integration} of the system with other 
+components of the robot software architecture, and finally what we call
 the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
@@
 
-The next paragraph details only briefly the exact meaning of these categories
-and introduce accordingly sub-categories. We kindly refer the interested reader
-to~\cite{lemaignan2012symbolic} for in-depth discussion of each of these
-dimensions of knowledge representation systems, as well as related literature.
-
+The next section details only briefly the exact meaning of these categories
+and, accordingly, introduces sub-categories. We kindly refer the interested 
+reader to~\cite{lemaignan2012symbolic} for an in-depth discussion of each of 
+these dimensions of knowledge representation systems as well as related 
+literature.
+% 
 Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
-the main contributions of each of the surveyed KRS with respect to these
-dimensions.
-
-\subsection{Expressiveness: What Can be Represented?}
+the main contributions of each of the surveyed systems with respect to these
+dimensions. \todo{MT: Can we list this taxonomy of capabilities as one of the 
+contributions of this article?}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Representational expressiveness}
 \label{sect|expressiveness}
 
-This first axis of analysis is its intrinsic expressive power. It answers the
-question: what can be possibly represented. When it explicitly exists, the
-\emph{language} of representation plays here an obvious role.
-
+This first axis of analysis is the intrinsic expressive power. It answers the
+question of what can possibly be represented. In case it explicitly exists, the
+\emph{language} of representation obviously plays an important role here.
+% 
 We subdivide the category \taxon{A}{Expressiveness} into five subcategories:
-\taxon{A.1}{Logic formalism} classifies the underlying logical formalism of the
+\taxon{A.1}{Logical formalism} classifies the underlying logical formalism of the
 KRS (Horn clauses like for Prolog-based systems, Description
@@
 \taxon{A.3}{OWA/CWA} reflects the design choice between the \emph{open world}
-(OWA) versus the \emph{closed world} (CWA) assumptions, which related to the
+(OWA) versus the \emph{closed world} (CWA) assumptions, which refers to the
 possibility of representing \emph{unknown} facts. \taxon{A.4}{Uncertainty}
-relates to the ability for the system to represent both the \emph{intrinsic}
+relates to the ability of the system to represent both the \emph{intrinsic}
 uncertainty of given facts, and the \emph{extrinsic} uncertainty due to
@@
 of \emph{representation}. Reasoning on uncertain facts is discussed in
-\taxon{C.4}{}. The last features of a KRS related to representation is
+\taxon{C.4}{}. The last feature of a KRS related to representation is
 \taxon{A.5}{Meta-cognition} that conveys the capability for a system to
@@
 
-%%%%%%%%%%
-\subsection{How things are represented?}
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Representation techniques}
 \label{sect|higher-level-domain-representation}
 
-This category focuses on questions that involve
-representational challenges (time, space, context) or require specific
-cognitive capabilities (theory of mind, introspection, memory).
-
+This category focuses on questions that involve representational challenges 
+(time, space, context) or require specific cognitive capabilities (theory of 
+mind, introspection, memory).
+% 
 The dimension \taxon{B.1}{Roles}, with its sub-categories \taxon{B.1.1}{Space},
@@
 (spatial relations, time and representation of actions). To give a few examples
-(in no way exhaustive), \emph{Spatial relations} covers techniques like
+(in no way exhaustive), \emph{spatial relations} cover techniques like
 topological maps, allocentric and egocentric frame projections. \emph{Time
@@
 chronicles~\cite{Ghallab1996} or the concept of \emph{fluent}. \emph{Actions}
-concern themselves with representation of \emph{pre-, post-conditions},
-\emph{thematic roles} or \emph{plans}.
+are concerned with representation of \emph{pre- and post-conditions},
+\emph{thematic roles} and \emph{plans}.
 
@@
 often related to the initial \emph{common-sense knowledge} available to the
-robot. \taxon{B.3}{Modality} refers to the ability, for the KRS, to represent
+robot. \taxon{B.3}{Modality} refers to the ability of the KRS to represent
 possible alternate worlds or belief states. A typical application of a
 \emph{modal} knowledge model is the implementation of a \emph{theory of mind},
-\ie an independent belief model for the other agents the robot interact with.
+\ie an independent belief model for other agents the robot interact with.
 \taxon{B.4}{Introspection} covers both \emph{self-knowledge} (what do I know
@@
 Finally, \taxon{B.5}{Memory} refers to the implementation of
-different kind of bio-inspired memory models like \emph{declarative},
+different kinds of (bio-inspired) memory models like \emph{declarative},
 \emph{procedural} or \emph{episodic}, and the attached processes
@@
 
-%%%%%%%%%%
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Reasoning techniques}
@@
 have found their place in robotics over the years. We propose to consider ten
-subcategories of \taxon{C}{Reasoning} that reflects in our opinion important
-features in regards to the general requirements of robotics (like decision
-making being tractable in pseudo-real time, being able to deal with partially
-unknown or uncertain domain or specific reasoning related to interaction with
-the physical world).
-
-Again, the descriptions of these ten subcategories hereafter are very
-superficial. Readers can refer to~\cite{lemaignan2012symbolic} for the details.
+subcategories of \taxon{C}{Reasoning} that reflect features we consider 
+important with respect to the general requirements of robotics (like tractable 
+decision making in pseudo-real time, dealing with partially unknown or uncertain 
+domains, or reasoning related to interaction with the physical world). Again, 
+the descriptions of these ten subcategories hereafter are very superficial. 
+Readers are referred to~\cite{lemaignan2012symbolic} for details.
 
@@
 \emph{concept satisfiability}, \emph{consistency checking} and \emph{instance
-checking}. \taxon{C.2}{Instantiation and structural alteration} describes how
-dynamic a system is with respect to its internal knowledge structure (allowing
-for modification of the knowledge structure, often called \emph{TBox}, may
-impact the capability for the system to \emph{learn} for instance).
-\taxon{C.3}{Lazy evaluation} describes the ability for a KRS to delay active
+checking}. \taxon{C.2}{Instantiation and structural alteration} describe how
+dynamic a system is with respect to its internal knowledge structure (admitting
+the modification of the knowledge structure, often called \emph{TBox}, may
+for instance influence the capability of a system to \emph{learn}).
+\taxon{C.3}{Lazy evaluation} describes the ability of a KRS to delay active
 knowledge acquisition or reasoning operations until the value is actually
@@
 \taxon{C.5}{Non-monotonicity} discusses how systems deal with non-monotonic
-inferences, \ie exceptions to a rule or inferences that lead to the retractation
+inferences, \ie exceptions to a rule or inferences that lead to the retraction
 of previously inferred facts. Techniques to deal with non-monotonic reasoning
 include for instance \emph{Default Logics}.  \taxon{C.6}{Presupposition
-accomodation} is the ability for a system to automatically create a context
-allowing to make sense of a proposition. This is tighly linked to the capability
-for a system to deal with \emph{under-specification}. \taxon{C.7}{Prediction and
-explanation} (with subcategories \taxon{C.7.1}{Projection},
-\taxon{C.7.2}{Legality}, \taxon{C.7.3}{Diagnosis} and
-\taxon{C.7.4}{Explanation}) relates to the capabilites for a system to predict
+accommodation} is the ability of a system to automatically create a context
+allowing to make sense of a proposition. This is tightly linked to the 
+capability of dealing with \emph{under-specification}. \taxon{C.7}{Prediction 
+and explanation} (with subcategories \taxon{C.7.1}{Projection}, 
+\taxon{C.7.2}{Legality}, \taxon{C.7.3}{Diagnosis} and \taxon{C.7.4}{Explanation}) 
+relates to the capabilities of a system to predict
 and project the outcomes of a sequence of actions onto the logical system, and
@@
 conclusion. Related to this category, \taxon{C.8}{Planning} discusses to which
-extend a KRS can be used for \emph{symbolic task planning}, \ie the ability for
+extend a KRS can be used for \emph{symbolic task planning}, \ie the ability of
 a robot to select a sequence of actions in order to reach a given final state.
-\taxon{C.9}{Physics-based reasoning} describes the capability for a symbolic
+\taxon{C.9}{Physics-based reasoning} describes the capability of a symbolic
 knowledge representation system to rely on physical reasoning (which may include
@@
 Finally, \taxon{C.10}{Learning} groups the various capabilities that lead to
-learning behaviours.
-
-%%%%%%%%%%%%%%%%%
-\subsection{Acquiring knowledge}
+learning behaviors.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Knowledge acquisition}
 
 In the survey inclusion criteria, we have chosen to only consider robotic
-systems that acquire knowledge by themselves, at runtime. The category
+systems that acquire knowledge by themselves at runtime. The category
 \taxon{D}{Acquisition} reviews the different approaches to knowledge acquisition
@@
 interaction (typically, language processing), and \taxon{D.1.3}{Linked
-Resources} relates to the connections and alignement issues with external
+Resources} relates to the connections and alignment issues with external
 (often Web-based) knowledge bases.
-
-\taxon{D.2}{Grounding} discusses the KRS capabilities in regards to attaching a
+% 
+\taxon{D.2}{Grounding} discusses the KRS' capabilities with regard to attaching a
 practical meaning to the symbol it manipulates. In the context of robotics, it
-mainly consists in building and maintaining a bi-directional link between
-sub-symbolic representations (sensors data, low-level actuation) and symbolic
+mainly consists of building and maintaining a bi-directional link between
+sub-symbolic representations (sensor data, low-level actuation) and symbolic
 representations.
-
-Finally, \taxon{D.3}{Instrinsic Motivation} describes the possible
+% 
+Finally, \taxon{D.3}{Intrinsic Motivation} describes the possible
 \emph{internal} mechanisms by which the robot seeks by itself to acquire more
@@
 
-%%%%%%%%%%%%%%%%%
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Practical integration in robotic architectures}
@@
 considered in isolation, and it seems important to also evaluate KRS from the
-perspective their integration into larger control architectures that includes
+perspective of their integration into larger control architectures that include
 perception routines, decision-making processes and actuation control.
@@
 from the perspective of the robot controller. Here also, many different
-approaches have been investigated, from classical server-client architecture, to
-DSL-like language extensions. \taxon{E.3}{Monitoring} details the support tools
-provided by the system for the developpers to monitor and introspect their
-system. This is especially important for knowledge representation systems
-because they often lie at the heart of the architecture, and act as hubs for
-many other components. \taxon{E.4}{Performances}, finally, evaluates the
-performances of the KRS, both in term of scalability and classification speed
-\taxon{E.4.1}{Raw performances}, and in term of cognitive abilities provided by
-the system \taxon{E.4.2}{Cognitive performances}. Note that performances of KRS
-are notoriously difficult to assess since on one hand, cognitive architectures
-for robotics are often build upon tightly interleaved modules which makes it
-difficult to test a KRS in isolation, and on the other hand, we lack standard
-benchmarks and scales for cognitive capabilities.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
+approaches have been investigated, from classical client-server architectures, 
+to domain-specific language extensions. \taxon{E.3}{Monitoring} details the 
+support tools provided by the system for the developers to monitor and 
+introspect their system. This is especially important for knowledge representation 
+systems because they often lie at the heart of the architecture and act as hubs 
+for many other components. \taxon{E.4}{Performance}, finally, evaluates the
+performance of the KRS -- the \taxon{E.4.1}{Raw performance} in terms of 
+scalability and classification speed, the \taxon{E.4.2}{Cognitive performance} 
+in terms of cognitive abilities provided by the system. Note that the performance 
+of a KRS is notoriously difficult to assess since, on the one hand, cognitive 
+architectures for robotics are often build upon tightly interleaved modules 
+which makes it difficult to test a KRS in isolation, and on the other hand, 
+because standard benchmarks and scales for cognitive capabilities are still 
+lacking.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Knowledge instantiation}
@@
 called the ABox), but also includes the knowledge structure itself (the TBox).
-
-The \taxon{F.1}{Design strategy} of the KRS can be \emph{top-down} (from an
-initial symbolic model down to the percepts), \emph{bottom-up} or  a mix of
+% 
+The \taxon{F.1}{Design strategy} of a KRS can be \emph{top-down} (from an
+initial symbolic model down to the percepts), \emph{bottom-up}, or a mix of
 both. The bottom-up and top-down strategies reflect how the knowledge structure
 itself is constructed: either from generic categories, typically extracted from
-standard upper-ontologies or by successive classification and refinement of
+standard upper-ontologies, or by successive classification and refinement of
 percepts.
-
-\taxon{F.2}{Common-sense and alignment} discusses if and how the KRS reuse
-existing pools of knowledge like \emph{upper-ontologies}, and how the question
+% 
+\taxon{F.2}{Common-sense and alignment} discusses if and how the KRS reuses
+existing pools of knowledge like \emph{upper ontologies}, and how the question
 of knowledge \emph{alignment} is handled. \taxon{F.3}{Metrics and quality
-criteria} includes qualitative and quantitative metrics that give an insight on
-the size, complexity and effectiveness of knowledge base available to the robot.
-These metrics include for instance the number of terms and assertions, but also
-the type of predicates that are used, or the computed Description Logics
-expressiveness (when applicable).  Qualitative evaluation of ontologies is a
-research field in its own right, and introduce criteria like \emph{accuracy,
-adaptability, clarity, completeness, computational efficiency, conciseness,
-consistency}. The last category is \taxon{F.4}{Granularity}, that accounts for
-the level of details considered in the knowledge representation system, that can
-go down to storing SIFT features~\cite{Suh2007} directly in the KRS.
+criteria} includes qualitative and quantitative metrics that provide insight 
+into the size, complexity and effectiveness of knowledge bases available to the 
+robot. These metrics include for instance the number of terms and assertions, 
+but also the type of predicates that are used, or the computed Description Logics
+expressiveness (if applicable). The qualitative evaluation of ontologies is a
+research field in its own right that has introduced criteria like \emph{accuracy,
+adaptability, clarity, completeness, computational efficiency, conciseness, and
+consistency}. The last category is the \taxon{F.4}{Granularity}, which accounts 
+for the level of details considered in the knowledge representation system 
+which can go down to storing SIFT features~\cite{Suh2007} directly in the KRS.
 
@@
 Table \ref{table|surveyed-systems} lists the knowledge representation
-systems that we have surveyed.
-
-This section briefly presents each of them.
+systems that we have surveyed. This section briefly presents each of them.
 
@@
 \midrule
-ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
-CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ.,
+ARMAR/Tapas & Formal & Holzapfel, Waibel \par (KIT Karlsruhe) & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
+CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Birmingham Univ.,
 DFKI Saarbrücken) & & Amodal proxies & \cite{jacobsson2008crossmodal} \\
@@
 Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
-{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
+{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU München) & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
 NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & NKRL & & \cite{Sabri2011} \\
@@
 Institute of Technology.
-
+% 
 Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
-knowledge. The later is split into \emph{lexical knowledge}, \emph{semantic
+knowledge. The latter is split into \emph{lexical knowledge}, \emph{semantic
 knowledge} and a database of identified objects (with their properties). The
 \emph{lexical knowledge} contains lexical and grammatical informations about
-the objects. The \emph{semantic knowledge} is organised into an ontology
+the objects. The \emph{semantic knowledge} is organized into an ontology
 relying on \emph{typed feature structures} (TFS~\cite{Carpenter1992}, a
@@
 superset of first-order logic).
-
+% 
 {\sc Tapas} has a strong focus on natural language grounding. It proceeds by
@@
 understand dialogue.
-
+% 
 Another focus is put on handling unknown words and objects. {\sc Tapas}
-provides routines to recognise unknown entities, and propose and interactive
-and iterative verbal process to categorise (including adding new categories)
+provides routines to recognize unknown entities, and propose and interactive
+and iterative verbal process to categorize (including adding new categories)
 those new concepts.
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{CAST knowledge model}
@@
 through a set of interconnected \emph{SA} (\emph{subarchitectures}). The CAS
-does not expose a central knowledge base as seen in other works. It represents
-instead knowledge as unrooted \emph{proxies}. Those proxies are formally
+does not expose a central knowledge base as seen in other works. It instead 
+represents knowledge as unrooted \emph{proxies}. Those proxies are formally
 defined in \cite{jacobsson2008crossmodal} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
@@
 entity.
-
-A union of proxies forms a global amodal representation of an entity, that can
-be explicitly shared and manipulated. Being not centralised, the knowledge
-model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
-the CAS architecture is tightly bound to the on-line grounding process (be it
+% 
+A union of proxies forms a global amodal representation of an entity that can
+be explicitly shared and manipulated. Not being centralized, the knowledge
+model can be qualified as \emph{ubiquitous}. Furthermore, each knowledge source 
+in the CAS architecture is tightly bound to the on-line grounding process (be it
 grounded in perception or in dialogue). While nothing seems to prevent it, no
 {\it a priori} knowledge (including common-sense knowledge) is used.
-
-Knowledge sharing is ensured by the event mechanism of CAST: modules can
-monitor proxies for alteration by other modules. Jacobsson et al. mention how
-this can apply to reinforcement learning: the vision module creates a proxy for
-an orange object. This proxy get monitored by a learning module. In parallel,
-the proxy is bound to an union by the natural language understanding module
-that add new a feature like \emph{"this object is a fruit"}. The learning
-module is called back, and can add this new information to its model.
-
+% 
+Knowledge sharing is ensured by the event mechanism of CAST -- modules can
+monitor proxies for alteration by other modules. Jacobsson et al. explain how
+this can be applied to reinforcement learning: The vision module creates a 
+proxy for, for example, an orange object. This proxy gets monitored by a 
+learning module. In parallel, the proxy can be bound to a union by the natural 
+language understanding module that may add new features like \emph{"this object 
+is a fruit"}. The learning module will then be called back and can add this 
+new information to its model.
+% 
 In the presented implementation, the CAST knowledge model does no allow for
@@
 fusion~\cite{Hawes2007a} is well studied, and the interaction with natural
-language processing~\cite{Kruijff2010, Kruijff2010a} is a particular emphasise
+language processing~\cite{Kruijff2010, Kruijff2010a} is a particular emphasis
 of the project.
-
+% 
 In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
 the context of spatial representations with the robot \emph{Dora}.
-
+% 
 \paragraph{Experiments} CAST has been used in several experiments, including
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{GSM}
@@
 robot.
-
+% 
 GSM does not rely on any formal language but rather on a layered data structure
-(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
+(Figure~\ref{fig|gsm}) that organizes the surrounding world into agents and
 relations between agents.  Each agent (any animate or inanimate object) is
@@
 stochastic model, and a discrete, categorical model.
-
-One notable feature of GSM is the \emph{bidirectionality} of the grounding
-process: not only sensor percepts are abstracted into categories suitable for
-human conversation, but human utterance (like ``There is a red ball in the
+% 
+One notable feature of GSM is the \emph{bi-directionality} of the grounding
+process: not only are sensor percepts abstracted into categories suitable for
+human conversation, but human utterances (like ``There is a red ball in the
 center of the table'') can also be turned into property descriptions. This
-basically enable the knowledge representation system of the robot to
+basically enables the knowledge representation system of the robot to
 \emph{imagine} entities.
-
+% 
 GSM also features several strategies for managing time and events.
-\emph{Moments} are created by storing timestamped snap-shots of GSM, and
+\emph{Moments} are created by storing timestamped snap-shots of the GSM, and
 \emph{event classifiers} allow to define and detect events.
@@
 tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
-an in-depth analysis of the performance of GSM by the mean of a standard
+an in-depth analysis of the performance of GSM by means of a standard
 psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{Ke Jia Project}
@@
 
-The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
-representation language with natural language processing, task planning and
-motion planning.
-
-Knowledge representation relies on \emph{Action Language C}, itself based on
+The Ke Jia project~\cite{Chen2010} integrates a knowledge representation 
+language with natural language processing, task planning and motion planning 
+on a mobile platform.
+% 
+Knowledge representation relies on the \emph{Action Language C}, itself based on
 \emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
@@
 task planning~\cite{Ji2011} and underspecified natural language processing.
-
+% 
 Amongst other features, the natural language processing capabilities of the
@@
 involving human-robot interaction with natural language. These tasks include a
-task with multiple \emph{pick \& carry} that are globally optimised, naive
+multiple \emph{pick \& carry} actions that are globally optimized, naive
 physics reasoning via taught rules or more complex scenarii with the robot
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{KnowRob}
@@
 
-{\sc KnowRob}~\cite{Tenorth2009a} is an integrated knowledge management system
-developed at the Technical University of Munich. It is build as a set of
-modules (figure~\ref{fig|knowrob} organised around a core reasoning system
-written in Prolog. This core module interfaces through Java/Prolog or
-C/Prolog APIs with external modules.
+{\sc KnowRob}~\cite{Tenorth2009a,tenorth13knowrob} is an integrated knowledge 
+management system originally developed at the Technische Universität München 
+that has become an open-source project that is maintained by the University
+of Bremen. It is build as a set of modules (Figure~\ref{fig|knowrob} that are 
+organized around a core reasoning system written in Prolog. This core module 
+interfaces with external modules through Java/Prolog or C/Prolog APIs.
 
@@
 
-Extension modules can plug into the system to provide specialised reasoning
+Extension modules can plug into the system to provide specialized reasoning
 capabilities or interfaces to external data sources, \eg to read object
@@
 \emph{KnowRob-Base} ontology is provided as a common-sense ontology, with a
-focus on household and kitchen domains. {\sc KnowRob} also store and reason on
-introspective knowledge through the \emph{Semantic Robot Description
-Language}~\cite{Kunze2011} that allow to represent symbolically the
-capabilities of the robot, and is used for planning.
+focus on household and kitchen domains. {\sc KnowRob} also stores and reasons 
+on introspective knowledge through the \emph{Semantic Robot Description
+Language}~\cite{Kunze2011} that allows to symbolically represent the robot's
+capabilities.
 
@@
 {\sc ProbCog}~\cite{Jain2009} is an effort to provide probabilistic reasoning
-based on bayesian networks, integration with naive physics reasoning has been
-studied~\cite{Kunze2011a}, automatic parsing of Web resources in semi-natural
-language has been also experimented with~\cite{Nyga2009}.
-
-\paragraph{Grounding} {\sc KnowRob} offers a mechanism called \emph{computables} that allow to
-evaluate certain predicates by calling external dedicated functions (for
-instance, the valuation of a proposition like \stmt{object1 isOn object2} is
-computed when required by calling a specific geometric reasoning module). In
+based on Bayesian Logic Networks. The integration with naive physics reasoning 
+has been studied~\cite{Kunze2011a}, and the automatic parsing of Web resources 
+in natural language has been also experimented with~\cite{tenorth10webinstructions}.
+
+\paragraph{Grounding} {\sc KnowRob} offers a mechanism called \emph{computables} 
+that allows to evaluate certain predicates by calling external dedicated functions 
+(for instance, the truth value of a proposition like \stmt{object1 isOn object2} 
+is computed at runtime by calling a specific geometric reasoning module). In
 combination with Prolog's lazy evaluation strategy, this supports a good
 scalability.
-
-Computables rely on various subsystem to evaluate. In particular, it relies on
-the CoP framework~\cite{Klank2009} and \emph{semantic maps}~\cite{Blodow2011}
+Computables rely on various subsystem to evaluate, for example the robot's 
+perception system~\cite{Klank2009} and \emph{semantic environment maps}~\cite{Blodow2011}
 for the recognition of objects and environment.
 
-%\paragraph{Integration with Decisional Layers} CRAM~\cite{Beetz2010}
-
-\paragraph{Experiments} {\sc KnowRob} has been deployed on several scenarii at
-the Technical University of Munich, on the PR2 robot and on a 2-arm custom
-mobile manipulator in the scope of the ``assistive kitchen''~\cite{Beetz2008}
-project. These experiments include retrieval and automated parsing of recipies
-from the Web, retrieval and manipulation of various kitchen tools, cooperation
-between two robots.
-
+\paragraph{Experiments} {\sc KnowRob} has been deployed in several scenarii 
+on the PR2 robot and on a 2-arm custom mobile manipulator in the scope of the 
+``assistive kitchen''~\cite{Beetz2008} project. These experiments included the
+retrieval and automated parsing of recipes from the Web, the retrieval and 
+manipulation of various kitchen tools, and cooperation between two robots.
+{\sc KnowRob} has been used as knowledge base for the RoboEarth project that 
+investigated the exchange of knowledge between robots via the 
+Internet~\cite{waibel11roboearth}.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{NKRL}
@@
 \emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
-While this language is developed since a long time by Zarri~\cite{Zarri1997,
-Zarri2008}, recent research direction include application to the robotic
-field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
-system, as it is primarily a language. However, it is used as the
-representation and reasoning mechanism for robots by Sabri et al.
+While this language has been developed for a long time by Zarri~\cite{Zarri1997,
+Zarri2008}, recent research aims at its application to robotics~\cite{Sabri2011}. 
+NKRL is not {\it per-se} a knowledge representation system, it is primarily 
+a language, but it is used as the representation and reasoning mechanism for 
+robots by Sabri et al.
 
@@
 two ontologies: an ontology of concepts $\Omega$ and an ontology of events
-$\Psi$. The ontology of events is made of action or situation templates.
-Templates are a set of predicates (MOVE, PRODUCE, RECEIVE, EXPERIENCE, BEHAVE,
-OWN and EXIST) associated to thematic roles. Grounding and reasoning with NKRL
+$\Psi$. The ontology of events is made up of templates for actions and situations.
+Templates are sets of predicates (MOVE, PRODUCE, RECEIVE, EXPERIENCE, BEHAVE,
+OWN and EXIST) associated with thematic roles. Grounding and reasoning with NKRL
 is based on template matching.
@@
 \paragraph{Experiments} The main scenario of development for NKRL-based robots
-is the Smart Home and monitoring of eldery people. Knowledge acquisition
-partially rely on ambient intelligence (RFID, pressure sensors in the chairs,
-etc.). The scenario is still being implementated.
-
+is a Smart Home that is monitoring elderly people. Knowledge acquisition
+partially relies on ambient intelligence (RFID, pressure sensors in the chairs,
+etc.). The scenario is still being implemented.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{OUR-K and OMRKF}
@@
 Framework~\cite{Suh2007} (OMRKF), is a knowledge representation system based on
-five inter-related \emph{classes} of knowledge (figure~\ref{fig|omrkf}). It
+five inter-related \emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It
 proposes a layered approach to knowledge representation that allows to 
-integrate the grounding process to the knowledge representation process. OUR-K
+integrate the grounding process and the knowledge representation process. OUR-K's
 knowledge model is implemented with a mix of Description Logics for the concept
@@
 
-    \caption{OUR-K organises knowledge into fives \emph{classes}, each composed
+    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
     of \emph{levels}. Figure based on~\cite{Lim2011}.}
@@
 Each level of knowledge is build as three stages of ontological realization: a
-\emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
+\emph{meta-concept} (the level itself, like ``temporal context'', ``behavior''
 or ``object feature''), a taxonomy of concepts inside this level (for instance
@@
 
-\paragraph{Representation} The environment is represented in OUR-K in the
-$spaces : Model$ knowledge level as a classical three layers mapping (metric,
-topological and semantic maps). Objects (in $objects : Model$) are localised in
-$spaces : Model$ through Voronoi nodes.
-
-The knowledge class $Context$ proposes an explicit statement of spatial context
-(mostly geometric relations between objects), temporal context and a more
-general \emph{high-level} context, inferred from spatial and temporal contexts.
-
-Finally, the $Activity$ knowledge class store compound actions in a HTN-like
-structure, exploited at run-time by a planner.
-
+\paragraph{Representation} The robot's environment is in OUR-K represented as 
+part of the $spaces : Model$ knowledge level as a classical three-layered map 
+(metric, topological and semantic map). Objects (in $objects : Model$) are 
+localized in the $spaces : Model$ through Voronoi nodes.
+% 
+The $Context$ knowledge class proposes an explicit statement of spatial context
+(mostly geometric relations between objects), temporal context, and a more
+general \emph{high-level} context inferred from spatial and temporal contexts.
+% 
+Finally, the $Activity$ knowledge class stores compound actions in an HTN-like
+structure that is to be exploited at run-time by a planner.
 \fxwarning{re-read the paper to improve the section...}
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{ORO}
@@
 
+\todo{nothing about ORO?}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{PEIS KR\&R}
@@
 
-{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim
-to binds autonomous robotics with ambient intelligence (network of sensors).
+{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that 
+aims at the integration of autonomous robots with ambient intelligence.
 \emph{PEIS} stands for \emph{Physically Embedded Intelligent System}: every
-robots or intelligent device in the environment is abstracted as a PEIS.
-
-Each PEIS physical component is running a \emph{PEIS Kernel} instance.
-Communication between instance relies on a custom P2P communication protocol.
-
+robots or intelligent device in the environment is abstracted as a PEIS. Each 
+PEIS physical component is running a \emph{PEIS Kernel} instance. Communication 
+between instances relies on a custom peer-to-peer communication protocol.
 The PEIS architecture allows for adding new abilities through software 
-components sharing the common \emph{tuple space}.
-
+components that are sharing the common \emph{tuple space} \todo{what is this 
+tuple space?}.
 We consider here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS
-KR\&R}, that includes symbolic representation and reasoning.
-
-\fxwarning{have a look to latest Daoutis paper!}
+KR\&R} which includes symbolic representation and reasoning.
+% 
+\fxwarning{have a look at the latest Daoutis paper!}
 
@@
 \paragraph{Knowledge model} The PEIS Knowledge representation system relies on
-the {\sc ResearchCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
-allows to represent first order logic sentences and has extensions for modal logics and higher order logics.
-
+the {\sc ResearchCyc} knowledge base and the {\sc CycL} language for representing 
+knowledge. The {\sc CycL} language allows to represent first-order logic 
+sentences and has extensions for modal logics and higher order logics.
+% 
 \fxwarning{Is modal logics and higher order logics actually used in PEIS?} 
-
+% 
 As a system relying on {\sc CycL}, contexts can be expressed as
@@
 \emph{microtheory} in which these statements are evaluated.
-
+% 
 \fxwarning{OWA/CWA?}
-
+% 
 \begin{figure}
@@
 \end{figure}
-
+% 
 The PEIS KR\&R system is deeply integrated to the general PEIS Ecology
@@
 
-\paragraph{Knowledge Acquisition} The primary source for knowledge acquisition
-is perception.  The PEIS ecosystem provides a SIFT-based object recogniser used
-in conjunction with ceiling cameras for object localisation.  Other perceptual
-modalities are available (like human tracking, ambient environment monitoring).
-
-A template-based natural language parsing system may also be used to add new
-assertions to the system.
-
-The system can ask the human for help to disambiguate between concept names.
-
-\paragraph{Anchoring} Daoutis et al. formalise the issue of anchoring as
+\paragraph{Knowledge Acquisition} The primary source of knowledge in the PEIS 
+system is perception. The PEIS ecosystem provides a SIFT-based object recognizer 
+used in conjunction with ceiling cameras for object localization.  Other 
+perceptual modalities are available, such as tracking of human motions or 
+ambient environment monitoring. A template-based natural language parsing 
+system can also be used to add new assertions to the knowledge base. The 
+system can ask a human for help to disambiguate between concept names.
+
+\paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
 finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
@@
 a set of percept's attributes, and $D(\Phi)$ the domain of these attributes.
-
-In the current implementation, object category (returned by the SIFT
-classifier), color, location, spatial relations (both topological -- \emph{at},
-\emph{near} -- and relative to the robot -- \emph{left}, \emph{behind}, etc.)
-and visibility are the five classes of extracted attributes.
-
-\paragraph{Integration in the robot architecture}
+% 
+In the current implementation, the object category (returned by the SIFT
+classifier), color, location, visibility and spatial relations (both topological 
+such as \emph{at} and \emph{near} and directional such as \emph{left of} or 
+\emph{behind}) are the five classes of extracted attributes.
+
+\paragraph{Integration into the robot architecture}
 \label{sect|peis-integration}
 
-The PEIS framework offers through the \emph{PEIS middleware} a practical way to
-insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
+The PEIS framework offers, through the \emph{PEIS middleware}, a practical way 
+to insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
 module can be seamlessly integrated into the PEIS ecosystem.
@@
 \paragraph{Experiments} Experiments involving PEIS take place in a Smart Home
-environment (\emph{PEIS Home}). The implemented case studies explore
+environment (the \emph{PEIS Home})and  explore
 dialogue-based interaction with the robot about known objects.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Synthesis: How Current Systems Tackle the Knowledge Representation
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
+\section{Synthesis: How Do Current Systems Tackle the Knowledge Representation
 Challenge?}
@@
 This section summarizes the different approaches to knowledge representation and
-reasoning that are surveyed in the previous section, and attempt to identify
+reasoning that are surveyed in the previous section, and attempts to identify
 research perspectives.
-
 First, the Table~\ref{table|contribution-by-systems} presents, system by system
@@
 have examined.
-
-Then, we go over again the nine ``challenges'' proposed by McCarthy and Roy (see
-section~\ref{sect|evaluation-literature}) to evaluate how far we are on the road
-to natural interaction and, to slightly paraphrase McCarthy, ``human-level
+Then, we revisit the nine ``challenges'' proposed by McCarthy and Roy (see
+Section~\ref{sect|evaluation-literature}) to evaluate how far we are on the 
+road to natural interaction and, to slightly paraphrase McCarthy, ``human-level
 robots''.
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{Logic formalisms, continuous world}
-
-McCarthy underlines the necessity of relying on mathematical logic, as the most
-fruitful formalism for machine intelligence. This is largely the case. Almost
-all the systems we have surveyed rely on logical formalisms, in several cases
-as a mix of declarative languages (often based on Decription Logics) and
-logical programming languages (Prolog, expectedly, but not only).
-
-The two exceptions (the CAST knowledge model and GSM) are however 
-interesting: CAST proposes a pervasive model of knowledge that is convincing from
-the grounding perspective (but likely suboptimal for deliberative tasks), and
-GSM encodes knowledge in an amodal (continuous, geometric) model while
-preserving features that are usually specific of symbolic models (like theory
-of mind or categorical knowledge).
+McCarthy underlines the necessity of relying on mathematical logic as the most
+fruitful formalism for machine intelligence. Almost all of the systems we have 
+surveyed rely on logical formalisms, in several cases a mix of declarative 
+languages such as Description Logics and logical programming languages such as 
+Prolog or Answer Set Programming.
+
+The two exceptions (the CAST knowledge model and GSM) are however interesting: 
+CAST proposes a pervasive model of knowledge that is convincing from the 
+grounding perspective (but likely suboptimal for deliberative tasks), and GSM 
+encodes knowledge in an amodal (continuous, geometric) model while preserving 
+features that are usually specific of symbolic models (like a theory of mind 
+or categorical knowledge).
 
@@
 attributes of the environment to insert them in the symbolic model ({\sc Tapas},
-Ke Jia, OUR-K, PEIS), thus skipping altogether an intermediate geometric model.
+Ke Jia, OUR-K, PEIS), thus skipping an intermediate geometric model altogether.
 ORO has tight (but unidirectional) links with an intermediate geometric
-representation (called {\sc Spark}), and {\sc KnowRob} adopt a top-down approach
-where local geometric models are set up on-demand to compute symbolic properties
+representation (called {\sc Spark}), and {\sc KnowRob} adopts a top-down approach
+where local geometric models are set up on demand to compute symbolic properties
 like relative locations.
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Management of uncertainty and approximation}
@@
 McCarthy also mentions the necessity to deal with \emph{approximate concepts
-and approximate theories} (that includes representing them and reasoning with
-them).
-
+and approximate theories}, which includes representing them and reasoning with
+them.
 The most notable attempt at modeling uncertainty down to the knowledge
@@
 difficult to overcome.
-
-It can be also noted that some effort do exist to provided probabilistic
+It can be also noted that some efforts do exist to provided probabilistic
 reasoning to Description Logics based systems (like the {\sc Pronto}
 reasoner~\cite{Klinov2008}), but this does not appear to be a major focus in the
-currently available tools.
-
+currently available tools. \todo{statistical relational models start to be 
+used in robotics for modeling spatial relations (De Raedt ICRA 14) or for 
+improving perception (Nyga ICRA 14)}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Non-monotonic reasoning}
+
+McCarthy assigns particular importance to non-monotonic reasoning. Systems 
+like Ke Jia explicitly tackle this issue. A monotonic system does not 
+theoretically allow for knowledge retraction during the reasoning process, 
+which is an important issue in a robotic context where the world model is 
+likely to change quite regularly. This is however only a practical issue
+if the reasoning process has to be \emph{continuously active} during the 
+robot's activity lifespan, which is rarely the case. It is often possible to
+stop the reasoner, alter the knowledge, and restart the inference process on
+a new domain.
+
+Non-monotonicity can be partially dealt with using appropriate time
+representation and reasoning methods (the reasoner then only takes into account
+statements that are set as valid for a given moment, as it is for instance used 
+in {\sc KnowRob}) or with dedicated reasoners for tasks requiring non-monotonic 
+reasoning like symbolic task planning \fxwarning{mention ORO + HATP}.
+Probabilistic reasoning also implicitly leads to non-monotonic
+reasoning by relying on a continuous description of the state of the
+world.
+
+Non-monotonicity is however of broader significance to the knowledge
+representation field, in particular for the representation of common-sense
+knowledge where a \emph{default} representation is currently sorely lacking.
+It is also related to what McCarthy calls \emph{elaboration tolerance}, the
+ability to extend \emph{on demand} the closed domain of interpretation for a
+given assertion to take new contextual knowledge into account.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Modeling of contexts}
+
+Both Roy and McCarthy highlight the importance of formalizing and reasoning 
+about contexts. Many of the KRS we have surveyed do at some point mention the 
+modeling of contexts, but no consistent interpretation, let alone theory, of 
+context management has clearly emerged. Several approaches for building 
+contextualized knowledge have been hinted in Section~\ref{sect|features} such 
+as symbolic environment interpretation, perspective taking and independent 
+mental states for each agent, grounded natural language resolution, and 
+self-awareness of an agent's own activities. Much remains to be done for a 
+robot to actually identify its current context as well as contexts that
+may be referred to.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
+\subsection{Reasoning about time, events and actions}
+
+While the representation of spatial roles (topology and placement with or without
+perspective-taking) is well studied and KRS usually integrate with spatial
+perception and reasoning components, the representation of time and its 
+integration into knowledge-based systems is less uniform.
+Two important tasks that require temporal reasoning are task planning and 
+sequence recognition (in particular, action recognition). Systems that directly
+integrate task planning (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
+fluents in {\sc KnowRob}) to represent time.
+Other approaches include storing snapshots of the complete knowledge base 
+(like in GSM) that are used to move back to past mental states of the robot 
+or of another agent. Integration of these techniques with memory mechanisms 
+(forgetting and reinforcement learning) remains mostly to be explored.
+Regarding events, several systems (GSM, NKRL, ORO) have adopted an {\em
+event-oriented} architecture where conditions (or templates) are used to
+trigger decisional and execution processes.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
+\subsection{Grounding in a multi-modal environment}
+
+Symbol grounding remains a focus of most of the developers of cognitive
+architectures for robots, and all the systems we have surveyed implement
+grounding strategies. Most of the systems
+adapt their grounding strategies to the sensing modality, so that grounding can
+not be considered as a single, well delimited process.
+
+Some systems (ORO, GSM) introduce an intermediate step in the grounding process
+by means of an amodal model of the environment that aggregates the perceptions 
+(or suppositions) in a single place. This enables geometric reasoning that 
+improves the observability of the system and takes all perceptual modalities 
+into account. The latter is particularly helpful for spatial perspective taking 
+in order to know what humans are looking at and where objects are.
+
+We note in addition that the development of high-level sensors is simplifying
+the grounding task. The most obvious example is the human tracking system
+provided with the low-cost Kinect family of devices that is now widely used
+in robotics labs. Not only does this system segment and compute the poses of 
+humans that enter its field of view, it also tracks them even in case they 
+are partially or completely occluded. This kind of high-level sensing device 
+does not remove the need of grounding one symbolic instance of a human with 
+the physical human, but the task is much simpler than it used to be.
+
+Many systems also tackle the difficult issue of natural-language processing.
+The CAST middleware, in particular, has been used in the European CoSy and CogX
+projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
+the interest in grounding verbal interaction in the cognitive robotics community, 
+and we are likely to reach important milestones in this field in the coming years.
+Merging other modalities (especially back-channel communication, deictic
+gestures and social gazes) also receives a lot of attention and is becoming more
+and more present as symbolic knowledge available to the control layers.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \begin{landscape}
@@
                                                                                & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 & +++ ({\em Dialogs} \cite{Lemaignan2011a})     &                               & + (template-based)             \\
-                                                                     & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{Nyga2009}})&                             &                                               &                               &                                \\
+                                                                     & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{tenorth10webinstructions}})&                             &                                               &                               &                                \\
                                                                          & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & ++ ({\em amodal model} \cite{Lemaignan2012})  & +++ (bottom-up)               & +++ \cite{Loutfi2008}          \\
@@
 
-\subsection{Non-monotonic reasoning}
-
-McCarthy gives a particular importance to non-monotonic reasoning. Systems like
-Ke Jia explicitly tackle this issue.
-
-A monotonic system does not theoretically allow for knowledge retraction during
-the reasoning, which is an important issue in the robotic context where the
-world model is likely to be often altered.  However it is a practical issue
-only if the reasoning process has to be \emph{continuous} during the whole
-robot's activity lifespan, which is rarely the case. It is often possible to
-stop the reasoner, alter the knowledge, and restart the inference process on a
-new domain.
-
-Non-monotonicity can also be partially dealt with with appropriate time
-representation and reasoning (the reasoner then only takes into account
-statements that are set as valid for a given moment) or with dedicated reasoners
-for tasks requiring non-monotonic reasoning like symbolic task planning
-\fxwarning{mention ORO + HATP}.
-
-Finally, probabilistic reasoning also implicitly leads to non-monotonic
-reasoning, by relying on a continuous description of the state of the
-world.
-
-Non-monotonicity is however of broader significance to the knowledge
-representation field, in particular for the representation of common-sense
-knowledge where \emph{default} representation is currently sorely lacking.
-
-It is also related to what McCarthy calls \emph{elaboration tolerance}: the
-ability to extend \emph{on demand} the closed domain of interpretation for a
-given assertion, to take into account new contextual knowledge.
-
-\subsection{Modeling of contexts}
-
-Both Roy and McCarthy mention the importance of formalizing and reasoning about
-contexts. Many of the KRS we have surveyed mention at some point the context
-modeling, but no consistent interpretation, let alone theory, of context
-management has clearly emerged.
-
-Several approaches for building contextualized knowledge have been however
-hinted in section~\ref{sect|features}: symbolic environment interpretation,
-perspective taking and independent mental state for each agent, grounded natural
-language resolution, self-awareness of its own activity. Much remains to be done
-for a robot to actually identify its current context as well as contexts that
-may be referred to.
-
-\subsection{Reasoning about time, events and actions}
-
-While representation of spatial roles (topology, placement, with or without
-perspective-taking) is well studied and KRS usually integrate with spatial
-perception and reasoning components, time representation and integration in
-knowledge systems is less uniform.
-
-The two main tasks that require time reasoning are task planning and sequence
-recognition (in particular, action recognition). Systems that directly
-integrate task planning (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
-fluents in {\sc KnowRob}) to represent time.
-
-Other approaches include storing snapshots of the knowledge state (like GSM),
-that is used to move back to past mental states (of the robot or of another
-agent).
-
-Integration of these techniques with memory mechanisms (forgetting and
-reinforcement learning) remains mostly to be explored.
-
-Regarding events, several systems (GSM, NKRL, ORO) have adopted an {\em
-event-oriented} architecture where conditions (or templates) are used to
-trigger decisional and execution processes.
-
-\subsection{Grounding in a multi-modal environment}
-
-Symbol grounding remains a focus for most of the developers of cognitive
-architectures for robots, and all the systems we have surveyed implement
-grounding strategies. Most of the systems
-adapt their grounding strategies to the sensing modality, so that grounding can
-not be considered as a single, well delimited process.
-
-Some systems (ORO, GSM) introduce an intermediate step in the grounding process
-by the mean of an amodal model of the environment that aggregates the
-perceptions (or suppositions) in a single place. This enables geometric
-reasoning that takes into account all the perceptual modalities (typically
-required for spatial perspective taking: we need to know where the humans are
-looking at, and also where are the objects), and also improve the observability
-of the system.
-
-We note in addition that the development of high-level sensors is simplifying
-the grounding task. The most obvious example is the human tracking system
-provided with the low-cost Kinect family of devices, and now used
-pervasively in the robotics lab. Not only this system segments and computes the
-pose of humans that enter its field of view, but it also tracks them (including
-in case they are partially or completely occluded). This kind of high level
-sensing device does not completely remove the need of grounding one symbolic
-instance of a human with the physical human, but the task is much simpler than
-it used to be.
-
-Many systems also tackle the difficult question of natural language processing.
-The CAST middleware, in particular, has been used in the European CoSy and CogX
-projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
-the interest for the grounding of verbal interaction within the cognitive
-robotics community, and we are likely to reach important milestones in this
-field in the coming years.
-
-Merging more modalities (especially, back-channel communication, deictic
-gestures and social gazes) also sparks a lot of attention, and is becoming more
-and more present as symbolic knowledge available to the control layers.
-
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Common-sense}
-
 McCarthy actually starts his list by affirming that intelligent systems must be
 able to \emph{``operate successfully in the common sense informatic
-situation''}. This question of the \emph{common-sense} is probably one of the
-toughest because what \emph{common-sense} mean is not very clear at first place
-(since, by definition, common-sense is, well, common-sense...). We feel,
+situation''}. This question of \emph{common sense} is probably one of the
+toughest because what \emph{common sense} means is not clear in the first place
+(since, by definition, common sense is, well, common sense...). We feel,
 however, that it is related to a diffuse cultural background, and one may even
-claim that, for a system to acquire common-sense reasoning is equivalent to
-solve the strong AI challenge (this is at least more or less McCarthy's opinion).
-
-Knowledge representation systems have usually a pragmatic approach to
-common-sense: they reuse knowledge bases stored on the Web. While {\sc KnowRob}
-is the only one explicitly working on common-sense reasoning, through
-physics-based reasoning, integration with Web databases or parsing of
-semi-structured Web documents), several other KRS rely on Web standards (OWL,
-OpenCyc) to represent their knowledge (Table~\ref{table|knowledge-sources} lists
-the sources of common-sense knowledge found in the surveyed systems). With the
-development of initiatives like OpenMind that encode with the same standards
-common-sense facts and rules, there is a strong potential for our robot to gain
+claim that for a system to acquire common sense reasoning is equivalent to
+solving the strong AI challenge (this is at least more or less McCarthy's opinion).
+
+Knowledge representation systems usually follow a pragmatic approach to
+common-sense knowledge by reusing existing knowledge bases, e.g. from the 
+Web. While {\sc KnowRob} is the only one explicitly working on common-sense 
+reasoning through physics-based reasoning, integration with Web databases and 
+parsing of semi-structured Web documents, several other KRS rely on Web 
+standards (OWL, OpenCyc) to represent their knowledge (Table~\ref{table|knowledge-sources} 
+lists the sources of common-sense knowledge found in the surveyed systems). With the
+development of initiatives like OpenMind\ref{TODO} that collect large amounts 
+of common-sense facts and rules, there is a strong potential for robots to gain
 common-sense knowledge from well-structured online resources in the coming
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Learning, representation of experience, introspection}
@@
 in conjunction with visual perception systems (mentioned for instance for the
-CAST project~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
-facts from informations automatically extracted from the Web~\cite{Nyga2009},
-can also be considered as a \emph{learning} system (it autonomously acquire
-knowledge). We lack however a formal study of learning strategies and
-techniques at the knowledge representation level.
-
-Representation and matching of past experience is a related topic. This ability
+CAST project~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its 
+pool of facts from informations automatically extracted from the 
+Web~\cite{tenorth10webinstructions}, can also be considered as a \emph{learning} 
+system (as one that autonomously acquires knowledge). We do however lack a 
+formal study of learning strategies and techniques at the knowledge 
+representation level.
+
+The representation and matching of past experience is a related topic. This ability
 is a key step for general action recognition, and is of particular importance
 for the robot to assess the state of the interaction with the human.
-
 While we already mentioned that several systems are able to reason on past
 states, we are not aware of existing implementation of algorithms to reflect on
-past experiences.
-
-This is itself related to introspection and meta-cognition: the shift towards
+past experiences. \todo{RACE, CRAM$_m$}
+
+This topic is related to introspection and meta-cognition: the shift towards
 explicit knowledge representation exemplified in the nine systems we have
-presented, has a major impact on the meta-cognitive capabilities of our robots.
+presented has major impact on the meta-cognitive capabilities of our robots.
 They can exhibit, manipulate and reason on their internal belief state (what we
-will call in the conclusion the \emph{cognitive observability}). We have still
-to discover all the possibilities that are open by this important cognitive
-ability.
-
+will later call the \emph{cognitive observability}). We still have to discover 
+all the possibilities that are open by this important cognitive ability.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 \subsection{Perspective-awareness}
@@
 Finally, Roy mentions \emph{the ability to take into account the human
-perspective}: this cognitive ability, that relates to the representation of
-different modalities, is explicitly advertised in ORO and GSM only: ORO
+perspective}. This cognitive ability that relates to the representation of
+different modalities is explicitly advertised in ORO and GSM only. ORO
 explicits the perspectives in different symbolic mental states while GSM
-recursively stores models for each agent.  We think however that several of the
-other systems could easily gain support for alternative belief models, leading
-the way to becoming perspective-aware KRS.
-
-\subsection{Conclusion: the Future Directions to Explore}
+recursively stores models for each agent.  We think however that several 
+of the other systems could easily gain support for alternative belief models, 
+leading the way to becoming perspective-aware KRS.
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Conclusion: Future Directions to Explore}
 

### 2014-04-21 letter-to-editor.tex
reworked text, added references

@@
 \usepackage{booktabs}
+\usepackage{color}
 
@@
 \begin{document}
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:} #1}}
 
@@
 Those systems or frameworks have rapidly grown in number over the last few
-years: Brachman and Levesque's book~\cite{brachman2004knowledge} in 2004, dozens
-of systems have emerged to fulfil the need for explicit, symbolic (and
+years: Since Brachman's and Levesque's book~\cite{brachman2004knowledge} in 
+2004, dozens of systems have emerged to fulfil the need for explicit, symbolic (and
 sub-symbolic) \emph{representation} and \emph{reasoning} on knowledge in
-cognitive architectures for robots.
+cognitive architectures for robots.\todo{MT: why do we survey only nine if there are dozens?}
 
 Because these systems have explored many different paths, from building upon purely
-logic languages, to proposing distributed, diffuse knowledge representation
+logic languages to proposing distributed, diffuse knowledge representation
 schemes; because more and more research groups that are looking for increased
-robot autonomy, look towards implementing their own framework for knowledge
+robot autonomy look into the implementation of their own framework for knowledge
 representation; because, also, no previous survey of these tools has ever been
-published to our best knowledge, we propose to write such a review of the
+published to the best of our knowledge, we propose to write such a review of the
 existing literature, starting with brief typology of main features to be found
-in those systems, followed by a systematic review of the existing tools.
+in those systems, followed by a systematic review of the existing approaches.
 
-This survey would be a joint work by Moritz Tenorth, from the University of
+This survey would be a joint work by Moritz Tenorth from the University of
 Bremen, Germany (Michael Beetz's group) and Séverin Lemaignan, currently working
@@
 suite for knowledge management for service robots. This framework has been
-deployed on many different robots, ranging from the Willow Garage's PR2 to
-[...?]. His different previous publications on this project have been cited over
-500 times, and he is regarded as an expert in this field.
+deployed on many different robots, ranging from Willow Garage's PR2 to several
+custom manipulation platforms, and is being used in several research groups 
+world-wide. His publications on this project have been cited over 800 times 
+and he is regarded as an expert in this field.
 
-Séverin Lemaignan is the author of the {\sc ORO} knowledge base, that has
+Séverin Lemaignan is the author of the {\sc ORO} knowledge base that has
 received special attention in the human-robot interaction community for being
-one of the first knowledge representation and reasoning system taking explicitly
-into account the cognitive models of the humans interacting with the robot. His
-platform has also been deployed on several robots, in at least five different
+one of the first knowledge representation and reasoning system that is explicitly
+taking the cognitive models of humans interacting with the robot into account. 
+His platform has also been deployed on several robots in at least five different
 laboratories. Séverin Lemaignan is regularly invited for talks on cognitive
@@
 \midrule
-ARMAR/Tapas &  Holzapfel, Waibel \par (Karlsruhe TH) &  \cite{Holzapfel2008}\\
+ARMAR/Tapas &  Holzapfel, Waibel \par (KIT Karlsruhe) &  \cite{Holzapfel2008}\\
 CAST Proxies &  Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) &  \cite{Jacobsson2008} \\
@@
 
-We hope that you may consider our proposal as interesting and relevant for a future issue of
-\emph{Transaction on Robotics}.
+We hope that you may consider our proposal as interesting and relevant for a 
+future issue of \emph{Transaction on Robotics}. If you require any additional 
+information, please feel free to contact us.
 
@@
 
-\ps{Séverin Lemaignan is with EPFL, CHILI Lab, Lausanne, Switzerland; Moritz Tenorth is with Institute for Artificial Intelligence, Universität Bremen, Germany.}
+\ps{Séverin Lemaignan is with the CHILI Lab, EPFL, Lausanne, Switzerland.\\ 
+    Moritz Tenorth is with the Institute for Artificial Intelligence, Universität Bremen, Germany.}
 

### 2014-04-28 letter-to-editor.tex
Final version letter to editor

@@
 years: Since Brachman's and Levesque's book~\cite{brachman2004knowledge} in 
-2004, dozens of systems have emerged to fulfil the need for explicit, symbolic (and
+2004, numerous systems have emerged to fulfil the need for explicit, symbolic (and
 sub-symbolic) \emph{representation} and \emph{reasoning} on knowledge in
-cognitive architectures for robots.\todo{MT: why do we survey only nine if there are dozens?}
+cognitive architectures for robots.
 
@@
 published to the best of our knowledge, we propose to write such a review of the
-existing literature, starting with brief typology of main features to be found
-in those systems, followed by a systematic review of the existing approaches.
+existing literature. We plan to start with a broad overview of the existing techniques,
+followed by a refined typology of the features to be found in knowledge
+representation and reasoning systems, eventually followed by an in-depth review
+of ten of these systems.
 
@@
 ARMAR/Tapas &  Holzapfel, Waibel \par (KIT Karlsruhe) &  \cite{Holzapfel2008}\\
-CAST Proxies &  Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) &  \cite{Jacobsson2008} \\
+CAST Proxies &  Wyatt, Hawes, Jacobsson, Kruijff \par (Brimingham Univ., DFKI Saarbrücken) &  \cite{jacobsson2008crossmodal} \\
 GOLOG & Levesque (Toronto Univ.) & \cite{levesque1997golog} \\
@@
 
-\closing{Yours Sincerly,}
+\closing{Yours Sincerely,}
 
-\ps{Séverin Lemaignan is with the CHILI Lab, EPFL, Lausanne, Switzerland.\\ 
-    Moritz Tenorth is with the Institute for Artificial Intelligence, Universität Bremen, Germany.}
+\ps{Séverin Lemaignan ({\tt severin.lemaignan@epfl.ch}) is with the CHILI Lab, EPFL, Lausanne, Switzerland.\\ 
+Moritz Tenorth ({\tt tenorth@cs.uni-bremen.de}) is with the Institute for Artificial Intelligence, Universität Bremen, Germany.}
 
